---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
```

# Reproducing the NoManagement Scenario

## NK Figure 2

To validate FVS runs, we reconstruct the table behind Figure 2. To do
this, we use the tool <https://automeris.io/wpd/> which aids in
estimating the tabular data behind a chart.

The project used to extract the table is in
[02_rScripts/AutoMeris_NK_Fig2.json]. The tool is imperfect, so we
apply some cleanup to produce an approximate table.

Note that all values are approximate. For example, all values should
align in 2005, but they do not; we ignore this, since small differences
are within the error bounds for parsing the chart.

```{r}
NK_Fig2_NoManagement <- read_csv(
    "../01_processedInputData/NK_Fig2_NoManagement.csv",
    col_names = c("Year","NoManagement"),
    # Read columns as double, since that is how they were written
    col_types = "dd"
  ) |>
  # Round columns to nearest integer, since that is what was intended
  mutate(Year=round(Year), NoManagement=round(NoManagement))
NK_Fig2_ClearcutHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ClearcutHigh.csv",
    col_names = c("Year","ClearcutHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutHigh=round(ClearcutHigh))
NK_Fig2_ClearcutLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ClearcutLow.csv",
    col_names = c("Year","ClearcutLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutLow=round(ClearcutLow))
NK_Fig2_ShelterwoodHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ShelterwoodHigh.csv",
    col_names = c("Year","ShelterwoodHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodHigh=round(ShelterwoodHigh))
NK_Fig2_ShelterwoodLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ShelterwoodLow.csv",
    col_names = c("Year","ShelterwoodLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodLow=round(ShelterwoodLow))
NK_Fig2_ITS_LowHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_LowHigh.csv",
    col_names = c("Year","ITS_LowHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowHigh=round(ITS_LowHigh))
NK_Fig2_ITS_LowLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_LowLow.csv",
    col_names = c("Year","ITS_LowLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowLow=round(ITS_LowLow))
NK_Fig2_ITS_HighHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_HighHigh.csv",
    col_names = c("Year","ITS_HighHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighHigh=round(ITS_HighHigh))
NK_Fig2_ITS_HighLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_HighLow.csv",
    col_names = c("Year","ITS_HighLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighLow=round(ITS_HighLow))

# Assemble a single table from 2005 - 2165
fig2 <- NK_Fig2_NoManagement |>
  full_join(NK_Fig2_ClearcutHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ClearcutLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighLow, by=join_by(Year)) |>
  filter(Year>=2005) |>
  # The last year in Fig2 is 2164; nudge it to 2165
  # to align with contemporary FVS runs.
  mutate(Year=if_else(Year==2164,2165,Year)) |>
  arrange(Year)

ggplot(
    data = melt(fig2, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("NK Fig2 Reconstruction") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 200)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,200,20))
```

## Configuring FVS

FVS Software with `Release date: 20240401` was used.

FIA data version:

```         
% sqlite3 SQLite_FIADB_ENTIRE.db 'select * from REF_FIADB_VERSION ORDER BY VERSION DESC limit 1;'
FIADB_1.9.2.00|RELEASE|THIS IS FIADB VERSION 9.2 PROD|FS_NIMS_FIADB_RMRS|2024-01-23 10:16:14|732862|||
```

## Stands

Stands can be found in the FVS UI using FVS_STAND_ID from the table
above.

-   Inventory Data Tables = FVS_StandInit_Plot
-   Variants = ne: Northeast
-   Find stand(s): = 239901700110

Clicking `Find` will then filter to the single matching stand, which can
be added to the run.

Repeat for each of the above stands.

## Time

FVS will not run prior to the inventory year; it's unclear what NK did
to get their chart to go back prior to 2005. We will instead use 2005 as
the common starting year.

NK ends in 2164; we adjust this to 2165 so that all cycles are 10 years
long. Note that FVS does not include the ending year unless it is at the
start of a new cycle, so we set the common ending year to 2166.

All events in NK take place on 10 year intervals, therefore we select
10 years as the reporting interval.

## Carbon

To set the carbon calculations to be metric (Metric tons of carbon per
hectare), go to Simulate -\> Components -\> Keywords, and select:

-   Extensions: Fire and Fuels Extension
-   Keywords: CarbCalc: Set the carbon accounting parameters.
-   Component title: CarbCalc: Metric
-   Biomass predictions: 0 = Use FFE biomass predictions
-   Units: 1 = Metric (metric tons carbon/hectare)
-   Note: Annual root decay rate (proportion per year) remains at its
    default value of 0.0425.

Select `Save in run` to apply to `Grp: All_FIA_Plots`.

## Run

The main output will be Carbon and fuels. Stand visualization, Tree
lists, and Inventory statistics may also be interesting.

## Automation

Do it in code:
<https://github.com/USDAForestService/ForestVegetationSimulator-Interface/tree/main/rFVS/R>

Instructions:
<https://github.com/USDAForestService/ForestVegetationSimulator-Interface/wiki/rFVS>

This documents such gems in the interface as:

> fvsGetRestartcode() Provides an R interface to Fortran function
> fvsGetRestartcode.

But, this example shows that they expect you to create your project in
the UI, then run it using the API:

```{r eval=FALSE}
setwd ("<FVS-interface>/rFVS/tests")
# Note: edit <open-fvs> to reflect your system. This directory string should be wherever 
# your keyword file and input data are located

# load the R functions using this method (if you don't have the rFVS package)
sdir = "../R"
# Note: this directory string should be wherever the rFVS functions are located
for (rf in dir (sdir)) source (paste(sdir,rf,sep="/"))

# or, if the rFVS package is installed, then just enter
library(rFVS)

# load the FVS library
fvsLoad("FVSie", bin="<open-fvs>/trunk/bin")
# Note: edit <open-fvs> to reflect your system. This directory string should 
# be wherever your FVS shared libraries are located

# specify the keyword file
fvsSetCmdLine("--keywordfile=iet01.key")

# Run the simulation as would be done using FVS as a program
fvsRun()
```

## Results

### By Plot

```{r}
NoManagement_Carbon_ByPlot <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByPlot.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  mutate(Standing_Carbon = Aboveground_Live + Standing_Dead) |>
  select(StandID, Year, Total_Stand_Carbon,
         Aboveground_Carbon, Standing_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead)
knitr::kable(NoManagement_Carbon_ByPlot)
```

```{r}
NoManagement_Carbon_ByPlot_ByYear <- NoManagement_Carbon_ByPlot |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Stand_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  # Join in the NoManagement scenario from fig2 for comparison
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

write_csv(NoManagement_Carbon_ByPlot_ByYear, "../03_outputs/NoManagement_Carbon_ByPlot_ByYear.csv")

ggplot(
    data = melt(NoManagement_Carbon_ByPlot_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Plot") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,300,20))
```

Examine residuals:

```{r}
Plots_RMSE <- tibble(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Total_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Aboveground_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Standing_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Standing_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Plots_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Plot") +
  geom_col()
```

Note that NK describe using Aboveground Carbon for their chart, but the
best fit is from Standing Carbon, with RMSE \~= 7.

```{r}
Plots_RMSE |> select(Standing_Carbon)
```

### By Subplot

```{r}
NoManagement_Carbon_BySubplot_ByYear <- read_csv(
    "../03_outputs/NoManagement_Carbon_BySubplot.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Total_Carbon = Total_Stand_Carbon) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  mutate(Standing_Carbon = Aboveground_Live + Standing_Dead) |>
  select(StandID, Year, Total_Carbon,
         Aboveground_Carbon, Standing_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead) |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

ggplot(
    data = melt(NoManagement_Carbon_BySubplot_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Subplot") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,300,20))
```

Examine residuals:

```{r}
Subplots_RMSE <- tibble(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Total_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Aboveground_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Standing_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Standing_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Subplots_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Subplot") +
  geom_col()
```

Note that all RMSE for the subplot projections are significantly worse
than for the plot projections.

### By Condition

Plot 360304303966 has 3 conditions; we include all three.

```{r}
NoManagement_Carbon_ByCondition_ByYear <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByCondition.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Total_Carbon = Total_Stand_Carbon) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  mutate(Standing_Carbon = Aboveground_Live + Standing_Dead) |>
  select(StandID, Year, Total_Carbon,
         Aboveground_Carbon, Standing_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead) |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

ggplot(
    data = melt(NoManagement_Carbon_ByCondition_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Condition") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,300,20))
```

Examine residuals:

```{r}
Conditions_RMSE <- tibble(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Total_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Aboveground_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  )),
  Standing_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Standing_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Conditions_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Condition") +
  geom_col()
```

RMSE for Standing Carbon by Condition is \~8, which is close to that for
Standing Carbon by Plot, but Standing Carbon by Plot has the lowest
RMSE:

```{r}
Conditions_RMSE |> select(Standing_Carbon)
```
