---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r include=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
```

# Reproducing the NoManagement Scenario

## Reconstructing NK Figure 2

To validate FVS runs, we reconstruct the table behind Figure 2. To do
this, we use the tool <https://automeris.io/wpd/> which aids in
estimating the tabular data behind a chart.

The project used to extract the table is in
[02_rScripts/AutoMeris_NK_Fig2.json]. The tool is imperfect, so we apply
some cleanup to produce an approximate table.

Note that all values are approximate. For example, all values should
align in 2005, but they do not; we ignore this, since small differences
are within the error bounds for parsing the chart.

```{r}
NK_Fig2_NoManagement <- read_csv(
    "../01_processedInputData/NK_Fig2_NoManagement.csv",
    col_names = c("Year","NoManagement"),
    # Read columns as double, since that is how they were written
    col_types = "dd"
  ) |>
  # Round columns to nearest integer, since that is what was intended
  mutate(Year=round(Year), NoManagement=round(NoManagement))
NK_Fig2_ClearcutHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ClearcutHigh.csv",
    col_names = c("Year","ClearcutHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutHigh=round(ClearcutHigh))
NK_Fig2_ClearcutLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ClearcutLow.csv",
    col_names = c("Year","ClearcutLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutLow=round(ClearcutLow))
NK_Fig2_ShelterwoodHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ShelterwoodHigh.csv",
    col_names = c("Year","ShelterwoodHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodHigh=round(ShelterwoodHigh))
NK_Fig2_ShelterwoodLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ShelterwoodLow.csv",
    col_names = c("Year","ShelterwoodLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodLow=round(ShelterwoodLow))
NK_Fig2_ITS_LowHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_LowHigh.csv",
    col_names = c("Year","ITS_LowHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowHigh=round(ITS_LowHigh))
NK_Fig2_ITS_LowLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_LowLow.csv",
    col_names = c("Year","ITS_LowLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowLow=round(ITS_LowLow))
NK_Fig2_ITS_HighHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_HighHigh.csv",
    col_names = c("Year","ITS_HighHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighHigh=round(ITS_HighHigh))
NK_Fig2_ITS_HighLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_HighLow.csv",
    col_names = c("Year","ITS_HighLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighLow=round(ITS_HighLow))

# Assemble a single table from 2005 - 2165
fig2 <- NK_Fig2_NoManagement |>
  full_join(NK_Fig2_ClearcutHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ClearcutLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighLow, by=join_by(Year)) |>
  filter(Year>=2005) |>
  # The last year in Fig2 is 2164; nudge it to 2165
  # to align with contemporary FVS runs.
  mutate(Year=if_else(Year==2164,2165,Year)) |>
  arrange(Year)

ggplot(
    data = melt(fig2, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("NK Fig2 Reconstruction") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 200)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,200,20))
```

## Configuring FVS

FVS Software with `Release date: 20240401` was used.

FIA data version:

```         
% sqlite3 SQLite_FIADB_ENTIRE.db 'select * from REF_FIADB_VERSION ORDER BY VERSION DESC limit 1;'
FIADB_1.9.2.00|RELEASE|THIS IS FIADB VERSION 9.2 PROD|FS_NIMS_FIADB_RMRS|2024-01-23 10:16:14|732862|||
```

### Stands

Stands can be found in the FVS UI using FVS_STAND_ID from the table
above.

-   Inventory Data Tables = FVS_StandInit_Plot
-   Variants = ne: Northeast
-   Find stand(s): = 239901700110

Clicking `Find` will then filter to the single matching stand, which can
be added to the run.

Repeat for each of the above stands.

### Time

FVS will not run prior to the inventory year; it's unclear what NK did
to get their chart to go back prior to 2005. We will instead use 2005 as
the common starting year.

NK ends in 2164; we adjust this to 2165 so that all cycles are 10 years
long. Note that FVS does not include the ending year unless it is at the
start of a new cycle, so we set the common ending year to 2166.

All events in NK take place on 10 year intervals, therefore we select 10
years as the reporting interval.

### Regeneration

NK regeneration rates are in NK Table 4, in seedlings per hectare.

In addition to scenario-specific regeneration, NK 2.4 states,
"Background regenration rates (intermediate to shade tolerant species
only), input at 10 year intervals, emulated natural regeneration within
stands, independent of forest management activities."

```{r}
table4 <- read_csv("../00_rawData/NK_Table_4.csv")

# Table4 uses scientific names; translate to get FVS Species Code
fvs_species_codes <-
  read_csv("../01_processedInputData/FVSne_Overview_Table_3.2.1.csv") |>
  select("Scientific Name", "Species Code", "Common Name")

# table4 has one column per species. We wish to add observations per species
# with different types, so pivot the table to have one row per species.
table4_rot <- table4 |>
  pivot_longer(cols = !`Management scenario`) |>
  pivot_wider(names_from = `Management scenario`) |>
  rename(`Scientific Name`=name) |>
  left_join(fvs_species_codes, by=join_by(`Scientific Name`))

# Values in NK are seedlings per hectare; FVS needs seedlings per acre
hectare_to_acre <- 2.471
table4_rot$Clearcut <- round(table4_rot$Clearcut * hectare_to_acre)
table4_rot$Shelterwood <- round(table4_rot$Shelterwood * hectare_to_acre)
table4_rot$`ITS_Low Retention` <- round(table4_rot$`ITS_Low Retention` * hectare_to_acre)
table4_rot$`ITS_High Retention` <- round(table4_rot$`ITS_High Retention` * hectare_to_acre)
table4_rot$Background <- round(table4_rot$Background * hectare_to_acre)

knitr::kable(
  table4_rot |>
  select(`Scientific Name`, Background, `Species Code`, `Common Name`) |>
  arrange(desc(Background))
)
```

In FVS, go to Simulate -\> Components -\> Management From Categories,
select "Planting & Natural Regeneration" From Components, select
"Plant/Natural with Partial Estab Model" Component title: "Baseline
Regen" Schedule the date of disturbance: Schedule by condition Create a
condition: Every cycle, every even cycle, ... Condition title: Every
cycle Years before condition can become true again: 0 The modulus of
cycle number: 1 = every cycle Number of years after condition is found
true: 0 Sprouting: On

Note that this leaves Percent survival at 100, and average age and
height empty.

In freeform, this looks like:

```         
Estab              0
Sprout
Natural            0        SM      1221      100.                             0
Natural            0        AB       610      100.                             0
Natural            0        EH       153      100.                             0
Natural            0        RS       153      100.                             0
Natural            0        YB       153      100.                             0
Natural            0        RM       153      100.                             0
```

Switch to freeform and add all the rows, fix the titles.

Select `Save in run` to apply to `Grp: All_FIA_Plots`.


### Carbon

To set the carbon calculations to be metric (Metric tons of carbon per
hectare), go to Simulate -\> Components -\> Keywords, and select:

-   Extensions: Fire and Fuels Extension
-   Keywords: CarbCalc: Set the carbon accounting parameters.
-   Component title: CarbCalc: Metric
-   Biomass predictions: 0 = Use FFE biomass predictions
-   Units: 1 = Metric (metric tons carbon/hectare)
-   Note: Annual root decay rate (proportion per year) remains at its
    default value of 0.0425.

Select `Save in run` to apply to `Grp: All_FIA_Plots`.

## Run FVS

The main output will be Carbon and fuels. Stand visualization, Tree
lists, and Inventory statistics may also be interesting.

## Automation

Do it in code:
<https://github.com/USDAForestService/ForestVegetationSimulator-Interface/tree/main/rFVS/R>

Instructions:
<https://github.com/USDAForestService/ForestVegetationSimulator-Interface/wiki/rFVS>

This documents such gems in the interface as:

> fvsGetRestartcode() Provides an R interface to Fortran function
> fvsGetRestartcode.

But, this example shows that they expect you to create your project in
the UI, then run it using the API:

```{r eval=FALSE}
setwd ("<FVS-interface>/rFVS/tests")
# Note: edit <open-fvs> to reflect your system. This directory string should be wherever 
# your keyword file and input data are located

# load the R functions using this method (if you don't have the rFVS package)
sdir = "../R"
# Note: this directory string should be wherever the rFVS functions are located
for (rf in dir (sdir)) source (paste(sdir,rf,sep="/"))

# or, if the rFVS package is installed, then just enter
library(rFVS)

# load the FVS library
fvsLoad("FVSie", bin="<open-fvs>/trunk/bin")
# Note: edit <open-fvs> to reflect your system. This directory string should 
# be wherever your FVS shared libraries are located

# specify the keyword file
fvsSetCmdLine("--keywordfile=iet01.key")

# Run the simulation as would be done using FVS as a program
fvsRun()
```

## Carbon Projection

### By Plot

```{r}
NoManagement_Carbon_ByPlot <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByPlot.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  mutate(Standing_Carbon = Aboveground_Live + Standing_Dead) |>
  select(StandID, Year, Total_Stand_Carbon,
         Aboveground_Carbon, Standing_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead)
knitr::kable(NoManagement_Carbon_ByPlot)
```

```{r}
NoManagement_Carbon_ByPlot_ByYear <- NoManagement_Carbon_ByPlot |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Stand_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  # Join in the NoManagement scenario from fig2 for comparison
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

write_csv(NoManagement_Carbon_ByPlot_ByYear, "../03_outputs/NoManagement_Carbon_ByPlot_ByYear.csv")

ggplot(
    data = melt(NoManagement_Carbon_ByPlot_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Plot") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,300,20))
```

Examine residuals:

```{r}
Plots_RMSE <- data.frame(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Total_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Aboveground_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Standing_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Standing_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Plots_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Plot") +
  geom_col()
```

Note that NK describe using Aboveground Carbon for their chart, but the
best fit is from Standing Carbon, with RMSE \~= 12.

```{r}
Plots_RMSE |> select(Standing_Carbon)
```

### By Subplot

```{r}
NoManagement_Carbon_BySubplot_ByYear <- read_csv(
    "../03_outputs/NoManagement_Carbon_BySubplot.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Total_Carbon = Total_Stand_Carbon) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  mutate(Standing_Carbon = Aboveground_Live + Standing_Dead) |>
  select(StandID, Year, Total_Carbon,
         Aboveground_Carbon, Standing_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead) |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

ggplot(
    data = melt(NoManagement_Carbon_BySubplot_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Subplot") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,300,20))
```

Examine residuals:

```{r}
Subplots_RMSE <- data.frame(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Total_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Aboveground_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Standing_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Standing_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Subplots_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Subplot") +
  geom_col()
```

Note that all RMSE for the subplot projections are significantly worse
than for the plot projections.

### By Condition

Plot 360304303966 has 3 conditions; we include all three.

```{r}
NoManagement_Carbon_ByCondition_ByYear <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByCondition.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Total_Carbon = Total_Stand_Carbon) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  mutate(Standing_Carbon = Aboveground_Live + Standing_Dead) |>
  select(StandID, Year, Total_Carbon,
         Aboveground_Carbon, Standing_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead) |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

ggplot(
    data = melt(NoManagement_Carbon_ByCondition_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Condition") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,300,20))
```

Examine residuals:

```{r}
Conditions_RMSE <- data.frame(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Total_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Aboveground_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  )),
  Standing_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Standing_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Conditions_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Condition") +
  geom_col()
```

RMSE for Standing Carbon by Condition is \~15, which is close to that for
Standing Carbon by Plot, but Standing Carbon by Plot has the lowest
RMSE:

```{r}
Conditions_RMSE |> select(Standing_Carbon)
```
## Carbon Measurement

### Plot surveys

When were the plots re-surveyed?

```{r}
nk_to_fia <- read_csv("../03_outputs/nk_to_fia.csv", col_types = "ciiiiiiiii") |>
  select(STATECD, UNITCD, COUNTYCD, PLOT)

fia = DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')
fia_cond = tbl(fia, 'COND')

matching_cond <- fia_cond |>
  right_join(
    nk_to_fia,
    by=join_by(STATECD, UNITCD, COUNTYCD, PLOT),
    copy=TRUE
  ) |>
  filter(INVYR >= 1999) |>
  collect()

DBI::dbDisconnect(fia)

knitr::kable(matching_cond)
```

```{r}
# Create ordinals for plots
#numbered_plots <- matching_cond |>
#  distinct(STATECD, UNITCD, COUNTYCD, PLOT, CONDID.x) |>
#  mutate(PLOT_NUM = row_number())

numbered_plots <- matching_cond |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT, CONDID) |>
  summarise(INVYR = min(INVYR), .groups = "keep") |>
  ungroup() |>
  arrange(INVYR) |>
  mutate(PLOT_NUM = row_number()) |>
  select(PLOT_NUM, STATECD, UNITCD, COUNTYCD, PLOT, CONDID)

inventory_years <- matching_cond |>
  left_join(numbered_plots, by=join_by(STATECD, UNITCD, COUNTYCD, PLOT, CONDID)) |>
  group_by(PLOT_NUM, INVYR)

ggplot(
    data = inventory_years,
    mapping = aes(x = INVYR, y = PLOT_NUM)
  ) +
  geom_point()

```
Some additional fields that are likely to be interesting:
- DSTRBCD1/2/3 - Disturbance codes - anything except harvest
Code Description
0 No visible disturbance.
10 Insect damage.
11 Insect damage to understory vegetation.
12 Insect damage to trees, including seedlings and saplings.
20 Disease damage.
21 Disease damage to understory vegetation.
22 Disease damage to trees, including seedlings and saplings.
30 Fire damage (from crown and ground fire, either prescribed or natural).
31 Ground fire damage.
32 Crown fire damage.
40 Animal damage.
41 Beaver (includes flooding caused by beaver).
42 Porcupine.
43 Deer/ungulate.
44 Bear (core optional ).
45 Rabbit (core optional ).
46 Domestic animal/livestock (includes grazing).
50 Weather damage.
51 Ice.
52 Wind (includes hurricane, tornado).
53 Flooding (weather induced).
54 Drought.
60 Vegetation (suppression, competition, vines).
70 Unknown / not sure / other.
80 Human-induced damage - any significant threshold of human-caused damage
not described in the DISTURBANCE codes or in the TREATMENT codes.
90 Geologic disturbances.
91 Landslide.
92 Avalanche track.
93 Volcanic blast zone.
94 Other geologic event.
95 Earth movement / avalanches.

- TRTCD1/2/3 - Treatment code - harvest

Code Description
00 No observable treatment.
10 Cutting - The removal of one or more trees from a stand.
20 Site preparation - Clearing, slash burning, chopping, disking, bedding, or other
practices clearly intended to prepare a site for either natural or artificial
regeneration.
30 Artificial regeneration - Following a disturbance or treatment (usually cutting), a new
stand where at least 50 percent of the live trees present resulted from planting or
direct seeding.
40 Natural regeneration - Following a disturbance or treatment (usually cutting), a new
stand where at least 50 percent of the live trees present (of any size) were
established through the growth of existing trees and/or natural seeding or
sprouting.
50 Other silvicultural treatment - The use of fertilizers, herbicides, girdling, pruning, or
other activities (not covered by codes 10-40) designed to improve the commercial
value of the residual stand; or chaining, which is a practice used on woodlands to
encourage wildlife forage.

```{r}
# Was anybody not surveyed?
nonsampled <- matching_cond |>
  filter(COND_NONSAMPLE_REASN_CD > 0) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, COND_NONSAMPLE_REASN_CD)
knitr::kable(nonsampled)
```

Ooh, shoot - Reason 2 is nonforest land; what's up with that?


```{r}
# Who had disturbance?
disturbed <- matching_cond |>
  filter(DSTRBCD1 > 0 | DSTRBCD2 > 0 | DSTRBCD3 > 0) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, DSTRBCD1, DSTRBCD2, DSTRBCD3)
knitr::kable(disturbed)
```

```{r}
# Who had treatment?
treated <- matching_cond |>
  filter(TRTCD1 > 0 | TRTCD2 > 0 | TRTCD3 > 0) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, TRTCD1, TRTCD2, , TRTCD3)
knitr::kable(treated)
```

Alright, we can combine these three (skip, disturb, harvest) into a single
table.
```{r}
plot_numbers = numbered_plots |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT, CONDID, .groups = "keep") |>
  summarize(PLOT_NUM = min(PLOT_NUM), .groups = "keep")
  

survey_status <- matching_cond |>
  mutate(STATUS = if_else(COND_NONSAMPLE_REASN_CD > 0, "Skipped", NA)) |>
  mutate(STATUS = if_else(!is.na(TRTCD1) & TRTCD1 > 0, "Harvested", STATUS)) |>
  mutate(STATUS = if_else(!is.na(DSTRBCD1) & DSTRBCD1 > 0, "Disturbed", STATUS)) |>
  mutate(STATUS = if_else(is.na(STATUS), "Grown", STATUS)) |>
  mutate(GROWN = if_else(STATUS == "Grown", INVYR, NA)) |>
  mutate(SKIPPED = if_else(STATUS == "Skipped", INVYR, NA)) |>
  mutate(HARVESTED = if_else(STATUS == "Harvested", INVYR, NA)) |>
  mutate(DISTURBED = if_else(STATUS == "Disturbed", INVYR, NA)) |>
  left_join(plot_numbers, by = join_by(STATECD, UNITCD, COUNTYCD, PLOT, CONDID)) |>
  select(
    PLOT_NUM, GROWN, HARVESTED, DISTURBED, SKIPPED
  )

ggplot(
    data = melt(survey_status, id.vars = "PLOT_NUM", variable.name="Status", value.name = "Year") |>
      filter(!is.na(Year)),
    mapping = aes(x = Year, y = PLOT_NUM, shape = Status, color = Status)
  ) +
  ggtitle("Plot Inventory Status") +
  theme(legend.title = element_blank()) +
  geom_point()
```


