---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Pre-load libraries that issue warnings on startup so these
# warnings don't corrupt the text later.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(dbplyr))
```

# Reproducing the NoManagement Scenario

## Reconstructing NK Figure 2

To validate FVS runs, we reconstruct the table behind Figure 2. To do
this, we use the tool <https://automeris.io/wpd/> which aids in
estimating the tabular data behind a chart.

The project used to extract the table is in
[02_rScripts/AutoMeris_NK_Fig2.json]. The tool is imperfect, so we apply
some cleanup to produce an approximate table.

Note that all values are approximate. For example, all values should
align in 2005, but they do not; we ignore this, since small differences
are within the error bounds for parsing the chart.

```{r}
NK_Fig2_NoManagement <- read_csv(
    "../01_processedInputData/NK_Fig2_NoManagement.csv",
    col_names = c("Year","NoManagement"),
    # Read columns as double, since that is how they were written
    col_types = "dd"
  ) |>
  # Round columns to nearest integer, since that is what was intended
  mutate(Year=round(Year), NoManagement=round(NoManagement))
NK_Fig2_ClearcutHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ClearcutHigh.csv",
    col_names = c("Year","ClearcutHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutHigh=round(ClearcutHigh))
NK_Fig2_ClearcutLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ClearcutLow.csv",
    col_names = c("Year","ClearcutLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutLow=round(ClearcutLow))
NK_Fig2_ShelterwoodHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ShelterwoodHigh.csv",
    col_names = c("Year","ShelterwoodHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodHigh=round(ShelterwoodHigh))
NK_Fig2_ShelterwoodLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ShelterwoodLow.csv",
    col_names = c("Year","ShelterwoodLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodLow=round(ShelterwoodLow))
NK_Fig2_ITS_LowHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_LowHigh.csv",
    col_names = c("Year","ITS_LowHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowHigh=round(ITS_LowHigh))
NK_Fig2_ITS_LowLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_LowLow.csv",
    col_names = c("Year","ITS_LowLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowLow=round(ITS_LowLow))
NK_Fig2_ITS_HighHigh <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_HighHigh.csv",
    col_names = c("Year","ITS_HighHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighHigh=round(ITS_HighHigh))
NK_Fig2_ITS_HighLow <- read_csv(
    "../01_processedInputData/NK_Fig2_ITS_HighLow.csv",
    col_names = c("Year","ITS_HighLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighLow=round(ITS_HighLow))

# Assemble a single table from 2005 - 2165
fig2 <- NK_Fig2_NoManagement |>
  full_join(NK_Fig2_ClearcutHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ClearcutLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighLow, by=join_by(Year)) |>
  filter(Year>=2005) |>
  # The last year in Fig2 is 2164; nudge it to 2165
  # to align with contemporary FVS runs.
  mutate(Year=if_else(Year==2164,2165,Year)) |>
  arrange(Year)

ggplot(
    data = melt(fig2, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("NK Fig2 Reconstruction") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 200)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,200,20))
```

## Configuring FVS

FVS Software with `Release date: 20240401` was used.

FIA data version:

```         
% sqlite3 SQLite_FIADB_ENTIRE.db 'select * from REF_FIADB_VERSION ORDER BY VERSION DESC limit 1;'
FIADB_1.9.2.00|RELEASE|THIS IS FIADB VERSION 9.2 PROD|FS_NIMS_FIADB_RMRS|2024-01-23 10:16:14|732862|||
```

### Stands

Stands can be found in the FVS UI using FVS_STAND_ID from the table
above.

-   Inventory Data Tables = FVS_StandInit_Plot
-   Variants = ne: Northeast
-   Groups: All_FIA_Plots
-   Find stand(s): = 239901700110

Clicking `Find` will then filter to the single matching stand, which can
be added to the run.

Repeat for each of the above stands.

### Time

FVS will not run prior to the inventory year; it's unclear what NK did
to get their chart to go back prior to 2005. We will instead use 2005 as
the common starting year.

NK ends in 2164; we adjust this to 2165 so that all cycles are 10 years
long. Note that FVS does not include the ending year unless it is at the
start of a new cycle, so we set the common ending year to 2166.

All events in NK take place on 10 year intervals, therefore we select 10
years as the reporting interval.

### Regeneration

NK regeneration rates are in NK Table 4, in seedlings per hectare.

In addition to scenario-specific regeneration, NK 2.4 states,
"Background regeneration rates (intermediate to shade tolerant species
only), input at 10 year intervals, emulated natural regeneration within
stands, independent of forest management activities."

This background regeneration rate is in NK Table 4.

```{r}
table4 <- read_csv(
  "../00_rawData/NK_Table_4.csv",
  col_types = cols(
    `Management scenario` = col_character(),
    .default = col_number(),
  )
)

# Table4 uses scientific names; translate to get FVS Species Code
fvs_species_codes <- read_csv(
    "../01_processedInputData/FVSne_Overview_Table_3.2.1.csv",
    col_types = "iiciccc"
  ) |>
  select("Scientific Name", "Species Code", "Common Name")

# table4 has one column per species. We wish to add observations per species
# with different types, so pivot the table to have one row per species.
table4_rot <- table4 |>
  pivot_longer(cols = !`Management scenario`) |>
  pivot_wider(names_from = `Management scenario`) |>
  rename(`Scientific Name`=name) |>
  left_join(fvs_species_codes, by=join_by(`Scientific Name`))

# Values in NK are seedlings per hectare; FVS needs seedlings per acre
hectare_to_acre <- 2.471
table4_rot$Clearcut <- round(table4_rot$Clearcut * hectare_to_acre)
table4_rot$Shelterwood <- round(table4_rot$Shelterwood * hectare_to_acre)
table4_rot$`ITS_Low Retention` <- round(table4_rot$`ITS_Low Retention` * hectare_to_acre)
table4_rot$`ITS_High Retention` <- round(table4_rot$`ITS_High Retention` * hectare_to_acre)
table4_rot$Background <- round(table4_rot$Background * hectare_to_acre)

knitr::kable(
  table4_rot |>
  select(`Scientific Name`, Background, `Species Code`, `Common Name`) |>
  arrange(desc(Background))
)
```

In FVS, go to Simulate -\> Components -\> Management From Categories,
select "Planting & Natural Regeneration" From Components, select
"Plant/Natural with Partial Estab Model"

- Component title: "Baseline Regen"
- Schedule the date of disturbance: Schedule by condition
- Create a condition: Every cycle, every even cycle, ...
- Condition title: Every cycle
- Years before condition can become true again: 0
- The modulus of cycle number: 1 = every cycle
- Number of years after condition is found true: 0
- Sprouting: On

Note that this leaves Percent survival at 100, and average age and
height empty.

In freeform, this looks like:

```         
Estab              0
Sprout
Natural            0        SM      1221      100.                             0
Natural            0        AB       610      100.                             0
Natural            0        EH       153      100.                             0
Natural            0        RS       153      100.                             0
Natural            0        YB       153      100.                             0
Natural            0        RM       153      100.                             0
```

Switch to freeform and add all the rows, fix the titles.

Select `Save in run` to apply to `Grp: All_FIA_Plots`.

### Carbon

To set the carbon calculations to be metric (Metric tons of carbon per
hectare), go to Simulate -\> Components -\> Keywords, and select:

-   Extensions: Fire and Fuels Extension
-   Keywords: CarbCalc: Set the carbon accounting parameters.
-   Component title: CarbCalc: Metric
-   Biomass predictions: 0 = Use FFE biomass predictions
-   Units: 1 = Metric (metric tons carbon/hectare)
-   Note: Annual root decay rate (proportion per year) remains at its
    default value of 0.0425.

Select `Save in run` to apply to `Grp: All_FIA_Plots`.

## Run FVS

The main output will be Carbon and fuels. Stand visualization, Tree
lists, and Inventory statistics may also be interesting.


## Carbon Projection

### By Plot

Read FVS output for carbon storage by plot. FVS column names are a bit
idiosyncratic, so clean those up.

NK states that Fig 2 sums carbon from aboveground live, standing dead,
and down dead. For comparison, we also compute standing carbon as
aboveground live plus standing dead.

```{r}
NoManagement_Carbon_ByPlot <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByPlot.csv",
    col_types = cols(
      StandID = col_character()
    )
  ) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  select(StandID, Year, Total_Stand_Carbon, Aboveground_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead)
knitr::kable(NoManagement_Carbon_ByPlot)
```

Plot these against NK Fig 2 "NoManagement" scenario. NK Fig 2 has three
different scales on the horizontal axis:

1.  From 1995 to 2005, major ticks are every 5 years
2.  From 2005 to 2155, major ticks are every 10 years
3.  From 2155 to 2164, major ticks are every 9 years

NK states that data prior to 2005 is projected from mean growth rate. To
simplify the chart, we omit data prior to 2005, place major tick marks
every 10 years, and nudge the end date from 2164 to 2165.

Note that NK Fig2 year 2164 shows an anomalous reduction in carbon
storage in the NoManagement scenario; it might be the right thing to do
to omit 2164 / 2165 as well.

```{r}
NoManagement_Carbon_ByPlot_ByYear <- NoManagement_Carbon_ByPlot |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Stand_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Standing_Carbon=mean(Standing_Carbon),
    Aboveground_Live=mean(Aboveground_Live),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  # Join in the NoManagement scenario from fig2 for comparison
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

#write_csv(NoManagement_Carbon_ByPlot_ByYear, "../03_outputs/NoManagement_Carbon_ByPlot_ByYear.csv")

ggplot(
    data = melt(NoManagement_Carbon_ByPlot_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Plot") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 320)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,320,20))
```

From the chart it seems evident that standing carbon is the best match
for the NoManagement scenario in NK Fig 2. We confirm this by computing
RMS error for the residuals between our FVS runs and the NK Fig 2 data:

```{r}
Plots_RMSE <- data.frame(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Total_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Aboveground_Carbon - NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Live = sqrt(mean(
    (NoManagement_Carbon_ByPlot_ByYear$Aboveground_Live -
       NoManagement_Carbon_ByPlot_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Plots_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Plot") +
  geom_col()
```

Note that NK describe using Aboveground Carbon for their chart, but the
best fit is from Aboveground Live, with RMSE \~= 12.

```{r}
Plots_RMSE |> select(Aboveground_Live)
```

### By Subplot

To confirm that NK used the FVS_StandInit_Plot table, we examine the
RMSE between projections from other tables and the NK Fig 2 values.

We repeat the above FVS run, but select table FVS_PlotInit_Plot and
select all subplots using the translated stand IDs. We apply the same
configuration to all stands as we did for FVS_StandInit_Plot. We load
and clean the results in the same manner.

```{r}
NoManagement_Carbon_BySubplot_ByYear <- read_csv(
    "../03_outputs/NoManagement_Carbon_BySubplot.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Total_Carbon = Total_Stand_Carbon) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  select(StandID, Year, Total_Carbon, Aboveground_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead) |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Aboveground_Live=mean(Aboveground_Live),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

ggplot(
    data = melt(NoManagement_Carbon_BySubplot_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Subplot") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 320)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,320,20))
```

Examine RMSE for the residuals:

```{r}
Subplots_RMSE <- data.frame(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Total_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Aboveground_Carbon - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Live = sqrt(mean(
    (NoManagement_Carbon_BySubplot_ByYear$Aboveground_Live - NoManagement_Carbon_BySubplot_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Subplots_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Subplot") +
  geom_col()
```

Note that all RMSE for the subplot projections are significantly worse
than for the plot projections.

### By Condition

We repeat the exercise using the FVS_StandInit_Cond table, which creates
FVS stands for corresponding FIA conditions.

We select all conditions using the translated stand IDs. Note that plot
360304303966 has 3 conditions; we include all three. We apply the same
configuration to all stands as we did for FVS_StandInit_Plot. We load
and clean the results in the same manner.

```{r}
NoManagement_Carbon_ByCondition_ByYear <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByCondition.csv",
    col_types = cols(StandID = col_character())
  ) |>
  rename(Total_Carbon = Total_Stand_Carbon) |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  rename(Down_Dead = Forest_Down_Dead_Wood) |>
  mutate(Aboveground_Dead = Standing_Dead + Down_Dead) |>
  mutate(Aboveground_Carbon = Aboveground_Live + Aboveground_Dead) |>
  select(StandID, Year, Total_Carbon, Aboveground_Carbon,
         Aboveground_Live, Aboveground_Dead, Standing_Dead, Down_Dead) |>
  group_by(Year) |>
  summarize(
    Total_Carbon=mean(Total_Carbon),
    Aboveground_Carbon=mean(Aboveground_Carbon),
    Aboveground_Live=mean(Aboveground_Live),
    .groups = "keep"
  ) |>
  filter(Year >= 2005) |>
  full_join(fig2 |> select(Year,NoManagement), by=join_by(Year)) |>
  rename(NK_Fig2_NoManagement = NoManagement)

ggplot(
    data = melt(NoManagement_Carbon_ByCondition_ByYear, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("FVS Carbon Projection by FIA Condition") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 320)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,320,20))
```

Examine RMSE for the residuals:

```{r}
Conditions_RMSE <- data.frame(
  Total_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Total_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Carbon = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Aboveground_Carbon - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  )),
  Aboveground_Live = sqrt(mean(
    (NoManagement_Carbon_ByCondition_ByYear$Aboveground_Live - NoManagement_Carbon_ByCondition_ByYear$NK_Fig2_NoManagement)^2
  ))
)
ggplot(
    data = melt(Conditions_RMSE, id = NULL, variable.name = "Series", value.name = "RMSE"),
    mapping = aes(x = Series, y = RMSE)
  ) +
  ggtitle("FVS Carbon Projection by FIA Condition") +
  geom_col()
```

RMSE for Aboveground Live by Condition is \~13, which is close to that
for Aboveground Live by Plot.

```{r}
Conditions_RMSE |> select(Aboveground_Live)
```

## Carbon Measurement

### Inventory Years

When were the plots re-surveyed?

The FIA PLOT table lists when plots were measured; the FIA COND table lists
observations from those measurements. We can use the above established
connection between FVS identifiers and FIA identifiers to find
subsequent measurement for the plots and conditions included in NK.

```{r}
# Load the translation table we created
nk_to_fia <- read_csv(
    "../03_outputs/nk_to_fia.csv",
    col_types = cols(
      `FIA plot code` = col_character(),
      .default = col_integer()
    )
  ) |>
  # Filter to only the columns needed to identify FIA conditions.
  select(`FIA plot code`, STATECD, UNITCD, COUNTYCD, PLOT)

# We also want the FVS_STAND_ID for these plots; graft that on
nk_to_fvs <- read_csv(
    "../03_outputs/nk_to_fvs.csv",
    col_types = cols(
      `FIA plot code` = col_character(),
      FVS_STAND_ID = col_character(),
      STAND_ID_PLOT = col_character(),
      STAND_ID_COND = col_character(),
      .default = col_integer()
    )
  ) |>
  # Filter to only the columns needed to identify FIA conditions.
  select(`FIA plot code`, FVS_STAND_ID)

nk_plot_ids <- nk_to_fia |>
  full_join(nk_to_fvs, by=join_by(`FIA plot code`)) |>
  # PLOT_ID is state, county, plot - like FVS_STAND_ID with no year
  mutate(PLOT_ID = sprintf(paste0(
    '%02d',  '%03d',   '%05d'),
    STATECD, COUNTYCD, PLOT
  ))

# Connect to the FIA tables
fia = DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')
fia_cond = tbl(fia, 'COND')
fia_plot = tbl(fia, 'PLOT')

# Create a dataframe containing the PLOT rows for all inventories of
# all plots used in NK.
matching_plot <- fia_plot |>
  semi_join(
    nk_plot_ids,
    by=join_by(STATECD, UNITCD, COUNTYCD, PLOT),
    copy=TRUE
  ) |>
  filter(INVYR >= 1999) |>
  collect()

# Create a dataframe containing the COND rows for all inventories of
# all conditions for all plots used in NK.
matching_cond <- fia_cond |>
  semi_join(
    nk_plot_ids,
    by=join_by(STATECD, UNITCD, COUNTYCD, PLOT),
    copy=TRUE
  ) |>
  filter(INVYR >= 1999) |>
  # Paste MEASYEAR on to COND, because everything in FVS works off MEASYEAR
  collect() |>
  left_join(
    matching_plot |> select(STATECD, UNITCD, COUNTYCD, PLOT, INVYR, MEASYEAR),
    by=join_by(STATECD, UNITCD, COUNTYCD, PLOT, INVYR)
  )

DBI::dbDisconnect(fia)

knitr::kable(matching_cond)
```

To aid visual analysis, we create arbitrary ordinals for plots; this
will allow us to chart the conditions for each plot in a single
horizontal band, against time, to create a visual timeline for plot
measurement.

```{r}
# Create ordinals for plots; to do this, we sort plots by:
# - the first year measured, earliest first; then
# - break ties by the second year measured, earliest year first; then
# - break ties by the third year inventoried, earliest year first; then
# - break ties by the fourth year inventoried, earliest year first
# This will put plots surveyed earlier towards the bottom of the chart,
# creating an "upward and to the right" visual that is generally easier to read.

# Find inventory years
year1 <- matching_cond |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) |>
  summarize(YEAR1 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(matching_cond, by=join_by(STATECD, UNITCD, COUNTYCD, PLOT))

year2 <- year1 |>
  filter(MEASYEAR > YEAR1) |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) |>
  summarize(YEAR2 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(year1, by=join_by(STATECD, UNITCD, COUNTYCD, PLOT))
  
year3 <- year2 |>
  filter(MEASYEAR > YEAR2) |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) |>
  summarize(YEAR3 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(year2, by=join_by(STATECD, UNITCD, COUNTYCD, PLOT))

year4 <- year3 |>
  filter(MEASYEAR > YEAR3) |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) |>
  summarize(YEAR4 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(year3, by=join_by(STATECD, UNITCD, COUNTYCD, PLOT))

# Compute plot numbers as ordinals according to inventory years
numbered_plots <- year4 |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) |>
  summarise(
    YEAR1 = min(YEAR1),
    YEAR2 = min(YEAR2),
    YEAR3 = min(YEAR3),
    YEAR4 = min(YEAR4),
    .groups = "keep"
  ) |>
  ungroup() |>
  arrange(YEAR1, YEAR2, YEAR3, YEAR4) |>
  mutate(PLOT_NUM = row_number()) |>
  select(PLOT_NUM, STATECD, UNITCD, COUNTYCD, PLOT)

# Paste plot numbers back on to conditions, then aggregate by plot
inventory_years <- matching_cond |>
  left_join(numbered_plots, by=join_by(STATECD, UNITCD, COUNTYCD, PLOT)) |>
  group_by(PLOT_NUM, MEASYEAR)

ggplot(
    data = inventory_years,
    mapping = aes(x = MEASYEAR, y = PLOT_NUM)
  ) +
  geom_point()

```

### Non-Sampled

Survey may be skipped for various reasons captured in
COND_STATUS_CD and COND_NONSAMPLE_REASN_CD. Codes other than 1 indicate
that forest was not inventoried in the condition.

Status Code Descriptions are given in [@Burrill2024]:

-   1 Accessible forest land - Land within the population of interest
    that can be occupied safely and has at least 10 percent canopy cover
    by live tally trees of any size or has had at least 10 percent
    canopy cover of live tally species in the past, based on the
    presence of stumps, snags, or other evidence. To qualify, the area
    must be at least 1.0 acre in size and 120.0 feet wide. Forest land
    includes transition zones, such as areas between forest and
    nonforest lands that meet the minimal tree canopy cover and forest
    areas adjacent to urban and built-up lands. Roadside, streamside,
    and shelterbelt strips of trees must have a width of at least 120
    feet and continuous length of at least 363 feet to qualify as forest
    land. Unimproved roads and trails, streams, and clearings in forest
    areas are classified as forest if they are less than 120 feet wide
    or less than an acre in size. Tree-covered areas in agricultural
    production settings, such as fruit orchards, or tree-covered areas
    in urban settings, such as city parks, are not considered forest
    land.
-   2 Nonforest land - Land that has less than 10 percent canopy cover
    of tally tree species of any size and, in the case of afforested
    land, fewer than 150 established trees per acre; or land that has
    sufficient canopy cover or stems, but is classified as nonforest
    land use (see criteria under PRESNFCD). Nonforest includes areas
    that have sufficient cover or live stems to meet the forest land
    definition, but do not meet the dimensional requirements. Note:
    Nonforest land includes "other wooded land" that has at least 5
    percent, but less than 10 percent, canopy cover of live tally tree
    species of any size or has had at least 5 percent, but less than 10
    percent, canopy cover of tally species in the recent past, based on
    the presence of stumps, snags, or other evidence. Other wooded land
    is recognized as a subset of nonforest land, and therefore is not
    currently considered a separate condition class. Other wooded land
    is not subject to nonforest use(s) that prevent normal tree
    regeneration and succession, such as regular mowing, intensive
    grazing, or recreation activities. In addition, other wooded land is
    classified according to the same nonforest land use rules as forest
    land (e.g., 6 percent cover in an urban setting is not considered
    other wooded land). Other wooded land is therefore defined as having
    \>5 percent and \<10 percent canopy cover at present, or evidence of
    such in the past, and PRESNFCD = 20, 40, 42, 43 or 45.
-   3 Noncensus water - Lakes, reservoirs, ponds, and similar bodies of
    water 1.0 acre to 4.5 acre in size. Rivers, streams, canals, etc.,
    30.0 feet to 200 feet wide. This definition was used in the 1990
    census and applied when the data became available. Earlier
    inventories defined noncensus water differently.
-   4 Census water - Lakes, reservoirs, ponds, and similar bodies of
    water 4.5 acre in size and larger; and rivers, streams, canals,
    etc., more than 200 feet wide.
-   5 Nonsampled, possibility of forest land - Any portion of a plot
    within accessible forest land that cannot be sampled is delineated
    as a separate condition. There is no minimum size requirement. The
    reason the condition was not sampled is provided in
    COND_NONSAMPLE_REASN_CD.

Nonsample Reason Code Descriptions are given in [@Burrill2024]:

-   01 Outside U.S. boundary - Condition class is outside the U.S.
    border.
-   02 Denied access area - Access to the condition class is denied by
    the legal owner, or by the owner of the only reasonable route to the
    condition class.
-   03 Hazardous situation - Condition class cannot be accessed because
    of a hazard or danger, for example cliffs, quarries, strip mines,
    illegal substance plantations, temporary high water, etc.
-   05 Lost data - The data file was discovered to be corrupt after a
    panel was completed and submitted for processing. Used for the
    single condition that is required for this plot. This code is for
    office use only.
-   06 Lost plot - Entire plot cannot be found. Used for the single
    condition that is required for this plot.
-   07 Wrong location - Previous plot can be found, but its placement is
    beyond the tolerance limits for plot location. Used for the single
    condition that is required for this plot.
-   08 Skipped visit - Entire plot skipped. Used for plots that are not
    completed prior to the time a panel is finished and submitted for
    processing. Used for the single condition that is required for this
    plot. This code is for office use only.
-   09 Dropped intensified plot - Intensified plot dropped due to a
    change in grid density. Used for the single condition that is
    required for this plot. This code used only by units engaged in
    intensification. This code is for office use only.
-   10 Other - Condition class not sampled due to a reason other than
    one of the specific reasons listed.
-   11 Ocean - Condition falls in ocean water below mean high tide line.

```{r}
# Were any conditions not sampled during an inventory?
skipped <- matching_cond |>
  filter(COND_STATUS_CD > 1) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, MEASYEAR, COND_STATUS_CD)
knitr::kable(skipped)
```

### Disturbance

Three fields, DSTRBCD1/2/3, capture Disturbance codes. These indicate
types of disturbance to the condition other than disturbances related to
harvest.

Code Descriptions are given in [@Burrill2024]:

-   0 No visible disturbance.
-   10 Insect damage.
-   11 Insect damage to understory vegetation.
-   12 Insect damage to trees, including seedlings and saplings.
-   20 Disease damage.
-   21 Disease damage to understory vegetation.
-   22 Disease damage to trees, including seedlings and saplings.
-   30 Fire damage (from crown and ground fire, either prescribed or
    natural).
-   31 Ground fire damage.
-   32 Crown fire damage.
-   40 Animal damage.
-   41 Beaver (includes flooding caused by beaver).
-   42 Porcupine.
-   43 Deer/ungulate.
-   44 Bear (core optional).
-   45 Rabbit (core optional).
-   46 Domestic animal/livestock (includes grazing).
-   50 Weather damage.
-   51 Ice.
-   52 Wind (includes hurricane, tornado).
-   53 Flooding (weather induced).
-   54 Drought.
-   60 Vegetation (suppression, competition, vines).
-   70 Unknown / not sure / other.
-   80 Human-induced damage - any significant threshold of human-caused
    damage not described in the DISTURBANCE codes or in the TREATMENT
    codes.
-   90 Geologic disturbances.
-   91 Landslide.
-   92 Avalanche track.
-   93 Volcanic blast zone.
-   94 Other geologic event.
-   95 Earth movement / avalanches.

Harvest-related disturbance is captured in the three columns TRTCD1/2/3.
These capture treatment code describing disturbance related to harvest.

Code Descriptions are given in [@Burrill2024]:

-   00 No observable treatment.
-   10 Cutting - The removal of one or more trees from a stand.
-   20 Site preparation - Clearing, slash burning, chopping, disking,
    bedding, or other practices clearly intended to prepare a site for
    either natural or artificial regeneration.
-   30 Artificial regeneration - Following a disturbance or treatment
    (usually cutting), a new stand where at least 50 percent of the live
    trees present resulted from planting or direct seeding.
-   40 Natural regeneration - Following a disturbance or treatment
    (usually cutting), a new stand where at least 50 percent of the live
    trees present (of any size) were established through the growth of
    existing trees and/or natural seeding or sprouting.
-   50 Other silvicultural treatment - The use of fertilizers,
    herbicides, girdling, pruning, or other activities (not covered by
    codes 10-40) designed to improve the commercial value of the
    residual stand; or chaining, which is a practice used on woodlands
    to encourage wildlife forage.


```{r}
# Who had disturbance?
disturbed <- matching_cond |>
  filter(DSTRBCD1 > 0 | DSTRBCD2 > 0 | DSTRBCD3 > 0) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, MEASYEAR, DSTRBCD1, DSTRBCD2, DSTRBCD3)
knitr::kable(disturbed)
```

```{r}
# Who had treatment?
treated <- matching_cond |>
  filter(TRTCD1 > 0 | TRTCD2 > 0 | TRTCD3 > 0) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, MEASYEAR, TRTCD1, TRTCD2, , TRTCD3)
knitr::kable(treated)
```

### Timeline

We can combine nonsample, disturbance, and harvest to create a single timeline
that shows plot status since the start of NK.

Note that status is per condition, and we are charting per plot. To accomodate,
we introduce horizontal offset so that the status of individual conditions
within a plot is visible in the chart.

```{r}
survey_status <- matching_cond |>

  # Offset INVYR by CONDID to get horizontal
  # displacement of conditions in the chart
  mutate(MEASYEAR = MEASYEAR + (CONDID - 1)/4) |>
  
  # Collect the various status indicators into a single STATUS column
  mutate(STATUS = if_else(COND_STATUS_CD > 1, "Skipped", NA)) |>
  mutate(STATUS = if_else(!is.na(TRTCD1) & TRTCD1 > 0, "Harvested", STATUS)) |>
  mutate(STATUS = if_else(!is.na(DSTRBCD1) & DSTRBCD1 > 0, "Disturbed", STATUS)) |>
  mutate(STATUS = if_else(is.na(STATUS), "Grown", STATUS)) |>
  
  # Use the single STATUS column to create separate columns for each status,
  # where the inventory year is the value for the column
  mutate(GROWN = if_else(STATUS == "Grown", MEASYEAR, NA)) |>
  mutate(SKIPPED = if_else(STATUS == "Skipped", MEASYEAR, NA)) |>
  mutate(HARVESTED = if_else(STATUS == "Harvested", MEASYEAR, NA)) |>
  mutate(DISTURBED = if_else(STATUS == "Disturbed", MEASYEAR, NA)) |>
  
  # Add plot numbers
  left_join(numbered_plots, by = join_by(STATECD, UNITCD, COUNTYCD, PLOT)) |>
  select(
    PLOT_NUM, GROWN, HARVESTED, DISTURBED, SKIPPED
  ) |>
  
  # Melt and arrange for ggplot
  melt(id.vars = "PLOT_NUM", variable.name="Status", value.name = "Year") |>
  filter(!is.na(Year)) |>
  select(PLOT_NUM, Year, Status) |>
  arrange(PLOT_NUM, Year, Status)

knitr::kable(survey_status)
```

```{r}
ggplot(
    data = survey_status,
    mapping = aes(x = Year, y = PLOT_NUM, shape = Status, color = Status)
  ) +
  ggtitle("Plot Inventory Timeline") +
  theme(legend.title = element_blank()) +
  geom_point()
```
### Grow-Only Plots

From this, it seems there are a subset of plots that are grow-only for
all subsequent inventories. We would like to compare the carbon in the FIA
inventoried plots with the carbon projected by FVS.

We start by identifying those plots that are grow-only.

```{r}
grow_only <- matching_cond |>
  group_by(STATECD, UNITCD, COUNTYCD, PLOT) |>
  summarize(
    COND_STATUS_CD = max(COND_STATUS_CD),
    TRTCD = max(TRTCD1),
    DSTRBCD = max(DSTRBCD1),
    .groups = "keep"
  ) |>
  ungroup() |>
  filter(COND_STATUS_CD == 1 & TRTCD == 0 & DSTRBCD == 0) |>
  left_join(numbered_plots, by = join_by(STATECD, UNITCD, COUNTYCD, PLOT)) |>
  left_join(nk_plot_ids, by = join_by(STATECD, UNITCD, COUNTYCD, PLOT))
  
knitr::kable(grow_only)
```

```{r}
ggplot(
    data = semi_join(survey_status, grow_only, by = join_by(PLOT_NUM)),
    mapping = aes(x = Year, y = PLOT_NUM, shape = Status, color = Status)
  ) +
  ggtitle("Grow-only Plot Inventory Timeline") +
  theme(legend.title = element_blank()) +
  geom_point() +
  coord_cartesian(xlim = c(1999, 2021), ylim = c(0, 31))

```
### Computing Carbon

[@fvsstaff2021] Module 2, section Audit 1, describes how to perform a zero-year
run of FVS. This applies the mapping from FIA to FVS, and can apply
e.g. the carbon model. We do not apply the regeneration adjustments, since
we are simply computing carbon from the FIA inventory data.
We do this for each individual grow-only plot, then
chart the observed vs. predicted carbon for just these plots.

```{r}
grown_fvs_stands <- matching_cond |>
  semi_join(grow_only, by = join_by(STATECD, UNITCD, COUNTYCD, PLOT)) |>
  select(STATECD, UNITCD, COUNTYCD, PLOT, MEASYEAR) |>
  left_join(
    nk_plot_ids,
    by = join_by(STATECD, UNITCD, COUNTYCD, PLOT)
  ) |>
  arrange(MEASYEAR, PLOT_ID)

knitr::kable(grown_fvs_stands)
```

Now plot measured vs projected

```{r}
comparable_carbon <- NoManagement_Carbon_ByPlot |>
  semi_join(grow_only |> rename(StandID = FVS_STAND_ID), by = join_by(StandID)) |>
  left_join(nk_plot_ids |> rename(StandID = FVS_STAND_ID), by=join_by(StandID)) |>
  mutate(Variable = "Projected") |>
  select(PLOT_ID, Year, Variable, Aboveground_Live) |>
  arrange(Year, PLOT_ID)
knitr::kable(comparable_carbon)
```


```{r}
ggplot(
    data = comparable_carbon,
    mapping = aes(x = Year, y = Aboveground_Live, color = PLOT_ID, style = Variable)
  ) +
  geom_point() +
  coord_cartesian(xlim = c(2000, 2025), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2000,2025,10))
```

```{r}
FIA_Measured_2002 <- read_csv("../03_outputs/FIA_Measured_2002.csv", 
    col_types = cols(StandID = col_character()))
FIA_Measured_2003 <- read_csv("../03_outputs/FIA_Measured_2003.csv", 
    col_types = cols(StandID = col_character()))
FIA_Measured_2004 <- read_csv("../03_outputs/FIA_Measured_2004.csv", 
    col_types = cols(StandID = col_character()))

FIA_Measured <- FIA_Measured_2002 |>
  union_all(FIA_Measured_2003) |>
  union_all(FIA_Measured_2004) |>
  mutate(Aboveground_Live = Aboveground_Total_Live) |>
  mutate(PLOT_ID = paste0(
     substring(StandID, 1, 2),
     substring(StandID, 5, 12)
  )) |>
  mutate(Variable = "Measured") |>
  select(PLOT_ID, Year, Variable, Aboveground_Live)
```

```{r}
knitr::kable(FIA_Measured)
```

```{r}
ggplot(
    data = comparable_carbon |> union_all(FIA_Measured),
    mapping = aes(x = Year, y = Aboveground_Live, color = PLOT_ID, shape = Variable)
  ) +
  geom_point(size = 2) +
  scale_shape_manual(values = c("Measured" = 3, "Projected" = 21)) +
  coord_cartesian(xlim = c(2000, 2025), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2000,2025,10))
```