---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Pre-load libraries that issue warnings on startup so these
# warnings don't corrupt the text later.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(dbplyr))
suppressPackageStartupMessages(library(measurements))
```

# Carbon Calculation

[@Herbert2023] use FIA measurement of biomass to determine carbon. FIA
also provides above-groud carbon in the TREE table, in CARBON_AG. This
metric is, of course, not available on plots that FVS grows. For
consistent carbon calculation on both measured and grown plots, we can
use the FVS Fire and Fuels extension (FFE-FVS), which contains a Carbon
Submodel, to compute carbon. We compare FFE carbon calculation with
Jenkins and FIA carbon calculation at the end of this section.

```{r}
# Load in-memory versions of various FIA tables, filtered to just
# those rows that match the plots used in NK.

# Load the translation table we created
nk_to_fia <- read_csv(
    "../03_outputs/nk_to_fia.csv",
    col_types = cols(
      `FIA plot code` = col_character(),
      .default = col_integer()
    )
  ) |>
  # Filter to only the columns needed to identify FIA conditions.
  select(`FIA plot code`, STATECD, COUNTYCD, PLOT, CONDID)

# We also want the FVS_STAND_ID for these plots; graft that on
nk_to_fvs <- read_csv(
    "../03_outputs/nk_to_fvs.csv",
    col_types = cols(
      `FIA plot code` = col_character(),
      FVS_STAND_ID = col_character(),
      STAND_ID_PLOT = col_character(),
      STAND_ID_COND = col_character(),
      .default = col_integer()
    )
  ) |>
  # Filter to only the columns needed to identify FIA conditions.
  select(`FIA plot code`, FVS_STAND_ID)

nk_plot_ids <- nk_to_fia |>
  full_join(nk_to_fvs, by=join_by(`FIA plot code`)) |>
  # STAND_ID is state, county, plot, cond - like FVS_STAND_ID with no year
  mutate(STAND_ID = sprintf(paste0(
    '%02d',  '%03d',   '%05d', '%01d'),
    STATECD, COUNTYCD, PLOT, CONDID
  ))

# Connect to the FIA tables
fia = DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')
fia_plot = tbl(fia, 'PLOT')
fia_cond = tbl(fia, 'COND')
fia_subplot = tbl(fia, 'SUBPLOT')
fia_subp_cond = tbl(fia, 'SUBP_COND')
fia_tree = tbl(fia, 'TREE')
fia_ref_forest_type = tbl(fia, 'REF_FOREST_TYPE') |>
  select(VALUE, MEANING)

# Create a dataframe containing the PLOT rows for all inventories of
# all plots used in NK.
matching_plot <- fia_plot |>
  filter(INVYR >= 1999) |>
  semi_join(
    nk_plot_ids,
    by=join_by(STATECD, COUNTYCD, PLOT),
    copy=TRUE
  ) |>
  collect()

# Create a dataframe containing the COND rows for all inventories of
# all conditions for all plots used in NK.
matching_cond <- fia_cond |>
  filter(INVYR >= 1999) |>
  semi_join(
    nk_plot_ids,
    # Omitting CONDID in the join will not filter out any conditions
    by=join_by(STATECD, COUNTYCD, PLOT),
    copy=TRUE
  ) |>
  # Dereference forest type codes
  left_join(fia_ref_forest_type, by = join_by(FORTYPCD == VALUE)) |>
  rename(FORTYPE = MEANING) |>
  # Paste MEASYEAR on to COND, because everything in FVS works off MEASYEAR
  collect() |>
  left_join(
    matching_plot |> select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR),
    by=join_by(STATECD, COUNTYCD, PLOT, INVYR)
  )

# Create a dataframe containing the SUBP_COND rows for all inventories of
# all subplot / condition intersections used in NK.
# Add condition and subplot status on for convenience.
matching_subp_cond <- fia_subp_cond |>
  filter(INVYR >= 1999) |>
  semi_join(
    nk_plot_ids,
    # Omitting CONDID in the join will not filter out any conditions
    by=join_by(STATECD, COUNTYCD, PLOT),
    copy=TRUE
  ) |>
  # Put COND_STATUS_CD on SUBP_COND for convenience
  left_join(
    fia_cond |> select(STATECD, COUNTYCD, PLOT, CONDID, INVYR, COND_STATUS_CD),
    by = join_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR)
  ) |>
  # Put SUBP_STATUS_CD on SUBP_COND for convenience
  left_join(
    fia_subplot |> select(STATECD, COUNTYCD, PLOT, SUBP, INVYR, SUBP_STATUS_CD),
    by = join_by(STATECD, COUNTYCD, PLOT, SUBP, INVYR)
  ) |>
  collect()

# Create a dataframe containing the TREE rows for all inventories of
# all conditions for all subplots for all plots used in NK.
matching_tree <- fia_tree |>
  filter(INVYR >= 1999) |>
  semi_join(
    nk_plot_ids,
    # Omitting CONDID in the join will not filter out any conditions
    # Omitting SUBP in the join will not filter out any subplots
    by = join_by(STATECD, COUNTYCD, PLOT),
    copy = TRUE
  ) |>
  collect()

DBI::dbDisconnect(fia)
```

## Inventory Years

The FIA PLOT table lists when plots were measured; the FIA COND table
lists observations from those measurements. We can use the above
established connection between FVS identifiers and FIA identifiers to
find subsequent measurement for the plots and conditions included in NK.

To aid visual analysis, we create arbitrary ordinals for conditions;
this will allow us to chart the conditions for each plot in a single
horizontal band, against time, to create a visual timeline for plot
measurement.

> ⚠️ TODO nik: Conditions are numbered according to where they appear in
> the plot. If the condition at the center of subplot 1 changes, then
> the characteristics of condition 1 will change, even if there are
> portions of the plot that are still in what used to be condition 1. To
> assign stand ordinals actually correctly, we would want to align
> conditions by non-volatile characteristics, such as forest type.

```{r}
# Create ordinals for conditions; to do this, we sort conditions by:
# - the first year measured, earliest first; then
# - break ties by the second year measured, earliest year first; then
# - break ties by the third year inventoried, earliest year first; then
# - break ties by the fourth year inventoried, earliest year first
# This will put plots surveyed earlier towards the bottom of the chart,
# creating an "upward and to the right" visual that is generally easier to read.
year1 <- matching_cond |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(YEAR1 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(matching_cond, by=join_by(STATECD, COUNTYCD, PLOT))

year2 <- year1 |>
  filter(MEASYEAR > YEAR1) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(YEAR2 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(year1, by=join_by(STATECD, COUNTYCD, PLOT))
  
year3 <- year2 |>
  filter(MEASYEAR > YEAR2) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(YEAR3 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(year2, by=join_by(STATECD, COUNTYCD, PLOT))

year4 <- year3 |>
  filter(MEASYEAR > YEAR3) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(YEAR4 = min(MEASYEAR), .groups = "keep") |>
  ungroup() |>
  right_join(year3, by=join_by(STATECD, COUNTYCD, PLOT))

# Compute stand numbers as ordinals according to inventory years
numbered_stands <- year4 |>
  group_by(STATECD, COUNTYCD, PLOT, CONDID) |>
  summarise(
    YEAR1 = min(YEAR1),
    YEAR2 = min(YEAR2),
    YEAR3 = min(YEAR3),
    YEAR4 = min(YEAR4),
    .groups = "keep"
  ) |>
  ungroup() |>
  arrange(YEAR1, YEAR2, YEAR3, YEAR4, CONDID) |>
  mutate(STAND_NUM = row_number()) |>
  select(STAND_NUM, STATECD, COUNTYCD, PLOT, CONDID)

# Paste stand numbers back on to conditions, then aggregate by plot
stand_inventory <- matching_cond |>
  left_join(numbered_stands, by=join_by(STATECD, COUNTYCD, PLOT, CONDID)) |>
  group_by(STAND_NUM, MEASYEAR)

ggplot(
    data = stand_inventory,
    mapping = aes(x = MEASYEAR, y = STAND_NUM)
  ) +
  geom_point()

```

## Plot Status

### Non-Sampled

Survey may be skipped for various reasons captured in COND_STATUS_CD and
COND_NONSAMPLE_REASN_CD. Codes other than 1 indicate that forest was not
inventoried in the condition.

Status Code Descriptions are given in [@Burrill2024]:

-   1 Accessible forest land
-   2 Nonforest land
-   3 Noncensus water
-   4 Census water
-   5 Nonsampled, possibility of forest land

Nonsample Reason Code Descriptions are given in [@Burrill2024]:

-   01 Outside U.S. boundary
-   02 Denied access area
-   03 Hazardous situation
-   05 Lost data
-   06 Lost plot
-   07 Wrong location
-   08 Skipped visit
-   09 Dropped intensified plot
-   10 Other
-   11 Ocean

```{r}
# Were any conditions not sampled during an inventory?
skipped <- matching_cond |>
  filter(COND_STATUS_CD > 1) |>
  select(STATECD, COUNTYCD, PLOT, CONDID, MEASYEAR, COND_STATUS_CD, COND_NONSAMPLE_REASN_CD)
knitr::kable(skipped)
```

### Disturbance

Three fields, DSTRBCD1/2/3, capture Disturbance codes. These indicate
types of disturbance to the condition other than disturbances related to
harvest.

Code Descriptions are given in [@Burrill2024]:

-   0 No visible disturbance.
-   10 Insect damage.
-   11 Insect damage to understory vegetation.
-   12 Insect damage to trees, including seedlings and saplings.
-   20 Disease damage.
-   21 Disease damage to understory vegetation.
-   22 Disease damage to trees, including seedlings and saplings.
-   30 Fire damage (from crown and ground fire, either prescribed or
    natural).
-   31 Ground fire damage.
-   32 Crown fire damage.
-   40 Animal damage.
-   41 Beaver (includes flooding caused by beaver).
-   42 Porcupine.
-   43 Deer/ungulate.
-   44 Bear (core optional).
-   45 Rabbit (core optional).
-   46 Domestic animal/livestock (includes grazing).
-   50 Weather damage.
-   51 Ice.
-   52 Wind (includes hurricane, tornado).
-   53 Flooding (weather induced).
-   54 Drought.
-   60 Vegetation (suppression, competition, vines).
-   70 Unknown / not sure / other.
-   80 Human-induced damage - any significant threshold of human-caused
    damage not described in the DISTURBANCE codes or in the TREATMENT
    codes.
-   90 Geologic disturbances.
-   91 Landslide.
-   92 Avalanche track.
-   93 Volcanic blast zone.
-   94 Other geologic event.
-   95 Earth movement / avalanches.

```{r}
# Who had disturbance?
disturbed <- matching_cond |>
  filter(DSTRBCD1 > 0 | DSTRBCD2 > 0 | DSTRBCD3 > 0) |>
  select(STATECD, COUNTYCD, PLOT, CONDID, MEASYEAR, DSTRBCD1, DSTRBCD2, DSTRBCD3)
knitr::kable(disturbed)
```

### Harvest

Harvest-related disturbance is captured in the three columns TRTCD1/2/3.
These capture treatment code describing disturbance related to harvest.

Code Descriptions are given in [@Burrill2024]:

-   00 No observable treatment.
-   10 Cutting
-   20 Site preparation
-   30 Artificial regeneration
-   40 Natural regeneration
-   50 Other silvicultural treatment

```{r}
# Who had treatment?
treated <- matching_cond |>
  filter(TRTCD1 > 0 | TRTCD2 > 0 | TRTCD3 > 0) |>
  select(STATECD, COUNTYCD, PLOT, MEASYEAR, CONDID, TRTCD1, TRTCD2, , TRTCD3)
knitr::kable(treated)
```

### Timeline

We can combine nonsample, disturbance, and harvest to create a single
timeline that shows plot status since the start of NK.

```{r}
stand_status <- matching_cond |>

  # Collect the various status indicators into a single STATUS column
  mutate(STATUS = if_else(COND_STATUS_CD > 1, "Skipped", NA)) |>
  mutate(STATUS = if_else(!is.na(TRTCD1) & TRTCD1 > 0, "Harvested", STATUS)) |>
  mutate(STATUS = if_else(!is.na(DSTRBCD1) & DSTRBCD1 > 0, "Disturbed", STATUS)) |>
  mutate(STATUS = if_else(is.na(STATUS), "Grown", STATUS)) |>
  
  # Use the single STATUS column to create separate columns for each status,
  # where the inventory year is the value for the column
  mutate(GROWN = if_else(STATUS == "Grown", MEASYEAR, NA)) |>
  mutate(SKIPPED = if_else(STATUS == "Skipped", MEASYEAR, NA)) |>
  mutate(HARVESTED = if_else(STATUS == "Harvested", MEASYEAR, NA)) |>
  mutate(DISTURBED = if_else(STATUS == "Disturbed", MEASYEAR, NA)) |>
  
  # Add stand numbers
  left_join(numbered_plots, by = join_by(STATECD, COUNTYCD, PLOT, CONDID)) |>
  select(
    STAND_NUM, GROWN, HARVESTED, DISTURBED, SKIPPED
  ) |>
  
  # Melt and arrange for ggplot
  melt(id.vars = "STAND_NUM", variable.name="Status", value.name = "Year") |>
  filter(!is.na(Year)) |>
  select(STAND_NUM, Year, Status) |>
  arrange(STAND_NUM, Year, Status)

knitr::kable(stand_status)
```

```{r}
ggplot(
    data = stand_status,
    mapping = aes(x = Year, y = STAND_NUM, shape = Status, color = Status)
  ) +
  ggtitle("Stand Inventory Timeline") +
  geom_point()
```

### Grow-Only Plots

[@Herbert2023] says "Where stands were coded as having ≥25% of the trees
or sample area affected by insects, disease, weather damage, or geologic
disturbances damage between inventory visits, we coded a binary,
non-fire disturbance attribute to have available as an explanatory
factor but did not drop them from the analysis set."

From this, it seems there are a subset of plots that are grow-only for
all subsequent inventories. We would like to compare the carbon in the
FIA inventoried plots with the carbon projected by FVS.

We start by identifying those plots that are grow-only. We need all
conditions in the plot to be grow-only for all inventory years, since
FIA and FVS may diverge in their understanding of condition composition
of plots over time.

```{r}
grow_only_plots <- matching_cond |>
  # Group to the plot level, not the cond level.
  group_by(STATECD, COUNTYCD, PLOT) |>
  # Across all status indicators, "higher-numbered "undisturbed" is the lowest status number,
  # so we can find disturbed stands by taking the max() of status across inventories.
  summarize(
    COND_STATUS_CD = max(COND_STATUS_CD),
    TRTCD = max(TRTCD1),
    DSTRBCD = max(DSTRBCD1),
    .groups = "keep"
  ) |>
  ungroup() |>
  filter(COND_STATUS_CD == 1 & TRTCD == 0 & DSTRBCD == 0)

grow_only_stands <- matching_cond |>
  semi_join(grow_only_plots, by = join_by(STATECD, COUNTYCD, PLOT)) |>
  left_join(numbered_stands, by = join_by(STATECD, COUNTYCD, PLOT, CONDID)) |>
  left_join(nk_plot_ids, by = join_by(STATECD, COUNTYCD, PLOT, CONDID))
  
knitr::kable(grow_only_stands)
```

```{r}
ggplot(
    data = semi_join(stand_status, grow_only_stands, by = join_by(STAND_NUM)),
    mapping = aes(x = Year, y = STAND_NUM, shape = Status, color = Status)
  ) +
  ggtitle("Grow-only Stand Inventory Timeline") +
  theme(legend.title = element_blank()) +
  geom_point() +
  coord_cartesian(xlim = c(1999, 2021), ylim = c(0, 31))

```

## Carbon Accounting

[@fvsstaff2021] Module 2, section Audit 1, describes how to perform a
zero-year run of FVS. This applies the mapping from FIA plot to FVS
stand, then grows the stand for zero years, which keeps the trees as
they are in the FIA plot. This does however produce all the FVS output
data structures for the stand, which enables application of e.g. the
carbon model. We do not apply the regeneration adjustments, since we are
simply computing carbon from the FIA inventory data. We do this for each
individual grow-only plot, then chart the observed vs. predicted carbon
for just these plots.

Note that plots change over time, and FIA's way of dealing with this is
to change the conditions over time. FVS does not change conditions as a
plot ages; it grows trees assuming that the forest type of the plot
remains unchanged. The result is that FVS grows the conditions as
inventoried at the start of the window, whereas FIA will adjust those
conditions over time. For both projected and measured carbon, we use all
conditions in the plot as inventoried, to compare carbon projection vs.
measurement for the entire plot land area.

### FFE

```{r}
grown_fvs_stands <- matching_cond |>
  # Note: Filter by plot, not by stand - we want all conditions for the plot,
  # which may change over time.
  semi_join(grow_only_plots, by = join_by(STATECD, COUNTYCD, PLOT)) |>
  select(STATECD, COUNTYCD, PLOT, CONDID, MEASYEAR, INVYR) |>
  left_join(
    matching_cond |>
      select(STATECD, COUNTYCD, PLOT, INVYR, CONDID, FORTYPE),
    by = join_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR)
  ) |>
  # We'd like to join on pre-computed stand IDs here, but since conditions
  # change over time, we'll have to re-compute them.
  # Can't do this: left_join(nk_plot_ids, by = join_by(STATECD, COUNTYCD, PLOT, CONDID)) |>

  # Re-compute FVS_STAND_ID to use the correct year
  mutate(FVS_STAND_ID=sprintf(paste0(
    '%02d' , '%02d'      , '%03d'  , '%05d', '%01d' ),
    STATECD, INVYR %% 100, COUNTYCD, PLOT  , CONDID
  )) |>
  # Re-compute STAND_ID, which omits the year
  mutate(STAND_ID=sprintf(paste0(
    '%02d' , '%03d'  , '%05d', '%01d' ),
    STATECD, COUNTYCD, PLOT  , CONDID
  )) |>

  arrange(MEASYEAR, STAND_ID)

knitr::kable(grown_fvs_stands)
```

Now plot measured vs projected

```{r}
NoManagement_Carbon_ByConditionYearly <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByConditionYearly.csv",
    col_types = cols(StandID = col_character())
  )
knitr::kable(NoManagement_Carbon_ByCondition)
```

```{r}
x <- NoManagement_Carbon_ByConditionYearly
knitr::kable(grow_only_stands)
```

```{r}

comparable_carbon <- NoManagement_Carbon_ByConditionYearly |>
  inner_join(grow_only_stands |> rename(StandID = FVS_STAND_ID), by = join_by(StandID)) |>
  left_join(nk_plot_ids |> rename(StandID = FVS_STAND_ID), by=join_by(StandID)) |>
  mutate(Variable = "Projected") |>
  rename(Aboveground_Live = Aboveground_Total_Live) |>
  select(STAND_NUM, Year, Variable, Aboveground_Live) |>
  arrange(Year, STAND_NUM)
knitr::kable(comparable_carbon)
```

```{r}
ggplot(
    data = comparable_carbon,
    mapping = aes(x = Year, y = Aboveground_Live, color = PLOT_ID, style = Variable)
  ) +
  geom_point() +
  coord_cartesian(xlim = c(2000, 2025), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2000,2025,10))
```

```{r}

measyears <- grown_fvs_stands |>
  select(MEASYEAR) |>
  group_by(MEASYEAR) |>
  summarize(.groups = "keep") |>
  ungroup()

FIA_Measured <- NULL
FIA_Measured_Jenkins <- NULL
for (measyear in measyears)
  fname <- paste0(
    "../03_outputs/FIA_Measured_",
    sprintf("%04d", measyear),
    ".csv"
  )
  FIA_Measured <- bind_rows(
      FIA_Measured,
      read_csv(fname, col_types = cols(StandID = col_character()))
    )
  fname <- paste0(
    "../03_outputs/FIA_Measured_",
    sprintf("%04d", measyear),
    "_Jenkins.csv"
  )
  FIA_Measured_Jenkins <- bind_rows(
      FIA_Measured_Jenkins,
      read_csv(fname, col_types = cols(StandID = col_character()))
    )
  

FIA_Measured <- FIA_Measured |>
  mutate(Aboveground_Live = Aboveground_Total_Live) |>
  mutate(PLOT_ID = paste0(
     substring(StandID, 1, 2),
     substring(StandID, 5, 12)
  )) |>
  mutate(Variable = "Measured") |>
  left_join(
    grown_fvs_stands |> select(PLOT_ID, MEASYEAR, FORTYPE),
    by = join_by(PLOT_ID, Year == MEASYEAR)
  ) |>
  select(PLOT_ID, Year, Variable, Aboveground_Live, FORTYPE)

FIA_Measured_Jenkins <- FIA_Measured_Jenkins |>
  mutate(Aboveground_Live = Aboveground_Total_Live) |>
  mutate(PLOT_ID = paste0(
     substring(StandID, 1, 2),
     substring(StandID, 5, 12)
  )) |>
  mutate(Variable = "Measured") |>
  left_join(
    grown_fvs_stands |> select(PLOT_ID, MEASYEAR, FORTYPE),
    by = join_by(PLOT_ID, Year == MEASYEAR)
  ) |>
  select(PLOT_ID, Year, Variable, Aboveground_Live, FORTYPE)
```

```{r}
knitr::kable(FIA_Measured)
```

```{r}
knitr::kable(FIA_Measured_Jenkins)
```

```{r}
projected_carbon <- read_csv(
    "../03_outputs/NoManagement_Carbon_ByPlotYearly.csv",
    col_types = cols(StandID = col_character())
  ) |>
  mutate(Aboveground_Live = Aboveground_Total_Live) |>
  mutate(PLOT_ID = paste0(
     substring(StandID, 1, 2),
     substring(StandID, 5, 12)
  )) |>
  mutate(Variable = "Projected") |>
  select(PLOT_ID, Year, Variable, Aboveground_Live) |>
  # filter down to only those plots/years that match measurement
  semi_join(FIA_Measured, by = join_by(PLOT_ID, Year))

knitr::kable(projected_carbon)

```

```{r}
ggplot(
    data = projected_carbon |>
      union_all(FIA_Measured |> select(PLOT_ID, Year, Variable, Aboveground_Live)),
    mapping = aes(x = Year, y = Aboveground_Live, color = PLOT_ID, shape = Variable)
  ) +
  geom_point(size = 2) +
  scale_shape_manual(values = c("Measured" = 3, "Projected" = 21)) +
  coord_cartesian(xlim = c(2000, 2025), ylim = c(0, 300)) +
  scale_x_continuous(breaks=seq(2000,2025,10)) +
  ggtitle("Measured vs. Projected Carbon: FVS NoManagement") +
  ylab("Carbon (Tonnes / Hectare)")
```

Let's look at error:

```{r}
projection_error <- projected_carbon |>
  mutate(Projected = Aboveground_Live) |>
  select(PLOT_ID, Year, Projected) |>
  full_join(FIA_Measured, by = join_by(PLOT_ID, Year)) |>
  mutate(Measured = Aboveground_Live) |>
  mutate(Error = Projected - Measured) |>
  select(PLOT_ID, Year, Error)

ggplot(
    data = projection_error
  ) +
  geom_point(
    size = 2,
    mapping = aes(x = Year, y = Error, color = PLOT_ID)
  ) +
  coord_cartesian(xlim = c(2000, 2025), ylim = c(-100, 100)) +
  scale_x_continuous(breaks=seq(2000,2025,10)) +
  ggtitle("Measured vs. Projected Carbon Error: FVS NoManagement") +
  ylab("Carbon (Tonnes / Hectare)") +
  geom_smooth(
    data=projection_error |> select(Year, Error),
    method=lm,
    mapping = aes(x=Year, y=Error),
    formula = 'y ~ x'
  )
```

```{r}
first_measure_year <- FIA_Measured |>
  group_by(PLOT_ID) |>
  summarize(first_measure_year = min(Year), .groups = "keep") |>
  select(PLOT_ID, first_measure_year) |>
  ungroup()
```

RMSE

```{r}
projection_rmse <- projection_error |>
  left_join(first_measure_year, by = join_by(PLOT_ID)) |>
#  mutate(Years = floor((Year - first_measure_year) / 5) * 5) |>
  mutate(Years = Year - first_measure_year) |>
  select(PLOT_ID, Years, Error) |>
  group_by(Years) |>
  summarize(RMSE = sqrt(mean(Error^2)), .groups = "keep") |>
  ungroup()

ggplot(data = projection_rmse) +
  ggtitle("FVS NoManagement\nMeasured vs. Projected RMSE") +
  geom_point(size = 2, mapping = aes(x = Years, y = RMSE)) +
  geom_smooth(
    data=projection_rmse |> select(Years, RMSE),
    method=lm,
    mapping = aes(x=Years, y=RMSE),
    formula = 'y ~ x'
  )

```

plot predicted vs. observed

```{r}

projected_observed <- projected_carbon |>
  mutate(Projected = Aboveground_Live) |>
  select(PLOT_ID, Year, Projected) |>
  full_join(FIA_Measured, by = join_by(PLOT_ID, Year)) |>
  rename(`Forest Type` = FORTYPE) |>
  full_join(first_measure_year, by = join_by(PLOT_ID)) |>
  mutate(Measured = Aboveground_Live) |>
  mutate(Years = Year - first_measure_year) |>
  select(PLOT_ID, Years, Projected, Measured, `Forest Type`)

ggplot(
    data = projected_observed
  ) +
  geom_point(
    size = 2,
    mapping = aes(x = Measured, y = Projected, color = Years, shape = `Forest Type`)
  ) +
  ggtitle("FVS NoManagement Scenario\nMeasured vs. Projected Carbon (Mg/ha)") +
  coord_cartesian(xlim = c(0, 200), ylim = c(0, 200)) +
  geom_line(
    data = tibble(x = c(0, 200), y=c(0, 200)),
    mapping = aes(x = x, y = y)
  )
```

### Jenkins

FVS provides two ways to compute carbon: FFE and Jenkins

```{r}

ffe_jenkins <- FIA_Measured |>
  mutate(FFE = Aboveground_Live) |>
  rename(`Forest Type` = FORTYPE) |>
  select(PLOT_ID, Year, FFE, `Forest Type`) |>
  full_join(FIA_Measured_Jenkins, by = join_by(PLOT_ID, Year)) |>
  mutate(Jenkins = Aboveground_Live) |>
  full_join(first_measure_year, by = join_by(PLOT_ID)) |>
  # mutate(Years = Year - first_measure_year) |>
  select(PLOT_ID, Year, FFE, Jenkins, `Forest Type`)

ggplot(
    data = ffe_jenkins
  ) +
  geom_point(
    size = 2,
    mapping = aes(x = FFE, y = Jenkins, color = Year, shape = `Forest Type`)
  ) +
  ggtitle("FVS Carbon Calculation\nFFE vs. Jenkins Carbon (Mg/ha)") +
  coord_cartesian(xlim = c(0, 200), ylim = c(0, 200)) +
  geom_line(
    data = tibble(x = c(0, 200), y=c(0, 200)),
    mapping = aes(x = x, y = y)
  )
```

### FIA via Expansion

FIA has carbon in the tree table, in CARBON_AG; we can sum this per
stand and compute a per-hectare value.

Method:

-   Determine how much of each plot we will use the inventory \*\*
    Remove plots, subplots, and conditions that are not inventoried
    forest
-   Determine scale factor for each plot \*\* A full plot is 4 subplots
    of 24\^2\*pi square feet each; scale this according to how much
    we're using
-   scale the carbon in each tree for the area we're using to get carbon
    per acre
-   sum across all trees

```{r}
subplot_area <- conv_unit(24^2 * pi, 'ft2', 'm2') # subplot area in square meters

useful_subp_cond <- matching_subp_cond |>
  semi_join(
    grow_only,
    by = join_by(STATECD, COUNTYCD, PLOT)
  ) |>
  filter(SUBP <= 4) |> # FIA introduces estimation subplots with ID > 4; ignore those
  filter(COND_STATUS_CD == 1) |> # Only use measured, forested conditions
  filter(SUBP_STATUS_CD == 1) # Only use measured, forested subplots

useful_area <- useful_subp_cond |>
  group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
  summarize(SUBPLOTS_INCLUDED = sum(SUBPCOND_PROP), .groups = "keep") |>
  ungroup() |>
  mutate(PLOT_AREA = SUBPLOTS_INCLUDED * subplot_area)
# TODO: Maybe filter out plots below a certain total area?
#       Example: a full plot is 672.4535 m2; filter out anything below
#       89.2% of that

knitr::kable(useful_area)
```

Carbon

```{r}
useful_trees <- matching_tree |>
  filter(!is.na(CARBON_AG)) |>
  semi_join(
    useful_subp_cond,
    by = join_by(STATECD, COUNTYCD, PLOT, SUBP, CONDID, INVYR)
  ) |>
  left_join(
    useful_area,
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  )

m2_per_hectare <- conv_unit(1, 'hectare', 'm2')

# We have inventoried trees in PLOT_AREA m2; scale to 1 ha
carbon_per_plot <- useful_trees |>
  # Note: FIA says that CARBON_AG needs to be scaled by TPA_UNADJ to get to
  # carbon per acre. That is a different approach to computing carbon.
  mutate(CARBON_AG_METRIC = m2_per_hectare * conv_unit(CARBON_AG, 'lbs', 'Mg') / PLOT_AREA) |>
  group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
  summarize(CARBON_AG_METRIC = sum(CARBON_AG_METRIC), .groups = "keep") |>
  ungroup() |>
  # Join to get PLOT_IDs
  left_join(
    nk_plot_ids,
    by = join_by(STATECD, COUNTYCD, PLOT)
  ) |>
  # Join to get MEASYEAR
  left_join(
    matching_plot |> select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR),
    by=join_by(STATECD, COUNTYCD, PLOT, INVYR)
  )


knitr::kable(carbon_per_plot)

```

```{r}
ffe_fia <- FIA_Measured |>
  mutate(FFE = Aboveground_Live) |>
  rename(`Forest Type` = FORTYPE) |>
  select(PLOT_ID, Year, FFE, `Forest Type`) |>
  full_join(carbon_per_plot |> mutate(Year = MEASYEAR), by = join_by(PLOT_ID, Year)) |>
  mutate(FIA = CARBON_AG_METRIC) |>
  full_join(first_measure_year, by = join_by(PLOT_ID)) |>
  mutate(Years = Year - first_measure_year) |>
  select(PLOT_ID, Years, FFE, FIA, `Forest Type`)

ggplot(
    data = ffe_fia
  ) +
  geom_point(
    size = 2,
    mapping = aes(x = FFE, y = FIA, color = Years, shape = `Forest Type`)
  ) +
  ggtitle("FVS Carbon Calculation\nFFE vs. FIA Carbon (Mg/ha)") +
  coord_cartesian(xlim = c(0, 200), ylim = c(0, 200)) +
  geom_line(
    data = tibble(x = c(0, 200), y=c(0, 200)),
    mapping = aes(x = x, y = y)
  )
#knitr::kable(ffe_fia)
```
