---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Pre-load libraries that issue warnings on startup so these
# warnings don't corrupt the text later.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(dbplyr))
suppressPackageStartupMessages(library(measurements))
```

## Grow-Only Plots in the Northeast

Find all FIA plots in the northeastern region that are grow-only.

Northeastern region:

*  SURVEY.RCSD == 24

Grow-only is described previously:

*  Only one condition: max(FIA.COND.CONDID) == 1
*  Survey was not skipped: COND.COND_STATUS_CD == 1
*  Plot was not disturbed: COND.DSTRBCD1 == 0 & COND.DSTRBCD2 == 0 & COND.DSTRBCD3 == 0
*  Plot was not treated: COND.TRTCD1 == 0 & TRTCD2 == 0 & TRTCD3 == 0

```{r}
fia <- DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')
fia_cond <- tbl(fia, 'COND')
fia_plot <- tbl(fia, 'PLOT') |>
  # Narrow and rename columns to facilitate join
  select(CN, SRV_CN) |>
  rename(PLT_CN = CN)
fia_survey <- tbl(fia, 'SURVEY') |>
  # Narrow and rename columns to facilitate join
  select(CN, RSCD) |>
  rename(SRV_CN = CN)

plot_grow_only <- fia_cond |>
  filter(INVYR >= 1999) |>
  left_join(fia_plot, by = join_by(PLT_CN)) |>
  left_join(fia_survey, by = join_by(SRV_CN)) |>
  filter(RSCD == 24) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(
    CONDID = max(CONDID, na.rm = TRUE),
    COND_STATUS_CD = max(COND_STATUS_CD, na.rm = TRUE),
    DSTRBCD1 = max(DSTRBCD1, na.rm = TRUE),
    DSTRBCD2 = max(DSTRBCD2, na.rm = TRUE),
    DSTRBCD3 = max(DSTRBCD3, na.rm = TRUE),
    TRTCD1 = max(TRTCD1, na.rm = TRUE),
    TRTCD2 = max(TRTCD2, na.rm = TRUE),
    TRTCD3 = max(TRTCD3, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  filter(CONDID == 1) |>
  filter(COND_STATUS_CD == 1) |>
  filter(
    (is.na(DSTRBCD1) | DSTRBCD1 == 0) & 
    (is.na(DSTRBCD2) | DSTRBCD2 == 0) &
    (is.na(DSTRBCD3) | DSTRBCD3 == 0)
  ) |>
  filter(
    (is.na(TRTCD1) | TRTCD1 == 0) &
    (is.na(TRTCD2) | TRTCD2 == 0) &
    (is.na(TRTCD3) | TRTCD3 == 0)
  ) |>
  select(STATECD, COUNTYCD, PLOT) |>
  collect()

DBI::dbDisconnect(fia)
remove(fia, fia_cond, fia_plot, fia_survey)

length(plot_grow_only$PLOT)
```

This filters to 4299 plots, most of which will have multiple inventories.
 
### FIA Carbon and BALIVE

We can gather BALIVE and CARBON_AG for these plots over time.

```{r}
fia <- DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')

measyear <- tbl(fia, 'PLOT') |>
  select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR)

forest_type <- tbl(fia, 'REF_FOREST_TYPE') |>
  select(VALUE, MEANING) |>
  rename(FORTYPCD = VALUE) |>
  rename(FORTYPE = MEANING)

tree_stats <- tbl(fia, 'TREE') |>
  group_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR) |>
  summarize(
    CARBON_AG = sum(CARBON_AG),
    CPA = sum(CARBON_AG * TPA_UNADJ),
    .groups = "keep"
  )

stand_stats <- tbl(fia, 'COND') |>
  select(STATECD, COUNTYCD, PLOT, INVYR, STDAGE, BALIVE, FORTYPCD) |>
  filter(INVYR >= 1999) |>
  semi_join(plot_grow_only, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
  summarize(
    BALIVE = sum(BALIVE),
    FORTYPCD = max(FORTYPCD),
    STDAGE = max(STDAGE),
    .groups = "keep"
  ) |>
  left_join(measyear, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(tree_stats, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(forest_type, by = join_by(FORTYPCD)) |>
  rename(`Forest Type` = FORTYPE) |>
  mutate(FORTYPCD = floor(FORTYPCD / 10) * 10) |>
  left_join(forest_type, by = join_by(FORTYPCD)) |>
  rename(`Forest Type Group` = FORTYPE) |>
  collect() |>
  mutate(BALIVE_METRIC = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare")) |>
  mutate(CARBON_METRIC = conv_multiunit(CPA, "lbs / acre", "Mg / hectare"))

DBI::dbDisconnect(fia)
remove(fia, measyear, forest_type, tree_stats)
```

Look at BA over time

```{r}
ggplot(
    data = stand_stats |> filter(STDAGE > 0),
    mapping = aes(x = STDAGE, y = BALIVE_METRIC)
  ) +
  geom_point(alpha = 0.25) +
  facet_wrap(~`Forest Type Group`) +
  ggtitle("Grow-Only Stands\nFIA Measured basal area (m2 / ha)")
```

```{r}
seen_types <- stand_stats |>
  filter(!is.na(CARBON_AG)) |>
  group_by(`Forest Type`) |>
  summarize(
    CARBON_AG_METRIC = sum(conv_unit(CARBON_AG, "lbs", "Mg")),
    .groups = "keep"
  )

seen_types |>
  ggplot(
    aes(y = fct_reorder(`Forest Type`, CARBON_AG_METRIC), x = CARBON_AG_METRIC)
#    aes(x = CARBON_AG_METRIC, y = `Forest Type`)
  ) +
  scale_x_log10() +
  geom_col() +
  ggtitle("Total Carbon by Forest Type (Mg)") +
  ylab("Forest Type") +
  xlab("Carbon")
```