---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Pre-load libraries that issue warnings on startup so these
# warnings don't corrupt the text later.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(dbplyr))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(measurements))
suppressPackageStartupMessages(library(maps))
```

```{r source-functions}
source('../R/functions.R')
```

## Extract FIA plots from LANDIS-II Initial Communities

Execute the filters to create a list of matching plots.

```{r parse-plots}
plots <- tibble(text = read_lines('~/Downloads/sub_imputed_map_2024-07-17.txt')) |>
  filter(str_detect(text, 'FIA PLT_CN')) |>
  mutate(PLT_CN = str_replace(text, '^.* FIA PLT_CN ([0-9]+) .*$', '\\1'))

nrow(plots)
```

### Locations

Where are these plots?

Fetch plots with measyear, invyr, latitude, longitude, and state
information. Filter to plots that can be run with the northeast variant,
and INVYR 2019

```{r plot_location}
fia <- DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')

# Survey has the research station code
fia_survey <- tbl(fia, 'SURVEY') |>
  select(CN, RSCD) |>
  rename(SRV_CN = CN)

# plot has invyr, measyr, lat and lon
fia_plot <- tbl(fia, 'PLOT') |>
  # Narrow and rename columns to facilitate join
  select(CN,STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, SRV_CN, ECOSUBCD, LAT, LON) |>
  rename(PLT_CN = CN)

# research station has state name and abbreviation
fia_ref_research_station <- tbl(fia, 'REF_RESEARCH_STATION') |>
  select(STATECD, STATE_NAME, STATE_ABBR)

plot_location <- fia_plot |>
  semi_join(plots, by = join_by(PLT_CN), copy = TRUE) |>
  left_join(fia_survey, by = join_by(SRV_CN)) |>
  left_join(fia_ref_research_station, by = join_by(STATECD)) |>
  northeastern_plots_filter() |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  # Pick just the last inventory year
  filter(INVYR == max(INVYR, na.rm = TRUE)) |> 
  ungroup() |>
  collect() |>
  rename(
    long = LON,
    lat = LAT
  ) |>
  mutate(FVS_STAND_ID=sprintf(paste0(
    '%02d',  '%02d',       '%03d',   '%05d'),
    STATECD, INVYR %% 100, COUNTYCD, PLOT
  )) |>
  mutate(group = sprintf('%02d%03d%05d', STATECD, COUNTYCD, PLOT))

dbDisconnect(fia)
remove(fia, fia_plot, fia_survey, fia_ref_research_station)
```

What's the most recent inventory?

```{r}
plot_location |>
  group_by() |>
  summarize(INVYR = max(INVYR), .groups = "keep")
```

How many plots are in each state?

```{r state_plot_counts}
state_plot_counts <- plot_location |>
  group_by(STATE_NAME) |>
  summarize(COUNT = n()) |>
  ungroup() |>
  arrange(desc(COUNT))
knitr::kable(state_plot_counts)
```

Show the plots on a map.

```{r map_states}
northeastern_states <- plot_location |>
  distinct(STATE_NAME) |>
  mutate(region = str_to_lower(STATE_NAME))
map_states <- map_data('state', northeastern_states$region)

ggplot(map_states, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + 
  geom_point(data = plot_location, color = alpha("blue", alpha = 0.5)) +
  coord_quickmap()
```

## Prescriptions

### Overview

We want to model a few different harvesting rates:

-  0%/yr, 5%/yr, 10%, 20%, and 40%.
-  Because the time step is 5 years, those rates needs to be changed to:
   -  5%/yr --\> 22.6%/5yr
   -  10%/yr --\> 41%/5yr
   -  20%/yr --\> 67%/5yr, and
   -  40%/yr --\> 92%/5yr.
   
These values are calculated by the equation 1 - (1 - rate/year)\^5

### Grow-Only

### Regeneration

```{r table4_rot}
table4 <- read_csv(
  "../00_rawData/NK_Table_4.csv",
  col_types = cols(
    `Management scenario` = col_character(),
    .default = col_number(),
  )
)

# Table4 uses scientific names; translate to get FVS Species Code
fvs_species_codes <- read_csv(
    "../01_processedInputData/FVSne_Overview_Table_3.2.1.csv",
    col_types = "iiciccc"
  ) |>
  select("Scientific Name", "Species Code", "Common Name")

# table4 has one column per species. We wish to add observations per species
# with different types, so pivot the table to have one row per species.
table4_rot <- table4 |>
  pivot_longer(cols = !`Management scenario`) |>
  pivot_wider(names_from = `Management scenario`) |>
  rename(`Scientific Name`=name) |>
  left_join(fvs_species_codes, by=join_by(`Scientific Name`))

# Values in NK are seedlings per hectare; FVS needs seedlings per acre
table4_rot$Clearcut <-
  round(conv_unit(table4_rot$Clearcut, 'hectare', 'acre'))
table4_rot$Shelterwood <-
  round(conv_unit(table4_rot$Shelterwood, 'hectare', 'acre'))
table4_rot$`ITS_Low Retention` <-
  round(conv_unit(table4_rot$`ITS_Low Retention`, 'hectare', 'acre'))
table4_rot$`ITS_High Retention` <-
  round(conv_unit(table4_rot$`ITS_High Retention`, 'hectare', 'acre'))
table4_rot$Background <-
  round(conv_unit(table4_rot$Background, 'hectare', 'acre'))
```

The regeneration section of a keywordfile looks like

```         
Natural            1        QA     37857       100                             0
Natural            1        PB     37857       100                             0
Natural            1        WA     20149       100                             0
Natural            1        YB     19998       100                             0
Natural            1        RM     19998       100                             0
Natural            1        SM     10991       100                             0
Natural            1        AB      4275       100                             0
Natural            1        EH      1067       100                             0
Natural            1        RS      1067       100                             0
```

From FVS Report INT-279, Appendix C, Page 30, re: the Natural keyword:

-   Field 1: Year that trees per acre reaches the density coded in
    field 3. Default is 1.0.
-   Field 2: Numeric species code, see table 1.
    *Note: Two-letter species code can also be used.*
-   Field 3: Trees per acre.
-   Field 4: Percent survival at the end of the cycle. A blank field is
    interpreted as 100 percent survival.
    *Note: Unclear why one would use a value \< 100 here, rather than a lower TPA.*
-   Field 5: Average seedling age in the year coded in field 1; default
    is 2 years.
-   Field 6: Average seedling height.
    *Note: There's a lot more on this; never seen it used.*
-   Field 7: Shade code.
    -   Blank or 0 = default = seedlings occur uniformly on plots
        throughout the stand.
    -   1 = seedlings occur more frequently on plots with more overstory
        basal area (in the shade).
    -   2 = seedlings occur more frequently on plots with less overstory
        basal area (in the sun).

Generate a natural regen section for Clearcut:

```{r natural_regen_clearcut}
natural_regen_clearcut <- table4_rot |>
  select(`Species Code`, Background) |>
  filter(!is.na(Background)) |>
  arrange(desc(Background)) |>
  mutate(
    KEYWORD = paste0(
      sprintf("%-10s", "Natural"),        # Keyword
      sprintf("%10d", 1),                 # Years to reach density.
      sprintf("%10s", `Species Code`),    # Species Code
      sprintf("%10d", Background),        # TPA
      sprintf("%10d", 100),               # Survival %
      sprintf("%10s", ""),                # Age
      sprintf("%10s", ""),                # Height
      sprintf("%10d", 0)                  # Shade Code
    )
  )
natural_regen_clearcut_str <- ''
for (kw in natural_regen_clearcut$KEYWORD) {
  natural_regen_clearcut_str <- paste0(natural_regen_clearcut_str, kw, "\n")
}
```

## Run FVS

### Keword File Generation

Gather stand IDs and CNs for keywordfile generation.

```{r keywordfile_section}
keywordfile_section <- function(Title, MgmtId, StandID, StandCN, InvYear, CutEff, Partition) {
  InvYear <- as.numeric(InvYear)
  TimeConfig <- NULL
  if (InvYear == 2022) {
    TimeConfig <- paste0(
      fvs_kwd1("InvYear", InvYear),
      fvs_kwd2("TimeInt", 0, 5),
      fvs_kwd1("NumCycle", 30)
    )
  } else {
    TimeConfig <- paste0(
      fvs_kwd1("InvYear", InvYear),
      fvs_kwd2("TimeInt", 0, 5),
      fvs_kwd2("TimeInt", 1, 2022 - InvYear),
      fvs_kwd1("NumCycle", 31)
    )
  }
  paste0(
    fvs_kwd0("StdIdent"),
    StandID, " ", Title, "\n",
    fvs_kwd0("StandCN"),
    StandCN, "\n",
    fvs_kwd0("MgmtId"),
    MgmtId, "\n",
    TimeConfig,
    fvs_kwd0("FMIn"), # Fire and Fuels Extension
    fvs_kwd1("CarbRept", 2),
    fvs_kwd1("CarbCut" , 2),
    fvs_kwd5("CarbCalc", 1, 1, 0.0425, 9, 11),
    fvs_kwd1("FuelOut" , 2),
    fvs_kwd1("FuelRept", 2),
    fvs_kwd0("End"), # FMIn
    fvs_kwd0("Database"), # Database extension
    fvs_kwd0("DSNIn"),
    "SQLite_FIADB_ENTIRE.db", "\n",
    fvs_kwd0("StandSQL"),
    "SELECT * FROM FVS_StandInit_Plot WHERE Stand_CN = '%Stand_CN%'", "\n",
    fvs_kwd0("EndSQL"), # StandSQL
    fvs_kwd0("TreeSQL"),
    "SELECT * FROM FVS_TreeInit_Plot WHERE Stand_CN = '%Stand_CN%'", "\n",
    fvs_kwd0("EndSQL"), # TreeSQL
    fvs_kwd0("DSNOut"),
    "FVS_", Title, "_", MgmtId, "_", Partition, ".db", "\n",
    fvs_kwd1("Summary",  2),
    fvs_kwd2("Computdb", 0, 1),
    fvs_kwd1("MisRpts",  2),
    fvs_kwd1("CarbReDB", 2),
    fvs_kwd1("FuelReDB", 2),
    fvs_kwd1("FuelsOut", 2),
    fvs_kwd0("End"), # Database
    # Condition: Every 20 years
    fvs_kwd1("If", 20),
    "year ge 0", "\n",
    fvs_kwd0("Then"),
    fvs_kwd6("ThinDBH",  0, 5, 999, CutEff, "All", 0),
    fvs_kwd1("Estab", 0),
    fvs_kwd0("Sprout"),
    natural_regen_clearcut_str,
    fvs_kwd0("End"), # Estab
    fvs_kwd0("EndIf"), # Every 20 years
    fvs_kwd0("Process")
  )
}
```

Find the stands we want to project; these are the grow-only stands
starting in their first measurement year.

```{r stands_projected}
first_measyear <- plot_location |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(
    INVYR = max(INVYR),
    MEASYEAR = max(MEASYEAR, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

fia <- DBI::dbConnect(RSQLite::SQLite(), '../00_rawData/SQLite_FIADB_ENTIRE.db')
fia_fvs_standinit_plot <- tbl(fia, 'FVS_StandInit_Plot') |>
  select(STAND_CN, STAND_ID)

stands_projected <- tbl(fia, 'PLOT') |>
  select(CN, STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, DESIGNCD) |>
  filter(DESIGNCD == 1) |>
  rename(STAND_CN = CN) |>
  semi_join(first_measyear, by = join_by(STATECD, COUNTYCD, PLOT, INVYR), copy = TRUE) |>
  left_join(fia_fvs_standinit_plot, by = join_by(STAND_CN)) |>
  collect()

dbDisconnect(fia)
remove(fia, fia_fvs_standinit_plot)
```

The first run is to project from the first measurement year to 2025.
These are given management ID "NONE".

```{r FVS_Repro_X100_part.key}
for (cuteff in c(100, 92, 67, 41, 22, 0)) {
  mgt_id <- sprintf('X%03d', cuteff)
  for (partition in 1:8) {
    filename <- paste0("../03_outputs/FVS_Northeastern_", mgt_id, "_", partition, ".key")
    write_lines(
      apply(
        stands_projected |> filter((as.numeric(STAND_ID) %% 8) == (partition - 1)),
        1,
        function(row) {
          keywordfile_section(
            "Northeastern",
            mgt_id,
            row['STAND_ID'],
            row['STAND_CN'],
            row['MEASYEAR'],
            cuteff / 100,
            partition
          )
        }
      ),
      filename
    )
    write_lines("Stop", filename, append = TRUE)
  }
}
```

### Run FVS.

```{r run_fvs}
# TODO nik: Run FVS directly from R.
# Batch file to run FVS in parallel:
# for %%p in (FVS_Northeastern_NONE_*.key) do start \FVS\FVSbin\FVSne.exe --keywordfile=%%p
```

## Results

Load the output from all FVS partitions into a single dataframe

```{r FVS_Carbon}
FVS_Carbon <- NULL
FVS_Summary2_East <- NULL

for (cuteff in c(100, 92, 67, 41, 22, 0)) {
  mgt_id <- sprintf('X%03d', cuteff)
  for (partition in 1:8) {
    fvs_out_db <- DBI::dbConnect(
      RSQLite::SQLite(),
      paste0('../03_outputs/FVS_Northeastern_', mgt_id, '_', partition, '.db')
    )
    fvs_carbon_tbl = tbl(fvs_out_db, 'FVS_Carbon') |>
      mutate(MgmtId = mgt_id) |>
      mutate(Partition = partition)
    fvs_summary2_east_tbl = tbl(fvs_out_db, 'FVS_Summary2_East') |>
      mutate(MgmtId = mgt_id) |>
      mutate(Partition = partition)
  
    FVS_Carbon <- bind_rows(FVS_Carbon, fvs_carbon_tbl |> collect())
    FVS_Summary2_East <- bind_rows(FVS_Summary2_East, fvs_summary2_east_tbl |> collect())
    
    dbDisconnect(fvs_out_db)
  }
}

remove(fvs_out_db, fvs_carbon_tbl, fvs_summary2_east_tbl)
```

Filter projections to corresponding observations

```{r projected_ba}
decode_mgmt_id <- function(mgmt_id) {
  if_else(
    mgmt_id == 'X100', 'Clearcut',
    if_else(mgmt_id == 'X092', '40%/yr',
      if_else(mgmt_id == 'X067', '20%/yr',
        if_else(mgmt_id == 'X041', '10%/yr',
          if_else(mgmt_id == 'X022', '05%/yr',
            'Grow Only'
          )
        )
      )
    )
  )
}

projected_ba <- FVS_Carbon |>
  select(StandID, MgmtId, Year, Aboveground_Total_Live) |>
  rename(Carbon = Aboveground_Total_Live) |>
  mutate(Prescription = decode_mgmt_id(MgmtId)) |> 
  group_by(Prescription, Year) |>
  summarize(Carbon = mean(Carbon), .groups = "keep") |>
  ungroup()
ggplot(projected_ba, aes(Year, Carbon, color = Prescription)) +
  geom_line() +
  coord_cartesian(ylim = c(0, 200)) +
  ylab('Aboveground Live Carbon (Mg/ha)') +
  ggtitle('FVS Carbon Projection vs. Harvest Prescription')

```


