---
title: "FVS Carbon Modeling Analysis"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

# Setup

```{r setup, include = FALSE}
library(targets)
library(tarchetypes)
library(crew)
```

# Introduction

In [@nunery2010], different management regimens were modeled to explore
their impact on carbon storage. This study used the Forest Vegetation
Simulator (FVS) to model carbon storage. Other models exist to predict
carbon storage; we seek to compare FVS with other models to understand
their different predictions for carbon storage under different
management regimens.

# Targets

This project uses [`targets`](https://docs.ropensci.org/targets) and
[Targets Markdown](https://books.ropensci.org/targets/markdown.html). You can
interact with the notebooks in the normal way for any R code; in addition, when
you knit the notebooks, file `_targets.R` and the folder `_targets_r/` will be
updated such that running `tar_make()` will do a minimal rebuild of the project.

## Targets Cleanup

Two commands are useful to clean up stale objects cached by targets:

- `tar_prune()` - removes from `_targets/` cached objects that are no longer defined in the project
- `tar_unscript()` - removes build scripts from `_targets_r/` so they can be rebuilt

`tar_prune()` it makes sense to run on every build; `tar_unscript()` is run
once, here, to ensure that knitting will produce a clean build.

```{r tar_unscript, include = FALSE}
tar_unscript()
```

## Targets Globals

This is the global configuration for Targets.

```{targets globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(
  packages = c(
    "tarchetypes",
    "tidyverse",
    "RSQLite",
    "measurements",
    "reshape2",
    "maps",
    "pbapply",
    "digest"
  ),
  controller = crew::crew_controller_local(workers = 8)
)
tar_source()
```

# Database Preparation

This project depends heavily on the SQLite version of the FIA data, distributed
as SQLite_FIADB_ENTIRE.db. Place a copy of this file in data/raw/
The database ships with most indexes missing, which makes most queries very slow.


```{targets fiadb}
# Cannot use tar_simple = TRUE for targets with format = "file"
tar_target(fiadb, fia_fiadb_indexed(), format = "file")
```

## Primary and Foreign Keys

Most tables with a simple primary key name that key "CN", for Control Number.
Many tables have foreign keys named <table abbreviation>_CN.
Create indexes on primary and foreign keys to speed up joins.

Note that some tables have a CN column that is not a primary key. Rather
than manage this, we create the primary key indexes witout the UNIQUE
constraint. This does not change the structure of the index, but may impact
query plans.

## Plot Identifiers

Many tables have the three columns that uniquely identify a plot:
STATECD, COUNTYCD, PLOT
Index those because we often want to look up all the records pertaining to
a plot by location.

If a table also has INVYR, this uniquely identifies a plot at a point of time.
Index those too.

# Species Crosswalk

Different systems use different species identifiers; create a single crosswalk
that connects them all.

## FVSne Species Identifiers

The FVSne Overview document describes the species codes used in the Northeast
variant of FVS in table 3.2.1.

```{targets fvsne_table_3_2_1}
# tar_file_read is a tarchetypes metatarget, so can't use tar_simple
tarchetypes::tar_file_read(
  fvsne_table_3_2_1,
  "data/raw/FVSne_Overview_Table_3.2.1.csv",
  read_csv(!!.x, col_types = "iiciccc")
)
```

```{r fvsne_table_3_2_1_preview, eval = FALSE}
knitr::kable(fvsne_table_3_2_1 |> head(5))
```

## Jenkins Species Group Names

FIADB has a table of species identifiers, but that table omits Jenkins
species group names. The FIADB Database Description has a list of Jenkins
Species Group names in table 11.5.17.

```{targets fia_table_11_5_17}
# tar_file_read is a tarchetypes metatarget, so can't use tar_simple
tarchetypes::tar_file_read(
  fia_table_11_5_17,
  "data/raw/FIADB_11.5.17_JENKINS_SPGRPCD.csv",
  read_csv(!!.x, col_types = cols(JENKINS_SPGRPCD = col_integer()))
)
```

```{r fia_table_11_5_17_preview, eval = FALSE}
knitr::kable(fia_table_11_5_17 |> head(5))
```

## FIADB Species Identifiers

FIADB spreads species information across tables REF_SPECIES and REF_SPECIES_GROUP.
We can merge those with FVSne and FIADB Jenkins Group information to form
a single species crosswalk.

```{targets fia_ref_species, tar_simple = TRUE}
con <- DBI::dbConnect(RSQLite::SQLite(), fiadb, flags = SQLITE_RO)
on.exit(dbDisconnect(con), add = TRUE, after = FALSE)

tbl(con, 'REF_SPECIES') |>
  select(SPCD, GENUS, SPECIES, E_SPGRPCD, JENKINS_SPGRPCD, SCIENTIFIC_NAME, COMMON_NAME) |>
  left_join(
    tbl(con, 'REF_SPECIES_GROUP') |> distinct(SPGRPCD, NAME),
    by = join_by(E_SPGRPCD == SPGRPCD)
  ) |>
  rename(SPGRPCD = E_SPGRPCD, SPGRP_NAME = NAME) |>
  collect()
```

```{r fia_ref_species_preview, eval = FALSE}
knitr::kable(fia_ref_species |> head(5))
```

## Hardwood and Softwood Identification

In addition, the FIADB User Guides Volume Database Description (version 9.2),
Appendix E: Tree Species Group Codes gives the following description of hard
vs. softwoods (numbers refer to SPCD values):

 - Softwoods: 1 - 24
 - Hardwoods: 25 - 48
 - Tropical and subtropical: 51 - 54
 - Urban: 55, 56
 
To decode Tropical, Subtropical, and Urban tree soft vs. hardwood, we observe
that SPGRP_NAME often has hardwood or softwood in the name, and, less
frequently, JENKINS_SPGRP_NAME does.

```{targets species_crosswalk, tar_simple = TRUE}
jenkins_spgrpcd_mixin <- fia_table_11_5_17 |>
  select(JENKINS_SPGRPCD, NAME) |>
  rename(JENKINS_SPGRP_NAME = NAME)

fvs_spcd_mixin <- fvsne_table_3_2_1 |>
  select(`FIA Code`, `Species Number`, `Species Code`) |>
  rename(
    SPCD = `FIA Code`,
    FVS_SPNO = `Species Number`,
    FVS_SPCD = `Species Code`
  )

fia_ref_species |>
  left_join(jenkins_spgrpcd_mixin, by = join_by(JENKINS_SPGRPCD)) |>
  left_join(fvs_spcd_mixin, by = join_by(SPCD)) |>
  mutate(HARD_SOFT = case_when(
    SPGRPCD %in% 1:24 ~ "SOFTWOOD",
    SPGRPCD %in% 25:48 ~ "HARDWOOD",
    grepl("hardwood", SPGRP_NAME) ~ "HARDWOOD",
    grepl("softwood", SPGRP_NAME) ~ "SOFTWOOD",
    grepl("hardwood", JENKINS_SPGRP_NAME) ~ "HARDWOOD",
    grepl("softwood", JENKINS_SPGRP_NAME) ~ "SOFTWOOD",
    .default = NA
  ))
```

```{r species_crosswalk_preview, eval = FALSE}
knitr::kable(species_crosswalk |> filter(!is.na(FVS_SPNO)) |> head(5))
```
