---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)
library(RSQLite)
library(dbplyr)
```


# Introduction

In [@nunery2010], different management regimens were modeled to explore
their impact on carbon storage. This study used the Forest Vegetation
Simulator (FVS) to model carbon storage. Other models exist to predict
carbon storage; we seek to compare FVS with other models to understand
their different predictions for carbon storage under different
management regimens.

# Database Preparation

This project depends heavily on the SQLite version of the FIA data, distributed
as SQLite_FIADB_ENTIRE.db. Place a copy of this file in data/raw/
The database ships with most indexes missing, which makes most queries very slow.

## Primary and Foreign Keys

Most tables with a simple primary key name that key "CN", for Control Number.
Many tables have foreign keys named <table abbreviation>_CN.
Create indexes on primary and foreign keys to speed up joins.

Note that some tables have a CN column that is not a primary key. Rather
than manage this, we create the primary key indexes witout the UNIQUE
constraint. This does not change the structure of the index, but may impact
query plans.

```{r primary_and_foreign_keys}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
# We need to perform a correlated subquery through a pragma
# (which looks like a table function); dbplyr doesn't support this.
# We also need to do a cross join, which is also not supported by dbplyr.
# So we run this query using raw DBI.
column_query <- "
  SELECT t.name AS table_name, c.name AS column_name
  FROM sqlite_schema t
  CROSS JOIN pragma_table_info((t.name)) c
  WHERE column_name = 'CN' OR column_name LIKE '%_CN'
  ORDER BY table_name, column_name;
"

columns <- DBI::dbGetQuery(fia, column_query)

idx_col <- function(row) {
  tbl <- row["table_name"]
  col <- row["column_name"]
  idx_name <- paste0("IDX_", tbl, "_", col)
  sql <- paste(
    "CREATE INDEX IF NOT EXISTS", idx_name, "ON", tbl, "(", col, ");",
    sep = ' '
  )
  DBI::dbExecute(fia, sql)
}

apply(columns, 1, idx_col) |>
  length()

dbDisconnect(fia)
remove(fia, column_query, columns, idx_col)
```

## Plot Identifiers

Many tables have the three columns that uniquely indentify a plot:
STATECD, COUNTYCD, PLOT
Index those because we often want to look up all the records pertaining to
a plot by location.

If a table also has INVYR, this uniquely identifies a plot at a point of time.
Index those too.

```{r plot_location}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
table_query <- "
  SELECT
    table_name,
    SUM(has_statecd) & SUM(has_countycd) & SUM(has_plot) AS has_plt,
    SUM(has_statecd) & SUM(has_countycd) & SUM(has_plot) & SUM(has_invyr) AS has_pltyr
  FROM (
    SELECT
      t.name AS table_name,
      IIF(c.name = 'STATECD', 1, 0) AS has_statecd,
      IIF(c.name = 'COUNTYCD', 1, 0) AS has_countycd,
      IIF(c.name = 'PLOT', 1, 0) AS has_plot,
      IIF(c.name == 'INVYR', 1, 0) AS has_invyr
    FROM sqlite_schema t
    CROSS JOIN pragma_table_info((t.name)) c
  )
  GROUP BY table_name
  HAVING has_plt
  ORDER BY table_name;
"

tables <- DBI::dbGetQuery(fia, table_query)

idx_tbl <- function(row) {
  tbl <- row["table_name"]
  idx_name <- paste0("IDX_", tbl, "_PLOT")
  sql <- paste(
    "CREATE INDEX IF NOT EXISTS", idx_name, "ON", tbl, "(STATECD, COUNTYCD, PLOT);",
    sep = ' '
  )
  DBI::dbExecute(fia, sql)
  plotyr <- row["has_pltyr"]
  if (plotyr == 1) {
    idx_name <- paste0("IDX_", tbl, "_PLOT_INVYR")
    sql <- paste(
      "CREATE INDEX IF NOT EXISTS", idx_name, "ON", tbl, "(STATECD, COUNTYCD, PLOT, INVYR);",
      sep = ' '
    )
    DBI::dbExecute(fia, sql)
  }
}

apply(tables, 1, idx_tbl) |>
  length()

dbDisconnect(fia)
remove(fia, table_query, tables, idx_tbl)
```


# Project Plan
```{r}
library(DiagrammeR)

create_graph() |>

  # =============
  # FIA provided stuff
  # =============

  
  # It all starts with FIADB
  add_node(
    label = "FIADB",
    type = "data",
    node_data = node_data(provider = "FIA")
  ) |>
  
  # FIA provides a converter to create FVS Stands
  add_node(
    label = "FIA2FVS",
    type = "process",
    node_data = node_data(provider = "FIA")
  ) |>
  add_edge(from = "FIADB", to = "FIA2FVS") |>
  add_node(
    label = "FVS Stand",
    type = "data",
    node_data = node_data(provider = "FIA")
  ) |>
  add_edge(from = "FIA2FVS", to = "FVS Stand") |>
  
  # FIA has a TPA-based carbon accounting method
  add_node(
    label = "FIA TPA\nCarbon Calc",
    type = "process",
    node_data = node_data(provider = "FIA")
  ) |>
  add_edge(from = "FIADB", to = "FIA TPA\nCarbon Calc") |>
  add_node(
    label = "TPA 0-year\nCarbon",
    type = "data",
    node_data = node_data(provider = "FIA")
  ) |>
  add_edge(from = "FIA TPA\nCarbon Calc", to = "TPA 0-year\nCarbon") |>
  
  # =============
  # FVS-provided stuff
  # =============
  
  # FVS 0-year projection
  add_node(
    label = "FVS 0-Year\nProjection",
    type = "process",
    node_data = node_data(provider = "FVS")
  ) |>
  add_edge(from = "FVS Stand", to = "FVS 0-Year\nProjection") |>
  add_node(
    label = "FVS 0-Year\nOutput",
    type = "data",
    node_data = node_data(provider = "FVS")
  ) |>
  add_edge(from = "FVS 0-Year\nProjection", to = "FVS 0-Year\nOutput") |>

  # FVS 20-year projection
  add_node(
    label = "FVS 20-Year\nProjection",
    type = "process",
    node_data = node_data(provider = "FVS")
  ) |>
  add_edge(from = "FVS Stand", to = "FVS 20-Year\nProjection") |>
  add_node(
    label = "FVS 20-Year\nOutput",
    type = "data",
    node_data = node_data(provider = "FVS")
  ) |>
  add_edge(from = "FVS 20-Year\nProjection", to = "FVS 20-Year\nOutput") |>
  
  # FVS FFE - Fire and Fuels Extension computes carbon
  add_node(
    label = "FVS FFE",
    type = "process",
    node_data = node_data(provider = "FVS")
  ) |>
  add_edge(from = "FVS 0-Year\nOutput", to = "FVS FFE") |>
  add_edge(from = "FVS 20-Year\nOutput", to = "FVS FFE") |>
  add_node(
    label = "FFE 0-Year\nCarbon",
    type = "data",
    node_data = node_data(provider="FVS")
  ) |>
  add_edge(from = "FVS FFE", to = "FFE 0-Year\nCarbon") |>
  add_node(
    label = "FFE 20-Year\nCarbon",
    type = "data",
    node_data = node_data(provider="FVS")
  ) |>
  add_edge(from = "FVS FFE", to = "FFE 20-Year\nCarbon") |>
  
  # FVS-Jenkins - Fire and Fuels Extension Jenkins method
  add_node(
    label = "FVS Jenkins",
    type = "process",
    node_data = node_data(provider = "FVS")
  ) |>
  add_edge(from = "FVS 0-Year\nOutput", to = "FVS Jenkins") |>
  add_edge(from = "FVS 20-Year\nOutput", to = "FVS Jenkins") |>
  add_node(
    label = "Jenkins 0-Year\nCarbon",
    type = "data",
    node_data = node_data(provider="FVS")
  ) |>
  add_edge(from = "FVS Jenkins", to = "Jenkins 0-Year\nCarbon") |>
  add_node(
    label = "Jenkins 20-Year\nCarbon",
    type = "data",
    node_data = node_data(provider="FVS")
  ) |>
  add_edge(from = "FVS Jenkins", to = "Jenkins 20-Year\nCarbon") |>

  
  # Harvard Forest does our own carbon computation directly from FIA
  add_node(
    label = "HF Carbon\nCalc",
    type = "process",
    node_data = node_data(provider="HF")
  ) |>
  add_edge(from = "FIADB", to = "HF Carbon\nCalc") |>
  add_node(
    label = "HF 0-Year\nCarbon",
    type = "data",
    node_data = node_data(provider = "HF")
  ) |>
  add_edge(from = "HF Carbon\nCalc", to = "HF 0-Year\nCarbon") |>
  

  # =============
  # Styling
  # =============

  # Edges are black
  set_edge_attrs(edge_attr = color, value = "black") |>
  
  # Nodes resize to accommodate their labels
  set_node_attrs(node_attr = fixedsize, value = FALSE) |>

  # Processes are rectangles
  select_nodes(conditions = type == "process") |>
  set_node_attrs_ws(node_attr = shape, value = "rectangle") |>
  clear_selection() |>
  
  # Data is rounded
  select_nodes(conditions = type == "data") |>
  set_node_attrs_ws(node_attr = shape, value = "rectangle") |>
  set_node_attrs_ws(node_attr = style, value = "rounded") |>
  clear_selection() |>
  
  # FIA-provided is green
  select_nodes(conditions = provider == "FIA") |>
  set_node_attrs_ws(node_attr = color, value = "green") |>
  clear_selection() |>
  
  # FVS-provided is red
  select_nodes(conditions = provider == "FVS") |>
  set_node_attrs_ws(node_attr = color, value = "red") |>
  clear_selection() |>

  # HF-provided is black
  select_nodes(conditions = provider == "HF") |>
  set_node_attrs_ws(node_attr = color, value = "black") |>
  clear_selection() |>


  render_graph(layout = "tree")
```
