---
title: "FVS Carbon Modeling Analysis"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

# Setup

```{r setup, include=FALSE}
library(targets)
library(tidyverse)
library(reshape2) # for melt()
library(ggplot2)
tar_source()
```

# Reconstructing NK Figure 2

To validate FVS runs, we reconstruct the table behind Figure 2. To do
this, we use the tool <https://automeris.io/wpd/> which aids in
estimating the tabular data behind a chart.

The project used to extract the table is in
[config/AutoMeris_NK_Fig2.json]. The tool is imperfect, so we apply
some cleanup to produce an approximate table.

Note that all values are approximate. For example, all values should
align in 2005, but they do not; we ignore this, since small differences
are within the error bounds for parsing the chart.

```{targets nk_fig_2}
nk_read_fig_2 <- function(dir) {
  nk_read_fig_2_helper <- function(dir, column) {
    filename <- file.path(dir, paste0("NK_Fig2_", column, ".csv"))
    read_csv(
      filename,
      col_names = c("Year", "value"),
      # Read columns as double, since that is how they were written
      col_types = "dd"
    ) |>
    # Round columns to nearest integer, since that is what was intended
    mutate(
      Year = round(Year),
      name = column,
      value = round(value)
    )
  }
  columns <- c(
    "NoManagement",
    "ClearcutHigh", "ClearcutLow",
    "ShelterwoodHigh", "ShelterwoodLow",
    "ITS_LowHigh", "ITS_LowLow", "ITS_HighHigh", "ITS_HighLow"
  )
  bind_rows(lapply(columns, \(column) nk_read_fig_2_helper(dir, column) )) |>
    pivot_wider(id_cols = "Year") |>
    # The last year in Fig2 is 2164; nudge it to 2165
    # to align with contemporary FVS runs.
    mutate(Year = if_else(Year == 2164, 2165, Year)) |>
    arrange(Year)
}

# By specifying the folder as the file source,
# we make nk_fig_2 depend on all files in the folder.
tarchetypes::tar_file_read(
  nk_fig_2,
  "data/raw/nk_fig2/",
  nk_read_fig_2(!!.x)
)
```

```{r nk_fig_2_preview, eval = FALSE}
ggplot(
    data = melt(nk_fig_2, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("NK Fig. 2 Reconstruction") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 200)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,200,20))
```


