---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
require(tidyverse)
require(reshape2) # for melt()
require(RSQLite)
require(htmltools)
require(rmarkdown)
require(dbplyr)
```

# Reconstructing NK Figure 2

To validate FVS runs, we reconstruct the table behind Figure 2. To do
this, we use the tool <https://automeris.io/wpd/> which aids in
estimating the tabular data behind a chart.

The project used to extract the table is in
[config/AutoMeris_NK_Fig2.json]. The tool is imperfect, so we apply
some cleanup to produce an approximate table.

Note that all values are approximate. For example, all values should
align in 2005, but they do not; we ignore this, since small differences
are within the error bounds for parsing the chart.

```{r}
NK_Fig2_NoManagement <- read_csv(
    "data/raw/NK_Fig2_NoManagement.csv",
    col_names = c("Year","NoManagement"),
    # Read columns as double, since that is how they were written
    col_types = "dd"
  ) |>
  # Round columns to nearest integer, since that is what was intended
  mutate(Year=round(Year), NoManagement=round(NoManagement))
NK_Fig2_ClearcutHigh <- read_csv(
    "data/raw/NK_Fig2_ClearcutHigh.csv",
    col_names = c("Year","ClearcutHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutHigh=round(ClearcutHigh))
NK_Fig2_ClearcutLow <- read_csv(
    "data/raw/NK_Fig2_ClearcutLow.csv",
    col_names = c("Year","ClearcutLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ClearcutLow=round(ClearcutLow))
NK_Fig2_ShelterwoodHigh <- read_csv(
    "data/raw/NK_Fig2_ShelterwoodHigh.csv",
    col_names = c("Year","ShelterwoodHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodHigh=round(ShelterwoodHigh))
NK_Fig2_ShelterwoodLow <- read_csv(
    "data/raw/NK_Fig2_ShelterwoodLow.csv",
    col_names = c("Year","ShelterwoodLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ShelterwoodLow=round(ShelterwoodLow))
NK_Fig2_ITS_LowHigh <- read_csv(
    "data/raw/NK_Fig2_ITS_LowHigh.csv",
    col_names = c("Year","ITS_LowHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowHigh=round(ITS_LowHigh))
NK_Fig2_ITS_LowLow <- read_csv(
    "data/raw/NK_Fig2_ITS_LowLow.csv",
    col_names = c("Year","ITS_LowLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_LowLow=round(ITS_LowLow))
NK_Fig2_ITS_HighHigh <- read_csv(
    "data/raw/NK_Fig2_ITS_HighHigh.csv",
    col_names = c("Year","ITS_HighHigh"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighHigh=round(ITS_HighHigh))
NK_Fig2_ITS_HighLow <- read_csv(
    "data/raw/NK_Fig2_ITS_HighLow.csv",
    col_names = c("Year","ITS_HighLow"),
    col_types = "dd"
  ) |>
  mutate(Year=round(Year), ITS_HighLow=round(ITS_HighLow))

# Assemble a single table from 2005 - 2165
fig2 <- NK_Fig2_NoManagement |>
  full_join(NK_Fig2_ClearcutHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ClearcutLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ShelterwoodLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_LowLow, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighHigh, by=join_by(Year)) |>
  full_join(NK_Fig2_ITS_HighLow, by=join_by(Year)) |>
  filter(Year>=2005) |>
  # The last year in Fig2 is 2164; nudge it to 2165
  # to align with contemporary FVS runs.
  mutate(Year=if_else(Year==2164,2165,Year)) |>
  arrange(Year)

write_csv(
  fig2,
  "data/intermediate/nk_fig2_reconstructed.csv"
)

ggplot(
    data = melt(fig2, id.vars = "Year"),
    mapping = aes(x = Year, y = value, color = variable)
  ) +
  ggtitle("NK Fig2 Reconstruction") +
  ylab("Carbon (Tonnes / Hectare)") +
  theme(legend.title = element_blank()) +
  geom_line() +
  coord_cartesian(xlim = c(2005, 2165), ylim = c(0, 200)) +
  scale_x_continuous(breaks=seq(2005,2165,20)) +
  scale_y_continuous(breaks=seq(0,200,20))
```


