---
title: "056_NRSCalib10"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

# Setup

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(targets)
  library(tarchetypes)
  library(tidyverse)
  library(dbplyr) # for window_order
  library(reshape2) # melt()
  library(measurements) # conv_unit and conv_multiunit
  library(maps)
  library(ggspatial) # for maps
  library(ggpmisc) # for stat_poly_eq
})
options(scipen = 9) # Use integer notation for numbers under 9 digits
tar_source()
```


## Load targets results from other documents

```{r load tar objects defined elsewhere, eval = FALSE}
tar_load(fiadb)
tar_load(species_crosswalk)
```

## Preload everything from 05_NRSGrowOnly.Rmd and 055_NRSCalibration.Rmd

```{r load 05 nrsgrowonly, eval = FALSE}
this_rmd_file <- get_this_rmd_file()
this_rmd_file |>
  str_replace(basename(this_rmd_file), "05_NRSGrowOnly.Rmd") |>
  tar_objects_defined_in_rmd() |>
  tar_load()
this_rmd_file |>
  str_replace(basename(this_rmd_file), "055_NRSCalibration.Rmd") |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```


## Preload targets results from this document

When tar_make() has already built the objects defined in this document,
run this block to pre-load them. You can then skip running ```{targets} blocks.

```{r load tar objects defined here, eval = FALSE}
get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```

# Longer Calibration Period

Calibration using 5 years of growth and mortality was unhelpful.
What happens if we use more years of data to calibrate?

FVSne is tuned to 10 years for everything, so that's a possible choice. For
the small tree height growth model, this is probably the right interval to use,
since small trees may outgrow the small tree range at larger intervals.

Most of the plots we're looking at have 20 or more years (4 or more inventories)
of data available, so 20 years is another possibility.

## Tree Survival & Mortality

What portion of trees survive for 20 years?

First, what portion of plots have observations over 20 years?
```{r plots with 4 inventories, eval = FALSE}
num_inventories <- nrsgro_plot |>
  distinct(STATECD, COUNTYCD, PLOT, MEASYEAR) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(
    INVENTORIES = n(),
    DURATION = max(MEASYEAR) - min(MEASYEAR),
    .groups = "keep"
  ) |>
  ungroup()
num_inventories |>
  filter(INVENTORIES > 3) |>
  nrow()
```

```{r plot num inventories, eval = FALSE}
num_inventories |>
  ggplot(aes(INVENTORIES)) +
  geom_histogram(binwidth = 1) +
  theme_bw()
```

So, 2220 plots have 4 or 5 inventories. Enough to work with.

```{r plot duration, eval = FALSE}
num_inventories |>
  filter(DURATION > 14) |>
  nrow()
num_inventories |>
  filter(DURATION > 19) |>
  nrow()
num_inventories |>
  ggplot(aes(DURATION)) +
  geom_histogram(binwidth = 1) +
  theme_bw()
```

So all plots have 10+ years (which makes sense, since that's what we selected for);
only 612 plots have duration of at least 20 years.

2199 have duration of at least 15 years.

Let's see how many trees are observed for 10 / 15 / 20 years.

Pulling all trees into memory might be too much; we can do this as:
For each plot,
get all trees for all time
For each tree,
find first live observation year, last live observation year, first dead
observation year

```{r tree observations, eval = FALSE}
tree_obs <- fia_trees_filtered(fiadb, nrsgro_plot, \(.data, con) {
  .data |>
    select(
      STATECD, COUNTYCD, PLOT, SUBP, TREE,
      INVYR, STATUSCD
    ) |>
    group_by(STATECD, COUNTYCD, PLOT, SUBP, TREE) |>
    window_order(INVYR) |>
    mutate(INVNUM = row_number()) |>
    summarize(
      INVYR = min(INVYR),
      FIRST_LIVE = min(INVNUM[STATUSCD == 1]),
      LAST_LIVE = max(INVNUM[STATUSCD == 1]),
      FIRST_DEAD = min(INVNUM[STATUSCD == 2]),
      .groups = "keep"
    ) |>
    ungroup()
  })
```

```{r plot tree obs, eval = FALSE}
tmp_tree_duration <- tree_obs |>
  filter(is.finite(FIRST_LIVE) & is.finite(LAST_LIVE)) |>
  mutate(DURATION = LAST_LIVE - FIRST_LIVE)
tmp_tree_duration |>
  ggplot(aes(DURATION)) +
  geom_histogram(binwidth = 1) +
  theme_bw()
```

118,788 trees that live for at least 10 years; only 18,181 that live for 20.
Oh, that's because some significant portion of the trees recruit after the
first inventory.

OK, let's ask a different question: how many trees are alive two inventories
after the first? And, how many trees are alive once, and dead two inventories
after the first?

```{r plot tree calib, eval = FALSE}
nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  summarize(INVYR = min(INVYR), .groups = "keep") |>
  ungroup() |>
  left_join(tree_obs, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  mutate(
    CALIB_CAT = case_when(
      LAST_LIVE >= 2 ~ "Survivor",
      FIRST_DEAD == 1 ~ "Dead at 5",
      FIRST_DEAD == 2 ~ "Dead at 10",
      !is.na(LAST_LIVE) ~ paste0("Vanished at ", LAST_LIVE),
      .default = "Phantom"
    )
  ) |>
  ggplot(aes(CALIB_CAT)) +
  geom_histogram(stat = "count") +
  geom_text(aes(y = after_stat(count + 2500), label = after_stat(count)), stat = "count") +
  theme_bw()
```

So it seems we should be able to use 10 years of survivors and dead trees.

We need two tables:

## Growth Table

```{targets nrsgro_ca10_calibration, tar_simple = TRUE}
tmp_plot_1 <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 1) |> 
  ungroup()

# All the trees we'll give to FVS for calibration
tmp_tre <- fia_trees(fiadb, tmp_plot_1) |>
  filter(STATUSCD == 1) |>
  select(CN, PLT_CN, STATECD, COUNTYCD, PLOT, INVYR, DIA, HT)

tmp_plot_2 <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 2) |>
  ungroup()

tmp_tre_grow_2 <- fia_trees(fiadb, tmp_plot_2) |>
  filter(STATUSCD == 1) |>
  select(CN, PREV_TRE_CN, DIA, HT) |>
  rename(
    DIA2 = DIA,
    HT2 = HT,
    PREV_TRE_CN = CN,
    CN = PREV_TRE_CN
  )

tmp_plot_3 <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 3) |>
  ungroup()

tmp_tre_grow_3 <- fia_trees(fiadb, tmp_plot_3) |>
  filter(STATUSCD == 1) |>
  select(PREV_TRE_CN, DIA, HT) |>
  rename(
    DIA3 = DIA,
    HT3 = HT
  )

tmp_tre_grow <- tmp_tre_grow_2 |>
  left_join(tmp_tre_grow_3, by = join_by(PREV_TRE_CN))

tree_growth <- tmp_tre |>
  left_join(tmp_tre_grow, by = join_by(CN))

# For FVS, we need:
# STAND_CN
# TREE_CN
# DG
# HTG
# The rest is assumed

tree_growth |>
  filter(!is.na(DIA3) & !is.na(HT3)) |>
  select(PLT_CN, CN, DIA3, HT3) |>
  rename(
    STAND_CN = PLT_CN,
    TREE_CN = CN,
    DG = DIA3,
    HTG = HT3
  )
```

## Mortality Table

We need trees that are alive at INVNUM==1, then dead in either INVNUM==2
or INVNUM==3. We can't just check at INVNUM==3 because trees drop out of the
inventory pretty fast once they're dead.

```{targets nrsgro_ca10_mortality, tar_simple = TRUE}
# All the trees we'll give to FVS for calibration
# There are 7 sets of trees:
# 1. Trees in INVNUM==1 that die in INVNUM==2
# 2. Long-term dead in INVNUM==2
# 3. Trees in INVNUM==1 that die in INVNUM==3
# 4. Long-term dead in INVNUM==3 that was missed in INVNUM==2
# 5. Trees in INVNUM==2 that die in INVNUM==3
# 6. Trees that establish and die in INVNUM==2
# 7. Trees that establish and die in INVNUM==3

# Trees in INVNUM==1
# In theory, we could mine some mortality from INVNUM==1
# (e.g. trees that ingrew and died in the past 5 years), but
# it's messy and unreliable so we skip it.
# We need dead trees in INVNUM==1 so we can check long-term mortality.
tmp_plot_1 <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 1) |>
  ungroup() |>
  select(STATECD, COUNTYCD, PLOT, INVYR, CN)

tmp_tre_1 <- fia_trees(fiadb, tmp_plot_1) |>
  select(
    CN, PREV_TRE_CN, PLT_CN,
    STATECD, COUNTYCD, PLOT, SUBP, CONDID, TREE, INVYR,
    TPA_UNADJ, SPCD, STATUSCD, PREV_STATUS_CD, DIA, HT
  ) |>
  rename(
    TRE_CN0 = PREV_TRE_CN,
    TRE_CN1 = CN,
    STAT0 = PREV_STATUS_CD,
    STAT1 = STATUSCD,
    TPA1 = TPA_UNADJ,
    DIA1 = DIA,
    HT1 = HT
  )

# Trees in INVNUM==2
# We are interrested in dead and live trees.
# This will include both recent and long-term mortality
tmp_plot_2 <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 2) |>
  ungroup()

tmp_tre_2 <- fia_trees(fiadb, tmp_plot_2) |>
  select(
    CN, PREV_TRE_CN, PLT_CN,
    STATECD, COUNTYCD, PLOT, SUBP, CONDID, TREE, INVYR,
    TPA_UNADJ, SPCD, STATUSCD, PREV_STATUS_CD, DIA, PREVDIA, HT
  ) |>
  rename(
    TRE_CN1 = PREV_TRE_CN,
    TRE_CN2 = CN,
    STAT1 = PREV_STATUS_CD,
    STAT2 = STATUSCD,
    TPA2 = TPA_UNADJ,
    DIA1 = PREVDIA,
    DIA2 = DIA,
    HT2 = HT
  )

# Trees in INVNUM==3
# We are interested only in dead trees.
# This will include both recent and long-term mortality
tmp_plot_3 <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 3) |>
  ungroup()

tmp_tre_3 <- fia_trees(fiadb, tmp_plot_3) |>
  filter(STATUSCD == 2) |>
  select(
    CN, PREV_TRE_CN, PLT_CN,
    STATECD, COUNTYCD, PLOT, SUBP, CONDID, TREE, INVYR,
    TPA_UNADJ, SPCD, STATUSCD, PREV_STATUS_CD, DIA, PREVDIA, HT
  ) |>
  rename(
    TRE_CN2 = PREV_TRE_CN,
    TRE_CN3 = CN,
    STAT2 = PREV_STATUS_CD,
    STAT3 = STATUSCD,
    TPA3 = TPA_UNADJ,
    DIA2 = PREVDIA,
    DIA3 = DIA,
    HT3 = HT
  )

# 1. Trees in INVNUM==1 that die in INVNUM==2
# These will have STAT1==1 and STAT2 == 2
# And
# 2. Long-term dead in INVNUM==2
# These will have STAT1==2 and STAT2 == 2
# 6. Trees that establish and die in INVNUM==2
# These will have STAT2==2 and TRE_CN1==NA
tmp_tre_mort_2 <- tmp_tre_2 |>
  filter(STAT2 == 2) |>
  left_join(
    tmp_tre_1 |> select(TRE_CN1, STAT1, TPA1, DIA1, HT1),
    by = join_by(TRE_CN1)
  ) |>
  mutate(
    STAT1 = coalesce(STAT1.y, STAT1.x),
    DIA1 = coalesce(DIA1.y, DIA1.x)
  ) |>
  select(!ends_with(".x") & !ends_with(".y")) |>
  filter(
    is.na(TRE_CN1) | # Case 6
      STAT1 == 1 |   # Case 1
      STAT1 == 2     # Case 2
  )

# 3. Trees in INVNUM==1 that die in INVNUM==3
# 4. Long-term dead in INVNUM==3 that was missed in INVNUM==2
# 5. Trees in INVNUM==2 that die in INVNUM==3
# 7. Trees that establish and die in INVNUM==3
tmp_tre_mort_3 <- tmp_tre_3 |>
  filter(STAT3 == 2) |>
  left_join(
    tmp_tre_2 |> select(TRE_CN2, STAT2, TPA2, DIA2, DIA1, HT2),
    by = join_by(TRE_CN2)
  ) |>
  mutate(
    STAT2 = coalesce(STAT2.y, STAT2.x),
    DIA2 = coalesce(DIA2.y, DIA2.x)
  ) |>
  select(!ends_with(".x") & !ends_with(".y")) |>
  filter(
    is.na(TRE_CN2) | # Case 7 and 4
      STAT2 == 1     # Case 3 and 5
  )

# Combine mortality from INVNUM==2 and INVNUM==3 into one table
tmp_tre_mort <- tmp_tre_mort_2 |>
  # Adapt schema
  mutate(
    TRE_CN3 = NA,
    TPA3 = NA,
    STAT3 = NA,
    DIA3 = NA,
    HT3 = NA
  ) |>
  union_all(
    tmp_tre_mort_3 |>
      mutate(
        STAT1 = NA,
        TRE_CN1 = NA,
        TPA1 = NA,
        DIA1 = NA,
        HT1 = NA
      )
  )

# No tree should be included from both INVYR==2 and INVYR==3
stopifnot(
  tmp_tre_mort |>
    group_by(STATECD, COUNTYCD, PLOT, SUBP, TREE) |>
    summarize(N = n(), .groups = "keep") |>
    ungroup() |>
    filter(N>1) |>
    nrow() == 0
)

# Add remaining info
tmp_cond_mixin <- fia_conds(
    fiadb,
    nrsgro_plot |>
      distinct(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  select(
    STATECD, COUNTYCD, PLOT, CONDID, INVYR, SLOPE, ASPECT
  )

tmp_tre_mort |>
  # Don't bother with TREE - this'll get clobbered later anyway.
  left_join(tmp_cond_mixin, by = join_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR)) |>
  select(-PLT_CN) |>
  left_join(
    tmp_plot_1 |>
      select(STATECD, COUNTYCD, PLOT, CN) |>
      rename(PLT_CN = CN),
    by = join_by(STATECD, COUNTYCD, PLOT)
  ) |>
  filter_add_stand_id() |>
  mutate(
    STAND_CN = PLT_CN,
    STAND_ID = STAND_ID,
    PLOT_CN = PLT_CN,
    PLOT_ID = SUBP,
    STANDPLOT_CN = paste0(STAND_CN, "_", SUBP),
    STANDPLOT_ID = paste0(STAND_ID, "_", SUBP),
    TREE_CN = coalesce(TRE_CN1, TRE_CN2, TRE_CN3),
    TREE_ID = TREE,
    TREE_COUNT = coalesce(TPA3, TPA2, TPA1),
    SPECIES = sprintf("%03d", SPCD),
    HISTORY = if_else(
        # Note: if observed dead in INVNUM==1, then long-term dead;
        # otherwise it's either recent dead or new establishment death
        (!is.na(STAT1) & STAT1 == 2), 8, 6
      ),
    DIAMETER = coalesce(DIA3, DIA2, DIA1),
    HT = coalesce(HT3, HT2, HT1),
    SLOPE = SLOPE,
    ASPECT = ASPECT
  ) |>
  select(
    STAND_CN,
    STAND_ID,
    PLOT_CN,
    PLOT_ID,
    STANDPLOT_CN,
    STANDPLOT_ID,
    TREE_CN,
    TREE_ID,
    TREE_COUNT,
    SPECIES,
    HISTORY,
    DIAMETER,
    HT,
    SLOPE,
    ASPECT
  )

```

## Regeneration Table

```{targets nrsgro_ca10_regen, tar_simple = TRUE}
# fvs_run wants establishment in the form:
# STAND_CN
# SPECIES (FVS_SPCD)
# DENSITY (TPA)
# HEIGHT (FT)
tmp_plot_cn <- nrsgro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  filter(row_number() == 1) |>
  ungroup() |>
  filter_add_stand_id() |>
  select(STAND_ID, CN) |>
  rename(STAND_CN = CN)

nrsgro_cal10_regen <- nrsgro_estab_rate |>
  filter_add_stand_id() |>
  left_join(
    tmp_plot_cn |> select(STAND_ID, STAND_CN),
    by = join_by(STAND_ID)
  ) |>
  select(STAND_CN, STATECD, COUNTYCD, PLOT, SPCD, RATE_PER_ACRE) |>
  mutate(RATE_PER_ACRE = floor(RATE_PER_ACRE * 10)) |> # 10-year timesteps
  rename(DENSITY = RATE_PER_ACRE) |>
  left_join(
    nrsgro_estab_height |>
      select(STATECD, COUNTYCD, PLOT, SPCD, HT),
    by = join_by(STATECD, COUNTYCD, PLOT, SPCD)
  ) |>
  mutate(HT = floor(HT)) |>
  rename(HEIGHT = HT) |>
  left_join(
    species_crosswalk |>
      select(SPCD, FVS_SPCD),
    by = join_by(SPCD)
  ) |>
  rename(SPECIES = FVS_SPCD) |>
  filter(!is.na(SPECIES) & !is.na(DENSITY) & !is.na(HEIGHT)) |>
  select(STAND_CN, SPECIES, DENSITY, HEIGHT)
```

# Run FVS

```{targets nrsgro_ca10}
tar_target(
  nrsgro_ca10,
  {
    # fvs_run wants a table of the form:
    # STAND_ID - arbitrary identifier for a stand
    # STAND_CN - CN for the plot
    # FIRST_YEAR - start of the projection; if this doesn't align with the actual
    #   year (MEASYEAR) of first survey, the stand will be projected from the
    #   survey year to the start of the projection.
    # LAST_YEAR - end of the projection
    timestep <- 10 # years; determined by FVSne variant
    
    plots_for_fvs <- nrsgro_plot |>
      group_by(STATECD, COUNTYCD, PLOT) |>
      arrange(INVYR) |>
      mutate(
        STAND_CN = if_else(row_number() == 1, CN, NA)
      ) |>
      summarize(
        STAND_CN = min(STAND_CN, na.rm = TRUE), # only one will not be NA
        FIRST_YEAR = min(MEASYEAR, na.rm = TRUE),
        LAST_YEAR = max(MEASYEAR, na.rm = TRUE),
        .groups = "keep"
      ) |>
      ungroup() |>
      filter_add_stand_id()
    

    fvsbin_dir <- "/fvs/fvsbin" # TODO: put these in a config file
    fvs_variant <- "fvsne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "NRSGrowOnly"
    mgmt_id <- "CA10"
    
    # We communicate with FVS through files. FVSOnline shows a model in which
    # a "project" (the inputs and outputs of a single FVS run) live in a
    # single directory; we follow that model.
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = fiadb,
      title = title,
      mgmt_id = mgmt_id,
      stands = plots_for_fvs,
      calibration = nrsgro_ca10_calibration,
      calib_mort = nrsgro_ca10_mortality,
      calib_years = 10,
      regen = nrsgro_ca10_regen,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see pattern, below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

# Load Results

## NRS GrowOnly CA10

Carbon

```{targets nrsgro_ca10_carbon, tar_simple = TRUE}
fvs_read_output(nrsgro_ca10, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets nrsgro_ca10_summary, tar_simple = TRUE}
fvs_read_output(nrsgro_ca10, "FVS_Summary2_East") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```
## Calibration Rate

Did we actually get calibration?

```{r nrsgro_calb_calibration cases, eval = FALSE}
single_seed <- nrsgro_ca10 |>
  group_by() |>
  filter(random_seed == min(random_seed)) |>
  ungroup()
nrsgro_ca10_cases <- fvs_read_output(single_seed, "FVS_Cases")
nrsgro_ca10_calibstats <- fvs_read_output(single_seed, "FVS_CalibStats")
num_cases <- nrsgro_ca10_cases |> distinct(CaseID) |> nrow()
num_cases_calibrated <- nrsgro_ca10_calibstats |> distinct(CaseID) |> nrow()
num_cases_calibrated / num_cases
```

61%, so about identical to 5-year calibration.

## Scale Factor


What kind of calibration did we get?
```{r plot calibstats, eval = FALSE}
nrsgro_ca10_calibstats |>
  mutate(Log10ScaleFactor = log10(ScaleFactor)) |>
  ggplot(aes(Log10ScaleFactor)) +
  geom_histogram(binwidth = 0.1) +
  geom_vline(xintercept = 0) +
  theme_bw()
```

looks about the same as 5 year.

# 5- vs 10-year calibration

```{targets nrsgro_ca10_vs_none, tar_simple = TRUE}
stand_mixin <- nrsgro_plot_stats |>
  filter_add_stand_id() |>
  group_by(STAND_ID) |>
  arrange(INVYR) |>
  filter(row_number() == 1) |>
  ungroup() |>
  select(STAND_ID, MEASYEAR, BALIVE_METRIC) |>
  rename(StandID = STAND_ID, StartYear = MEASYEAR)

none_carbon_tmp <- nrsgro_none_carbon |>
  rename(Carbon = Aboveground_Total_Live) |>
  select(StandID, Year, Carbon) |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(Carbon_Delta = Carbon - lag(Carbon)) |>
  ungroup() |>
  filter(!is.na(Carbon_Delta))

ca10_carbon_tmp <- nrsgro_ca10_carbon |>
  rename(Carbon = Aboveground_Total_Live) |>
  select(StandID, Year, Carbon) |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(Carbon_Delta = Carbon - lag(Carbon)) |>
  ungroup() |>
  filter(!is.na(Carbon_Delta))

nrsgro_ca10_vs_none <- none_carbon_tmp |>
  rename(Carbon_None = Carbon, Carbon_Delta_None = Carbon_Delta) |>
  left_join(ca10_carbon_tmp, by = join_by(StandID, Year)) |>
  rename(Carbon_Ca10 = Carbon, Carbon_Delta_Ca10 = Carbon_Delta) |>
  left_join(stand_mixin, by = join_by(StandID)) |>
  mutate(
    Carbon_Diff = Carbon_Ca10 - Carbon_None,
    ProjectionYears = Year - StartYear,
    Carbon_Flux_Ca10 = -(Carbon_Delta_Ca10 / ProjectionYears),
    Carbon_Flux_None = -(Carbon_Delta_None / ProjectionYears),
    Carbon_Flux_Residual = Carbon_Flux_Ca10 - Carbon_Flux_None
  )
```


```{r plot nrsgro_ca10_vs_none flux, eval = FALSE}
nrsgro_ca10_vs_none |>
  group_by(StandID) |>
  arrange(desc(ProjectionYears)) |>
  filter(row_number() == 1) |>
  ungroup() |>
  ggplot(aes(Carbon_Flux_None, Carbon_Flux_Ca10)) +
  geom_point(size = 0.1) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.25) +
  geom_abline(slope = 1, intercept = 0, color = "gray", linewidth = 0.25) +
  #coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  ggtitle("10-Year Calibrated vs. Uncalibrated Annual Carbon Flux") +
  xlab(bquote("Uncalibrated Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% year^-1))) +
  ylab(bquote("Calibrated Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% year^-1))) +
  labs(caption = "With 10-Year Calibration")
```

# Projected vs. Measured

```{targets nrsgro_ca10_proj_vs_meas, tar_simple = TRUE}
stand_mixin <- nrsgro_plot_stats |>
  filter_add_stand_id() |>
  group_by(STAND_ID) |>
  arrange(INVYR) |>
  filter(row_number() == 1) |>
  ungroup() |>
  select(STAND_ID, MEASYEAR, BALIVE_METRIC) |>
  rename(StandID = STAND_ID, StartYear = MEASYEAR)

srvy_carbon_tmp <- nrsgro_srvy_carbon |>
  rename(Carbon = Aboveground_Total_Live) |>
  select(StandID, Year, Carbon) |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(Carbon_Delta = Carbon - lag(Carbon)) |>
  ungroup() |>
  filter(!is.na(Carbon_Delta))

srvy_summary_tmp <- nrsgro_srvy_summary |>
  mutate(BA_METRIC = conv_multiunit(BA, "ft2 / acre", "m2 / hectare")) |>
  select(StandID, Year, BA_METRIC) |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(
    BA_Delta = BA_METRIC - lag(BA_METRIC)
  ) |>
  ungroup() |>
  filter(!is.na(BA_Delta))

ca10_carbon_tmp <- nrsgro_ca10_carbon |>
  rename(Carbon = Aboveground_Total_Live) |>
  select(StandID, Year, Carbon) |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(Carbon_Delta = Carbon - lag(Carbon)) |>
  ungroup() |>
  filter(!is.na(Carbon_Delta))

ca10_summary_tmp <- nrsgro_ca10_summary |>
  mutate(BA_METRIC = conv_multiunit(BA, "ft2 / acre", "m2 / hectare")) |>
  select(StandID, Year, BA_METRIC) |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(
    BA_Delta = BA_METRIC - lag(BA_METRIC)
  ) |>
  ungroup() |>
  filter(!is.na(BA_Delta))

nrsgro_ca10_proj_vs_meas <- ca10_carbon_tmp |>
  group_by(StandID) |>
  arrange(desc(Year)) |>
  filter(row_number() == 1) |>
  ungroup() |>
  rename(Carbon_Calb = Carbon, Carbon_Delta_Calb = Carbon_Delta) |>
  left_join(ca10_summary_tmp, by = join_by(StandID, Year)) |>
  rename(
    BA_Calb = BA_METRIC,
    BA_Delta_Calb = BA_Delta
  ) |>
  left_join(srvy_carbon_tmp, by = join_by(StandID, Year)) |>
  rename(Carbon_Srvy = Carbon, Carbon_Delta_Srvy = Carbon_Delta) |>
  left_join(srvy_summary_tmp, by = join_by(StandID, Year)) |>
  rename(
    BA_Srvy = BA_METRIC,
    BA_Delta_Srvy = BA_Delta
  ) |>
  left_join(stand_mixin, by = join_by(StandID)) |>
  mutate(
    BA_Diff = BA_Calb - BA_Srvy,
    Carbon_Diff = Carbon_Calb - Carbon_Srvy,
    ProjectionYears = Year - StartYear,
    BAI_Calb = BA_Delta_Calb / ProjectionYears,
    BAI_Srvy = BA_Delta_Srvy / ProjectionYears,
    BAI_Residual = BAI_Calb - BAI_Srvy,
    Carbon_Flux_Calb = -(Carbon_Delta_Calb / ProjectionYears),
    Carbon_Flux_Srvy = -(Carbon_Delta_Srvy / ProjectionYears),
    Carbon_Flux_Residual = Carbon_Flux_Calb - Carbon_Flux_Srvy
  )
```

```{r plot nrsgro_ca10_proj_vs_meas flux, eval = FALSE}
nrsgro_ca10_proj_vs_meas |>
  ggplot(aes(Carbon_Flux_Srvy, Carbon_Flux_Calb)) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.25) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.25) +
  geom_abline(slope = 1, intercept = 0, color = "gray", linewidth = 0.25) +
  geom_point(size = 0.1, alpha = 0.25) +
  coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.5, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  ggtitle("FVS Grow-Only\nProjected vs. Measured Annual Carbon Flux") +
  xlab(bquote("Measured Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% year^-1))) +
  ylab(bquote("Projected Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% year^-1))) +
  labs(caption = "With 10-Year Calibration")
```

```{r plot carbon flux residual vs balive, eval = FALSE}
nrsgro_ca10_proj_vs_meas |>
  ggplot(aes(BALIVE_METRIC, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.25) +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("Carbon Flux Residual" ~ (`Mg C` %.% hectare^-1 %.% year^-1))) +
  ggtitle("Carbon Flux Residual vs. Stand Starting Basal Area (10-year Calibration)") +
#  labs(caption = "With 10-year calibration") +
  scale_y_continuous(limits = c(-10, 5))
```


```{r plot bai residual vs balive, eval = FALSE}
nrsgro_ca10_proj_vs_meas |>
  ggplot(aes(BALIVE_METRIC, BAI_Residual)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.25) +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("BAI Residual" ~ (m^2 %.% hectare^-1 %.% year^-1))) +
  ggtitle("BAI Residual vs. Stand Starting Basal Area (10-year Calibration)")
#  labs(caption = "With 10-year calibration") +
  #scale_y_continuous(limits = c(-10, 5))
```


```{r basal_area_trend, eval = FALSE}
ggplot(
  nrsgro_ca10_proj_vs_meas |>
    mutate(
      BA_Range = cut(BALIVE_METRIC, 0:19 * 5)
    ),# |>
#    group_by(FOREST_TYPE_GROUP) |>
#    mutate(FOREST_TYPE_GROUP = paste0(FOREST_TYPE_GROUP, ", n=", n())) |>
#    ungroup(),
  aes(x = BA_Range, y = Carbon_Flux_Residual)
) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(outlier.size = 0.1, varwidth = TRUE) +
  scale_fill_discrete(name = "Forest Type Group") +
  ggtitle("Carbon Flux Residual vs. Measured BA") +
  labs(caption = "With 10-Year Calibration") +
  ylab(bquote("Carbon Flux Residual "(`Mg C` %.% ha^-1 %.% yr^-1))) +
  xlab(bquote("Starting Stand BA "(m^2 %.% ha^-1))) +
  theme_bw() +
  coord_cartesian(ylim = c(-5, 5))
```
