---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---
# Setup

```{r setup, include=FALSE}
library(targets)
library(tarchetypes)
library(tidyverse)
library(reshape2) # melt()
library(dbplyr) # window_order()
library(measurements) # conv_unit and conv_multiunit
library(maps)
library(ggspatial)
library(pbapply) # Progress bar for long operations
options(scipen = 9) # Use integer notation for numbers under 9 digits
tar_source()
```

## Load targets results from other documents

```{r load tar objects}
tar_load(fiadb)
tar_load(fvsne_states)
tar_load(species_crosswalk)
```

## Preload targets results from this document

When tar_make() has already built the objects defined in this document,
run this block to pre-load them. You can then skip running ```{targets} blocks.

```{r load tar objects defined here, eval = FALSE}
get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```

# Harvested Plots in the Northeast

Find all FIA plots that:

* are in the northeastern region, and
* use the modern plot design, and
* have at least 10 years of measurements, and
* are fully forested, and
* is not disturbed ("disturbed" is separate from harvest), and
* is harvested, and
* is not seeded, fertilized, herbicided, etc. post-harvest, and
* is measured both pre- and post-harvest, and
* have a single condition for the entire plot

"Is harvested" is as marked on the condition by FIA. Note that conditions with
harvested trees are not necessarily marked as harvested. To reconcile this, we
observe that TREE.DSTRBCD1 says only "significant disturbance" is registered,
and that "significant disturbance" is disturbance that causes "mortality or
damage to 25 percent of the trees in the condition" Testing confirms that when
the number of stems (TPA_UNADJ) represented by TREE.STATUSCD == 3 is >=25% of
the pre-harvest number of stems on the plot, the condition is marked as
harvested. To be consistent across all kinds of disturbance, we use the FIA
definitions for all disturbance, including harvest, and therefore do NOT mark a
plot as harvested even if there are harvested trees.


```{targets nrs_plots_prsc_all, tar_simple=TRUE}
fia_plots_filtered(fiadb, filter = \(.data, con) {
  .data |>
    filter(INVYR >= 1999) |>
    filter_plots_fvsne(con) |>
    filter_plots_trees(con) |>
    filter_plots_undisturbed(con) |>
    filter_plots_unfertilized(con) |>
    # harvested is implied by measured_pre_post_harvest
    #filter_plots_harvested(con)
    filter_plots_measured_pre_post_harvest(con)
})
```

## Stand Stats

Fetch more info about these stands.

```{targets nrs_plots_prsc_stats_all, tar_simple = TRUE}
fia_plots_filtered(fiadb, nrs_plots_prsc_all, \(.data, con) {
    plots <- .data |> distinct(STATECD, COUNTYCD, PLOT, INVYR)
    plots_join_by <- join_by(STATECD, COUNTYCD, PLOT, INVYR)

    forest_type <- tbl(con, "REF_FOREST_TYPE") |>
      select(VALUE, MEANING) |>
      rename(FORTYPCD = VALUE) |>
      rename(FORTYPE = MEANING)
    
    tree_stats <- tbl(con, "TREE") |>
      inner_join(plots, by = plots_join_by) |>
      select(STATECD, COUNTYCD, PLOT, INVYR, DIA, CARBON_AG, TPA_UNADJ) |>
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      summarize(
        CARBON_AG = sum(CARBON_AG, na.rm = TRUE),
        CPA = sum(CARBON_AG * TPA_UNADJ, na.rm = TRUE),
        # Restrict to trees that match those used in QMD computation, below.
        BA_TREES = sum(if_else(DIA >= 1, TPA_UNADJ, 0), na.rm = TRUE),
        .groups = "keep"
      )
    
    cond_stats <- tbl(con, "COND") |>
      inner_join(plots, by = plots_join_by) |>
      select(STATECD, COUNTYCD, PLOT, INVYR, STDAGE, BALIVE, FORTYPCD) |>
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      # QMD = sqrt(sum(DIA^2) / n)
      # Which is equivalent to
      # QMD = sqrt(sum(BALIVE * TPA_UNADJ) / (n * k)), where n is number of trees,
      # and k is π/576 ≅ 0.005454 for B in square feet and QMD in inches;
      # and k is π/40000 ≅ 0.0000785 for B in square meters and QMD in centimeters.
      # (see https://www.sciencedirect.com/science/article/pii/S2197562023000453 ,
      # https://doi.org/10.1016/j.fecs.2023.100114 )
      # When computing QMD from BA we need to use the same trees for BA and n;
      # FIADB data dictionary 2.5.51 BALIVE says "Basal area in square feet per
      # acre of all live trees ω1.0 inch d.b.h/d.r.c sampled in the condition."
      summarize(
        BALIVE = sum(BALIVE, na.rm = TRUE),
        FORTYPCD = max(FORTYPCD, na.rm = TRUE),
        STDAGE = max(STDAGE, na.rm = TRUE),
        .groups = "keep"
      ) |>
      ungroup()
    
    harvest_mixin <- tbl(con, "COND") |>
      inner_join(plots, by = plots_join_by) |>
      # Label each condition for harvest
      mutate(HARVEST =
         (!is.na(TRTCD1) & TRTCD1 == 10) |
         (!is.na(TRTCD2) & TRTCD2 == 10) |
         (!is.na(TRTCD3) & TRTCD3 == 10)
      ) |>
      # Collapse conditions up to plot level
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      summarize(HARVEST = any(HARVEST, na.rm = TRUE)) |>
      ungroup() |>
      # Give ordinals to plot inventories
      group_by(STATECD, COUNTYCD, PLOT) |>
      window_order(INVYR) |>
      mutate(
        INVNUM = row_number(),
        HARVEST = HARVEST & (INVNUM > 1) # Don't consider harvest in the first inventory
      ) |>
      ungroup() |>
      select(STATECD, COUNTYCD, PLOT, INVYR, HARVEST, INVNUM)
    
    pre_harvest_mixin <- harvest_mixin |>
      filter(HARVEST == 1) |>
      mutate(INVNUM = INVNUM - 1) |> # the inventory prior to harvest is pre-harvest
      rename(PRE_HARVEST = HARVEST) |>
      select(STATECD, COUNTYCD, PLOT, INVNUM, PRE_HARVEST)
    
    .data |>
      select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR) |>
      left_join(cond_stats, by = plots_join_by) |>
      left_join(tree_stats, by = plots_join_by) |>
      left_join(harvest_mixin, by = plots_join_by) |>
      left_join(pre_harvest_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVNUM)) |>
      left_join(forest_type, by = join_by(FORTYPCD)) |>
      rename(FOREST_TYPE = FORTYPE) |>
      mutate(
        PRE_HARVEST = if_else(is.na(PRE_HARVEST), 0, PRE_HARVEST),
        STDAGE = if_else(STDAGE < 0, NA, STDAGE),
        FRTYGRCD = floor(FORTYPCD / 10) * 10
      ) |>
      left_join(forest_type, by = join_by(FRTYGRCD == FORTYPCD)) |>
      rename(FOREST_TYPE_GROUP = FORTYPE)
  }) |>
  filter_decode_forest_type_group() |>
  mutate(
    BALIVE_METRIC = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    QMD = sqrt(BALIVE / (BA_TREES * (pi / 576))),
    QMD_METRIC = sqrt(BALIVE_METRIC / (BA_TREES * (pi / 40000))),
    CARBON_METRIC = conv_multiunit(CPA, "lbs / acre", "Mg / hectare")
  ) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(
    BALIVE_START = if_else(MEASYEAR == min(MEASYEAR, na.rm = TRUE), BALIVE_METRIC, NA),
    BALIVE_DELTA = BALIVE_METRIC - max(BALIVE_START, na.rm = TRUE),
    YEARS = MEASYEAR - min(MEASYEAR, na.rm = TRUE)
  ) |> 
  ungroup()
```

## Stand Locations

Where are they?

```{r map_by_ecoregion, eval = FALSE}
plot_location_tmp <- nrs_plots_prsc_all |>
  left_join(
    nrs_plots_prsc_stats_all |> select(STATECD, COUNTYCD, PLOT, INVYR, FOREST_TYPE_GROUP),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == min(INVYR, na.rm = TRUE)) |>
  ungroup() |>
  select(CN, LAT, LON, FOREST_TYPE_GROUP) |>
  rename(
    group = CN,
    lat = LAT,
    long = LON
  ) |>
  group_by(FOREST_TYPE_GROUP) |>
  mutate(
    n = n(),
    group_name = paste0(FOREST_TYPE_GROUP, ", n=", n())
  ) |>
  ungroup()

fvsne_states_map <- fvsne_states |>
  distinct(STATE_NAME) |>
  rename(region = STATE_NAME) |>
  _$region |>
  map_data('state', region = _)


ggplot(fvsne_states_map, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + 
  geom_point(
    plot_location_tmp,
    mapping = aes(long, lat, group = group, color = fct_reorder(group_name, desc(n)))
  ) +
  coord_sf(crs = 4326) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
    width = unit(1, "cm"),
    pad_x = unit(0.75, "cm"),
    pad_y = unit(0.5, "cm")
  ) +
  theme_bw() +
  scale_color_discrete(name = "Forest Type Group") +
  ggtitle("FIA Harvested Plots using FVSne")

remove(plot_location_tmp)
```

# Harvested Trees

Find the trees that were removed during harvest.

Note that FIA.TREE.STATUSCD == 3 indicates that a tree has been harvested,
but not all harvested trees have STATUSCD == 3. Therefore we approximate
harvested trees by finding all trees that exist in a pre-harvest inventory,
but do not exist (or are marked as harvested) in the corresponding harvest
inventory.

The goal is not to exactly replicate a harvest, but to tell FVSne which
trees should be removed before the next timestep.

## Trees removed in harvest

Fetch trees that were standing pre-harvest, that are not standing post-harvest.

```{targets nrs_trees_prsc_removed, tar_simple = TRUE}
tree_in_pre_harvest_plots <- fia_trees_filtered(
    fiadb,
    nrs_plots_prsc_stats_all |> filter(PRE_HARVEST == 1),
    filter = \(.data, com) {
      .data |>
        filter(STATUSCD == 1 | STATUSCD == 2) |> # live or dead trees
        select(CN, STATECD, COUNTYCD, PLOT, SUBP, TREE, INVYR, MORTYR, STATUSCD)  
  }) |>
  left_join(
    nrs_plots_prsc_stats_all |> select(STATECD, COUNTYCD, PLOT, INVYR, INVNUM),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  )

tree_in_harvested_plots <- fia_trees_filtered(
    fiadb,
    nrs_plots_prsc_stats_all |> filter(HARVEST == 1),
    filter = \(.data, con) {
      .data |>
        filter(STATUSCD == 1 | STATUSCD == 2) |> # live and dead trees are still there
        select(STATECD, COUNTYCD, PLOT, SUBP, TREE, INVYR)
    }
  ) |>
  left_join(
    nrs_plots_prsc_stats_all |> select(STATECD, COUNTYCD, PLOT, INVYR, INVNUM),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  # Move back a timestep so we can join the future onto the past
  mutate(
    INVNUM = INVNUM - 1
  )

tree_in_pre_harvest_plots |>
  anti_join(
    tree_in_harvested_plots,
    by = join_by(STATECD, COUNTYCD, PLOT, SUBP, TREE, INVNUM)
  )
```

## FIA Harvested Trees

Fetch trees that FIA says were harvested.

```{r tree_fia_harvested, eval = FALSE}
tree_fia_harvested <- fia_trees_filtered(
    fiadb,
    nrs_plots_prsc_stats_all |> filter(HARVEST == 1),
    filter = \(.data, con) {
      .data |>
        filter(STATUSCD == 3) |>
        select(STATECD, COUNTYCD, PLOT, SUBP, TREE, INVYR, STATUSCD)
    }
  )
```

## Validation

Were all the trees that FIA says were harvested actually removed?
```{r tree_harvested_but_not_removed, eval = FALSE}
tree_harvested_but_not_removed <- tree_fia_harvested |>
  anti_join(nrs_trees_prsc_removed, by = join_by(STATECD, COUNTYCD, PLOT, SUBP, TREE))
stopifnot(tree_harvested_but_not_removed |> nrow() == 0)
```

It is expected that FIA did not mark all removed trees as harvested.
```{r tree_removed_but_not_harvested, eval = FALSE}
tree_removed_but_not_harvested <- nrs_trees_prsc_removed |>
  anti_join(tree_fia_harvested, by = join_by(STATECD, COUNTYCD, PLOT, SUBP, TREE))
tree_removed_but_not_harvested |> nrow()
```

All harvested trees should be on harvested plots.
```{r stopifnot_tree_removed_from_non_harvested_plots, eval = FALSE}
stopifnot(
  nrs_trees_prsc_removed |>
    distinct(STATECD, COUNTYCD, PLOT, INVNUM) |>
    # INVNUM is the last inventory in which the tree was standing,
    # so it is marked harvested in the next inventory
    mutate(INVNUM = INVNUM + 1) |>
    anti_join(
      nrs_plots_prsc_stats_all |> filter(HARVEST == 1),
      by = join_by(STATECD, COUNTYCD, PLOT, INVNUM)
    ) |>
    nrow() == 0
)
```

There are some "harvested" plots that do not have trees removed.
```{r plot_harvested_without_tree_removed, eval = FALSE}
plot_harvested_without_tree_removed <- nrs_plots_prsc_all |>
  anti_join(
    nrs_trees_prsc_removed |>
      distinct(STATECD, COUNTYCD, PLOT),
    join_by(STATECD, COUNTYCD, PLOT)
  )
plot_harvested_without_tree_removed |> nrow()
```

## Remove Harvested Plots Without Removed Trees

Remove those plots from consideration.

```{targets nrs_plots_prsc, tar_simple = TRUE}
nrs_plots_prsc_all |>
  semi_join(
    nrs_trees_prsc_removed |> distinct(STATECD, COUNTYCD, PLOT),
    by = join_by(STATECD, COUNTYCD, PLOT)
  )
```
```{targets nrs_plots_prsc_stats, tar_simple = TRUE}
nrs_plots_prsc_stats_all |>
  semi_join(
    nrs_trees_prsc_removed |> distinct(STATECD, COUNTYCD, PLOT),
    by = join_by(STATECD, COUNTYCD, PLOT)
  )
```
What's left?
```{r map_by_ecoregion2}
plot_data <- nrs_plots_prsc |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(row_number() == 1) |>
  ungroup() |>
  rename(lat = LAT, long = LON, group = CN) |>
  left_join(
    nrs_plots_prsc_stats |> select(STATECD, COUNTYCD, PLOT, INVYR, FOREST_TYPE, FOREST_TYPE_GROUP),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  group_by(FOREST_TYPE_GROUP) |>
  mutate(
    N = n(),
    group = paste0(FOREST_TYPE_GROUP, ", n=", n())
  ) |>
  ungroup()

map_data('state', fvsne_states$STATE_NAME) |> 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + 
  geom_point(
    data = plot_data,
    aes(color = fct_reorder(group, desc(N)))
  ) +
  coord_sf(crs = 4326) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
    width = unit(1, "cm"),
    pad_x = unit(0.75, "cm"),
    pad_y = unit(0.5, "cm")
  ) +
  theme_bw() +
  scale_color_discrete(name = "Forest Type")
```
# Harvest Prescription for FVS

When running FVS, we will need:

* CutList / CutLiDB - list of trees harvested
* TreeList / TreeLiDB - list of trees; probably we want cycle 0 captured
* ATrtList / ATrtLiDB - After treatment (post treelist, pre harvest) tree list

## Prescription Table

Prescription table:

- Find all years in which trees were removed.
  - For trees marked as harvested, the estimated year of harvest is in MORTYR
    - 1 of 5580 records has MORTYR
    - this is not worth handling for one tree
  - For trees removed from conditions marked as harvested, the estimated year(s)
    of harvest are in TRTYR1/2/3
    - when there are multiple harvest years, we can't tell which trees are
      harvested in which year
    - when there are multiple harvest years, pick the latest one
  - if there are trees left that don't have a harvest year, put their harvest
    the year before measurement. This will be pessimistic with respect to regen,
    but fits the overall distribution
- Create ordinals for removal.
  - Each plot needs to have < 10 years in which removals occur to keep life easy
  - sort the years and number them
- Consolidate harvests
  - we don't want lots of short cycles, but
  - Essential FVS figure 3-1 shows that thinning happens once per cycle,
    no matter how many years are involved

We have access to:
- Stand_CN - ignored, but used for filtering
- Stand_ID - ignored
- PLOT_CN - ignored
- PLOT_ID
- TREE_CN - ignored, but can be used for join
- TREE_ID

```{r tree_prescription}
trtyr_mixin <- stand_stats |>
  filter(HARVEST == TRUE) |>
  mutate(
    TRTYR = if_else(is.na(TRTYR), MEASYEAR - 1, TRTYR),
    TRTNUM = if_else(is.na(TRTNUM), 1, TRTNUM),
    INVNUM = INVNUM - 1
  ) |>
  select(CN, STATECD, COUNTYCD, PLOT, INVNUM, TRTYR, TRTNUM) |>
  rename(STAND_CN = CN)


tree_prescription <- tree_removed_in_harvest |>
  inner_join(
    trtyr_mixin,
    by = join_by(STATECD, COUNTYCD, PLOT, INVNUM)
  ) |>
  rename(TREE_CN = CN) |>
  mutate(PRESCRIPTION = TRTNUM + 1) |> # PRESCRIPTION range is 2-9
  select(TREE_CN, PRESCRIPTION)
```

We can create a map from year to treatment number as:
```
stand_stats |> filter(!is.na(TRTYR)) |> select(STAND_CN, TRTYR, TRTNUM)
```

## Compose FVS Database

FVS fetches stand and tree data from a SQLite database. Unfortunately,
the way it queries the database prevents doing "intereting" things,
such as attaching a second database for prescription overrides. So we need
to write an entire database with the information FVS needs, which is
the FVS_StandInit_Plot and FVS_TreeInit_Plot tables.

```{r fvs_starting_stands}
fvs_starting_stands <- stand_stats |>
  filter(MEASYEAR == MIN_MEASYEAR)
```

Write a database for use just by this run.

```{r fvs_nrsharvest_prsc.db}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
fvs <- DBI::dbConnect(RSQLite::SQLite(), 'data/fvs/FVS_NRSHarvest_PRSC.db')

fvs_standinit_plot <- tbl(fia, "FVS_StandInit_Plot") |>
  semi_join(
    fvs_starting_stands |> select(CN) |> rename(STAND_CN = CN),
    by = join_by(STAND_CN),
    copy = TRUE
  ) |>
  collect()
DBI::dbWriteTable(fvs, "FVS_StandInit_Plot", fvs_standinit_plot, overwrite = TRUE)

fvs_treeinit_plot <- tbl(fia, "FVS_TreeInit_Plot") |>
  semi_join(
    fvs_starting_stands |> select(CN) |> rename(STAND_CN = CN),
    by = join_by(STAND_CN),
    copy = TRUE
  ) |>
  collect() |>
  select(!PRESCRIPTION) |>
  left_join(tree_prescription, by = join_by(TREE_CN))
DBI::dbWriteTable(fvs, "FVS_TreeInit_Plot", fvs_treeinit_plot, overwrite = TRUE)

DBI::dbDisconnect(fvs)
DBI::dbDisconnect(fia)
rm(fia, fvs, fvs_standinit_plot, fvs_treeinit_plot)
```

## Thin Prescription Table

Create a table describing thinning prescriptions by year

```{r fvs_thinprsc}
fvs_ThinPRSC <- function(rows) {
  thinprsc <- function(row) {
    year <- row["year"]
    if ("percent" %in% names(row)) {
      percent <- row["percent"]
    } else {
      percent <- 1
    }
    prsc <- row["prescription"]
    fvs_kwd3("ThinPRSC", year, percent, prsc)
  }
  apply(rows, 1, thinprsc)
}
```

```{r keyfile_thin_section}
keyfile_thin_section <- function(StandCN) {
  # StandCN referes to STAND_CN in stand_stats
  # swizzle that to CN, which is what trtyr_mixin uses
  stand_cn <- stand_stats |>
    filter(STAND_CN == StandCN) |>
    select(CN) |>
    rename(STAND_CN = CN)
  pertinent <- trtyr_mixin |>
    semi_join(stand_cn, by = join_by(STAND_CN)) |>
    filter(!is.na(TRTNUM)) |>
    rename(year = TRTYR) |>
    mutate(prescription = TRTNUM + 1) |>
    select(year, prescription)
  if (nrow(pertinent) > 0) {
    fvs_ThinPRSC(pertinent)
  } else {
    ''
  }
}
```


## Keyword File Definition

```{r keywordfile_section}
keywordfile_section <- function(
    Title, MgmtId, StandID, StandCN,
    FirstYear, LastYear, Timestep,
    Partition, RandSeed = NA
  ) {
  if (!is.na(RandSeed)) {
    StdIdent <- c(
      fvs_kwd0("StdIdent"),
      paste0(StandID, "_Y", Timestep, "_R", RandSeed, " ", Title)
    )
    # Careful: RannSeed != RandSeed
    RannSeed <- fvs_kwd1("RANNSEED", RandSeed)
  } else {
    StdIdent <- c(
      fvs_kwd0("StdIdent"),
      # RannSeed isn't 0, but 0 is an illegal random seed so a fine flag value
      paste0(StandID, "_Y", Timestep, "_R0", " ", Title))
    # Careful: RannSeed != RandSeed
    RannSeed <- c()
  }
  TimeConfig <- fvs_TimeConfig(FirstYear, LastYear, Timestep)
  if (FirstYear == LastYear) {
    Thin <- c() # Don't do thinning for 0-year runs
  } else {
    Thin <- keyfile_thin_section(StandCN)
  }
  c(
    StdIdent,
    fvs_kwd0("StandCN"),
    StandCN,
    fvs_kwd0("MgmtId"),
    MgmtId,
    TimeConfig,
    RannSeed,
    fvs_kwd1("CutList", 0),
    fvs_kwd1("ATrtList", 0),
    fvs_kwd1("TreeList", 0),
    fvs_kwd0("FMIn"), # Begin
    fvs_kwd0("CarbRept"),
    fvs_kwd0("CarbCut"),
    fvs_kwd5("CarbCalc", 1, 1, 0.0425, 9, 11),
    fvs_kwd0("FuelOut"),
    fvs_kwd0("FuelRept"),
    fvs_kwd0("End"), # FMIn
    fvs_kwd0("Database"), # Begin
    fvs_kwd0("DSNIn"),
    paste0("FVS_", Title, "_", MgmtId, ".db"),
    fvs_kwd0("StandSQL"), # Begin
    "SELECT * FROM FVS_StandInit_Plot WHERE Stand_CN = '%Stand_CN%'",
    fvs_kwd0("EndSQL"), # StandSQL
    fvs_kwd0("TreeSQL"), # Begin
    "SELECT * FROM FVS_TreeInit_Plot WHERE Stand_CN = '%Stand_CN%'",
    fvs_kwd0("EndSQL"), # TreeSQL
    fvs_kwd0("DSNOut"),
    paste0("FVS_", Title, "_", MgmtId, "_", Partition, ".db"),
    fvs_kwd1("Summary",  2),
    fvs_kwd2("Computdb", 0, 1),
    fvs_kwd1("MisRpts",  2),
    fvs_kwd1("ATrtLiDB", 2),
    fvs_kwd1("CarbReDB", 2),
    fvs_kwd1("CutLiDB", 2),
    fvs_kwd1("FuelReDB", 2),
    fvs_kwd1("FuelsOut", 2),
    fvs_kwd1("TreeLiDB", 2),
    fvs_kwd0("End"), # Database
    Thin,
    #Regen,
    fvs_kwd0("Process")
  )
}

```

## Harvest Keyword File Generation

```{r fvs_stands_prsc}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
stand_id_mixin <- tbl(fia, 'FVS_StandInit_Plot') |>
  semi_join(
    fvs_starting_stands |>
      select(CN) |>
      rename(STAND_CN = CN),
    by = join_by(STAND_CN),
    copy = TRUE
  ) |>
  select(STAND_CN, STAND_ID) |>
  rename(CN = STAND_CN) |>
  collect()

fvs_stands_prsc <- fvs_starting_stands |>
  left_join(stand_id_mixin, by = join_by(CN))

dbDisconnect(fia)
remove(fia, stand_id_mixin)

# Every stand must have a STAND_ID
stopifnot(fvs_stands_prsc |> filter(is.na(STAND_ID)) |> nrow() == 0)
```

```{r NRSHarvest PRSC}
num_partitions <- 60 # The number of simultaneous processes to use
timestep <- 10
title <- "NRSHarvest"
mgmtid <- "PRSC"
for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".key")
  unlink(filename)
#    print(paste0("Writing ", filename))
  apply(
    expand.grid(rannseed = c(1, 3, 5)),
    1,
    \(runconfig) {
      rannseed <- runconfig["rannseed"]
      apply(
        fvs_stands_prsc |> 
          filter((as.numeric(STAND_ID) %% num_partitions) == (partition - 1)) |>
          arrange(STAND_ID, MIN_MEASYEAR),
        1,
        function(row) {
          write_lines(
            keywordfile_section(
              title,
              mgmtid,
              row['STAND_ID'],
              row['STAND_CN'],
              row['MIN_MEASYEAR'],
              row['MAX_MEASYEAR'],
              timestep,
              partition,
              rannseed
            ),
            filename,
            append = TRUE
          )
        }
      )
    }
  )
  write_lines("Stop", filename, append = TRUE)
}
```

## Survey Keyword File Generation

We also do 0-year projections of the same stands in every measurement year
to provide a comparison.

```{r fvs_stands_srvy}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
fia_fvs_standinit_plot <- tbl(fia, 'FVS_StandInit_Plot') |>
  select(STAND_CN, STAND_ID)

fvs_stands_srvy <- tbl(fia, 'PLOT') |>
  semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  select(CN, STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, DESIGNCD) |>
  rename(STAND_CN = CN) |>
  left_join(fia_fvs_standinit_plot, by = join_by(STAND_CN)) |>
  collect()

dbDisconnect(fia)
remove(fia, fia_fvs_standinit_plot)
```

The second run is to do a 0-year projection for each measurement year. These
are given management ID "SRVY".

```{r NRSGrowOnly SRVY}
# We're running 0-year timesteps, but FVS requires us to run at least one
# year, so we run 0-year timesteps for 1 year.
timestep <- 1
title <- "NRSHarvest"
mgmtid <- "SRVY"
for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".key")
  unlink(filename)
  #print(paste0("Writing ", filename))
  apply(
    fvs_stands_srvy |>
      filter((as.numeric(STAND_ID) %% num_partitions) == (partition - 1)) |>
      arrange(STAND_ID, MEASYEAR),
    1,
    function(row) {
      write_lines(
        keywordfile_section(
          title,
          mgmtid,
          row['STAND_ID'],
          row['STAND_CN'],
          row['MEASYEAR'],
          row['MEASYEAR'],
          timestep,
          partition
        ),
        filename,
        append = TRUE
      )
    }
  )
  write_lines("Stop", filename, append = TRUE)
}
```
# Run FVS

> TODO nik: do this automatically; that'll be easier using targets

# Results

## Fault Analysis

### Parse Faults

What errors and warnings do we get?
```{r fvs warnings, cache=TRUE}
mgmtid <- "PRSC"
fvs_warnings <- NULL
for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".out")
  fvs_warnings <- bind_rows(
    fvs_warnings,
    read_lines(filename) |>
      as.data.frame(nm="text") |>
      # Filter to just initial STAND ID and WARNING lines
      filter(
        grepl("^ +STAND ID=", text) |
          grepl("^\\*{8} ", text)
      ) |>
      # Parse out the STAND ID and fill it into the following lines
      mutate(
        STAND_ID = case_when(
          grepl("^ +STAND ID=", text) ~ sub("^ +STAND ID= *([^ ]+) .*$", "\\1", text)
        )
      ) |>
      fill(STAND_ID) |>
      # Remove everything but warnings (which now have STAND_ID)
      filter(
        grepl("^\\*{8} ", text)
      )
  )
}
```

Find the distinct faults

```{r distinct fvs warnings}
fvs_warnings_distinct <- fvs_warnings |>
  # Most faults have some leading text, the tag ERROR: or WARNING:,
  # then descriptive text; if there's a number in the descriptive text,
  # it's probably pulled from the data, so we chop off the descriptive
  # text starting at the first number to find the truly distinct warnings
  mutate(
    PROBLEM = case_when(
      grepl("WARNING:", text) ~ sub("^[*. ]*(.*WARNING: [^0-9]+)([0-9].*)?$", "\\1", text),
      grepl("ERROR:", text) ~ sub("^[*. ]*(.*ERROR: [^0-9]+)([0-9].*)?$", "\\1", text),
      .default = text
    ),
    PROBLEM = sub("\\s+$", "", PROBLEM)
  ) |>
  distinct(PROBLEM) |>
  arrange(PROBLEM)
knitr::kable(fvs_warnings_distinct)
```
### Plot Count Mismatch

Plot counts do not match data on the design record could indicate a data prep
error.

```{r keyword context}
debug_stand_id <- fvs_warnings |>
  filter(grepl("FVS09 WARNING: ", text)) |>
  head(1) |>
  pull("STAND_ID")
fvs_warnings |>
  filter(STAND_ID == debug_stand_id)
```

From Essential FVS:
```
This
warning message is printed when either the plot count or nonstockable count determined
from reading the tree data records differs from the values coded on the DESIGN record.
```

Which partition is it in?
```{r keyword context partition}
as.numeric(sub("^([^_]+)_.*$", "\\1", debug_stand_id)) %% num_partitions
```

so the 16th partition, which is partition 17.

Here's the keyword file section:
```
StdIdent  
420301305916_Y10_R1 NRSHarvest
StandCN   
58377688010538
MgmtId    
PRSC
InvYear         2003
TimeInt            0        10
TimeInt            1         7
TimeInt            3         1
NumCycle           3
RANNSEED           1
FMIn      
CarbRept  
CarbCut   
CarbCalc           1         1    0.0425         9        11
FuelOut   
FuelRept  
End       
Database  
DSNIn     
FVS_NRSHarvest_PRSC.db
StandSQL  
SELECT * FROM FVS_StandInit_Plot WHERE Stand_CN = '%Stand_CN%'
EndSQL    
TreeSQL   
SELECT * FROM FVS_TreeInit_Plot WHERE Stand_CN = '%Stand_CN%'
EndSQL    
DSNOut    
FVS_NRSHarvest_PRSC_17.db
Summary            2
Computdb           0         1
MisRpts            2
CarbReDB           2
FuelReDB           2
FuelsOut           2
End       
ThinPRSC        2006         1         3
Process   
```

from the .out file, when processing TreeSQL:
> NUMBER ROWS PROCESSED:   18

The number of plots should match the number of distinct SUBP in FIA:
```{r debug_fia_plot}
debug_fia_plot <- fvs_stands_prsc |>
  filter(STAND_ID == sub("^([^_]+)_.*$", "\\1", debug_stand_id)) |>
  select(STATECD, COUNTYCD, PLOT, INVYR)
tree_removed_in_harvest |>
  semi_join(debug_fia_plot, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  distinct(SUBP) |>
  nrow()
```

This is probably benign - it means there were trees on only 2 out of 4 subplots
at the start of the run.

How common is this?
```{r fvs09 warnings}
fvs_warnings |>
  filter(grepl("FVS09 WARNING: ", text)) |>
  nrow()
```

This happens on 9 out of 400 plots; ignore.

### One Thousand Trees

The TPA fault may indicate a keyword file generation error; let's track it down.
```{r one thousand trees}
debug_stand_id <- fvs_warnings |>
  filter(grepl(" TPA ", text)) |>
  head(1) |>
  pull("STAND_ID")
fvs_warnings |>
  filter(STAND_ID == debug_stand_id)
```

a'ight, what's the species?

```{r fvs species 9}
debug_species <- fvs_warnings |>
  filter(grepl("FVS40 WARNING:", text) & grepl("SPECIES=", text)) |>
  mutate(FVS_SPNO = as.integer(sub("^.*SPECIES=\\s*(\\d+) .*$", "\\1", text))) |>
  group_by(FVS_SPNO) |>
  summarize(N = n(), .groups = "keep") |>
  ungroup() |>
  arrange(desc(N)) |>
  left_join(species_mixin, by = join_by(FVS_SPNO))
debug_species
```

These are plots where the number of seedlings / saplings on the microplots
tally to >1000 TPA.

I think we note that these exist, and ignore them.

How often does this occur?

```{r fvs tpa warnings}
fvs_warnings |>
  filter(grepl(" TPA ", text)) |>
  nrow()
```

27 out of 400 plots.

We can decide to not worry about them.

## Load Results

Load the output from all FVS partitions into a single database.
Particularly with the tree list, these start to get big, so we want to 
put them into SQLite where we can quickly read in just those portions
we're interested in.

```{r consolidate_results}
mgmt_ids <- c("PRSC", "SRVY")
partitions <- 1:num_partitions

fvs_output_partition <- function(title, mgmtid, partition) {
  paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".db")
}

fvs_output_consolidated <- function(title, mgmtid) {
  paste0("data/fvs/FVS_", title, "_", mgmtid, "_Results.db")
}

fvs_consolidate_output <- function(title, mgmtid, partitions) {
  dest <- dbConnect(RSQLite::SQLite(), fvs_output_consolidated(title, mgmtid))
  on.exit(dbDisconnect(dest), add = TRUE, after = FALSE)

  pblapply(partitions, \(partition) {
    part <- dbConnect(RSQLite::SQLite(), fvs_output_partition(title, mgmtid, partition))
    on.exit(dbDisconnect(part), add = TRUE, after = FALSE)
    
    tables <- dbListTables(part)
    lapply(tables, \(table) {
      dbWriteTable(dest, table, tbl(part, table) |> collect(), append = TRUE)
    })
  })
}

invisible(
  lapply(mgmt_ids, \(mgmtid) {
    filename <- fvs_output_consolidated(title, mgmtid)
    print(paste0("Writing ", filename))
    if (file.exists(filename)) {
      unlink(filename)
    }
    # Notable SQLite restrictions:
    # - One statement per dbExecute - so, no semicolon-separated statements
    # - Maximum of 10 connections at a time
    # - Not all tables exist in all outputs; e.g. FVS_Error only exists
    #   if an error was observed
    # Rather than try to be clever and have SQLite copy the tables,
    # it's easier (and about as fast) to pull them into R one partition
    # at a time, and write the data out to the consolidated database
    fvs_consolidate_output(title, mgmtid, partitions)
  })
)
rm(fvs_output_partition, fvs_output_consolidated, fvs_consolidate_output)
```


```{r FVS_Carbon, cache=TRUE}
load_result_table <- function(title, mgmtid, table) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_Results.db")
  stopifnot(file.exists(filename))
  fvs_out_db <- DBI::dbConnect(RSQLite::SQLite(), filename)
  on.exit(dbDisconnect(fvs_out_db), add = TRUE, after = FALSE)

  tbl(fvs_out_db, table) |>
    collect() |>
    mutate(
      Title = title,
      MgmtId = mgmtid,
      Timestep = sub("^.*_Y([0-9]+).*$", "\\1", StandID), # Extract timestep
      RanNSeed = sub("^.*_R([0-9]+).*$", "\\1", StandID), # Extract random seed
      StandID = sub("^([^_]+)_.*$", "\\1", StandID), # Extract stand ID
      Partition = (as.numeric(StandID) %% num_partitions) + 1
    )
}

print("Loading Carbon results")
FVS_Carbon <- bind_rows(
  load_result_table(title, "PRSC", "FVS_Carbon"),
  load_result_table(title, "SRVY", "FVS_Carbon")
)

print("Loading result Summary")
FVS_Summary2_East <- bind_rows(
  load_result_table(title, "PRSC", "FVS_Summary2_East"),
  load_result_table(title, "SRVY", "FVS_Summary2_East")
)

rm(load_result_table)
```
Consolidate outputs; average the multiple runs with different random seeds.

```{r FVS_Carbon_PRSC}
FVS_Carbon_PRSC <- FVS_Carbon |>
  filter(MgmtId == "PRSC") |>
  group_by(StandID, Year) |>
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    Total_Stand_Carbon = mean(Total_Stand_Carbon),
    .groups = 'keep'
  ) |>
  ungroup()

# There is no projection for the observed run, so no need to average
FVS_Carbon_SRVY <- FVS_Carbon |>  
  filter(MgmtId == "SRVY")
```

### Sanity Check

Sanity check: did we get the outputs we requested?

```{r sanity-check-expected-output}
matched_carbon_tmp <- fvs_stands_prsc |>
  left_join(
    FVS_Carbon_PRSC |>
      select(StandID, Year, Aboveground_Total_Live) |> 
      rename(
        STAND_ID = StandID,
        MIN_MEASYEAR = Year,
        Starting_Carbon = Aboveground_Total_Live
      ),
    by = join_by(STAND_ID, MIN_MEASYEAR)
  ) |>
  left_join(
    FVS_Carbon_PRSC |>
      select(StandID, Year, Aboveground_Total_Live) |>
      rename(
        STAND_ID = StandID,
        MAX_MEASYEAR = Year,
        Ending_Carbon = Aboveground_Total_Live
        ),
    by = join_by(STAND_ID, MAX_MEASYEAR)
  ) |>
  mutate(partition = (as.numeric(STAND_ID) %% num_partitions) + 1)
missing_start <- matched_carbon_tmp |>
  filter(is.na(Starting_Carbon)) |> 
  nrow()
print(missing_start)
missing_end <- matched_carbon_tmp |>
  filter(is.na(Ending_Carbon)) |> 
  nrow()
print(missing_end)
stopifnot(missing_start == 0)
stopifnot(missing_end == 0)
remove(
  matched_carbon_tmp,
  missing_start,
  missing_end
)
```

### Projected Carbon

```{r plot aboveground carbon}
ggplot(
    FVS_Carbon_PRSC,
    aes(x = Year, y = Aboveground_Total_Live)
  ) +
  geom_bin2d(bins = 25) +
  #geom_abline(intercept = 0, slope = 1) +
  #geom_smooth(
  #  method = lm,
  #  formula = 'y ~ x',
  #  color = 'brown',
  #  linetype = 'dashed',
  #  fullrange = TRUE,
  #  se = FALSE
  #) +
  scale_fill_continuous(name = "Number of Plots", transform = "log10") +
  #facet_wrap(~FOREST_TYPE_GROUP) +
  theme_bw() +
  #theme(aspect.ratio = 1) +
  ggtitle("FVS Projected Aboveground Carbon") +
  #xlab(bquote("Measured Annual Carbon Flux" ~(`Mg C` %.% ha^-1 %.% yr^-1))) +
  ylab(bquote("Projected Carbon" ~(`Mg C` %.% ha^-1))) #+

```

### Projected Carbon Flux

```{r plot carbon flux}
all_years <- tibble(Year = min(FVS_Carbon_PRSC$Year):max(FVS_Carbon_PRSC$Year))
num_years <- all_years |> nrow()

FVS_Carbon_PRSC |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(
    C_Flux = Aboveground_Total_Live - lag(Aboveground_Total_Live),
    Prev_Year = lag(Year),
    Annual_C_Flux = C_Flux / (Year - Prev_Year),
    Min_Year = min(Year),
    Max_Year = max(Year)
  ) |>
  ungroup() |>
  full_join(
    expand.grid(
      StandID = unique(FVS_Carbon_PRSC$StandID),
      Year = all_years$Year
    ),
    by = join_by(StandID, Year)
  ) |>
  arrange(StandID, Year) |>
  fill(StandID, Annual_C_Flux, .direction = "up") |>
  fill(Min_Year, Max_Year, .direction = "updown") |>
  filter(Year >= Min_Year & Year <= Max_Year) |>
  ggplot(
    aes(x = Year, y = Annual_C_Flux)
  ) +
  geom_bin2d(binwidth = c(1, 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_continuous(name = "Number of Plots", transform = "log10") +
  theme_bw() +
  ggtitle("FVS Projected Annual Carbon Flux for Harvested Plots") +
  ylab(bquote("Projected Carbon Flux" ~(`Mg C` %.% ha^-1 %.% yr^-1))) #+;

```
