---
title: "10_NRSBias"
output: html_document
---

#Setup

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(targets)
  library(tarchetypes)
  library(tidyverse)
  library(performance) # for check_model()
  library(car) # for Anova()
  library(dbplyr) # for window_order
  library(reshape2) # melt()
  library(measurements) # conv_unit and conv_multiunit
  library(maps)
  library(ggspatial) # for maps
  library(ggpmisc) # for stat_poly_eq
})
options(scipen = 4) # Use integer notation for numbers under 9 digits
tar_source()
```

The package meteoland can compute solar radiation from lat, lon, elevation,
and some other things.
https://emf-creaf.github.io/meteolandbook/solarradiation.html
There's a bunch of solar radiation functions; one of them shuld be useful:
https://emf-creaf.github.io/meteoland/reference/radiation_julianDay.html
probably either netRadiaton or solarRadiation - these describe incoming
solar radiation in units of bquote(MJ %.% m^-2 %.% day^-1)


## Load targets results from other documents

```{r load tar objects defined elsewhere, eval = FALSE}
tar_load(fiadb)
tar_load(species_crosswalk)
```

# Context

## Independent Variables

What variables do we have?
```{r nrsgro_plot}
tar_load(nrsgro_plot)
tar_load(nrsgro_plot_stats)
```

```{r independent_vars}
tmp_plot_stats <- nrsgro_plot_stats |>
  filter_add_stand_id() |>
  group_by(STAND_ID) |>
  arrange(INVYR) |>
  filter(row_number() == 1) |>
  ungroup() |>
  mutate(
    FOREST_TYPE = as.factor(FOREST_TYPE),
    FOREST_TYPE_GROUP = as.factor(FOREST_TYPE_GROUP)
  ) |>
  select(
    STAND_ID,
    BALIVE_METRIC, BA_TREES, QMD,
    FOREST_TYPE_GROUP, FOREST_TYPE, STDAGE
  ) |>
  # Fix the name of this so it's not confusing
  rename(TPA = BA_TREES) |>
  # Compute Stand Density Index; this is not relative stand density, it's the
  # absolute number. See FIADB data dictionary sectino 2.5.139 SDI_RMRS,
  # and VanderSchaaf, C.L., 2013. Reineke’s stand density index: a quantitative and non-unitless
  # measure of stand density.
  # In: Guldin, James M., ed. 2013. Proceedings of the 15th biennial southern silvicultural
  # research conference. e-Gen. Tech. Rep. SRS-GTR-175. Asheville, NC:
  # US Department of Agriculture, Forest Service, Southern Research Station.
  # 577-579. (Vol. 175, pp. 577-579).
  # Here we use 1.605 as the exponent of Reineke's equation, as is commonly reported.
  mutate(SDI = 10^(log10(TPA) + 1.605*log10(QMD) - 1.605))

independent_vars <- nrsgro_plot |>
  filter_add_stand_id() |>
  group_by(STAND_ID) |>
  arrange(INVYR) |>
  mutate(ELEV = max(ELEV, na.rm = TRUE)) |> # some ELEV are NA; replace them
  filter(row_number() == 1) |>
  ungroup() |>
  select(STAND_ID, LAT, LON, ELEV, ECOSUBCD) |>
  mutate(
    ECOCD = substr(ECOSUBCD, 1, nchar(ECOSUBCD) - 1),
    ECOSUBCD = as.factor(ECOSUBCD),
    ECOCD = as.factor(ECOCD)
  ) |>
  left_join(tmp_plot_stats, by = join_by(STAND_ID))  |>
  mutate(
    LAT_CTR = LAT - mean(LAT, na.rm = TRUE),
    LON_CTR = LON - mean(LON, na.rm = TRUE),
    ELEV_CTR = ELEV - mean(ELEV, na.rm = TRUE),
    BALIVE_CTR = BALIVE_METRIC - mean(BALIVE_METRIC, na.rm = TRUE),
    TPA_CTR = TPA - mean(TPA, na.rm = TRUE),
    QMD_CTR = QMD - mean(QMD, na.rm = TRUE),
    SDI_CTR = SDI - mean(SDI, na.rm = TRUE),
    STDAGE_CTR = STDAGE - mean(STDAGE, na.rm = TRUE)
  )
# TODO: we need slope and apsect - these are on the COND
```

```{r plot indep vars}
independent_vars |>
#  filter(FOREST_TYPE_GROUP == "Spruce / fir") |>
  filter(TPA > 10) |>
  ggplot(aes(TPA, BALIVE_METRIC, color = STDAGE, size = QMD)) +
  geom_point(position = "jitter") +
#  facet_wrap("FOREST_TYPE_GROUP") +
  scale_x_continuous(transform = "log10") +
  scale_y_continuous(transform = "log10") +
  scale_color_continuous(transform = "log10")
```

## Dependent Variable(s)

We want to predict bias in predicted carbon storage, probably over 10 years, maybe 20 or 25.

Let's use the Calb10 data for this.

Question: do we want this per stand, or per tree?

Per stand:
- It's what FVS gives us as primary output
- It's what we'll use in analysis of COP

Per tree:
- Isolates the effects of multiple species
- Don't have an easy way to get per-tree carbon from FVS
  - Can correct bias in BAI instead; is that good enough?
- Don't have an easy way to compute carbon
  - Can we use rFIA?
- Don't have an easy way to aggregate to stand level
  - Can we use rFIA?

```{r dependent variables}
tar_load(nrsgro_ca10_proj_vs_meas)
# We're only interested in the error of the final projection
dependent_variables <- nrsgro_ca10_proj_vs_meas |>
  group_by(StandID) |>
  arrange(desc(ProjectionYears)) |>
  filter(row_number() == 1) |>
  ungroup() |>
  select(-BALIVE_METRIC)
```

## Observations

Combine independent and dependent variables into observations
```{r obs}
obs <- dependent_variables |>
  left_join(independent_vars, by = join_by(StandID == STAND_ID))
```

```{r plot obs hist}
obs |>
  ggplot(aes(Carbon_Flux_Residual)) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(name = bquote("Carbon Flux Residual" ~ (`Mg C` %.% ha^-1 %.% year^-1))) +
  scale_y_continuous(name = "Number of Plots") +
  theme_bw() +
  ggtitle("Histogram of Carbon Flux Residual")
```

# Carbon Flux Bias Correction

## Linearity

Let's look at the independent variables and see if their (rough) relationship
with annual carbon flux is linear

### Latitude

```{r lat linearity}
obs |>
  ggplot(aes(LAT, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```

### Longitude

```{r lon linearity}
obs |>
  ggplot(aes(LON, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```

### Elevation

```{r elev linearity}
obs |>
  ggplot(aes(ELEV, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```

### Basal Area

```{r balive linearity}
obs |>
  mutate(BALIVE_METRIC = cut(BALIVE_METRIC, 0:19 * 5, include.lowest = TRUE)) |>
  ggplot(aes(BALIVE_METRIC, Carbon_Flux_Residual)) +
  geom_boxplot(size=0.25, outlier.size = 0.1) +
#  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw()
```

### TPA

```{r TPA linearity}
obs |>
  ggplot(aes(TPA, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```

### Stand Age

```{r stdage linearity}
obs |>
  ggplot(aes(STDAGE, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```

### QMD

```{r qmd linearity}
obs |>
  ggplot(aes(QMD, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```

### SDI

```{r sdi linearity}
obs |>
  ggplot(aes(SDI, Carbon_Flux_Residual)) +
  geom_point(size=0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.25, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue")
```


## Model All Independent Variables

```{r multiple linear regression}
m <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ LAT + LON + ELEV + BALIVE_METRIC + TPA + QMD + SDI + STDAGE + FOREST_TYPE_GROUP + ProjectionYears
)
summary(m)
Anova(m)
```

```{r check model m}
check_model(m)
```

### Centered

Same, but centered

```{r m_ctr}
m_ctr <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ LAT_CTR + LON_CTR + ELEV_CTR + BALIVE_CTR + TPA_CTR + QMD_CTR + SDI_CTR + STDAGE_CTR + ProjectionYears + FOREST_TYPE_GROUP
)
summary(m_ctr)
Anova(m_ctr)
```

check it
```{r check m_ctr}
check_model(m_ctr)
```

Looks like linearity is not respected, something to explore.
Homgeneity of variance is also a scatterplot, not sure what's up there.
Posterior predictive check looks bad - a spike vs. a bell curve. Not sure how to fix that.

## Model P<0.001 Variables

It looks like LON, TPA, and STDAGE aren't contributing, so let's take them out.
Furthermore, BALIVE and SDI conflict with each other (high colinearity).
So keep only one of them. Since SDI is comptued from BA and TPA, let's just
keep BALIVE.

```{r m2}
m2 <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ LAT + ELEV + BALIVE_METRIC + QMD + ProjectionYears +
    FOREST_TYPE_GROUP
)
summary(m2)
Anova(m2)
```

```{r check m2}
check_model(m2)
```

### SDI

Out of curiosity, let's see what happens if we keep SDI instead of BALIVE.

```{r m2 with sdi}
m2_sdi <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ LAT + ELEV + SDI + QMD + ProjectionYears + FOREST_TYPE_GROUP
)
summary(m2_sdi)
Anova(m2_sdi)

```

```{r check m2_sdi}
check_model(m2_sdi)
```

Pretty much the same.

### ECOCD

Out of curiosity, what happens if we swap FOREST_TYPE_GROUP for ECOCD?

```{r m2 with ecocd}
m2_ecocd <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ LAT + ELEV + SDI + QMD + ProjectionYears + ECOCD
)
summary(m2_ecocd)
Anova(m2_ecocd)

```

```{r check m2_ecocd}
check_model(m2_ecocd)
```


It does not find ECOCD interesting, and we can't plot the summary when using
ECOCD, so we'll ignore it.

## Scaled Variables

Let's try scaling some things.

```{r m3}
m3 <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ sin(LAT * pi / 180) + ELEV + sqrt(BALIVE_METRIC) + QMD + ProjectionYears + FOREST_TYPE_GROUP
)
summary(m3)
Anova(m3)
```

```{r check m3}
check_model(m3)
```

## No Projection Years

I really want a model that's independent of projection years. We know it'll only
be valid up to ~20 years.
Let's start with all the independent vars again.

```{r m4}
m4 <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ LAT + LON + ELEV + BALIVE_METRIC + TPA + QMD + SDI + STDAGE + FOREST_TYPE_GROUP
)
summary(m4)
Anova(m4)
```

Well that's completely different.

```{r check m4}
check_model(m4)
```

Once again, BALIVE and SDI are colinear, so we pick BALIVE.

```{r m5}
m5 <- lm(
  data = obs,
  formula = Carbon_Flux_Residual ~ ELEV + BALIVE_METRIC + FOREST_TYPE_GROUP
)
summary(m5)
print("-----")
Anova(m5)
```


```{r check m5}
check_model(m5)
```

## Apply Carbon Flux Bias Correction

```{r apply m5}
bias_adjusted <- obs |>
  select(-ends_with("_CTR")) |>
  mutate(Carbon_Flux_Bias = unname(predict(m5, obs))) |>
  mutate(
    Carbon_Flux_Adjusted = Carbon_Flux_Calb - Carbon_Flux_Bias,
    Carbon_Flux_BAR = Carbon_Flux_Adjusted - Carbon_Flux_Srvy
  )
```

```{r plot apply m5}
bias_adjusted |>
  select(c("FOREST_TYPE_GROUP", "Carbon_Flux_Residual", "Carbon_Flux_BAR")) |>
  pivot_longer(cols = c("Carbon_Flux_Residual", "Carbon_Flux_BAR")) |>
  ggplot(aes(value, fill = name)) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.25) +
  geom_histogram(binwidth = 0.1, position = "dodge") +
  scale_fill_brewer(name = "Series", labels = c("Adjusted", "Original"), palette="Set1") +
  coord_cartesian(xlim = c(-5, 5)) +
  #facet_wrap("FOREST_TYPE_GROUP") +
  theme_bw() +
  scale_x_continuous(name = bquote("Carbon Flux Residual" ~ (`Mg C` %.% hectare^-1 %.% year^-1))) +
  scale_y_continuous(name = "Number of Plots") +
  ggtitle("Bias-Adjusted Carbon Flux Residuals")
```

### Residual by Starting Stand BA

```{r plot carbon flux residual vs balive, eval = FALSE}
bias_adjusted |>
  mutate(BALIVE_METRIC = cut(BALIVE_METRIC, 0:19 * 5, include.lowest = TRUE)) |>
  ggplot(aes(BALIVE_METRIC, Carbon_Flux_BAR)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("Carbon Flux Residual" ~ (`Mg C` %.% hectare^-1 %.% year^-1))) +
  ggtitle("Carbon Flux Residual vs. Stand Starting Basal Area (with bias correction)") +
  geom_boxplot(outlier.size = 0.1, varwidth = TRUE) +
  coord_cartesian(ylim = c(-5, 5))

```

```{r plot carbon flux residual vs balive, eval = FALSE}
bias_adjusted |>
  ggplot(aes(BALIVE_METRIC, Carbon_Flux_BAR)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.25) +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("Carbon Flux Residual" ~ (`Mg C` %.% hectare^-1 %.% year^-1))) +
  ggtitle("Carbon Flux Residual vs. Stand Starting Basal Area (Bias Corrected)") +
#  labs(caption = "With 10-year calibration") +
  scale_y_continuous(limits = c(-10, 5))
```

### Bias-Corrected vs. Measured Flux

```{r proj_vs_meas_carbon_flux, eval = FALSE}
bias_adjusted |>
  ggplot(aes(Carbon_Flux_Srvy, Carbon_Flux_Adjusted)) +
  geom_hline(yintercept = 0, linewidth = 0.25, color = "black") +
  geom_vline(xintercept = 0, linewidth = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linewidth = 0.25, linetype = "dashed", color = "lightgray") +
  #geom_bin2d(binwidth = 0.5) +
  #scale_fill_viridis_c() +
  geom_point(size = 0.25, alpha = 0.25) +
  # stat_poly_* plots and annotates a fit; they need to be configured to match
  stat_poly_line(
    formula = y ~ x,
    fullrange = TRUE,
    se = FALSE,
    color = "blue",
    linetype = "dashed",
    linewidth = 0.5
  ) +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, size = 3, color = "blue") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  ggtitle("Bias-Corrected vs. Measured Carbon Flux") +
  xlab(bquote("Measured Annual Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% yr^-1))) +
  ylab(bquote("Projected Annual Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% yr^-1))) +
  coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5))
```

# BAI Bias Correction

## Model BAI Bias

```{r m5_bai}
m5_bai <- lm(
  data = obs,
  formula = BAI_Residual ~ ELEV + BALIVE_METRIC + FOREST_TYPE_GROUP
)
summary(m5_bai)
print("-----")
Anova(m5_bai)
```


```{r check m5_bai}
check_model(m5_bai)
```

## Apply BAI Bias Correction

```{r apply m5_bai}
bai_adjusted <- obs |>
  select(-ends_with("_CTR")) |>
  mutate(BAI_Bias = unname(predict(m5_bai, obs))) |>
  mutate(
    BAI_Adjusted = BAI_Calb - BAI_Bias,
    BAI_BAR = BAI_Adjusted - BAI_Srvy,
    BA_Calb_Adjusted = BA_Calb - BAI_Bias * ProjectionYears,
    BA_Delta_Adjusted = BA_Calb_Adjusted - BA_Srvy
  )
```

```{r plot apply m5_bai}
bai_adjusted |>
  select(c("FOREST_TYPE_GROUP", "BAI_Residual", "BAI_BAR")) |>
  pivot_longer(cols = c("BAI_Residual", "BAI_BAR")) |>
  ggplot(aes(value, fill = name)) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.25) +
  geom_histogram(binwidth = 0.1, position = "dodge") +
  scale_fill_brewer(name = "Series", labels = c("Adjusted", "Original"), palette="Set1") +
  #coord_cartesian(xlim = c(-5, 5)) +
  #facet_wrap("FOREST_TYPE_GROUP") +
  theme_bw() +
  scale_x_continuous(name = bquote("BAI Residual" ~ (m^2 %.% hectare^-1 %.% year^-1))) +
  scale_y_continuous(name = "Number of Plots") +
  ggtitle("BAI Residuals with Correction")
```

## Residual by Starting Stand BA

```{r plot carbon flux residual vs balive, eval = FALSE}
bai_adjusted |>
  mutate(BALIVE_METRIC = cut(BALIVE_METRIC, 0:19 * 5, include.lowest = TRUE)) |>
  ggplot(aes(BALIVE_METRIC, BAI_BAR)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("BAI Residual" ~ (m^2 %.% hectare^-1 %.% year^-1))) +
  ggtitle("BAI Residual vs. Stand Starting Basal Area (with bias correction)") +
  geom_boxplot(outlier.size = 0.1, varwidth = TRUE) +
  coord_cartesian(ylim = c(-5, 5))
```

```{r plot carbon flux residual vs balive, eval = FALSE}
bai_adjusted |>
  ggplot(aes(BALIVE_METRIC, BAI_BAR)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("BAI Residual" ~ (m^2 %.% hectare^-1 %.% year^-1))) +
  ggtitle("BAI Residual vs. Stand Starting Basal Area (Corrected)") #+
#  labs(caption = "With 10-year calibration") +
  #scale_y_continuous(limits = c(-5, 5))
```

## Quadratic BAI

What happens if we fit a quadratic model just to startign BA?

```{r m6}
m6 <- lm(
  data = obs,
  formula = BAI_Residual ~ sqrt(BALIVE_METRIC) + FOREST_TYPE_GROUP
)
summary(m6)
Anova(m6)
```

```{r apply m6}
bai_m6 <- obs |>
  select(-ends_with("_CTR")) |>
  mutate(BAI_Bias = unname(predict(m6, obs))) |>
  mutate(
    BAI_Adjusted = BAI_Calb - BAI_Bias,
    BAI_BAR = BAI_Adjusted - BAI_Srvy
  )
```


```{r m6, eval = FALSE}
bai_m6 |>
  ggplot(aes(BALIVE_METRIC, BAI_BAR)) +
  geom_point(size=0.25) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 1, color = "blue") +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x, color = "blue") +
  theme_bw() +
  xlab(bquote("Starting Basal Area" ~ (m^2 %.% hectare^-1))) +
  ylab(bquote("BAI Residual" ~ (m^2 %.% hectare^-1 %.% year^-1))) +
  ggtitle("BAI Residual vs. Stand Starting Basal Area (Bias Corrected)")
#  labs(caption = "With 10-year calibration") +
#  scale_y_continuous(limits = c(-5, 5))
```


So, no benefit to using the quadratic there.

## BA Error with BAI Correction

If we apply BAI correction, what does that do to BA error?

```{r ba error}
bai_adjusted |>
  select(StandID, BA_Delta_Calb, BA_Delta_Adjusted) |>
  pivot_longer(
    cols = c("BA_Delta_Calb", "BA_Delta_Adjusted"),
    names_to = "Series",
    values_to = "BA_Error"
  ) |>
  ggplot(aes(BA_Error, fill = Series)) +
  geom_histogram(binwidth = 2.5, position = "dodge") +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.25) +
  theme_bw() +
  xlab(bquote("Projection Error in Basal Area" ~ (m^2 %.% ha^-1))) +
  ylab("Number of Plots") +
  scale_fill_brewer(name = "Series", labels = c("Bias Corrected", "Original"), palette="Set1") +
  ggtitle("Basal Area Projection Error with Bias Correction")
```

How does this look on a map?

```{r map bai error}
tar_load(fvsne_states)

fvsne_states_map <- fvsne_states |>
  distinct(STATE_NAME) |>
  rename(region = STATE_NAME) |>
  _$region |>
  map_data("state", region = _)

nrsgro_plot_locations <- nrsgro_plot |>
  filter(!is.na(LAT) & !is.na(LON)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(row_number() == 1) |>
  ungroup() |>
  rename(lat = LAT, long = LON) |>
  filter_add_stand_id() |>
  select(STAND_ID, lat, long) |>
  left_join(
    bai_adjusted |>
      select(StandID, BA_Delta_Calb, BA_Delta_Adjusted),
    by = join_by(STAND_ID == StandID)
  )

ggplot(fvsne_states_map, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  geom_point(
    nrsgro_plot_locations,
    mapping = aes(group = STAND_ID, color = BA_Delta_Calb),
    size = 0.25
  ) +
  coord_sf(crs = 4326) +
  annotation_scale(location = "br") +
  annotation_north_arrow(
    location = "tl",
    width = unit(1, "cm"),
    pad_x = unit(0.75, "cm"),
    pad_y = unit(0.5, "cm")
  ) +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_viridis_c(name = "BA Residual", limits = c(-30, 30)) +
  ggtitle("Basal Area Residual (uncorrected)")

ggplot(fvsne_states_map, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  geom_point(
    nrsgro_plot_locations,
    mapping = aes(group = STAND_ID, color = BA_Delta_Adjusted),
    size = 0.25
  ) +
  coord_sf(crs = 4326) +
  annotation_scale(location = "br") +
  annotation_north_arrow(
    location = "tl",
    width = unit(1, "cm"),
    pad_x = unit(0.75, "cm"),
    pad_y = unit(0.5, "cm")
  ) +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_viridis_c(name = "BA Residual", limits = c(-30, 30)) +
  ggtitle("Basal Area Residual (corrected)")


```

Pick a plot; show:
- FIA measured BA
- FVS projected BA
- Annualized FIA BAI
- Annualized FVS BAI
- Bias corrected FVS BA

```{r single stand bias correction}
sample_stand <- nrsgro_plot_stats |>
  filter(YEARS > 20) |>
  slice_sample(n=1) |>
  select(STATECD, COUNTYCD, PLOT)

tmp_fia_measured <- sample_stand |>
  inner_join(nrsgro_plot_stats, by = join_by(STATECD, COUNTYCD, PLOT)) |>
  filter_add_stand_id() |>
  select(STAND_ID, MEASYEAR, BALIVE_METRIC) |>
  rename(StandID = STAND_ID, Year = MEASYEAR, FIA_Observed = BALIVE_METRIC) |>
  mutate(FVS_Projected = NA, FVS_Corrected = NA)

tmp_fvs_measured <- sample_stand |>
  filter_add_stand_id() |>
  rename(StandID = STAND_ID) |>
  select(StandID) |>
  inner_join(bai_adjusted, by = join_by(StandID)) |>
  rename(FVS_Projected = BA_Calb) |>
  select(StandID, Year, FVS_Projected) |>
  mutate(FIA_Observed = NA, FVS_Corrected = NA)

tmp_fvs_corrected <- sample_stand |>
  filter_add_stand_id() |>
  rename(StandID = STAND_ID) |>
  select(StandID) |>
  inner_join(bai_adjusted, by = join_by(StandID)) |>
  rename(FVS_Corrected = BA_Calb_Adjusted) |>
  select(StandID, Year, FVS_Corrected) |>
  mutate(FIA_Observed = NA, FVS_Projected = NA)

tmp_fia_measured |>
  union_all(tmp_fvs_measured) |>
  union_all(tmp_fvs_corrected) |>
  pivot_longer(
    cols = c("FIA_Observed", "FVS_Projected", "FVS_Corrected"),
    names_to = "Series"
  ) |>
  filter(!is.na(value)) |>
  ggplot(aes(Year, value, color = Series)) +
  geom_point() +
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw() +
  ggtitle("Basal Area Projection and Correction") +
  xlab("Year") +
  ylab(bquote("Basal Area" ~ (m^2 %.% ha^-1)))
```

# Per-Timestep BA

Can we use per-timestep BA?

Let's start by seeing how well our projected and measured timesteps line up.

```{r timestep alignment}
tar_load(nrsgro_ca10_summary)
tar_load(nrsgro_srvy_summary)

timestep_alignment <- nrsgro_ca10_summary |>
  select(StandID, Year) |>
  group_by(StandID) |>
  arrange(desc(Year)) |>
  mutate(StepSize = lag(Year) - Year) |>
  ungroup() |>
  rename(Projection_Year = Year) |>
  left_join(
    nrsgro_srvy_summary |>
      select(StandID, Year) |>
      rename(Measurement_Year = Year),
    by = join_by(StandID),
    relationship = "many-to-many"
  ) |>
  mutate(YearDelta = abs(Projection_Year - Measurement_Year)) |>
  group_by(StandID, Projection_Year) |>
  arrange(YearDelta) |>
  filter(row_number() == 1) |>
  ungroup()

timestep_alignment |>
  #filter(StepSize == 10) |>
  filter(!is.na(StepSize)) |>
  ggplot(aes(StepSize)) +
  geom_histogram(binwidth = 1) +
  scale_y_continuous() #transform = "log10")
```

```{r}
timestep_alignment |>
  filter(StepSize == 10) |>
  filter(YearDelta == 0) |>
  nrow() /
  timestep_alignment |>
  filter(StepSize == 10) |>
  nrow()
```

So, 40% of timesteps are length 10, and 70% of timesteps with length 10 land
within 1 year of a measurement.

But I think I wouldn't want to do it this way. I'd want to do a run landing on
each survey year where I can get a 10-year timestep in there.

Proposal 1:
- Since our model of BAI bias is independent of projection years, we can assume
  that it's good up to 4 timesteps.
- For longer runs:
  - Run up to 4 timesteps.
  - Apply BA bias correction proportionately to each tree record
  - Set up the next run, up to 4 timesteps.
    - This will use the new BA as the starting BA
- For harvest:
  - Run up to the pre-harvest timestep.
  - Apply BA bias correction proportionately to each tree record
  - Set up and run a 1-year harvest run (removal will use the corrected biomass)
    - Assume that all harvest impact is stored in the tables that we'll feed 
      back in as input, e.g. down dead, woody debris, etc.
  - Start subsequent post-harvest runs from the post-harvest plot, using the
    post-harvest BA
    - These will be runs with 10-year timesteps
    - No need to re-align with the original series

Proposal 2:
- For each plot, find all surveys that are at least 10 years from the initial survey
- Run FVS in an n-10-1 timestep configuration for each of those (plot, invyr)s
  - These runs will incorporate both regen and 10-year future calibration
- Compute a model of bias for 10-year timesteps given conditions at start and end
  of the 10-year timestep.
  - This doesn't work - we don't have real-world conditions at the start of the
    10-year timestep.
  - We could get close by bias-correcting the short run
  - Which might need a separate model of FVS bias for short runs, which
  - we could build for individual inventories
- For longer runs:
  - Run a 10-year timestep
  - Apply BA bias correction proportionately to each tree record
  - Set up the next 10-year timestep

