---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---
# Setup

```{r setup, include=FALSE}
library(tidyverse)
library(reshape2) # melt()
library(RSQLite)
library(dbplyr) # window_order()
library(measurements) # conv_unit and conv_multiunit
library(maps)
library(pbapply) # Progress bar for long operations
options(scipen = 9) # Use integer notation for numbers under 9 digits
```

```{r source-functions}
source('R/functions.R')
```

## Use Cache

Some datasets take a long time to compute; these are normally cached and
read from cache. Set `USE_CACHE` to `FALSE` to re-compute everything, and
`TRUE` to use cached values if available.

```{r USE_CACHE}
USE_CACHE <- TRUE
```

## Species Crosswalk

Different systems use different ways of encoding species. Here's a single
crosswalk for them all.

```{r species_mixin, cache=TRUE}
species_mixin <- read_rds("data/intermediate/species_mixin.rds")
```

# Harvested Plots in the Northeast

Find all FIA plots that:

* are in the northeastern region, and
* use the modern plot design, and
* have at least 10 years of measurements, and
* are fully forested, and
* is not disturbed ("disturbed" is separate from harvest), and
* is harvested, and
* is not seeded, fertilized, herbicided, etc. post-harvest, and
* is measured both pre- and post-harvest, and
* have a single condition for the entire plot

```{r plot_harvested, cache=TRUE}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

harvest_mixin <- tbl(fia, 'TREE') |>
  filter(STATUSCD == 3) |>
  distinct(STATECD, COUNTYCD, PLOT, INVYR) |>
  union(
    tbl(fia, 'COND') |>
      filter((TRTCD1 == 10) | (TRTCD2 == 10) | (TRTCD3 == 10)) |>
      distinct(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  mutate(HARVEST = 1) 

fia_cond <- tbl(fia, 'COND') |>
  select(
    STATECD, COUNTYCD, PLOT, PLT_CN, CONDID, INVYR,
    COND_STATUS_CD, BALIVE, DSTRBCD1, DSTRBCD2, DSTRBCD3,
    TRTCD1, TRTYR1, TRTCD2, TRTYR2, TRTCD3, TRTYR3
  )

fia_plot <- tbl(fia, 'PLOT') |>
  # Narrow and rename columns to facilitate join
  select(CN, DESIGNCD, SRV_CN, MEASYEAR, ECOSUBCD) |>
  rename(PLT_CN = CN)

fia_plotgeom <- tbl(fia, 'PLOTGEOM') |>
  # Narrow and rename columns to facilitate join
  select(CN, FVS_VARIANT) |>
  rename(PLT_CN = CN)

plot_harvested <- fia_cond |>
  left_join(fia_plot, by = join_by(PLT_CN)) |>
  left_join(fia_plotgeom, by = join_by(PLT_CN)) |>
  fvsne_plots_filter() |>
  modern_plots_filter() |>
  filter(INVYR >= 1999) |>
  long_measurement_filter() |>
  forested_plots_filter() |>
  single_condition_plots_filter() |>
  undisturbed_plots_filter() |>
  no_unnatural_regen_filter() |> 
  # harvested_plots_filter() |>
  left_join(harvest_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  # Special case:
  # - If a plot is harvested, AND
  # - the first harvest is also the first inventory, THEN
  # - we want to not process the first harvest as a harvest, BECAUSE
  # - we don't know which trees were on the plot prior to that,
  # - which means we can't figure out natural regen
  mutate(
    HARVEST = case_when(
      INVYR == min(INVYR, na.rm = TRUE) ~ 0,
      is.na(HARVEST) ~ 0,
      .default = HARVEST
    )
  ) |>
  ungroup() |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(NUM_HARVESTS = sum(HARVEST, na.rm = TRUE)) |>
  filter(NUM_HARVESTS > 0) |>
  ungroup() |>
  # measured_pre_post_harvest_filter() |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(
    MIN_MEASYEAR = min(MEASYEAR, na.rm = TRUE),
    MAX_MEASYEAR = max(MEASYEAR, na.rm = TRUE),
    MIN_HRVYR = min(if_else(HARVEST == 1, MEASYEAR, 9999), na.rm = TRUE),
    MAX_HRVYR = max(if_else(HARVEST == 1, MEASYEAR, 0), na.rm = TRUE),
  ) |> 
  filter(
    (MIN_MEASYEAR < MIN_HRVYR) &
    (MAX_MEASYEAR - MAX_HRVYR >= 10)
  ) |> 
  ungroup() |>
  distinct(STATECD, COUNTYCD, PLOT) |>
  collect()

DBI::dbDisconnect(fia)
remove(fia, fia_cond, fia_plot, fia_plotgeom)
nrow(plot_harvested)
```

## Stand Stats

Fetch more info about these stands.

```{r stand_stats, cache=TRUE}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

tree_mixin <- tbl(fia, 'TREE') |>
  select(STATECD, COUNTYCD, PLOT, CONDID, INVYR, DIA, CARBON_AG, TPA_UNADJ) |>
  group_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR) |>
  summarize(
    CARBON_AG = sum(CARBON_AG, na.rm = TRUE),
    CPA = sum(CARBON_AG * TPA_UNADJ, na.rm = TRUE),
    BA_TREES = sum(if_else(DIA >= 1, TPA_UNADJ, 0), na.rm = TRUE),
    .groups = "keep"
  )

harvest_mixin <- tbl(fia, 'TREE') |>
  semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  filter(STATUSCD == 3) |>
  distinct(STATECD, COUNTYCD, PLOT, INVYR) |>
  mutate(HARVEST_TREE = TRUE) |>
  full_join(
    tbl(fia, 'COND') |>
      semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
      filter((TRTCD1 == 10) | (TRTCD2 == 10) | (TRTCD3 == 10)) |>
      distinct(STATECD, COUNTYCD, PLOT, INVYR) |>
      mutate(HARVEST_COND = TRUE),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  mutate(HARVEST = TRUE)

trtyr_mixin <- tbl(fia, "COND") |>
  semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  select(
    STATECD, COUNTYCD, PLOT, INVYR,
    TRTCD1, TRTYR1, TRTCD2, TRTYR2, TRTCD3, TRTYR3
  ) |>
  # If harvest is performed in multiple years, we can't know which trees
  # were harvested in which year. In theory, this would be in MORTYR, but
  # 1 in 5000 harvested tree records has MORTYR filled in. Therefore, we
  # retain only the latest harvest year - this will be pessimistic with
  # respect to regeneration.
  mutate(
    TRTYR1 = if_else(TRTCD1 == 10, TRTYR1, NA),
    TRTYR2 = if_else(TRTCD2 == 10, TRTYR2, NA),
    TRTYR3 = if_else(TRTCD3 == 10, TRTYR3, NA),
    TRTYR = case_when(
      !is.na(TRTYR3) ~ TRTYR3,
      !is.na(TRTYR2) ~ TRTYR2,
      .default = TRTYR1
    )
  ) |>
  select(STATECD, COUNTYCD, PLOT, INVYR, TRTYR) |>
  filter(!is.na(TRTYR)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  window_order(TRTYR) |>
  mutate(TRTNUM = row_number()) |>
  ungroup()

cond_mixin <- tbl(fia, 'COND') |>
  semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  select(
    STATECD, COUNTYCD, PLOT, INVYR, STDAGE, BALIVE, FORTYPCD
  ) |>
  group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
  summarize(
    BALIVE = sum(BALIVE, na.rm = TRUE),
    FORTYPCD = max(FORTYPCD, na.rm = TRUE),
    STDAGE = max(STDAGE, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  left_join(harvest_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(
    HARVEST = case_when(
      INVYR == min(INVYR, na.rm = TRUE) ~ FALSE,
      is.na(HARVEST) ~ FALSE,
      .default = HARVEST
    ),
    HARVEST_TREE = if_else(!is.na(HARVEST_TREE), HARVEST_TREE, FALSE),
    HARVEST_COND = if_else(!is.na(HARVEST_COND), HARVEST_COND, FALSE),
  ) |>
  ungroup()

forest_type_mixin <- tbl(fia, 'REF_FOREST_TYPE') |>
  select(VALUE, MEANING) |>
  rename(FORTYPCD = VALUE) |>
  rename(FORTYPE = MEANING)

# research station has state name and abbreviation
state_mixin <- tbl(fia, 'REF_RESEARCH_STATION') |>
  distinct(STATECD, STATE_NAME, STATE_ABBR)

stand_stats_partial <- tbl(fia, 'PLOT') |>
  semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  filter(INVYR >= 1999) |>
  select(CN, STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, DESIGNCD, ECOSUBCD, LAT, LON) |>
  left_join(tree_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(cond_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(trtyr_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(forest_type_mixin, by = join_by(FORTYPCD)) |>
  rename(`Forest Type` = FORTYPE) |>
  mutate(FORTYPCD = floor(FORTYPCD / 10) * 10) |>
  left_join(forest_type_mixin, by = join_by(FORTYPCD)) |>
  rename(`Forest Type Group` = FORTYPE) |>
  left_join(state_mixin, by = join_by(STATECD)) |>
  collect() |>
  rename(
    long = LON,
    lat = LAT
  ) |>
  mutate(
    `Forest Type Group` = str_replace(`Forest Type Group`, ' group', ''),
    BALIVE_METRIC = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    QMD = sqrt(BALIVE / (BA_TREES * (pi / 576))),
    QMD_METRIC = sqrt(BALIVE_METRIC / (BA_TREES * (pi / 40000))),
    CARBON_METRIC = conv_multiunit(CPA, "lbs / acre", "Mg / hectare"),
    # If we're just stripping off subregion, could also do this as
    #   substr(ECOSUBCD, 1, nchar(ECOSUBCD) - 1)
    # but using sub() gives us more flexibility.
    ECOCD = sub(
      x = ECOSUBCD,
      # M? - Mountain
      # \d+ - Region
      # [A-Z] - Ecoregion
      # [a-z] - Ecosubregion
      pattern = "(M?)(\\d+)([A-Z])([a-z])",
      replacement = "\\1\\2\\3"
    )
  ) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(INVYR) |>
  mutate(
    INVNUM = row_number(),
    BALIVE_START = if_else(row_number() == 1, BALIVE_METRIC, NA),
    STAND_CN = if_else(row_number() == 1, CN, NA),
    MIN_MEASYEAR = min(MEASYEAR, na.rm = TRUE),
    MAX_MEASYEAR = max(MEASYEAR, na.rm = TRUE)
  ) |> 
  ungroup() |>
  arrange(STATECD, COUNTYCD, PLOT, INVYR) |>
  fill(BALIVE_START, STAND_CN) |>
  mutate(
    BALIVE_DELTA = BALIVE_METRIC - BALIVE_START,
    YEARS = MEASYEAR - MIN_MEASYEAR,
    group = sprintf('%02d%03d%05d', STATECD, COUNTYCD, PLOT)
  )

# Figure out which inventories are pre-harvest
pre_harvest_mixin <- stand_stats_partial |>
  filter(HARVEST == 1) |>
  mutate(INVNUM = INVNUM - 1) |> # the inventory prior to harvest is pre-harvest
  rename(PRE_HARVEST = HARVEST) |>
  select(STATECD, COUNTYCD, PLOT, INVNUM, PRE_HARVEST)

stand_stats <- stand_stats_partial |>
  left_join(pre_harvest_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVNUM)) |>
  mutate(
    PRE_HARVEST = if_else(is.na(PRE_HARVEST), 0, PRE_HARVEST)
  )

DBI::dbDisconnect(fia)
remove(fia, cond_mixin, trtyr_mixin, forest_type_mixin, tree_mixin,
       state_mixin, pre_harvest_mixin, harvest_mixin, stand_stats_partial)
```

## Stand Locations

Where are they?

```{r fvsne_states, cache=TRUE}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

fvs_variant_mixin <- tbl(fia, 'PLOTGEOM') |>
  # Narrow and rename columns to facilitate join
  select(CN, FVS_VARIANT)

# research station has state name and abbreviation
state_mixin <- tbl(fia, 'REF_RESEARCH_STATION') |>
  distinct(STATECD, STATE_NAME, STATE_ABBR)

fvsne_states <- tbl(fia, 'PLOT') |>
  left_join(fvs_variant_mixin, by = join_by(CN)) |>
  fvsne_plots_filter() |>
  distinct(STATECD) |>
  left_join(state_mixin, by = join_by(STATECD)) |>
  mutate(region = str_to_lower(STATE_NAME)) |>
  collect()

DBI::dbDisconnect(fia)

remove(fia, fvs_variant_mixin, state_mixin)
```

```{r map_by_ecoregion}
map_data('state', fvsne_states$region) |> 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + 
  geom_point(
    data = stand_stats |>
      group_by(STATECD, COUNTYCD, PLOT) |>
      filter(row_number() == 1),
    aes(color = ECOCD)
  ) +
  coord_quickmap() +
  theme_bw() +
  scale_color_discrete(name = "Ecoregion")
```

# Harvested Trees

Find the trees that were removed during harvest.

Note that FIA.TREE.STATUSCD == 3 indicates that a tree has been harvested,
but not all harvested trees have STATUSCD == 3. Therefore we approximate
harvested trees by finding all trees that exist in a pre-harvest inventory,
but do not exist (or are marked as harvested) in the corresponding harvest
inventory.

## Trees removed in harvest

Fetch trees that were standing pre-harvest, that are not standing post-harvest.

```{r tree_removed_in_harvest, cache=TRUE}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

tree_in_pre_harvest_plots <- tbl(fia, "TREE") |>
  semi_join(
    stand_stats |>
      filter(PRE_HARVEST == 1) |>
      select(STATECD, COUNTYCD, PLOT, INVYR),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR),
    copy = TRUE
  ) |>
  filter(STATUSCD == 1 | STATUSCD == 2) |> # live or dead trees
  select(CN, STATECD, COUNTYCD, PLOT, SUBP, TREE, INVYR, MORTYR, STATUSCD) |>
  collect() |>
  left_join(
    stand_stats |> select(STATECD, COUNTYCD, PLOT, INVYR, INVNUM),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  )

tree_in_harvested_plots <- tbl(fia, "TREE") |>
  semi_join(
    stand_stats |>
      filter(HARVEST == 1) |>
      select(STATECD, COUNTYCD, PLOT, INVYR),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR),
    copy = TRUE
  ) |>
  filter(STATUSCD == 1 | STATUSCD == 2) |> # live and dead trees are still there
  select(STATECD, COUNTYCD, PLOT, SUBP, TREE, INVYR) |>
  collect() |>
  left_join(
    stand_stats |> select(STATECD, COUNTYCD, PLOT, INVYR, INVNUM),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  mutate(
    INVNUM = INVNUM - 1
  )

tree_removed_in_harvest <- tree_in_pre_harvest_plots |>
  anti_join(
    tree_in_harvested_plots,
    by = join_by(STATECD, COUNTYCD, PLOT, SUBP, TREE, INVNUM)
  )

DBI::dbDisconnect(fia)

remove(fia, tree_in_harvested_plots, tree_in_pre_harvest_plots)
```

## FIA Harvested Trees

Fetch trees that FIA says were harvested.

```{r tree_fia_harvested, cache=TRUE}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

plot_mixin <- tbl(fia, 'PLOT') |>
  select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR)
  
tree_fia_harvested <- tbl(fia, 'TREE') |>
  semi_join(
    stand_stats |> filter(HARVEST == 1),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR),
    copy = TRUE
  ) |>
  select(STATECD, COUNTYCD, PLOT, SUBP, TREE, INVYR, STATUSCD) |>
  filter(STATUSCD == 3) |>
  left_join(plot_mixin, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  collect()

DBI::dbDisconnect(fia)

remove(fia, plot_mixin)
```

## Validation

Were all the trees that FIA says were harvested actually removed?
```{r tree_harvested_but_not_removed}
tree_harvested_but_not_removed <- tree_fia_harvested |>
  anti_join(tree_removed_in_harvest, by = join_by(STATECD, COUNTYCD, PLOT, SUBP, TREE))
stopifnot(tree_harvested_but_not_removed |> nrow() == 0)
```

It is expected that FIA did not mark all removed trees as harvested.
```{r tree_removed_but_not_harvested}
tree_removed_but_not_harvested <- tree_removed_in_harvest |>
  anti_join(tree_fia_harvested, by = join_by(STATECD, COUNTYCD, PLOT, SUBP, TREE))
tree_removed_but_not_harvested |> nrow()
```

All harvested trees should be on harvested plots.
```{r stopifnot_tree_removed_from_non_harvested_plots}
stopifnot(
  tree_removed_in_harvest |>
    distinct(STATECD, COUNTYCD, PLOT, INVNUM) |>
    # INVNUM is the last inventory in which the tree was standing,
    # so it is marked harvested in the next inventory
    mutate(INVNUM = INVNUM + 1) |>
    anti_join(
      stand_stats |> filter(HARVEST == 1),
      by = join_by(STATECD, COUNTYCD, PLOT, INVNUM)
    ) |>
    nrow() == 0
)
```

There are some plots that do not have trees removed.
```{r plot_harvested_without_tree_removed}
plot_harvested_without_tree_removed <- plot_harvested |>
  anti_join(
    tree_removed_in_harvest |>
      distinct(STATECD, COUNTYCD, PLOT),
    join_by(STATECD, COUNTYCD, PLOT)
  )
plot_harvested_without_tree_removed |> nrow()
```

## Remove Harvested Plots Without Removed Trees

Remove those plots from consideration.
```{r ignore plots with no trees removed}
join_cols <- plot_harvested_without_tree_removed |>
  names()
plot_harvested <- plot_harvested |>
  anti_join(plot_harvested_without_tree_removed, join_cols)
stand_stats <- stand_stats |>
  anti_join(plot_harvested_without_tree_removed, join_cols)
rm(join_cols)
```

What's left?
```{r map_by_ecoregion2}
map_data('state', fvsne_states$region) |> 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + 
  geom_point(
    data = stand_stats |>
      group_by(STATECD, COUNTYCD, PLOT) |>
      filter(row_number() == 1),
    aes(color = ECOCD)
  ) +
  coord_quickmap() +
  theme_bw() +
  scale_color_discrete(name = "Ecoregion")
```
# Harvest Prescription for FVS

When running FVS, we will need:

* CutList / CutLiDB - list of trees harvested
* TreeList / TreeLiDB - list of trees; probably we want cycle 0 captured
* ATrtList / ATrtLiDB - After treatment (post treelist, pre harvest) tree list

## Prescription Table

Prescription table:

- Find all years in which trees were removed.
  - For trees marked as harvested, the estimated year of harvest is in MORTYR
    - 1 of 5580 records has MORTYR
    - this is not worth handling for one tree
  - For trees removed from conditions marked as harvested, the estimated year(s)
    of harvest are in TRTYR1/2/3
    - when there are multiple harvest years, we can't tell which trees are
      harvested in which year
    - when there are multiple harvest years, pick the latest one
  - if there are trees left that don't have a harvest year, put their harvest
    the year before measurement. This will be pessimistic with respect to regen,
    but fits the overall distribution
- Create ordinals for removal.
  - Each plot needs to have < 10 years in which removals occur to keep life easy
  - sort the years and number them
- Consolidate harvests
  - we don't want lots of short cycles, but
  - Essential FVS figure 3-1 shows that thinning happens once per cycle,
    no matter how many years are involved

We have access to:
- Stand_CN - ignored, but used for filtering
- Stand_ID - ignored
- PLOT_CN - ignored
- PLOT_ID
- TREE_CN - ignored, but can be used for join
- TREE_ID

```{r tree_prescription}
trtyr_mixin <- stand_stats |>
  filter(HARVEST == TRUE) |>
  mutate(
    TRTYR = if_else(is.na(TRTYR), MEASYEAR - 1, TRTYR),
    TRTNUM = if_else(is.na(TRTNUM), 1, TRTNUM),
    INVNUM = INVNUM - 1
  ) |>
  select(CN, STATECD, COUNTYCD, PLOT, INVNUM, TRTYR, TRTNUM) |>
  rename(STAND_CN = CN)


tree_prescription <- tree_removed_in_harvest |>
  inner_join(
    trtyr_mixin,
    by = join_by(STATECD, COUNTYCD, PLOT, INVNUM)
  ) |>
  rename(TREE_CN = CN) |>
  mutate(PRESCRIPTION = TRTNUM + 1) |> # PRESCRIPTION range is 2-9
  select(TREE_CN, PRESCRIPTION)
```

We can create a map from year to treatment number as:
```
stand_stats |> filter(!is.na(TRTYR)) |> select(STAND_CN, TRTYR, TRTNUM)
```

## Compose FVS Database

FVS fetches stand and tree data from a SQLite database. Unfortunately,
the way it queries the database prevents doing "intereting" things,
such as attaching a second database for prescription overrides. So we need
to write an entire database with the information FVS needs, which is
the FVS_StandInit_Plot and FVS_TreeInit_Plot tables.

```{r fvs_starting_stands}
fvs_starting_stands <- stand_stats |>
  filter(MEASYEAR == MIN_MEASYEAR)
```

Write a database for use just by this run.

```{r fvs_nrsharvest_prsc.db}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
fvs <- DBI::dbConnect(RSQLite::SQLite(), 'data/fvs/FVS_NRSHarvest_PRSC.db')

fvs_standinit_plot <- tbl(fia, "FVS_StandInit_Plot") |>
  semi_join(
    fvs_starting_stands |> select(CN) |> rename(STAND_CN = CN),
    by = join_by(STAND_CN),
    copy = TRUE
  ) |>
  collect()
DBI::dbWriteTable(fvs, "FVS_StandInit_Plot", fvs_standinit_plot, overwrite = TRUE)

fvs_treeinit_plot <- tbl(fia, "FVS_TreeInit_Plot") |>
  semi_join(
    fvs_starting_stands |> select(CN) |> rename(STAND_CN = CN),
    by = join_by(STAND_CN),
    copy = TRUE
  ) |>
  collect() |>
  select(!PRESCRIPTION) |>
  left_join(tree_prescription, by = join_by(TREE_CN))
DBI::dbWriteTable(fvs, "FVS_TreeInit_Plot", fvs_treeinit_plot, overwrite = TRUE)

DBI::dbDisconnect(fvs)
DBI::dbDisconnect(fia)
rm(fia, fvs, fvs_standinit_plot, fvs_treeinit_plot)
```

## Thin Prescription Table

Create a table describing thinning prescriptions by year

```{r fvs_thinprsc}
fvs_ThinPRSC <- function(rows) {
  thinprsc <- function(row) {
    year <- row["year"]
    if ("percent" %in% names(row)) {
      percent <- row["percent"]
    } else {
      percent <- 1
    }
    prsc <- row["prescription"]
    fvs_kwd3("ThinPRSC", year, percent, prsc)
  }
  apply(rows, 1, thinprsc)
}
```

```{r keyfile_thin_section}
keyfile_thin_section <- function(StandCN) {
  # StandCN referes to STAND_CN in stand_stats
  # swizzle that to CN, which is what trtyr_mixin uses
  stand_cn <- stand_stats |>
    filter(STAND_CN == StandCN) |>
    select(CN) |>
    rename(STAND_CN = CN)
  pertinent <- trtyr_mixin |>
    semi_join(stand_cn, by = join_by(STAND_CN)) |>
    filter(!is.na(TRTNUM)) |>
    rename(year = TRTYR) |>
    mutate(prescription = TRTNUM + 1) |>
    select(year, prescription)
  if (nrow(pertinent) > 0) {
    fvs_ThinPRSC(pertinent)
  } else {
    ''
  }
}
```


## Keyword File Definition

```{r keywordfile_section}
keywordfile_section <- function(
    Title, MgmtId, StandID, StandCN,
    FirstYear, LastYear, Timestep,
    Partition, RandSeed = NA
  ) {
  if (!is.na(RandSeed)) {
    StdIdent <- c(
      fvs_kwd0("StdIdent"),
      paste0(StandID, "_Y", Timestep, "_R", RandSeed, " ", Title)
    )
    # Careful: RannSeed != RandSeed
    RannSeed <- fvs_kwd1("RANNSEED", RandSeed)
  } else {
    StdIdent <- c(
      fvs_kwd0("StdIdent"),
      # RannSeed isn't 0, but 0 is an illegal random seed so a fine flag value
      paste0(StandID, "_Y", Timestep, "_R0", " ", Title))
    # Careful: RannSeed != RandSeed
    RannSeed <- c()
  }
  TimeConfig <- fvs_TimeConfig(FirstYear, LastYear, Timestep)
  if (FirstYear == LastYear) {
    Thin <- c() # Don't do thinning for 0-year runs
  } else {
    Thin <- keyfile_thin_section(StandCN)
  }
  c(
    StdIdent,
    fvs_kwd0("StandCN"),
    StandCN,
    fvs_kwd0("MgmtId"),
    MgmtId,
    TimeConfig,
    RannSeed,
    fvs_kwd0("FMIn"), # Begin
    fvs_kwd0("CarbRept"),
    fvs_kwd0("CarbCut"),
    fvs_kwd5("CarbCalc", 1, 1, 0.0425, 9, 11),
    fvs_kwd0("FuelOut"),
    fvs_kwd0("FuelRept"),
    fvs_kwd0("End"), # FMIn
    fvs_kwd0("Database"), # Begin
    fvs_kwd0("DSNIn"),
    paste0("FVS_", Title, "_", MgmtId, ".db"),
    fvs_kwd0("StandSQL"), # Begin
    "SELECT * FROM FVS_StandInit_Plot WHERE Stand_CN = '%Stand_CN%'",
    fvs_kwd0("EndSQL"), # StandSQL
    fvs_kwd0("TreeSQL"), # Begin
    "SELECT * FROM FVS_TreeInit_Plot WHERE Stand_CN = '%Stand_CN%'",
    fvs_kwd0("EndSQL"), # TreeSQL
    fvs_kwd0("DSNOut"),
    paste0("FVS_", Title, "_", MgmtId, "_", Partition, ".db"),
    fvs_kwd1("Summary",  2),
    fvs_kwd2("Computdb", 0, 1),
    fvs_kwd1("MisRpts",  2),
    fvs_kwd1("CarbReDB", 2),
    fvs_kwd1("FuelReDB", 2),
    fvs_kwd1("FuelsOut", 2),
    fvs_kwd0("End"), # Database
    Thin,
    #Regen,
    fvs_kwd0("Process")
  )
}

```

## Harvest Keyword File Generation

```{r fvs_stands_prsc}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
stand_id_mixin <- tbl(fia, 'FVS_StandInit_Plot') |>
  semi_join(
    fvs_starting_stands |>
      select(CN) |>
      rename(STAND_CN = CN),
    by = join_by(STAND_CN),
    copy = TRUE
  ) |>
  select(STAND_CN, STAND_ID) |>
  rename(CN = STAND_CN) |>
  collect()

fvs_stands_prsc <- fvs_starting_stands |>
  left_join(stand_id_mixin, by = join_by(CN))

dbDisconnect(fia)
remove(fia, stand_id_mixin)

# Every stand must have a STAND_ID
stopifnot(fvs_stands_prsc |> filter(is.na(STAND_ID)) |> nrow() == 0)
```

```{r NRSHarvest PRSC}
num_partitions <- 60 # The number of simultaneous processes to use
timestep <- 10
title <- "NRSHarvest"
mgmtid <- "PRSC"
for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".key")
  unlink(filename)
#    print(paste0("Writing ", filename))
  apply(
    expand.grid(rannseed = c(1, 3, 5)),
    1,
    \(runconfig) {
      rannseed <- runconfig["rannseed"]
      apply(
        fvs_stands_prsc |> 
          filter((as.numeric(STAND_ID) %% num_partitions) == (partition - 1)) |>
          arrange(STAND_ID, MIN_MEASYEAR),
        1,
        function(row) {
          write_lines(
            keywordfile_section(
              title,
              mgmtid,
              row['STAND_ID'],
              row['STAND_CN'],
              row['MIN_MEASYEAR'],
              row['MAX_MEASYEAR'],
              timestep,
              partition,
              rannseed
            ),
            filename,
            append = TRUE
          )
        }
      )
    }
  )
  write_lines("Stop", filename, append = TRUE)
}
```

## Survey Keyword File Generation

We also do 0-year projections of the same stands in every measurement year
to provide a comparison.

```{r fvs_stands_srvy}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
fia_fvs_standinit_plot <- tbl(fia, 'FVS_StandInit_Plot') |>
  select(STAND_CN, STAND_ID)

fvs_stands_srvy <- tbl(fia, 'PLOT') |>
  semi_join(plot_harvested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  select(CN, STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, DESIGNCD) |>
  rename(STAND_CN = CN) |>
  left_join(fia_fvs_standinit_plot, by = join_by(STAND_CN)) |>
  collect()

dbDisconnect(fia)
remove(fia, fia_fvs_standinit_plot)
```

The second run is to do a 0-year projection for each measurement year. These
are given management ID "SRVY".

```{r NRSGrowOnly SRVY}
# We're running 0-year timesteps, but FVS requires us to run at least one
# year, so we run 0-year timesteps for 1 year.
timestep <- 1
title <- "NRSHarvest"
mgmtid <- "SRVY"
for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".key")
  unlink(filename)
  #print(paste0("Writing ", filename))
  apply(
    fvs_stands_srvy |>
      filter((as.numeric(STAND_ID) %% num_partitions) == (partition - 1)) |>
      arrange(STAND_ID, MEASYEAR),
    1,
    function(row) {
      write_lines(
        keywordfile_section(
          title,
          mgmtid,
          row['STAND_ID'],
          row['STAND_CN'],
          row['MEASYEAR'],
          row['MEASYEAR'],
          timestep,
          partition
        ),
        filename,
        append = TRUE
      )
    }
  )
  write_lines("Stop", filename, append = TRUE)
}
```
# Run FVS

> TODO nik: do this automatically; that'll be easier using targets

# Results

## Fault Analysis

### Parse Faults

What errors and warnings do we get?
```{r fvs warnings, cache=TRUE}
mgmtid <- "PRSC"
fvs_warnings <- NULL
for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".out")
  fvs_warnings <- bind_rows(
    fvs_warnings,
    read_lines(filename) |>
      as.data.frame(nm="text") |>
      # Filter to just initial STAND ID and WARNING lines
      filter(
        grepl("^ +STAND ID=", text) |
          grepl("^\\*{8} ", text)
      ) |>
      # Parse out the STAND ID and fill it into the following lines
      mutate(
        STAND_ID = case_when(
          grepl("^ +STAND ID=", text) ~ sub("^ +STAND ID= *([^ ]+) .*$", "\\1", text)
        )
      ) |>
      fill(STAND_ID) |>
      # Remove everything but warnings (which now have STAND_ID)
      filter(
        grepl("^\\*{8} ", text)
      )
  )
}
```

Find the distinct faults

```{r distinct fvs warnings}
fvs_warnings_distinct <- fvs_warnings |>
  # Most faults have some leading text, the tag ERROR: or WARNING:,
  # then descriptive text; if there's a number in the descriptive text,
  # it's probably pulled from the data, so we chop off the descriptive
  # text starting at the first number to find the truly distinct warnings
  mutate(
    PROBLEM = case_when(
      grepl("WARNING:", text) ~ sub("^[*. ]*(.*WARNING: [^0-9]+)([0-9].*)?$", "\\1", text),
      grepl("ERROR:", text) ~ sub("^[*. ]*(.*ERROR: [^0-9]+)([0-9].*)?$", "\\1", text),
      .default = text
    ),
    PROBLEM = sub("\\s+$", "", PROBLEM)
  ) |>
  distinct(PROBLEM) |>
  arrange(PROBLEM)
knitr::kable(fvs_warnings_distinct)
```
### Plot Count Mismatch

Plot counts do not match data on the design record could indicate a data prep
error.

```{r keyword context}
debug_stand_id <- fvs_warnings |>
  filter(grepl("FVS09 WARNING: ", text)) |>
  head(1) |>
  pull("STAND_ID")
fvs_warnings |>
  filter(STAND_ID == debug_stand_id)
```

From Essential FVS:
```
This
warning message is printed when either the plot count or nonstockable count determined
from reading the tree data records differs from the values coded on the DESIGN record.
```

Which partition is it in?
```{r keyword context partition}
as.numeric(sub("^([^_]+)_.*$", "\\1", debug_stand_id)) %% num_partitions
```

so the 16th partition, which is partition 17.

Here's the keyword file section:
```
StdIdent  
420301305916_Y10_R1 NRSHarvest
StandCN   
58377688010538
MgmtId    
PRSC
InvYear         2003
TimeInt            0        10
TimeInt            1         7
TimeInt            3         1
NumCycle           3
RANNSEED           1
FMIn      
CarbRept  
CarbCut   
CarbCalc           1         1    0.0425         9        11
FuelOut   
FuelRept  
End       
Database  
DSNIn     
FVS_NRSHarvest_PRSC.db
StandSQL  
SELECT * FROM FVS_StandInit_Plot WHERE Stand_CN = '%Stand_CN%'
EndSQL    
TreeSQL   
SELECT * FROM FVS_TreeInit_Plot WHERE Stand_CN = '%Stand_CN%'
EndSQL    
DSNOut    
FVS_NRSHarvest_PRSC_17.db
Summary            2
Computdb           0         1
MisRpts            2
CarbReDB           2
FuelReDB           2
FuelsOut           2
End       
ThinPRSC        2006         1         3
Process   
```

from the .out file, when processing TreeSQL:
> NUMBER ROWS PROCESSED:   18

The number of plots should match the number of distinct SUBP in FIA:
```{r debug_fia_plot}
debug_fia_plot <- fvs_stands_prsc |>
  filter(STAND_ID == sub("^([^_]+)_.*$", "\\1", debug_stand_id)) |>
  select(STATECD, COUNTYCD, PLOT, INVYR)
tree_removed_in_harvest |>
  semi_join(debug_fia_plot, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  distinct(SUBP) |>
  nrow()
```

This is probably benign - it means there were trees on only 2 out of 4 subplots
at the start of the run.

How common is this?
```{r fvs09 warnings}
fvs_warnings |>
  filter(grepl("FVS09 WARNING: ", text)) |>
  nrow()
```

This happens on 9 out of 400 plots; ignore.

### One Thousand Trees

The TPA fault may indicate a keyword file generation error; let's track it down.
```{r one thousand trees}
debug_stand_id <- fvs_warnings |>
  filter(grepl(" TPA ", text)) |>
  head(1) |>
  pull("STAND_ID")
fvs_warnings |>
  filter(STAND_ID == debug_stand_id)
```

a'ight, what's the species?

```{r fvs species 9}
debug_species <- fvs_warnings |>
  filter(grepl("FVS40 WARNING:", text) & grepl("SPECIES=", text)) |>
  mutate(FVS_SPNO = as.integer(sub("^.*SPECIES=\\s*(\\d+) .*$", "\\1", text))) |>
  group_by(FVS_SPNO) |>
  summarize(N = n(), .groups = "keep") |>
  ungroup() |>
  arrange(desc(N)) |>
  left_join(species_mixin, by = join_by(FVS_SPNO))
debug_species
```

These are plots where the number of seedlings / saplings on the microplots
tally to >1000 TPA.

I think we note that these exist, and ignore them.

How often does this occur?

```{r fvs tpa warnings}
fvs_warnings |>
  filter(grepl(" TPA ", text)) |>
  nrow()
```

27 out of 400 plots.

We can decide to not worry about them.

## Load Results

Load the output from all FVS partitions into a single dataframe

```{r FVS_Carbon, cache=TRUE}
FVS_Carbon <- NULL
FVS_Summary2_East <- NULL

for (partition in 1:num_partitions) {
  filename <- paste0("data/fvs/FVS_", title, "_", mgmtid, "_", partition, ".db")
  if (file.exists(filename)) {
    #print(paste0("Reading ", filename))
    fvs_out_db <- DBI::dbConnect(RSQLite::SQLite(), filename)
    fvs_carbon_tbl = tbl(fvs_out_db, 'FVS_Carbon') |>
      collect() |>
      mutate(
        Title = title,
        MgmtId = mgmtid,
        Timestep = sub("^.*_Y([0-9]+).*$", "\\1", StandID), # Extract timestep
        RanNSeed = sub("^.*_R([0-9]+).*$", "\\1", StandID), # Extract random seed
        Partition = partition,
        StandID = sub("^([^_]+)_.*$", "\\1", StandID) # Extract stand ID
      )
    FVS_Carbon <- bind_rows(FVS_Carbon, fvs_carbon_tbl)

    fvs_summary2_east_tbl = tbl(fvs_out_db, 'FVS_Summary2_East') |>
      collect() |>
      mutate(
        Title = title,
        MgmtId = mgmtid,
        Timestep = sub("^.*_Y([^_]+)_.*$", "\\1", StandID), # Extract timestep
        RanNSeed = sub("^.*_R([^_]+)_.*$", "\\1", StandID), # Extract random seed
        Partition = partition,
        StandID = sub("^([^_]+)_.*$", "\\1", StandID) # Extract stand ID
      )
      FVS_Summary2_East <- bind_rows(FVS_Summary2_East, fvs_summary2_east_tbl)
  
      dbDisconnect(fvs_out_db)
  } else {
    print(paste0("Skipping ", filename))
  }
}

remove(fvs_out_db, fvs_carbon_tbl, fvs_summary2_east_tbl)
```
Consolidate outputs; average the multiple runs with different random seeds.

```{r FVS_Carbon_NONE}
FVS_Carbon_PRSC <- FVS_Carbon |>
  group_by(StandID, Year) |>
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    Total_Stand_Carbon = mean(Total_Stand_Carbon),
    .groups = 'keep'
  ) |>
  ungroup()

# There is no projection for the observed run, so no need to average
#FVS_Carbon_SRVY <- FVS_Carbon |>  
#  filter(MgmtId == "SRVY")
```

### Sanity Check

Sanity check: did we get the outputs we requested?

```{r sanity-check-expected-output}
matched_carbon_tmp <- fvs_stands_prsc |>
  left_join(
    FVS_Carbon_PRSC |>
      select(StandID, Year, Aboveground_Total_Live) |> 
      rename(
        STAND_ID = StandID,
        MIN_MEASYEAR = Year,
        Starting_Carbon = Aboveground_Total_Live
      ),
    by = join_by(STAND_ID, MIN_MEASYEAR)
  ) |>
  left_join(
    FVS_Carbon_PRSC |>
      select(StandID, Year, Aboveground_Total_Live) |>
      rename(
        STAND_ID = StandID,
        MAX_MEASYEAR = Year,
        Ending_Carbon = Aboveground_Total_Live
        ),
    by = join_by(STAND_ID, MAX_MEASYEAR)
  ) |>
  mutate(partition = (as.numeric(STAND_ID) %% num_partitions) + 1)
missing_start <- matched_carbon_tmp |>
  filter(is.na(Starting_Carbon)) |> 
  nrow()
print(missing_start)
missing_end <- matched_carbon_tmp |>
  filter(is.na(Ending_Carbon)) |> 
  nrow()
print(missing_end)
stopifnot(missing_start == 0)
stopifnot(missing_end == 0)
remove(
  matched_carbon_tmp,
  missing_start,
  missing_end
)
```

### Projected Carbon

```{r plot aboveground carbon}
ggplot(
    FVS_Carbon_PRSC,
    aes(x = Year, y = Aboveground_Total_Live)
  ) +
  geom_bin2d(bins = 25) +
  #geom_abline(intercept = 0, slope = 1) +
  #geom_smooth(
  #  method = lm,
  #  formula = 'y ~ x',
  #  color = 'brown',
  #  linetype = 'dashed',
  #  fullrange = TRUE,
  #  se = FALSE
  #) +
  scale_fill_continuous(name = "Number of Plots", transform = "log10") +
  #facet_wrap(~`Forest Type Group`) +
  theme_bw() +
  #theme(aspect.ratio = 1) +
  ggtitle("FVS Projected Aboveground Carbon") +
  #xlab(bquote("Measured Annual Carbon Flux" ~(`Mg C` %.% ha^-1 %.% yr^-1))) +
  ylab(bquote("Projected Carbon" ~(`Mg C` %.% ha^-1))) #+

```

### Projected Carbon Flux

```{r plot carbon flux}
all_years <- tibble(Year = min(FVS_Carbon_PRSC$Year):max(FVS_Carbon_PRSC$Year))
num_years <- all_years |> nrow()

FVS_Carbon_PRSC |>
  group_by(StandID) |>
  arrange(Year) |>
  mutate(
    C_Flux = Aboveground_Total_Live - lag(Aboveground_Total_Live),
    Prev_Year = lag(Year)
  ) |>
  ungroup() |>
  mutate(
    Annual_C_Flux = C_Flux / (Year - Prev_Year)
  ) |>
  group_by(StandID) |>
  mutate(Min_Year = min(Year), Max_Year = max(Year)) |>
  ungroup() |>
  full_join(all_years, by = join_by(Year)) |>
  arrange(Year) |>
  fill(Min_Year, Max_Year, .direction = "updown") |>
  filter(Year >= Min_Year & Year <= Max_Year) |>
  fill(StandID, Annual_C_Flux, .direction = "up") |>
  ggplot(
    aes(x = Year, y = Annual_C_Flux)
  ) +
  geom_bin2d(binwidth = c(1, 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  #geom_smooth(
  #  method = lm,
  #  formula = 'y ~ x',
  #  color = 'brown',
  #  linetype = 'dashed',
  #  fullrange = TRUE,
  #  se = FALSE
  #) +
  scale_fill_continuous(name = "Number of Plots", transform = "log10") +
  #facet_wrap(~`Forest Type Group`) +
  theme_bw() +
  #theme(aspect.ratio = 1) +
  ggtitle("FVS Projected Annual Carbon Flux for Harvested Plots") +
  #xlab(bquote("Measured Annual Carbon Flux" ~(`Mg C` %.% ha^-1 %.% yr^-1))) +
  ylab(bquote("Projected Carbon Flux" ~(`Mg C` %.% ha^-1 %.% yr^-1))) #+;

```
