---
title: "Nunery Keeton Revisited"
author: "Nikolaus Bates-Haus"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Pre-load libraries that issue warnings on startup so these
# warnings don't corrupt the text later.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(reshape2)) # for melt()
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(dbplyr))
suppressPackageStartupMessages(library(measurements))
suppressPackageStartupMessages(library(maps))
```

```{r source-functions}
source('R/functions.R')
```

```{r plot_forested}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')
fia_cond <- tbl(fia, 'COND') |>
  select(
    STATECD, COUNTYCD, PLOT, PLT_CN, CONDID, INVYR,
    COND_STATUS_CD, BALIVE, DSTRBCD1, DSTRBCD2, DSTRBCD3,
    TRTCD1, TRTYR1, TRTCD2, TRTYR2, TRTCD3, TRTYR3
  )
fia_plot <- tbl(fia, 'PLOT') |>
  # Narrow and rename columns to facilitate join
  select(CN, DESIGNCD, SRV_CN, MEASYEAR, ECOSUBCD) |>
  rename(PLT_CN = CN)
fia_survey <- tbl(fia, 'SURVEY') |>
  # Narrow and rename columns to facilitate join
  select(CN, RSCD) |>
  rename(SRV_CN = CN)

plot_forested <- fia_cond |>
  left_join(fia_plot, by = join_by(PLT_CN)) |>
  left_join(fia_survey, by = join_by(SRV_CN)) |>
  northeastern_plots_filter() |>
  modern_plots_filter() |>
  long_measurement_filter() |>
  forested_plots_filter() |>
  undisturbed_plots_filter() |>
  harvested_plots_filter() |>
  no_unnatural_regen_filter() |> 
  measured_pre_post_harvest_filter() |>
  single_condition_plots_filter() |>
  distinct(STATECD, COUNTYCD, PLOT) |>
  collect()

dbDisconnect(fia)
remove(fia, fia_cond, fia_plot, fia_survey)

write_rds(plot_forested, '03_outputs/plot_grow_only.rds')

nrow(plot_forested)
```

Only 204 plots; fewer than expected!

Where are they?

```{r plot_location}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

# Survey has the research station code
fia_survey <- tbl(fia, 'SURVEY') |>
  select(CN, RSCD) |>
  rename(SRV_CN = CN)

# plot has invyr, measyr, lat and lon
fia_plot <- tbl(fia, 'PLOT') |>
  # Narrow and rename columns to facilitate join
  select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, SRV_CN, ECOSUBCD, LAT, LON)

# research station has state name and abbreviation
fia_ref_research_station <- tbl(fia, 'REF_RESEARCH_STATION') |>
  select(STATECD, STATE_NAME, STATE_ABBR)

plot_location <- fia_plot |>
  semi_join(plot_forested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  left_join(fia_survey, by = join_by(SRV_CN)) |>
  left_join(fia_ref_research_station, by = join_by(STATECD)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  # Pick just the first inventory year
  filter(min_rank(INVYR) == 1) |> 
  ungroup() |>
  collect() |>
  rename(
    long = LON,
    lat = LAT
  ) |>
  mutate(group = sprintf('%02d%03d%05d', STATECD, COUNTYCD, PLOT))

dbDisconnect(fia)
remove(fia, fia_plot, fia_survey, fia_ref_research_station)
```

```{r map_by_location}
northeastern_states <- plot_location |>
  distinct(STATE_NAME) |>
  mutate(region = str_to_lower(STATE_NAME))
map_data('state', northeastern_states$region) |> 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + 
  geom_point(data = plot_location, color = alpha("blue", alpha = 0.5)) +
  coord_quickmap()
remove(northeastern_states)
```

```{r stand_stats}
fia <- DBI::dbConnect(RSQLite::SQLite(), 'data/raw/SQLite_FIADB_ENTIRE.db')

measyear <- tbl(fia, 'PLOT') |>
  select(STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR, DESIGNCD, ECOSUBCD)

forest_type <- tbl(fia, 'REF_FOREST_TYPE') |>
  select(VALUE, MEANING) |>
  rename(FORTYPCD = VALUE) |>
  rename(FORTYPE = MEANING)

tree_stats <- tbl(fia, 'TREE') |>
  select(STATECD, COUNTYCD, PLOT, CONDID, INVYR, DIA, CARBON_AG, TPA_UNADJ) |>
  group_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR) |>
  summarize(
    CARBON_AG = sum(CARBON_AG, na.rm = TRUE),
    CPA = sum(CARBON_AG * TPA_UNADJ, na.rm = TRUE),
    BA_TREES = sum(if_else(DIA >= 1, TPA_UNADJ, 0), na.rm = TRUE),
    .groups = "keep"
  )

stand_stats <- tbl(fia, 'COND') |>
  select(
    STATECD, COUNTYCD, PLOT, INVYR, STDAGE, BALIVE, FORTYPCD,
    TRTCD1, TRTYR1, TRTCD2, TRTYR2, TRTCD3, TRTYR3
  ) |>
  semi_join(plot_forested, by = join_by(STATECD, COUNTYCD, PLOT), copy = TRUE) |>
  left_join(measyear, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  measured_pre_post_harvest_filter() |> # Adds interesting fields
  group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
  summarize(
    BALIVE = sum(BALIVE, na.rm = TRUE),
    FORTYPCD = max(FORTYPCD, na.rm = TRUE),
    STDAGE = max(STDAGE, na.rm = TRUE),
    MIN_MEASYEAR = min(MIN_MEASYEAR, na.rm = TRUE),
    MAX_MEASYEAR = max(MAX_MEASYEAR, na.rm = TRUE),
    MIN_HRVYR = min(MIN_HRVYR, na.rm = TRUE),
    MAX_HRVYR = max(MAX_HRVYR, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  left_join(measyear, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(tree_stats, by = join_by(STATECD, COUNTYCD, PLOT, INVYR)) |>
  left_join(forest_type, by = join_by(FORTYPCD)) |>
  modern_plots_filter() |>
  rename(`Forest Type` = FORTYPE) |>
  mutate(FORTYPCD = floor(FORTYPCD / 10) * 10) |>
  left_join(forest_type, by = join_by(FORTYPCD)) |>
  rename(`Forest Type Group` = FORTYPE) |>
  collect() |>
  mutate(
    `Forest Type Group` = str_replace(`Forest Type Group`, ' group', ''),
    BALIVE_METRIC = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    QMD = sqrt(BALIVE / (BA_TREES * (pi / 576))),
    QMD_METRIC = sqrt(BALIVE_METRIC / (BA_TREES * (pi / 40000))),
    CARBON_METRIC = conv_multiunit(CPA, "lbs / acre", "Mg / hectare")
  ) |> 
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(
    BALIVE_START = if_else(MEASYEAR == min(MEASYEAR, na.rm = TRUE), BALIVE_METRIC, NA),
    BALIVE_DELTA = BALIVE_METRIC - max(BALIVE_START, na.rm = TRUE),
    YEARS = MEASYEAR - min(MEASYEAR, na.rm = TRUE)
  ) |> 
  ungroup()

dbDisconnect(fia)
remove(fia, measyear, forest_type, tree_stats)
```

> TODO nik: in FVS_TreeInit_Plot, column PRESCRIPTION FLOAT is use for ThinPRSC.
> Should be able to identify in which year a tree is harvested and put the year
> in PRESCRIPTION, then run with a ThinPRSC to remove the trees prescribed for
> each year.
> Will need to find the database and modify it in-place.
> Good news: TREE_CN is TREE.CN (and, presumably, PLOT_CN is PLOT.CN, etc)
> Can a tree be indirectly harvested, e.g. TPA_UNADJ is reduced?

Find the trees that were removed.

```{r removed_trees}

```
