---
title: "11_CFIGrowOnly"
output: html_document
---

# Setup

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(targets)
  library(tarchetypes)
  library(tidyverse)
  library(readxl)
  library(terra)
  library(measurements)
})
tar_source()
```

## Load targets results from other documents

```{r load tar objects}
tar_load(fiadb)
tar_load(nrshrv_plot)
tar_load(nrshrv_plot_stats)
tar_load(species_crosswalk)
tar_load(fvs_num_partitions)
tar_load(fvs_partition)
tar_load(fvs_randseed)
```

## Preload targets results from this document

When tar_make() has already built the objects defined in this document,
run this block to pre-load them. You can then skip running ```{targets} blocks.

```{r load tar objects defined here, eval = FALSE}
get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```


# Stands

Use the raw data from Department of Conservation & Recreation (DCR) Office of Watershed Management, Division of Water Supply Protection (DWSP)'s Continuous Forest Inventory (CFI).

We need stands (plots), visits (inventories), disturbances, and the lookup
table for disturbance codes.

In addition, (A. Barker Plotkin, et al. 2025) did a thorough analysis of the CFI data,
identifying 283 plots that were consistently surveyed and not relocated
from 1970 - 2020. We follow ABP where possible for this analysis.

```{targets cfiabp_trees}
tarchetypes::tar_file_read(
  cfiabp_trees,
  "data/raw/DWSP_CFI/trees283-6-defpropasp.csv",
  read_csv(!!.x)
)
```

```{targets tblDWSPCFIPlotsComplete}
tarchetypes::tar_file_read(
  tblDWSPCFIPlotsComplete,
  "data/raw/DWSP_CFI/tblDWSPCFIPlotsComplete.xlsx",
  read_excel(!!.x)
)
```

```{targets tblSuppPlotDisturbanceCodes}
tarchetypes::tar_file_read(
  tblSuppPlotDisturbanceCodes,
  "data/raw/DWSP_CFI/tblSuppPlotDisturbanceCodes.xlsx",
  read_excel(!!.x)
)
```

```{targets tblDWSPCFIPlotVisitDisturbances}
tarchetypes::tar_file_read(
  tblDWSPCFIPlotVisitDisturbances,
  "data/raw/DWSP_CFI/tblDWSPCFIPlotVisitDisturbances.xlsx",
  read_excel(!!.x) |>
    # Fix column name
    rename(
      PlotVisitDisturbanceYearCorrected = PlotVisitDisturbandYearCorrected,
      MasterPlotVisitID = PlotVisitID
    ) |>
    left_join(
      tblSuppPlotDisturbanceCodes |>
        select(DisturbanceCode, DisturbanceDesc),
      by = join_by(PlotVisitDisturbanceCode == DisturbanceCode)
    ) |>
    mutate(
      DisturbanceDesc = as.factor(DisturbanceDesc)
    )
)
```

```{targets tblDWSPCFIPlotVisitsComplete}
tarchetypes::tar_file_read(
  tblDWSPCFIPlotVisitsComplete,
  "data/raw/DWSP_CFI/tblDWSPCFIPlotVisitsComplete.xlsx",
  read_excel(!!.x)
)
```

```{targets tblSuppTreeStatusCodes}
tarchetypes::tar_file_read(
  tblSuppTreeStatusCodes,
  "data/raw/DWSP_CFI/tblSuppTreeStatusCodes.xlsx",
  read_excel(
    !!.x,
    col_types = c(
      "numeric", "text", "numeric",
      "numeric", "numeric", "numeric",
      "numeric", "numeric", "numeric",
      "numeric", "numeric", "numeric",
      "numeric", "numeric", "numeric",
      "numeric", "numeric", "numeric",
      "numeric", "numeric"
    )
  )
)

```

```{targets qryDWSPCFIPlotVisitTreeDetail}
tarchetypes::tar_file_read(
  qryDWSPCFIPlotVisitTreeDetail,
  "data/raw/DWSP_CFI/qryDWSPCFIPlotVisitTreeDetail_Partial_HF.xlsx",
  read_excel(!!.x) |>
    # Fix column names
    rename(
      MasterTreeID = VisitMasterTreeID,
      MasterPlotVisitID = VisitIDNumberDetail
    ) |>
    # Fix column types - giving read_excel the right types causes it to hang
    mutate(
      MasterTreeID = as.numeric(MasterTreeID),
      VisitTreeNumberDetail = as.numeric(VisitTreeNumberDetail)
    ) |>
    left_join(
      tblSuppTreeStatusCodes |>
        select(StatusCode, TreeStatusDesc) |>
        rename(VisitTreeStatusCode = StatusCode),
      by = join_by(VisitTreeStatusCode)
    )
)
```

```{targets tblDWSPCFITreesComplete}
tarchetypes::tar_file_read(
  tblDWSPCFITreesComplete,
  "data/raw/DWSP_CFI/tblDWSPCFITreesComplete.xlsx",
  read_excel(!!.x)
)
```

## Elevation

LiDAR elevation maps are available from https://www.mass.gov/info-details/massgis-data-lidar-terrain-data#overview
These are in ERDAS IMAGINE (IMG) format.

Rather than download all tiles we can fetch the 31 tile index and determine
which tiles the plots are in.

```{r MassGIS DEM Tiles, eval=FALSE}
massgis_dem_tile_index <- vect(
  "data/raw/MassGIS/LIDARINDEX_DEMMOSTILES_POLY/LIDARINDEX_DEMMOSTILES_POLY.shp"
)
tblDWSPCFIPlotsComplete |>
  # Remove rows without lat/lon
  filter(!is.na(GPSLatitude) & !is.na(GPSLongitude)) |>
  # select an arbitrary row with lat/lon for each plot
  # we assume that lat/lon is consistent over time
  group_by(MasterPlotID) |>
  filter(row_number() == 1) |>
  ungroup() |>
  # Reshape the data for terra::vect()
  select(MasterPlotID, GPSLatitude, GPSLongitude) |>
  rename(
    lat = GPSLatitude,
    lon = GPSLongitude
  ) |>
  vect(crs = "EPSG:4326") |>
  # Re-project to match the tile index
  project(massgis_dem_tile_index) |>
  intersect(massgis_dem_tile_index) |>
  as.data.frame() |>
  distinct(TILE) |>
  arrange(TILE)
```

So, we need tiles 4, 12, 13 and 14.
Download those and unpack into data/raw/MassGIS.

Use terra::extract() to extract the elevation in feet for each plot from
the appropriate tile.

```{targets cfigro_plot_elevft, tar_simple = TRUE}
# Files for tile NN are named
# MassGIS/LidarElevation_YYYYtoZZZZ_NN/Lidar_Elevation_YYYYtoZZZZ_NN.img
# We'll keep this as a list and extract() each tile separately - this is
# much, MUCH faster than merging tiles with terra::merge() or terra::m
massgis_dem_tiles <- lapply(
  list.files("data/raw/MassGIS", pattern = "Lidar_Elevation_*"),
  \(tile) { rast(file.path("data/raw/MassGIS", tile, paste0(tile, ".img"))) }
)

tmp_plot_location <- tblDWSPCFIPlotsComplete |>
  # Remove rows without lat/lon
  filter(!is.na(GPSLatitude) & !is.na(GPSLongitude)) |>
  # select an arbitrary row with lat/lon for each plot
  # we assume that lat/lon is consistent over time
  group_by(MasterPlotID) |>
  filter(row_number() == 1) |>
  ungroup() |>
  # Reshape the data for terra::vect()
  select(MasterPlotID, GPSLatitude, GPSLongitude) |>
  rename(
    lat = GPSLatitude,
    lon = GPSLongitude
  ) |>
  vect(crs = "EPSG:4326") |>
  # Match the projection of the first tile
  project(massgis_dem_tiles[[1]])

# Build a dataframe with two columns:
# MasterPlotID, ELEVFT
# with one row per MasterPlotID
as.data.frame(tmp_plot_location) |>
  mutate(
    # Call extract() on each available tile; for plots that overlap
    # with that tile, we'll get back elevation, otherwise NA.
    # Coalesce across tiles to get the non-NA elevation for
    # each plot.
    ELEVFT = do.call(
      coalesce,
      lapply(massgis_dem_tiles, \(x) extract(x, tmp_plot_location)$FEET)
    )
  )
```

## County

MA county data is available from https://www.mass.gov/info-details/massgis-data-counties

Download those and unpack into data/raw/MassGIS/counties.

Use terra::intersect() to extract the county for each plot.


```{targets cfigro_plot_county, tar_simple = TRUE}
massgis_county <- vect("data/raw/MassGIS/counties/COUNTIES_POLY.shp")

tblDWSPCFIPlotsComplete |>
  # Remove rows without lat/lon
  filter(!is.na(GPSLatitude) & !is.na(GPSLongitude)) |>
  # select an arbitrary row with lat/lon for each plot
  # we assume that lat/lon is consistent over time
  group_by(MasterPlotID) |>
  filter(row_number() == 1) |>
  ungroup() |>
  # Reshape the data for terra::vect()
  select(MasterPlotID, GPSLatitude, GPSLongitude) |>
  rename(
    lat = GPSLatitude,
    lon = GPSLongitude
  ) |>
  vect(crs = "EPSG:4326") |>
  # Match the projection of the first tile
  project(massgis_county) |>
  intersect(massgis_county) |>
  as.data.frame() |>
  select(MasterPlotID, COUNTY)
```
## Watershed

The DCR Division of Water Supply Protection (DWSP) maintains CFI plots
in the various reservoir watersheds as a means to manage water quality.

The plots in the Quabbin watershed (Watershed == 1) have a particularly
long history of high-quality data collection and responsible harvest, so
we will focus on just this watershed.

ABP plots are filtered to the Quabbin watershed.

## Relocation

Many plots were relocated prior to 1970, while retaining the same
MasterPlotID. To avoid analysis across this disjunction in trees,
we restrict analysis to 1970 and after.

ABP trees are filtered to non-relocated plot visits.

## Harvest

We would like to follow the FIA definition of harvest:

> 2.5.44 TRTCD1
> ... The area affected by the treatment must be at least 1 acre in size.

Sadly, we don't have access to the area affected.

For each tree, however, we do have access to information on whether it was cut.

FIA also defines harvest as:

> Appendix H: Damage Agent Codes and Thresholds
> 71000, Harvest, Removal of 10% cubic volume.

We likewise do not have volume, but can approximate this.

### FIA Observed

Question: is "Harvest == removal of >= 10% of biomass" consistent with what
is recorded in FIA? We have noted previously that FIA may mark a plot as
harvested when no trees are cut, and may mark a plot as not harvested when
many of its trees are cut, so cutting is an imperfect predictor of fia harvest.

FIA doesn't record biomass, but does record carbon, which is a good proxy.

```{r fia havest analysis, eval=FALSE}
# Sum the carbon of trees that were live pre-harvest
pre_harvest_carbon <- fia_trees(
    fiadb,
    nrshrv_plot_stats |> filter(PRE_HARVEST == 1)
  ) |>
  filter(STATUSCD == 1) |>
  filter_add_stand_id() |>
  select(STAND_ID, CARBON_AG, DIA, TPA_UNADJ) |>
  group_by(STAND_ID) |>
  summarize(
    CARBON_AG = sum(CARBON_AG, na.rm = TRUE),
    BALIVE_AG = sum(pi * (DIA / 2)^2, na.rm = TRUE)
  ) |>
  ungroup()

# Harvested trees don't get biomass, so we need to find their
# previous selves and get the biomass from there.
harvested_carbon <- fia_trees_by_cn(
    fiadb,
    fia_trees_filtered(
      fiadb,
      nrshrv_plot_stats |> filter(HARVEST == 1),
      \(.data, con) {
        .data |>
          filter(STATUSCD == 3) |>
          select(PREV_TRE_CN) |>
          rename(CN = PREV_TRE_CN)
      }
    )
  ) |>
  filter_add_stand_id() |>
  select(STAND_ID, CARBON_AG, DIA, TPA_UNADJ) |>
  group_by(STAND_ID) |>
  summarize(
    CARBON_HRV = sum(CARBON_AG, na.rm = TRUE),
    BALIVE_HRV = sum(pi * (DIA / 2)^2, na.rm = TRUE)
  ) |>
  ungroup()

fia_harvest_intensity <- pre_harvest_carbon |>
  inner_join(harvested_carbon, by = join_by(STAND_ID)) |>
  mutate(
    CARBON_PCT = 100 * CARBON_HRV / CARBON_AG,
    BALIVE_PCT = 100 * BALIVE_HRV / BALIVE_AG
  )

fia_harvest_intensity |>
  ggplot(aes(CARBON_PCT)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = 10) +
  theme_bw() +
  ggtitle("Carbon Removal % in FIA harvested plots")

fia_harvest_intensity |>
  ggplot(aes(BALIVE_PCT)) +
  geom_histogram(binwidth = 5) +
  geom_vline(xintercept = 10) +
  theme_bw() +
  ggtitle("Basal Area Removal % in FIA harvested plots")

```

So, the vast majority of harvest is >= 10% of carbon. The distribution of
removal of basal area is about the same, so we'll use %BA removed as a proxy.

### DWSP CFI Tree Status

Let's look at individual tree status to help understand plot status:

```{r plot cfi tree status code, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(TreeStatusDesc = ordered(
    TreeStatusDesc,
    levels = c("Live tree >=5.6\"", "standing dead", "fallen dead", "Out of population")
  )) |>
  ggplot(aes(VisitCycle, fill = fct_rev(TreeStatusDesc))) +
  geom_bar(position = "stack") +
  scale_fill_viridis_d(name = "Tree Status") +
  theme_bw() +
  ggtitle("DWSP CFI Tree Status by Visit Cycle") +
  ylab("Number of Trees") +
  xlab("Visit Cycle")
```

So TreeStatusDesc uses codes 1-4, back-propagated to prior visit cycles.

### ABP 2025 Tree Status

```{r plot abp tree status code, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    StatusB = if_else(is.na(StatusB), "X", StatusB),
    StatusDesc = case_match(
      StatusB,
      "L" ~ "Live",
      "R" ~ "Recruit",
      "D" ~ "Dead",
      "C" ~ "Cut",
      "X" ~ "Out of Population"
    ),
    StatusDesc = ordered(
      StatusDesc,
      levels = c("Recruit", "Live", "Dead", "Cut", "Out of Population")
    )
  ) |>
  ggplot(aes(VisitCycle, fill = fct_rev(StatusDesc))) +
  geom_bar(position = "stack") +
  scale_fill_viridis_d(name = "Tree Status") +
  theme_bw() +
  ggtitle("ABP 2025 Tree Status by Visit Cycle") +
  ylab("Number of Trees") +
  xlab("Visit Cycle")
```

### FIA vs. CFI Harvest

CFI defines CutAnyTime, which presumably includes pre-1970 harvest.
CutAnyTime seems to mean CutAnyTimePreviously, since individual plots
have multiple values.

We compute two harvest indicators:
- FIA Harvest - is more than 10% of the BA in "Cut" trees
- CFI Harvest - is CutSinceLastVisit ever TRUE

```{targets cfigro_plot_harvested, tar_simple = TRUE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_abp(cfiabp_trees) |>
  cfi_harvested()
```


```{r plot cfigro_plot_harvested, eval = FALSE}
cfigro_plot_harvested |>
  mutate(
    HarvestType = case_when(
      FIAHarvested & CFIHarvested ~ "Both",
      FIAHarvested ~ "FIA",
      CFIHarvested ~ "CFI",
      .default = "Neither"
    )
  ) |>
  ggplot(aes(HarvestType)) +
  geom_bar() +
  geom_text(aes(y = after_stat(count + 5), label = after_stat(count)), stat = "count") +
  theme_bw() +
  ggtitle("FIA vs CFI Plot Harvest Status") +
  ylab("Number of Plots")
```

Defined in cfi_harvested()

## Disturbance

We would like to follow the FIA definition of disturbance:

> 2.5.38 DSTRBCD1
>
> The area affected by the disturbance must be at least 1 acre in size. A
> significant level of disturbance (mortality or damage to 25 percent of the trees
> in the condition) is required.

Sadly, we don't have access to the area affected, or to damage separate from
mortality.

CFI does provide a disturbance flag, and starting in 2010 a disturbance
intensity.

Disturbance Severity code:

- 1 - Light
- 2 - Moderate
- 3 - Severe

Ignoring light disturbance seems in-line with FIA definition, so we will
flag plots with Moderate to Severe disturbance.

Notable disturbance codes:

- 0: None
- 10: Harvest Cut

It looks like disturbances > 10 are not currently used. We also wish to
distinguish between harvest and other disturbance, so we won't mark things
with distrubance code 10 as disturbed.

Note that we need PlotVisits to get from Disturbance to Plots.

### Harvest vs. Non-Harvest Disturbance Types

```{r disturbance types, eval = FALSE}
tblSuppPlotDisturbanceCodes |>
  mutate(DisturbanceCategory = case_when(
    DisturbanceCode == 0 ~ "None",
    DisturbanceCode %in% c(8, 9, 11, 12, 13, 14, 17, 18, 19, 20, 21, 22) ~
      "Harvest Related",
    .default = "Non-Harvest"
    )
  ) |>
  select(DisturbanceCode, DisturbanceCategory, DisturbanceDesc)
```

### Disturbance History

```{r plot disturbance type, eval = FALSE}
tblDWSPCFIPlotVisitDisturbances |>
  left_join(
    tblDWSPCFIPlotVisitsComplete |>
      distinct(MasterPlotVisitID, MasterPlotID, VisitYear, VisitCycle),
    by = join_by(MasterPlotVisitID)
  ) |>
  # Remove harvest and harvest-related disturbances
  filter(!PlotVisitDisturbanceCode %in% c(0, 8, 9, 11, 12, 13, 14, 17, 18, 19, 20, 21, 22)) |>
  filter(
    PlotVisitDisturbanceYearCorrected >= VisitYear - 10 &
      PlotVisitDisturbanceYearCorrected < 9999 &
      (is.na(PlotVisitDisturbanceSeverity) | PlotVisitDisturbanceSeverity > 1)
  ) |>
  inner_join(
    cfiabp_trees |> distinct(MasterPlotID, VisitCycle),
    by = join_by(MasterPlotID, VisitCycle)
  ) |>
  mutate(DisturbanceDesc = factor(DisturbanceDesc)) |>
  ggplot(aes(PlotVisitDisturbanceYearCorrected, fill = DisturbanceDesc)) +
  geom_bar() +
  ggtitle("CFI Plot Disturbance Type by Year") +
  xlab("Year of Disturbance") +
  ylab("Number of Plots") +
  theme_bw()
```

### CFI Plot Disturbance

```{targets cfigro_plot_disturbed, tar_simple = TRUE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_abp(cfiabp_trees) |>
  cfi_disturbed(tblDWSPCFIPlotVisitDisturbances)
```

```{r plot cfigro_plot_disturbed, eval = FALSE}
cfigro_plot_disturbed |>
  ggplot(aes(Disturbed)) +
  geom_bar()
```

Cool, cool. So, how many plots are unharvested and undisturbed?

```{r plot plot status, eval = FALSE}
cfiabp_trees |>
  distinct(MasterPlotID) |>
  left_join(cfigro_plot_harvested, by = join_by(MasterPlotID)) |>
  left_join(cfigro_plot_disturbed, by = join_by(MasterPlotID)) |>
  mutate(
    Status = case_when(
      FIAHarvested & Disturbed ~ "Harvested & Disturbed",
      FIAHarvested ~ "Harvested",
      Disturbed ~ "Disturbed",
      .default = "Neither"
    )
  ) |>
  ggplot(aes(Status)) +
  geom_bar() +
  geom_text(aes(y = after_stat(count + 5), label = after_stat(count)), stat = "count") +
  theme_bw() +
  ggtitle("CFI Plots by Status") +
  ylab("Number of Plots")
```


## Final Formatting

put it in the right shape for FVS.
When FVS processes a stand from FIA data, it says things like
(omitting WAS IGNORED columns):

```
SELECT * FROM FVS_StandInit_Plot WHERE Stand_CN = '%Stand_CN%'

STAND-LEVEL DATA BASE READ:
  INV_YEAR:               2006 - VisitYear 
  REGION:                    9 - Keep it 9
  FOREST:                   20 - Nearest national forest / ranger district, 20 is Green Mountains
   COMPOSITE LOC:          920 - - See FVSne_Overview 3.1 Location Codes
  LOCATION:                920 - - See FVSne_Overview 3.1 Location Codes
  LONGITUDE:          -73.3379 - GPSLatitude
  LATITUDE:            41.2770 - GPSLongitude
  AGE:                      74 - ??? No Age
  ASPECT:                  6.0 - Aspect
  SLOPE:                  5.00 - Slope; CFI is probably %; FVS wants percent
  ELEVFT:                520.0 CONVERTED TO:    5.2 - join from MassGIS data
  BASAL_AREA_FACTOR:     -24.0 - plot radius for fixed-radius plot; 52.7
  INV_PLOT_SIZE:          300. - plot radius for fixed-size microplot; 6
  BRK_DBH:                 5.0 - Sapling / Tree DBH break; 6 in Audrey's data, 5 in CFI
  NUM_PLOTS:                 4 - Number of subplots; 1
  SAM_WT:          6321.345703 - Sample Weight (acres) - not used by FVS
  PHYSIO_REGION:            23 - maybe from fiadb 2.3.35 Physiographic Class; use forest type instead
  STATE:                     9 - State - fill in properly - 25
  COUNTY:                    1 - County - fill in properly - use MassGIS counties
END OF DATA BASE READ.
```

From the manual:
- Plots are 52.7' fixed-radius plots, which is 0.2 acre (0.2003011)
- Four, six-foot radius subplots are taken on each CFI pot along the cardinal directions, 26' from plot center.
- Size classes:
  - 1 - less than one foot in height
  - 2 - 1' to 4.5' in height
  - 3 - 4.5' to 1.0" DBH
  - 4 - 1" to 5" dbh
- Seedlings less than one foot in height (class 1) and thos 1' to 4.5' in height
  (class 2) are counted individually up to ten and in groups of ten (11 to 20 = 20, etc.)
  when their number exceeds ten.
- For the other sizes [sic] classes (3 and 4), individual stems are recorded.

Site index is recorded.

It looks like forest type / ecoregion will be used if present

ForestTypeDesc might be available - see appendix.

STAND_CN - MasterPlotVisitID
STAND_ID - MasterPlotID
VARIANT - NE
INV_YEAR / INV_MONTH / INV_DAY - VisitMeasuredDate
LATITUDE / LONGITUDE - GPSLatitude / GPSLongitude
DATUM - NA
REGION - 20
FOREST - 920
DISTRICT - ?
ECOREGION - can maybe get this
AGE - ?
ASPECT - ASPECT
SLOPE - SLOPE
ELEVATION - ?
ELEVFT - ELEVATION
INV_PLOT_SIZE - -52.7
BRK_DBH - 5
NUM_PLOTS - 1
SITE_SPECIES - ?
SITE_INDEX - ?
SITE_INDEX_BASE_AG - ?
FOREST_TYPE_FIA - ?
STATE - 25
COUNTY - COUNTY


# Trees


We have a couple more tables to look at to see if we can find everything:


```
SELECT * FROM FVS_TreeInit_Plot WHERE Stand_CN = '%Stand_CN%'
  PLOT_ID            WAS USED - MasterPlotID
  TREE_ID            WAS USED - MasterTreeID
  TREE_COUNT         WAS USED - stemsha.L6
  HISTORY            WAS USED - status.prior6
  SPECIES            WAS USED - genusp
  DIAMETER           WAS USED - dbh.cm
  HT                 WAS USED - :-(
  HTTOPK             WAS USED - :-(
  CRRATIO            WAS USED - :-( - maybe in supplement; see appendix
  DAMAGE1            WAS USED - X
  SEVERITY1          WAS USED - X
  DAMAGE2            WAS USED - X
  SEVERITY2          WAS USED - X
  DAMAGE3            WAS USED - X
  SEVERITY3          WAS USED - X
  TREEVALUE          WAS USED - X - tree removal priority
  PRESCRIPTION       WAS USED - X - tree removal group
  AGE                WAS USED - X
  SLOPE              WAS USED - Slope
  ASPECT             WAS USED - Aspect
  PV_CODE            WAS USED - ecoregion? - not used
  TOPOCODE           WAS USED - ?
  SITEPREP           WAS USED - ?
  DG                 WAS USED - Diameter growth - dbh.prior.cm
  HTG                WAS USED - height growth - X
NUMBER ROWS PROCESSED:   30
```
