---
title: "12_CFICOP"
output: html_document
---

# Setup

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(targets)
  library(tarchetypes)
  library(tidyverse)
  library(readxl)
  library(terra)
  library(sf)
  library(rnaturalearth)
  library(measurements)
  library(mlr3verse) # Machine Learning in R framework, larger universe
  library(ggspatial) # annotation_scale, annotation_north_arrow
  library(ggfx) # for fancy ggplot stuff like with_blur
  library(patchwork) # for n-up ggplot
})
tar_source()

## Load targets results from other documents ----

tar_load(fiadb)
tar_load(species_crosswalk)
tar_load(fvs_num_partitions)
tar_load(fvs_partition)
tar_load(fvs_randseed)
tar_load(fvsne_states)
tar_load(cfigro_bai_model)
tar_load(cfigro_estab)
tar_load(cfigro_none_growth)

## Preload everything from 11_CFIGrowOnly.Rmd ----

get_this_rmd_file() |>
  str_replace(basename(get_this_rmd_file()), "11_CFIGrowOnly.Rmd") |>
  tar_objects_defined_in_rmd() |>
  tar_load()

## Preload targets results from this document ----

# When tar_make() has already built the objects defined in this document,
# run this block to pre-load them. You can then skip running ```{targets}
# blocks.

get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()

## Eve's Colors ----

#colors1 <- c("indianred3", "chocolate2", "goldenrod2", "yellow3", "olivedrab")
eves_colors <- c(
  "Other" = "#B4B432",
  "Fraxinus" = "#CD5C5C",
  "Betula" = "#8B4513",
  "Tsuga" = "#5F9EA0",
  "Acer" = "#DEB887",
  "Pinus" = "#556B2F",
  "Quercus" = "#D2691E"
)
#colors3 <- c("darkorange4", "goldenrod3", "darkolivegreen")

## My Colors ----
series_colors <- RColorBrewer::brewer.pal(5, "Set1")
```

# MODEL ALL\* THE PLOTS!!!

We want to: - Model all the plots grow-only - Apply the model of
residuals to get an idea of how much FVS might have over-projected -
Look at harvest, and how much carbon is actually on the landscape -
Multiply by area to get an idea of what COP would be, etc.

-   We don't have harvest information except on Audrey's plots, so we'll
    actually just use her 283 plots. So it's not *all* of all the plots,
    but it's all the plots for which we have good data.

## Stands

From Barker Plotkin 2025:

> The study area encompasses the ca. 23,000 ha Quabbin Reservoir
> Watershed Forest (‘Quabbin’) in central Massachusetts, USA.

> The land area of the Quabbin is 88% forested. Oaks are part of nearly
> all forest types, and oak-dominated forest types comprise 39% of the
> forest.

> The spatial and temporal scale of our study allows a detailed
> examination of trends in tree demography throughout this heterogeneous
> landscape. This contrasts with prior studies that have relied on the
> spatially extensive but low sampling density of the USDA Forest
> Service’s Forest Inventory and Analysis (FIA) plot network (Fei et
> al., 2011, Morin and Liebhold, 2016). The study area is monitored
> using Continuous Forest Inventory (CFI) plots, with one plot per
> ∼80 ha, in contrast to one plot per ∼2400 ha for FIA. These plots have
> been sampled once per decade by the state of Massachusetts Department
> of Conservation and Recreation since 1960. Oak-dominated forest types
> comprise about 40 % of the forest in this landscape, with northern red
> oak (Q. rubra), black oak (Q. velutina), and white oak (Q. alba) the
> main oak species (Division of Water Supply Protection, 2018).

FIA plot density is one plot per 6,000 acres

A base hexagon positioned over the conterminous United States
was subdivided into approximately 28,000
hexagons whose centers are about 16.9 mi (27
km) apart (White et d. 1992). One field plot was
selected for each hexagon, usually the existing
FIA plot closest to the center of the hexagon.
Each of the hexagons was assigned to one of
four panels; a panel corresponds to a given
measurement year of the cycle. After the fourth
panel is measured, the cycle is repeated. One of
the advantages of this framework is that it is
unlikely to be aligned with regularly spaced
landscape features.

Because of these desirable features, we explored the possibility of
using the FHM framework as the basis for the FLA annual inventory
sampling framework. To meet its mandated
maximum sampling errors, the FIA program
requires a sampling intensity of one plot per
approximately 6,000 acres (M. H. Hansen
1998, pers. cornrn.). By creating a new lattice of
hexagonal cells where each hexagon is 1 /27
the size of an FHM hexagon, the desired sampling intensity
is achieved (A. R. Olsen 1998,
pers. comm.). The size of each FIA hexagon is
5,937.2 acres.

> Timber harvesting is part of the watershed management; silvicultural
> objectives and methods have varied over time. During the 50-year study
> period, thinning and shelterwood preparatory cuts comprised most of
> the silviculture activities from the 1960s to 1980s. In the 1990s, the
> selection system (single-tree or groups \<0.1 ha) was more common,
> with larger group selection (0.1–0.8 ha openings), shelterwood
> removals, and seed tree harvests becoming part of the mix in the
> 2000s. Clearcuts (openings \>0.8 ha) were rare throughout the study
> period. Foresters have conducted an average of 400 ha (1.7 % of the
> total area) of timber harvest each year since the 1960s (Division of
> Water Supply Protection, 2018); some areas have had multiple entries
> and others none. Tree regeneration was severely limited by deer
> browsing; a controlled deer hunting program, initiated in 1991, has
> resulted in recovery of tree regeneration (Division of Water Supply
> Protection, 2018).

> We utilized data from a network of permanent plots established in 1960
> across the Quabbin and censused each decade. Most of the plots are
> 0.08 ha (0.2 acres), with plots that are partially non-forest (e.g.
> road; permanent field) pro-rated for their forested size (e.g., a plot
> that was 25 % road would be 0.06 ha). The plots are located on a grid,
> spaced ∼800 m (0.5 mile) apart. For this analysis, we excluded plots
> that were missing one or more census measurement. Of the 401 plots in
> the complete Quabbin dataset, 283 were identified as having been
> sampled every 10 years from 1970–2020.

> Forty-two tree species were present in the 283 plots at some point
> during 1970–2020. Since oaks are the focal taxon for this study and
> the main host for spongy moth, we grouped the oak species together as
> ‘oaks’ and the remaining 37 species as ‘other.’ We acknowledge the
> silvical differences among the white oak (section Quercus) and red oak
> (section Lobatae) groups, and the varying average longevity of the
> different oak species (Loehle, 1987). Most of the oaks (53 %) were
> Quercus rubra (L.), with Q. velutina (Lam.) contributing 25 %, Q. alba
> (L.) 17 %, and the remaining 5 % Q. coccinea (Muenchh.) and Q. prinus
> (L.). Averaged across all censuses, the oaks together comprised 27 %
> of the total live tree stems in the plots. The most common other
> species were Pinus strobus (L.) (22 %), Acer rubrum (L.) (19 %), and
> Tsuga canadensis (L.) Carr. (9 %). We classified each tree as
> ‘susceptible’, ‘resistant’, or ‘immune’ to spongy moth defoliation
> based on Liebhold et al. (1995). For the six censuses, oaks comprised
> 88–93 % of all trees classified as ‘susceptible’ to spongy moth
> defoliation.

```{r}
knitr::download_image("https://ars.els-cdn.com/content/image/1-s2.0-S0378112725000155-gr3_lrg.jpg")
```

```{r}
knitr::download_image("https://ars.els-cdn.com/content/image/1-s2.0-S0378112725000155-gr4_lrg.jpg")
```

DCR Land Management Plan 2017, table 3-6 "Cover Types by Watershed",
page 40, gives breakdown for Quabbin:

```{r}
# Values are in acres
quabbin_cover_types <- c(
  "White pine" = 6518,
  "White pine / Hemlock" = 2586,
  "White pine / Hardwood" = 7901,
  "White pine / Oak" = 3770,
  "White pine / pitch pine" = 9,
  "Hemlock" = 1654,
  "Hemlock /  Hardwood" = 2922,
  "Red / white spruce" = 79,
  "Larch (tamarack)" = 5,
  "Red pine" = 1550,
  "Northern red oak" = 6907,
  "Oak / hardwood" = 8673,
  "Oak, mixed: dry site" = 7005,
  "Black birch/red maple/cherry" = 1617,
  "Poplar/grey birch" = 225,
  "Northern hardwoods" = 1973,
  "Wooded wet - deciduous" = 732,
  "Wooded wet – coniferous" = 188,
  "Wooded wet – mixed" = 418,
  "Red maple" = 1028
) |>
  as.data.frame() |>
  set_names("acres") |>
  rownames_to_column("cover_type")
# Forested total from the report
quabbin_forested_total <- 55760
stopifnot(sum(quabbin_cover_types$acres) == quabbin_forested_total)

# TODO: Map Quabbin cover types to FIA forest types, by hand
quabbin_cover_types |>
  mutate(forest_type = case_when(
    grepl("^White pine", cover_type) ~ "White pine",
    grepl("^Hemlock", cover_type) ~ "Hemlock",
    grepl("[Oo]ak|[Mm]aple|[Hh]ardwood|[Bb]irch|[Dd]eciduous", cover_type) ~ "Hardwood",
    .default = "Other"
  ))
```

## Plots

cfigro_plot has all 283 plots.

## Trees

cfigro_tree has trees for all 283 plots.

### All the plots?

We only need trees in 1970; can we use all the plots?

```{r all the plots, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
#  cfi_abp(cfiabp_trees) |>
  filter(VisitCycle == 1970) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  distinct(MasterPlotID) |>
  nrow()
```

There are 337 plots with trees in 1970; the 65 additional plots dont's
seem worth the extra effort.

For the manuscript, "Many plot locations in the Quabbin watershed were
adjusted in 1970, creating a discontinuity in observed data. This study
exampines the period from 1970 - 2020, during which plot location is
consistent. There were 337 plots with trees in 1970; of these, detailed
information on harvest and disturbance is available on 277
[Barker-Plotkin 2024]."

### Size Distribution in 1970

```{r, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  mutate(
    diameter_class = ordered(as.integer(VisitTreeDIAM / 2) * 2),
    diameter_class = fct_lump(diameter_class, n = 7, other_level = "> 16")
  ) |>
  filter(
    diameter_class >= 6 # TPA below 6" can't really be trusted
  ) |>
  group_by(MasterPlotID, diameter_class) |>
  summarize(value = n() * 5, .groups = "drop") |>
  ungroup() |>
  ggplot(aes(diameter_class, value)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 400)) +
  ggtitle("Tree Size Distribution in 1970") +
  xlab("Diameter at Breast Height (inches)") +
  ylab(bquote(trees %.% acre^-1))
```

## Regeneration

We ignored harvest in CFIGRO; can we continue to ignore it?

```{r plot establishment rate, eval = FALSE}
cfigro_estab |>
  mutate(STAND_CN = as.character(STAND_CN)) |>
  group_by(STAND_CN) |>
  summarize(DENSITY = sum(DENSITY), .groups = "keep") |>
  left_join(cfigro_plot |> select(STAND_CN, STAND_ID), by = join_by(STAND_CN)) |>
  mutate(MasterPlotID = as.numeric(STAND_ID)) |>
  left_join(cfigro_plot_harvested, by = join_by(MasterPlotID)) |>
  left_join(cfigro_plot_disturbed, by = join_by(MasterPlotID)) |>
  mutate(Status = case_when(
    CFIDisturbed & CFIHarvested ~ "Disturbed & Harvested",
    CFIDisturbed ~ "Disturbed",
    CFIHarvested ~ "Harvested",
    .default = "Grow-Only"
  )) |>
  group_by(Status) |>
  mutate(Status = paste0(Status, ", n=", n())) |>
  ggplot(aes(Status, DENSITY)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Establishment rate by plot status") +
  ylab(bquote("Establishment rate" ~ (stems %.% acre^-1 %.% year^-1)))
#  scale_y_continuous(transform = "log10")
  
```

non-Grow-Only plots show double and more the establishment rate of
grow-only plots.

Establishment comes from saplings / seedlings, which come from seeds,
which either come into the plot from outside or come from mature trees
on the plot. All of these plots are co-located, so they can trade seeds
with each other. For available seeds from trees on the plot, we can
match plots by BA of the major species, then use the establishment from
the matched / sibling plot.

### Plot Matching

Start simple - minimize the delta in the three major species. We can
leave "Other" alone.

```{targets cficop_plot_sibling, tar_simple = TRUE}
tmp_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(VisitCycle == 1970) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  # Quercus rubra and Quercus velutina are often conflated in surveys; instead
  # of treating them as two separate important species, combine them into one and
  # treat them as a single species.
  # Replace Quercus velutina (837) with Quercus rubra (833)
  mutate(
    SpeciesCode = if_else(SpeciesCode == 837, 833, SpeciesCode)
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, SCIENTIFIC_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    TPA = 5,
    BA = VisitTreeDIAM
#    BA = pi * (VisitTreeDIAM / 2)^2 * 5
  )

tmp_by_plot <- tmp_trees |>
  group_by(MasterPlotID) |>
  summarize(
    TPA_Plot = sum(TPA, na.rm = TRUE),
    BA_Plot = sum(BA, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ungroup() |>
  select(MasterPlotID, TPA_Plot, BA_Plot)

tmp_top_species <- c(
  "Acer rubrum", "Quercus rubra", "Pinus strobus"
)

tmp_plot_fingerprint <- tmp_trees |>
  group_by(MasterPlotID, SCIENTIFIC_NAME) |>
  summarize(
    TPA_Species = sum(TPA, na.rm = TRUE),
    BA_Species = sum(BA, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  left_join(tmp_by_plot, by = join_by(MasterPlotID)) |>
  mutate(
    TPA_Frac = TPA_Species / TPA_Plot,
    BA_Frac = BA_Species / BA_Plot
  ) |>
  select(-ends_with("_Species")) |>
#  filter(BA_Frac > 0.1) |>
  filter(SCIENTIFIC_NAME %in% tmp_top_species) |>
  group_by(MasterPlotID) |>
  mutate(
    Species = do.call(
      paste,
      append(sort(SCIENTIFIC_NAME), list(sep = ", "))
    )
  ) |>
  ungroup() |>
  mutate(SCIENTIFIC_NAME = gsub(" ", "_", SCIENTIFIC_NAME)) |>
  pivot_wider(names_from = "SCIENTIFIC_NAME", values_from = c("TPA_Frac", "BA_Frac"))

tmp_plots_grow_only <- cfigro_plot_harvested |>
  filter(!CFIHarvested) |>
  inner_join(
    cfigro_plot_disturbed |> filter(!CFIDisturbed),
    by = join_by(MasterPlotID)
  ) |>
  select(MasterPlotID)

pct_diff <- function(a, b) {
  abs(a - b) / ((a + b) / 2)
}

most_similar_plot <- function(target_plots) {
  target_plots <- target_plots |>
    # Penalize things that differ on whether they have
    # a species on the landscape
    replace_na(list(
      BA_Frac_Pinus_strobus = -1,
      BA_Frac_Acer_rubrum = -1,
      BA_Frac_Quercus_rubra = -1
    ))

  lapply(1:nrow(target_plots), \(n) {
    # Minimize RMS difference in BA
    ba_plot <- target_plots[[n, "BA_Plot"]]
    ba_pist <- target_plots[[n, "BA_Frac_Pinus_strobus"]]
    ba_acru <- target_plots[[n, "BA_Frac_Acer_rubrum"]]
    ba_quru <- target_plots[[n, "BA_Frac_Quercus_rubra"]]
    tmp_plot_fingerprint |>
      inner_join(tmp_plots_grow_only, by = join_by(MasterPlotID)) |>
      replace_na(list(
        BA_Frac_Pinus_strobus = -1,
        BA_Frac_Acer_rubrum = -1,
        BA_Frac_Quercus_rubra = -1
      )) |>
      mutate(
        BA_RMSE = sqrt(
          pct_diff(BA_Plot, ba_plot)^2 +
            (BA_Frac_Pinus_strobus - ba_pist)^2 +
            (BA_Frac_Acer_rubrum - ba_acru)^2 +
            (BA_Frac_Quercus_rubra - ba_quru)^2
        )
      ) |>
      arrange(BA_RMSE) |>
      head(1) |>
      pull(MasterPlotID)    
  }) |>
    unlist()
}

tmp_plot_fingerprint |>
  mutate(sibling_plot = most_similar_plot(tmp_plot_fingerprint)) |>
  select(MasterPlotID, sibling_plot)
```

How's that looks:

```{r plot sibling plots, eval = FALSE}
tmp_top_species <- c(
  "Acer rubrum", "Quercus rubra", "Pinus strobus"
)

tmp_species <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(VisitCycle == 1970) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  mutate(
    SpeciesCode = if_else(SpeciesCode == 837, 833, SpeciesCode)
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, SCIENTIFIC_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(SCIENTIFIC_NAME %in% tmp_top_species) |>
  distinct(MasterPlotID, SCIENTIFIC_NAME) |>
  group_by(MasterPlotID) |>
  summarize(
    Species = do.call(
      paste,
      append(sort(SCIENTIFIC_NAME), list(sep = ", "))
    ),
    .groups = "drop"
  ) |>
  select(MasterPlotID, Species)

matched_plots <- cficop_plot_sibling |>
  mutate(
    status = if_else(MasterPlotID == sibling_plot, "sibling", "disturbed")
  ) |>
  left_join(
    cfigro_plot,
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  rename(
    lat = LATITUDE,
    long = LONGITUDE
  ) |>
  mutate(
    group = as.character(sibling_plot)
  ) |>
  select(
    MasterPlotID, status, lat, long, group
  ) |>
  left_join(
    tmp_species, by = join_by(MasterPlotID)
  ) |>
  group_by(group) |>
  mutate(n = n(), group = paste0(group, ", n=", n())) |>
  ungroup() |>
  filter(n > 12)
#  filter(n > 6)

matched_plot_group_order <- matched_plots |>
  distinct(group, n) |>
  arrange(desc(n))

matched_plots$group <- ordered(matched_plots$group, levels = matched_plot_group_order$group)

cfi_bbox <- st_as_sfc(
  st_bbox(vect(matched_plots, geom = c("long", "lat")))
)
sf_use_s2(FALSE)
st_crs(cfi_bbox) <- "WGS84"
lakes_north_america <- ne_download(
  scale = 10, type = "lakes_north_america", category = "physical"
)
lakes <- lakes_north_america[
  st_intersects(lakes_north_america, cfi_bbox, sparse = FALSE),
]
rivers_north_america <- ne_download(
  scale = 10, type = "rivers_north_america", category = "physical"
)
rivers <- rivers_north_america[
  st_intersects(rivers_north_america, cfi_bbox, sparse = FALSE),
]
ggplot() +
  geom_sf(data = lakes, fill = "blue", inherit.aes = FALSE) +
#  geom_sf(data = rivers, fill = "blue", inherit.aes = FALSE) +
  geom_point(
    data = matched_plots |> filter(status == "sibling"),
    aes(long, lat, color = Species, shape = status), size = 2.5
  ) +
  geom_point(
    data = matched_plots |> filter(status == "disturbed"),
    aes(long, lat, color = Species, shape = status), size = 1.75
  ) +
  facet_wrap("group") +
#  scale_color_discrete(guide = "none") +
  scale_shape_manual(values = list(sibling = 17, disturbed = 16)) +
#  annotation_scale(location = "br") +
#  annotation_north_arrow(
#    location = "tl",
#    width = unit(1, "cm"),
#    pad_x = unit(0.75, "cm"),
#    pad_y = unit(0.5, "cm")
#  ) +
  #scale_x_continuous(name = NULL, labels = NULL) +
  #scale_y_continuous(name = NULL, labels = NULL) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ggtitle("Matching harvested and disturbed CFI plots to grow-only plots")

```

Create a table to facilitate plotting plots on a plot

```{r plots for plotting, eval = FALSE}
tmp_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(VisitCycle == 1970) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  # Quercus rubra and Quercus velutina are often conflated in surveys; instead
  # of treating them as two separate important species, combine them into one and
  # treat them as a single species.
  # Replace Quercus velutina (837) with Quercus rubra (833)
  mutate(
    SpeciesCode = if_else(SpeciesCode == 837, 833, SpeciesCode)
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, SCIENTIFIC_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    TPA = 5,
    BA = pi * (VisitTreeDIAM / 2)^2 * 5
  )

tmp_by_plot <- tmp_trees |>
  group_by(MasterPlotID) |>
  summarize(
    TPA_Plot = sum(TPA, na.rm = TRUE),
    BA_Plot = sum(BA, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ungroup() |>
  select(MasterPlotID, TPA_Plot, BA_Plot)

tmp_top_species <- c(
  "Acer rubrum", "Quercus rubra", "Pinus strobus"
)

tmp_plot_fingerprint <- tmp_trees |>
  group_by(MasterPlotID, SCIENTIFIC_NAME) |>
  summarize(
    TPA_Species = sum(TPA, na.rm = TRUE),
    BA_Species = sum(BA, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  left_join(tmp_by_plot, by = join_by(MasterPlotID)) |>
  mutate(
    TPA_Frac = TPA_Species / TPA_Plot,
    BA_Frac = BA_Species / BA_Plot
  ) |>
  select(-ends_with("_Species")) |>
#  filter(BA_Frac > 0.1) |>
  filter(SCIENTIFIC_NAME %in% tmp_top_species) |>
  group_by(MasterPlotID) |>
  mutate(
    Species = do.call(
      paste,
      append(sort(SCIENTIFIC_NAME), list(sep = ", "))
    )
  ) |>
  ungroup() |>
  mutate(SCIENTIFIC_NAME = gsub(" ", "_", SCIENTIFIC_NAME)) |>
  pivot_wider(names_from = "SCIENTIFIC_NAME", values_from = c("TPA_Frac", "BA_Frac")) |>
  replace_na(list(
    BA_Frac_Acer_rubrum = 0,
    BA_Frac_Pinus_strobus = 0,
    BA_Frac_Quercus_rubra = 0
  )) |>
  mutate(
    Dominant_Species = case_when(
      BA_Frac_Acer_rubrum > BA_Frac_Pinus_strobus &
        BA_Frac_Acer_rubrum > BA_Frac_Quercus_rubra ~
        "Acer rubrum",
      BA_Frac_Pinus_strobus > BA_Frac_Acer_rubrum &
        BA_Frac_Pinus_strobus > BA_Frac_Quercus_rubra ~
        "Pinus strobus",
      .default = "Quercus rubra"
    )
  ) |>
  left_join(
    cfigro_plot_harvested,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    cfigro_plot_disturbed,
    by = join_by(MasterPlotID)
  ) |>
  mutate(
    GrowOnly = !CFIHarvested & !CFIDisturbed
  ) |>
  left_join(
    cfigro_plot |>
      filter(INV_YEAR == 1970) |>
      select(MasterPlotID = STAND_ID, Lat = LATITUDE, Lon = LONGITUDE),
    by = join_by(MasterPlotID)
  ) |>
  select(
    MasterPlotID, Lat, Lon, Dominant_Species, GrowOnly
  )

```

### Copy CFIGRO Estab

We'll duplicate cfigro_estab from the reference plots to their sibling
plots.

```{targets cficop_estab, tar_simple = TRUE}
# Each plot has a sibling plot;
# find the estab associated with the sibling plot.
cficop_plot_sibling |>
  left_join(
    cfigro_plot |> select(STAND_CN, MasterPlotID = STAND_ID),
    by = join_by(sibling_plot == MasterPlotID)
  ) |>
  mutate(STAND_CN = as.numeric(STAND_CN)) |>
  left_join(
    cfigro_estab,
    by = join_by(STAND_CN),
    # Each row on the left is a stand; many stands have the same STAND_CN.
    # Each row on the right is a stand-species; many rows on the right will
    # match a single stand. Therefore this is a many-to-many join.
    relationship = "many-to-many"
  ) |>
  # STAND_CN is for the sibling plot; swap that out
  # with the STAND_CN for the matched plot.
  select(-STAND_CN) |>
  left_join(
    cfigro_plot |> select(STAND_CN, MasterPlotID = STAND_ID),
    by = join_by(MasterPlotID)
  )
```

### Growth

cfigro_none_growth is tree-specific; for each tree within a plot,
provide the diameter and height 10 years from the simulation start time.

### Mortality

cfigro_none_death is tree-specific; for each tree within a plot, provide
the status at simulation start time.

## Harvested, Disturbed, No Trees

Let's get a breakdown of plots by whether they had no trees in
1970, were harvested, disturbed, or both

```{r}
# All the plots we're using
cficop_plot_sibling |>
  left_join(cfigro_plot_disturbed, by = join_by(MasterPlotID)) |>
  left_join(cfigro_plot_harvested, by = join_by(MasterPlotID)) |>
  mutate(
    Harvested = as.ordered(if_else(CFIHarvested, "Yes", "No")),
    Disturbed = as.ordered(if_else(CFIDisturbed, "Yes", "No"))
  ) |>
  count(Harvested, Disturbed, name = "Count") |>
  ggplot(aes(Harvested, Disturbed, fill = Count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Count), color = "white") +
  coord_fixed() +
  theme_minimal(base_size = 20)
```


## Watershed Area

I think I need to get a polygon for the watershed / districts, then use
land cover to find the forested area ... in 1970. Which I can't. Unless
they have literature that tells me.

But we can start with, what if it's 100% forested.

```{targets cficop_sam_wt, tar_simple = TRUE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  group_by(MasterPlotID) |>
  arrange(VisitCycle, MasterTreeID) |>
  filter(row_number() == 1) |>
  ungroup() |>
  left_join(
    tblSuppPlotForestrySubBasins,
    by = join_by(DWSPForestrySubWatershed == MasterForestryBasinID)
  ) |>
  group_by(EQDISTRICT, EQSUBDISTRICT) |>
  mutate(
    TotalAcres = max(DWSPOwnedAcres, na.rm = TRUE),
    NumPlots = n(),
    AcresPerPlot = TotalAcres / NumPlots,
    .groups = "keep"
  ) |>
  ungroup() |>
  select(
    EQDISTRICT,
    EQSUBDISTRICT,
    MasterPlotID,
    TotalAcres,
    NumPlots,
    AcresPerPlot
  )
```

```{r plot scaling cfi plots to watershed, eval = FALSE}
cficop_sam_wt |>
  group_by(EQSUBDISTRICT) |>
  filter(row_number() == 1) |>
  ggplot(aes(EQSUBDISTRICT, NumPlots, size = AcresPerPlot)) +
  geom_point() +
  coord_cartesian(ylim = c(0, 80)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -15, hjust = 0, vjust = 1)) +
  ggtitle("Scaling CFI plots to the Watershed") +
  xlab("Subwatershed") +
  ylab("Number of Plots")
```

```{r map cfi plot scaling to watershed, eval = FALSE}
tmp_state_map <- fvsne_states |>
  filter(STATE_NAME == "Massachusetts") |>
  rename(region = STATE_NAME) |>
  _$region |>
  map_data("state", region = _)

tmp_plot_loc <- cfigro_plot |>
  rename(lat = LATITUDE, long = LONGITUDE, group = STAND_CN) |>
  left_join(
    cficop_sam_wt |> select(MasterPlotID, EQSUBDISTRICT, AcresPerPlot),
    by = join_by(STAND_ID == MasterPlotID)
  )

ggplot(tmp_state_map, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  geom_point(
    data = tmp_plot_loc,
    aes(size = AcresPerPlot, color = EQSUBDISTRICT)
  ) +
  coord_cartesian(xlim = c(-72.45, -72.15), ylim = c(42.25, 42.55)) +
  theme_bw()
```

It looks, however, like we want to treat the entire area as one thing.
What's the total area?

```{r total acres, eval = FALSE}
print(277*0.2)
tblSuppPlotForestrySubBasins |>
  filter(Watershed == 1) |>
  summarize(acres = sum(DWSPOwnedAcres)) |>
  pull(acres)
```

So, by area, we've sampled 277\*0.2 = 55.4 acres out of 58657 acres, or
about 0.1%.

Oh, but probably we only want the subwatersheds that plots are actually
present on, so:

```{r sample weight, eval = FALSE}
cficop_sam_wt |>
  group_by(EQSUBDISTRICT) |>
  filter(row_number() == 1) |>
  ungroup() |>
  summarize(acres = sum(TotalAcres)) |>
  pull(acres)
```

so 19522, or roughly 20,000 acres, or less than half the total.

Or maybe we want not just ABP plots, but all Quabbin plots

```{r total acres another way, eval = FALSE}
tblSuppPlotForestrySubBasins |>
  filter(Watershed == 1) |>
  semi_join(
    tblDWSPCFIPlotsComplete |> distinct(DWSPForestrySubWatershed),
    by = join_by(MasterForestryBasinID == DWSPForestrySubWatershed)
  ) |>
  summarize(acres = sum(DWSPOwnedAcres)) |>
  pull(acres)
```

56816 acres.

# CFICOP_NONE

## Run FVS

```{targets cficop_none}
tar_target(
  cficop_none,
  {
    timestep <- 10 # years; determined by FVSne variant
    fvsbin_dir <- "/fvs/fvsbin" # TODO: put these in a config file
    fvs_variant <- "fvsne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "NONE"
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = cfigro_plot |>
        # Filter out the 6 plots that do not get a sibling
        semi_join(
          cficop_plot_sibling,
          by = join_by(STAND_ID == MasterPlotID)
        ),
#        left_join(
#          cficop_sam_wt |> select(STAND_ID = MasterPlotID, SAM_WT = AcresPerPlot),
#          by = join_by(STAND_ID)
#        ),
      trees = cfigro_trees,
      regen = cficop_estab,
#      calibration = cfigro_none_growth,
#      calib_mort = cfigro_none_death,
#      calib_years = 10,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

## Load Results

Carbon

```{targets cficop_none_carbon, tar_simple = TRUE}
fvs_read_output(cficop_none, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_none_summary, tar_simple = TRUE}
fvs_read_output(cficop_none, "FVS_Summary2_East") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

## BA Over Time

```{r plot ba over time, eval = FALSE}
cficop_none_summary |>
  filter(Year %% 10 == 0) |>
#  group_by(Year) |>
#  summarize(BA = sum(BA) / n()) |>
#  ungroup() |>
  mutate(Series = "Projected") |>
  union_all(
    cfigro_srvy_summary |>
      filter(Year %% 10 == 0) |>
#      group_by(Year) |>
#      summarize(BA = sum(BA) / n()) |>
#      ungroup() |>
      mutate(Series = "Observed")
  ) |>
  mutate(
    Year = as.character(Year),
    BA = conv_multiunit(BA, "ft2 / acre", "m2 / hectare")
  ) |>
  ggplot(aes(Year, BA, fill = Series)) +
  geom_boxplot(outlier.size = 0.5) +
  theme_bw() +
  ggtitle("CFI Grow-Only Projection") +
  ylab(bquote("Basal Area " ~ (m^2 %.% hectare^-1)))

```

## Carbon Over Time

```{r plot carbon over time, eval = FALSE}
cficop_none_carbon |>
#  group_by(Year) |>
#  summarize(Aboveground_Total_Live = sum(Aboveground_Total_Live) / n()) |>
#  ungroup() |>
  mutate(Series = paste0("Projection, n=", n_distinct(StandID))) |>
  union_all(
    cfigro_srvy_carbon |>
#      group_by(Year) |>
#      summarize(Aboveground_Total_Live = sum(Aboveground_Total_Live) / n()) |>
#      ungroup() |>
      mutate(Series = paste0("Reference, n=", n_distinct(StandID)))
  ) |>
  mutate(Year = as.character(Year)) |>
  ggplot(aes(Year, Aboveground_Total_Live, fill = Series)) +
  geom_boxplot(outlier.size = 0.5) +
  theme_bw() +
  ggtitle("CFI Grow-Only Projection") +
  ylab(bquote("Aboveground Live Carbon " ~ (`Mg C` %.% hectare^-1)))
```

## Error Model

We need: cfigro_bai_task\$select( c("Acer_rubrum", "LAT",
"Pinus_strobus", "Projected_BA", "Projected_Tph", "Quercus_rubra",
"SLOPE") )

```{targets cficop_none_bai, tar_simple = TRUE}
tmp_top_species <- c("Acer rubrum", "Pinus strobus", "Quercus rubra")

tmp_species_cmp <- fvs_species_composition(cficop_none) |>
  filter(Year == 2020) |>
  mutate(SpeciesCode = as.numeric(SpeciesFIA)) |>
  # Combine black oak and red oak
  mutate(
    SpeciesCode = if_else(SpeciesCode == 837, 833, SpeciesCode)
  ) |>
  group_by(StandID, Year, SpeciesCode, random_seed) |>
  summarize(BA = sum(BA), TPA = sum(TPA), .groups = "drop") |>
  left_join(
    species_crosswalk |> select(SPCD, SCIENTIFIC_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    TPH = conv_unit(TPA, "hectare", "acre"),
    BA = conv_multiunit(BA, "ft2 / acre", "m2 / hectare"),
  )

tmp_by_plot <- tmp_species_cmp |>
  group_by(StandID, random_seed) |>
  summarize(
    TPA = sum(TPA),
    BA = sum(BA),
    .groups = "drop"
  )

# Top species by plot
# (only considering tmp_top_species)
tmp_plot_species <- tmp_species_cmp |>
  group_by(StandID, SCIENTIFIC_NAME, random_seed) |>
  summarize(BA = sum(BA), .groups = "drop") |>
  ungroup() |>
  left_join(
    tmp_by_plot |>
      select(StandID, random_seed, BA_Plot = BA),
    by = join_by(StandID, random_seed)
  ) |>
  mutate(BA_Frac = BA / BA_Plot) |>
  filter(SCIENTIFIC_NAME %in% tmp_top_species) |>
  mutate(
    SCIENTIFIC_NAME = sub(" ", "_", SCIENTIFIC_NAME)
  ) |>
  pivot_wider(
    id_cols = c("StandID", "random_seed"),
    names_from = SCIENTIFIC_NAME,
    values_from = BA_Frac,
    values_fill = 0
  )

# cficop_none_summary aggregates over random_seed; we need random_seed
# in the output.
tmp_none_summary <- fvs_read_output(cficop_none, "FVS_Summary2_East")

tmp_plot_starting_ba <- tmp_none_summary |>
  filter(Year == 1970) |>
  mutate(
    Starting_BA = conv_multiunit(BA, "ft2 / acre", "m2 / hectare"),
    Starting_Tph = conv_unit(Tpa, "hectare", "acre")
  ) |>
  select(StandID, random_seed, Starting_BA, Starting_Tph)

tmp_plot_projected_ba <- tmp_none_summary |>
  filter(Year == 2020) |>
  mutate(
    Projected_BA = conv_multiunit(BA, "ft2 / acre", "m2 / hectare"),
    Projected_Tph = conv_unit(Tpa, "hectare", "acre")
  ) |>
  select(StandID, random_seed, Projected_BA, Projected_Tph)

# cficop_none_carbon aggregates over random_seed; we need random_seed
# in the output.
tmp_none_carbon <- fvs_read_output(cficop_none, "FVS_Carbon")

tmp_plot_starting_carbon <- tmp_none_carbon |>
  filter(Year == 1970) |>
  rename(Starting_Carbon = Aboveground_Total_Live) |>
  select(StandID, random_seed, Starting_Carbon)

tmp_plot_projected_carbon <- tmp_none_carbon |>
  filter(Year == 2020) |>
  rename(Projected_Carbon = Aboveground_Total_Live) |>
  select(StandID, random_seed, Projected_Carbon)

cficop_none_bai <- tmp_plot_species |>
  left_join(
    cfigro_plot |>
      filter(INV_YEAR == 1970) |>
      mutate(StandID = as.character(STAND_ID)) |>
      select(StandID, LAT = LATITUDE, ASPECT, SLOPE, ELEV = ELEVFT),
    by = join_by(StandID)
  ) |>
  left_join(
    tmp_plot_starting_ba,
    by = join_by(StandID, random_seed)
  ) |>
  left_join(
    tmp_plot_projected_ba,
    by = join_by(StandID, random_seed)
  ) |>
  left_join(
    tmp_plot_starting_carbon,
    by = join_by(StandID, random_seed)
  ) |>
  left_join(
    tmp_plot_projected_carbon,
    by = join_by(StandID, random_seed)
  ) |>
  mutate(
    BAI_Error = cfigro_bai_model$predict_newdata(pick(everything()))$response,
    Carbon_Flux_Error = cfigro_aglcf_model$predict_newdata(pick(everything()))$response
  ) |>
  mutate(
    Projected_BAI = (Projected_BA - Starting_BA) / (2020 - 1970),
    Adjusted_BAI = Projected_BAI - BAI_Error,
    Projected_AGLCF = (Projected_Carbon - Starting_Carbon) / (2020 - 1970),
    Adjusted_AGLCF = Projected_AGLCF - Carbon_Flux_Error
  )
```

Plot it:

```{r plot cficop none bai, eval = FALSE}
cficop_none_bai |>
  rename(
    `FVS Projected` = Projected_BAI,
    `Corrected` = Adjusted_BAI
  ) |>
  pivot_longer(c("FVS Projected", "Corrected")) |>
  ggplot(aes(value, fill = name)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_discrete(name = "Series") +
  xlab(bquote("Basal Area Increment" ~ (m^2 %.% hectare^-1 %.% year^-1))) +
  ylab("Density") +
  ggtitle("CFI Quabbin All Plots 50 year BAI Projection")
```

```{r plot cficop none aglcf, eval = FALSE}
cficop_none_bai |>
  rename(
    `FVS Projected` = Projected_AGLCF,
    `Corrected` = Adjusted_AGLCF
  ) |>
  pivot_longer(c("FVS Projected", "Corrected")) |>
  ggplot(aes(value, fill = name)) +
  geom_density(alpha = 0.5) +
  theme_bw() +
  scale_fill_discrete(name = "Series") +
  xlab(bquote("Aboveground Live Carbon Flux" ~ (`Mg C` %.% hectare^-1 %.% year^-1))) +
  ylab("Density") +
  ggtitle("CFI Quabbin All Plots 50 year Aboveground Live Carbon Flux Projection")
```

## Error Distribution

```{r plot cficop error distribution, eval = FALSE}
cficop_none_bai |>
  mutate(Series = "Modeled") |>
  ggplot(aes(Carbon_Flux_Error, fill = Series)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.25) +
  theme_bw() +
  xlab(bquote("Aboveground Live Carbon Flux Residual" ~ (`Mg C` %.% ha^-1 %.% year^-1))) +
  ylab("Density") +
  scale_fill_brewer(palette="Set1") +
  ggtitle("Modeled AGL Carbon Flux Residual")
```

# CFICOP_DFLT

Like CFIGRO_NONE, but use default growth and mortality.

## Regeneration

Ignore plot-specificc regeneration. use a spatial average over the first
survey period to see how many trees of each species establish in each
plot.

Regeneration in Quabbin is fraught. Regen was suppressed in areas from
1960 into the 1990s by excessive deer browse; hunting was re-introduced
in the 1990s to reduce the deer population.

### 1980 Model Year

Can we use trees that establish in the 1970-1980 cycle to represent
"what would a manager's perspective be on establishment in 1970?".

Start by finding all the trees that establish:

```{r cfiabp_estab, eval = FALSE}
cfiabp_estab <- qryDWSPCFIPlotVisitTreeDetail |>
  filter(VisitTreeStatusCode == 1) |>
  left_join(
    tblDWSPCFIPlotVisitsComplete |>
      select(MasterPlotID, VisitCycle, MasterPlotVisitID),
    by = join_by(MasterPlotVisitID)
  ) |>
  left_join( 
    tblDWSPCFITreesComplete |> select(MasterTreeID, SpeciesCode),
    by = join_by(MasterTreeID)
  ) |>
  inner_join(
    cfiabp_trees |>
      distinct(MasterPlotID, VisitCycle, MasterTreeID, StatusB),
    by = join_by(MasterPlotID, VisitCycle, MasterTreeID)
  ) |>
  filter(StatusB == "R")
```

### CFI Establishment By Year

Let's look at establishment rate - this will be how many trees establish
per 0.2 acre plot in a 10-year interval, on grow-only plots.

```{r establisment rate, eval = FALSE}
cfiabp_estab |>
  left_join(
    cfigro_plot_harvested,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    cfigro_plot_disturbed,
    by = join_by(MasterPlotID)
  ) |>
  filter(!CFIHarvested & !CFIDisturbed) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(rate = n(), .groups = "keep") |>
  ungroup() |>
  # Convert rate to trees per acre per year
  mutate(rate = rate * 5 / 10) |>
  ggplot(aes(factor(VisitCycle), rate)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Grow-Only Quabbin CFI Plot Establisment Rate") +
  ylab("New 5\"dbh trees per acre per year") +
  xlab("Inventory Year")
  
```

This says "no" - 1980 establishment rate is roughly 5x that of 2020.
1990 is still double.

```{r establishment per cycle, eval = FALSE}
cfiabp_estab |>
  left_join(
    cfigro_plot_harvested,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    cfigro_plot_disturbed,
    by = join_by(MasterPlotID)
  ) |>
  filter(!CFIHarvested & !CFIDisturbed) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(rate = n(), .groups = "keep") |>
  ungroup() |>
  mutate(rate = rate * 5 / 10) |>
  group_by(VisitCycle) |>
  summarize(rate = mean(rate))
```

### CFI Grow-Only Recruitment

Maybe we can use Hunt's data for seedling establishment. To do this,
we'd create a stems-per-acre NATURAL block that omits height; the
omission of height says these are seedlings, and FVS should figure
things out from there.

```{r establishment by species, eval = FALSE}
cfiabp_estab |>
  left_join(
    cfigro_plot_harvested,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    cfigro_plot_disturbed,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    species_crosswalk,
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(GENUS = if_else(GENUS == "Tree", "Other", GENUS)) |>
  filter(!CFIHarvested & !CFIDisturbed) |>
  mutate(GENUS = fct_lump(GENUS, n = 5, other_level = "Other")) |>
  ggplot(aes(VisitCycle, fill = GENUS, group = GENUS)) +
  geom_bar() +
  theme_bw() +
  ggtitle("Quabbin CFI Recruitment in Grow-Only Plots") +
  ylab("Number of Trees Recruited")
```

### CFI All Recruitment

```{r establishment by species in all plots, eval = FALSE}
cfiabp_estab |>
  left_join(
    cfigro_plot_harvested,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    cfigro_plot_disturbed,
    by = join_by(MasterPlotID)
  ) |>
  left_join(
    species_crosswalk,
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(GENUS = if_else(GENUS == "Tree", "Other", GENUS)) |>
#  filter(!CFIHarvested & !CFIDisturbed) |>
  mutate(GENUS = fct_lump(GENUS, n = 5, other_level = "Other")) |>
  ggplot(aes(VisitCycle, fill = GENUS, group = GENUS)) +
  geom_bar() +
  theme_bw() +
  ggtitle("Quabbin CFI Recruitment in All Plots") +
  ylab("Number of Trees Recruited")
```

### Fred Hunt's Table 13

Fred Hunt's 1961 Master's Thesis at UMass Amherst describes a general
rate of 254 stems of "establishment" per acre; we can use this number,
distributed according to his table 13, to set up background
establishment.

Table 13. Reproduction by Species

Here is a reproduction of Fred Hunt's table 13 showing reproduction in
1960, from Brian Keevan, personal communication, June 12 2025.

Keevan notes: \> "Note that Hunt decided to call anything from 3 foot
tall to 4.9” \> diameter ‘reproduction’."

Hunt left some species unspecified; we substitute the most common in
Quabbin for these: - Red maple = red maple - Oak = northern red oak -
white pine = eastern white pine - black birch = sweet birch - white ash
= white ash - red pine = red pine - other birch = paper birch - other
hardwoods = black cherry - sugar maple = sugar maple - spruce = red
spruce

```{targets cficop_dflt_estab, tar_simple = TRUE}
tibble(
  Species = c(
    "White pine", "Hemlock", "Spruce", "Red Pine",
    "Sugar Maple", "Red maple", "Oak", "Black birch",
    "Other birch", "White ash", "Other hardwoods"
  ),
  Stems_Per_Acre = c(
    45, 13, 4, 10,
    8, 63, 54, 21,
    10, 16, 10
  ),
  Percent = c(
    10, 5, 2, 4,
    3, 25, 21, 8,
    4, 6, 4
  )
) |>
  mutate(
    SPCD = case_match(
      Species,
      "Red maple" ~ 316,
      "Oak" ~ 833,
      "White pine" ~ 129,
      "Black birch" ~ 372,
      "White ash" ~ 541,
      "Hemlock" ~ 261,
      "Red Pine" ~ 125,
      "Other birch" ~ 375,
      "Other hardwoods" ~ 762,
      "Sugar Maple" ~ 318,
      "Spruce" ~ 97
    )
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SPCD)
  ) |>
  select(SPECIES = FVS_SPCD, DENSITY = Stems_Per_Acre) |>
  mutate(STAND_CN = NA, YEAR = NA) |>
  arrange(desc(DENSITY))
```

## Run FVS

This should be simple; there's no calibration.

```{targets cficop_dflt}
tar_target(
  cficop_dflt,
  {
    timestep <- 10 # years; determined by FVSne variant
    fvsbin_dir <- "/fvs/fvsbin" # TODO: put these in a config file
    fvs_variant <- "fvsne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "DFLT"
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = cfigro_plot,
      trees = cfigro_trees,
      regen = cficop_dflt_estab,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

## Load Results

Carbon

```{targets cficop_dflt_carbon, tar_simple = TRUE}
fvs_read_output(cficop_dflt, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_dflt_summary, tar_simple = TRUE}
fvs_read_output(cficop_dflt, "FVS_Summary2_East") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

# CFICOP_SRVY

Look at how things actually grew, so we can compare to CFICOP_HVST.

## Run FVS

```{targets cficop_srvy}
tar_target(
  cficop_srvy,
  {
    visits <- cfiabp_trees |>
      distinct(MasterPlotID, VisitCycle) |>
      left_join(
        tblDWSPCFIPlotVisitsComplete |>
          select(MasterPlotID, VisitCycle, MasterPlotVisitID, VisitYear),
        by = join_by(MasterPlotID, VisitCycle)
      ) |>
      mutate(MasterPlotVisitID = as.character(MasterPlotVisitID)) |>
      select(
        STAND_CN = MasterPlotVisitID,
        STAND_ID = MasterPlotID,
        INV_YEAR = VisitYear
      )
    
    # Despite its name, cfigro_plot has all plots in it
    plots_for_fvs <- cfigro_plot |>
      select(-STAND_CN, -INV_YEAR) |>
      full_join(visits, by = join_by(STAND_ID)) |>
      mutate(FIRST_YEAR = INV_YEAR, LAST_YEAR = INV_YEAR)

    # No establishment for survey runs

    fvsbin_dir <- "/fvs/fvsbin" # TODO: put these in a config file
    fvs_variant <- "fvsne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "SRVY"

    # We communicate with FVS through files. FVSOnline shows a model in which
    # a "project" (the inputs and outputs of a single FVS run) live in a
    # single directory; we follow that model.
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = plots_for_fvs,
      trees = cfigro_trees,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition
    )
  },
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here:
  pattern = map(fvs_partition)
)
```

## Load Results

Carbon

```{targets cficop_srvy_carbon, tar_simple = TRUE}
fvs_read_output(cficop_srvy, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_srvy_summary, tar_simple = TRUE}
fvs_read_output(cficop_srvy, "FVS_Summary2_East") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

# CFICOP_HVST

Let's look at the distribution of harvest, and what we'll need to model.

## Observed Harvest

We're interested in a few things:

1.  What trees get cut? (when / where / species / size / tpa)
2.  What trees are left? (when / where / species / size / tpa)

```{r observed harvest ba, eval = FALSE}
tmp_observed_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(StatusB == "C" | Status6 == "B") |>
  filter(!is.na(basal_area)) |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

tmp_observed_harvest |>
  ggplot(aes(VisitCycle, basal_area, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("Harvested BA Per Plot") +
  ylab(bquote("Harvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

tmp_observed_harvest |>
  ggplot(aes(VisitCycle, tpa, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
#  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Harvested TPA Per Plot") +
  ylab(bquote("Harvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

```

Are any stands harvested in multiple cycles?

```{r observed multiple harvest, eval = FALSE}
tmp_observed_harvest |>
  distinct(MasterPlotID, VisitCycle) |>
  group_by(MasterPlotID) |>
  summarise(num_harvests = n()) |>
  ungroup() |>
  filter(num_harvests > 1) |>
  nrow()
```

So, 54 plots are harvested more than once.

Any stands harvested multiple times in a single cycle?

```{r observed multiple year cut, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    YearCut = VisitCycle - YearsSinceLastCut
  ) |>
  distinct(MasterPlotID, VisitCycle, YearCut) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(num_harvests = n(), .groups = "keep") |>
  ungroup() |>
  filter(num_harvests > 1) |>
  nrow()
```

Stand 10202 was harvested in 1980 and 1982.

No; when a plot is harvested during a visitcycle, all trees are
harvested in the same year.

## Harvest Parameters

### Who got left behind?

Unharvested: - Filter to plot/cycles with harvest - sum the ba of what's
left

```{r observed unharvested ba, eval = FALSE}

# Which plots are harvested, in which visit cycles?

tmp_harvest_history <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_abp(cfiabp_trees) |>
  # cfi_harvested() |> # Can't use this, doesn't retain VisitCycle
  group_by(MasterPlotID, VisitCycle) |>
  summarize(
    CFIHarvested = any(CutSinceLastVisit, na.rm = TRUE),
    .groups = "drop"
  )

# What trees are left behind after harvest?

tmp_observed_remainders <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | Status6 == "L") |>
  filter(!is.na(basal_area)) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

tmp_observed_remainders |>
  ggplot(aes(VisitCycle, basal_area, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("Unharvested BA Per Plot") +
  ylab(bquote("Unarvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

tmp_observed_remainders |>
  ggplot(aes(VisitCycle, tpa, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(5, 100)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
#  scale_y_continuous(transform = "log10") +
  ggtitle("Unharvested TPA Per Plot") +
  ylab(bquote("Unarvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

tmp_observed_remainders |>
  ggplot(aes(VisitCycle, tpa)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(1, 50)) +
  theme_bw() +
#  scale_fill_manual(values = eves_colors) +
#  scale_y_continuous(transform = "log10") +
  ggtitle("Unharvested TPA Per Plot") +
  ylab(bquote("Unarvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

```

### Harvest Efficiency

What fraction of trees were harvested?

> Nik is here take the 1980 pulse (delta in 6" fron 1970 - 1980?) and
> turn those into extra trees in 1970? 4" trees? 2/3 the height? 2/3 the
> DBH? (2/3 because 4 is 2/3 of 6) also, i feel like i was looking at a
> spot in harvest where we're looking at fvs trees without summing tpa
> Also also, instead of ThinDBH, try ThinBTA - thin from below to a TPA
> target. - DBH range - everything - TPA target - what's left on the
> plot - Efficiency - calculate ratio of initial to residual tpa in each
> diameter class, take the max? mean? median? ThinBTA doesn't allow
> species selection; ThinDBH does, can maybe use ThinDBH with Cuteff to
> achieve the same goal? Note that ThinDBH ignores CutEFF if target TPA
> or BA is specified. ThinQFA also allows species to be specified; can
> thin to TPA, specified in supplemental record.

```{r observed harvested ba, eval = FALSE}
# What got harvested?

tmp_observed_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(StatusB == "C" | Status6 == "B") |>
  filter(!is.na(basal_area)) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    hvst_basal_area = sum(basal_area, na.rm = TRUE),
    hvst_tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

# Which plots are harvested, in which visit cycles?

tmp_harvest_history <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_abp(cfiabp_trees) |>
  # cfi_harvested() |> # Can't use this, doesn't retain VisitCycle
  group_by(MasterPlotID, VisitCycle) |>
  summarize(
    CFIHarvested = any(CutSinceLastVisit, na.rm = TRUE),
    .groups = "drop"
  )

# What trees are left behind after harvest?

tmp_observed_remainders <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | Status6 == "L") |>
  filter(!is.na(basal_area)) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    rmdr_basal_area = sum(basal_area, na.rm = TRUE),
    rmdr_tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

tmp_harvest_fraction <- tmp_observed_harvest |>
  left_join(
    tmp_observed_remainders,
    by = join_by(MasterPlotID, VisitCycle, GENUS)
  ) |>
  replace_na(list(
    rmdr_basal_area = 0,
    rmdr_tpa = 0,
    hvst_basal_area = 0,
    hvst_tpa = 0
  )) |>
  mutate(
    tpa_fraction = hvst_tpa / (hvst_tpa + rmdr_tpa),
    ba_fraction = hvst_basal_area / (hvst_basal_area + rmdr_basal_area)
  ) |>
  mutate(
    group = sprintf("%s %s", VisitCycle, GENUS)
  )

tmp_harvest_fraction |>
  ggplot(aes(VisitCycle, tpa_fraction, fill = GENUS, group = group)) +
  geom_boxplot() +
  scale_fill_manual(values = eves_colors)


```

### 1980 Pulse

Let's see how many extra trees recruit in 1980

```{r cut in 1980, eval = FALSE}
excess_recruits <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees and recruits
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    diameter_class = as.integer(VisitTreeDIAM / 2) * 2
  ) |>
  filter(
    diameter_class == 6
  ) |>
  group_by(MasterPlotID, FVS_SPCD, VisitCycle) |>
  summarize(recruits = n(), .groups = "keep") |>
  left_join(
    cficop_estab |> select(MasterPlotID, SPECIES, DENSITY),
    by = join_by(MasterPlotID, FVS_SPCD == SPECIES)
  ) |>
  mutate(excess_recruits = recruits - DENSITY) |>
  ungroup() |>
  filter(!is.na(excess_recruits) & excess_recruits >= 0) |>
  mutate(Year = ordered(VisitCycle))

excess_recruits |>
  ggplot(aes(Year, excess_recruits, group = Year)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 5))
```

Rather than derive a principled way to turn 1980 ingrowth into 1970
saplings, we can represent the 1970 sapling population by actual 1980
ingrowth. By 1980, the 1970 seedling should have created a sapling pool
that FVS can track, so we don't need to do anything beyond 1980. So
let's add 1980 establishment as a 1-off. We use actual trees and height
when available.

```{targets cficop_hvst_estab, tar_simple = TRUE}
# Find 1970 trees
tmp_trees_1970 <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(
    VisitCycle == 1970
  ) |>
  select(
    MasterPlotID, MasterTreeID
  )

tmp_estab_1980 <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  # Remove any trees that were around in 1970
  anti_join(
    tmp_trees_1970,
    by = join_by(MasterPlotID, MasterTreeID)
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  # Live trees and recruits
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    diameter_class = as.integer(VisitTreeDIAM / 2) * 2
  ) |>
  filter(
    VisitCycle == 1980
  ) |>
  group_by(FVS_SPCD) |>
  # Impute mean height for missing values
  mutate(
    VisitTreeTotalHeight = if_else(
      is.na(VisitTreeTotalHeight) | VisitTreeTotalHeight == 0,
      mean(VisitTreeTotalHeight, na.rm = TRUE),
      VisitTreeTotalHeight
    ),
    DENSITY = 5  # trees per acre
  ) |>
  left_join(
    cfigro_plot |> select(STAND_ID, STAND_CN),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  select(
    STAND_ID = MasterPlotID,
    STAND_CN,
    YEAR = VisitCycle,
    SPECIES = FVS_SPCD,
    DENSITY,
    HEIGHT = VisitTreeTotalHeight
  )

cficop_hvst_estab <- cficop_dflt_estab |>
  mutate(
    STAND_ID = NA,
    HEIGHT = 0
  ) |>
  union_all(
    tmp_estab_1980
  )
```

### Diameter Class

What's the distribution of trees left behind, by diameter class?

```{r harvest by diameter class, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = ordered(as.integer(conv_unit(dbh_prior, "cm", "in") / 5) * 5)
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | StatusB == "R" | Status6 == "L" | Status6 == "R") |>
  filter(!is.na(diameter_class)) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(basal_area = sum(basal_area, na.rm = TRUE), .groups = "keep") |>
  ungroup() |>
  ggplot(aes(VisitCycle, basal_area, fill = diameter_class)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() +
  scale_fill_discrete() +
  ggtitle("Unharvested BA Per Plot by 5\" diameter class") +
  ylab(bquote("Unharvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

```{r remaining tpa by diameter class, eval = FALSE}
remaining_tpa_by_diameter_class <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = as.integer(conv_unit(dbh_prior, "cm", "in") / 2) * 2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | StatusB == "R" | Status6 == "L" | Status6 == "R") |>
  filter(!is.na(diameter_class)) |>
  filter(diameter_class <= 40) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(
    tpa = n() * 5, .groups = "keep"
  ) |>
  ungroup()

remaining_tpa_by_diameter_class |>
  mutate(
    VisitCycle = ordered(VisitCycle),
    diameter_class = ordered(diameter_class)
  ) |>
  ggplot(aes(VisitCycle, tpa, fill = diameter_class)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(5, 100)) +
  theme_bw() +
  scale_y_continuous(transform = "log2") +
  scale_fill_discrete() +
  ggtitle("Unharvested TPA Per Plot by 2\" diameter class") +
  ylab(bquote("Unharvested TPA" ~ (stems %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

ideal_tpa_by_diameter_class <- tibble(
  diameter_class = seq(3, 35, 2),
  tpa = sapply(0:16, \(i) as.integer(42 / 1.4^i))
)


remaining_tpa_by_diameter_class |>
  ggplot(aes(diameter_class, tpa)) +
  geom_boxplot(
    aes(group = diameter_class),
    width = 0.7,
    outlier.size = 0.1
  ) +
  geom_step(
    data = ideal_tpa_by_diameter_class,
    color = "red"
  ) +
  scale_x_continuous(breaks = seq(2,40,2)) +
  coord_cartesian(xlim = c(6, 20), ylim = c(0, 100)) +
  annotate("text", x = 18, y = 0, label = "DBq, q=1.4", color = "red") +
  theme_bw() +
  ggtitle("Unharvested TPA Per Plot by 2\" diameter class") +
  ylab(bquote("Unharvested TPA" ~ (stems %.% acre^-1))) +
  xlab("Diameter Class")

```

```{r remaining tpa by diameter class 2, eval = FALSE}
remaining_tpa_by_diameter_class <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = as.integer(conv_unit(dbh_prior, "cm", "in") / 2) * 2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | StatusB == "R" | Status6 == "L" | Status6 == "R") |>
  filter(!is.na(diameter_class)) |>
  filter(diameter_class <= 40) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(
    tpa = n() * 5, .groups = "keep"
  ) |>
  ungroup()

starting_tpa <- remaining_tpa_by_diameter_class |>
  filter(diameter_class == 6) |>
  ungroup() |>
  select(MasterPlotID, VisitCycle, starting_tpa = tpa) |>
  mutate(
    starting_tpa = 25 * floor(starting_tpa / 25),
    tpa_class = if_else(
      starting_tpa < 75, as.character(starting_tpa),
      "75+"
    )
  )

ideal_tpa_by_diameter_class <- lapply(
  c(0, 25, 50, 75), \(x) {
    tibble(
      diameter_class = seq(5, 35, 2),
      tpa = sapply(0:15, \(i) as.integer((x + 12.5) / 1.4^i)),
      starting_tpa = x
    ) |>
      mutate(
        tpa_class = if_else(
          starting_tpa < 75, as.character(starting_tpa),
          "75+"
        )
      )
  }
) |>
  do.call(bind_rows, args = _)

remaining_tpa_by_diameter_class |>
  left_join(
    starting_tpa, by = join_by(MasterPlotID, VisitCycle)
  ) |>
  replace_na(list(tpa_class = "0")) |>
  mutate(group = paste(tpa_class, diameter_class)) |>
  ggplot(aes(diameter_class, tpa)) +
  geom_boxplot(
    aes(fill = tpa_class, group = group),
    outlier.size = 0.1
  ) +
  geom_step(
    data = ideal_tpa_by_diameter_class,
    aes(color = tpa_class),
    show.legend = FALSE
  ) +
  scale_x_continuous(breaks = seq(2,40,2)) +
  coord_cartesian(xlim = c(6, 20), ylim = c(0, 100)) +
  theme_bw() +
  labs(fill = "6\" TPA") +
  ggtitle("Unharvested TPA Per Plot by 2\" diameter class") +
  ylab(bquote("Unharvested TPA" ~ (stems %.% acre^-1))) +
  xlab("Diameter Class")
```

### Before / After Ratio

```{r harvest ratio by diameter class, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = ordered(as.integer(conv_unit(dbh_prior, "cm", "in") / 5) * 5)
  ) |>
  # Live trees, not recruits
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(
    StatusB == "L" | StatusB == "R" | StatusB == "C" |
      Status6 == "L" | Status6 == "R" | Status6 == "C"
  ) |>
  filter(!is.na(diameter_class)) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(
    pre_harvest_ba = sum(basal_area, na.rm = TRUE),
    harvested_ba = sum(if_else(StatusB == "C" | Status6 == "C", 0, basal_area), na.rm = TRUE),
    harvest_fraction = harvested_ba / pre_harvest_ba,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, harvest_fraction, fill = diameter_class)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() +
  scale_fill_discrete() +
  ggtitle("Harvested BA Ratio Per Plot by 5\" diameter class") +
  ylab(bquote("Harvested BA" ~ (fraction))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

```{r harvest ratio by genus, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(
    StatusB == "L" | StatusB == "R" | StatusB == "C" |
      Status6 == "L" | Status6 == "R" | Status6 == "C"
  ) |>
  filter(!is.na(basal_area)) |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    VisitCycle = ordered(VisitCycle),
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other")
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    pre_harvest_ba = sum(basal_area, na.rm = TRUE),
    harvested_ba = sum(if_else(StatusB == "C" | Status6 == "C", 0, basal_area), na.rm = TRUE),
    harvest_fraction = harvested_ba / pre_harvest_ba,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, harvest_fraction, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("Harvested BA Ratio Per Plot by Genus") +
  ylab(bquote("Harvested BA" ~ (fraction))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

### Theoretical Q

```{r theoretical description of what is on the landscape, eval = FALSE}
x <- tibble(
  diameter_class = seq(6, 40, 2),
  tpa = sapply(0:17, \(i) (40 / 1.4^i))
)

x |>
  ggplot(aes(diameter_class, tpa)) +
  geom_line() +
  geom_point()
```

## Harvest by DBH

After talking with folks, we thought we were going to harvest by basal
area by diameter class by species by plot by time.

FVS provides the THINDBH keyword to schedule thinning from throughout a
dbh range for any or all species to a residual basal area or trees per
acre target.

We will provide multiple THINDBH keywords per stand per timestep,
describing the target residual trees per acre of that species in that
dbh range for that stand; the rest will be harvested. Ideally, we'd
specify the number of trees to remove, but between ingrowth, mortality,
and multiple rounds of harvest, we don't actually know how many trees
FVS will have in the stand at any given time, so we can't compute the
residual number of trees that would cause the right number to be
removed.

We could also do this by telling FVS the fraction of trees to remove; we
know how many were harvested and when, but we don't know how many trees
were on the landscape at the time of harvest, so we can't compute the
correct fraction. By looking at cficop_dflt_treelist and
cficop_none_treelist, we can get a pretty good idea of what FVS thinks
is there.

### Harvest List

Let's build a list of the number of live trees after each harvest:

We need: - In what plot, in what year, what species was harvested - for
the plot/year/species combo, what trees were left behind? - lump those
left behind into dbh classes for ThinDBH - Trees on the landscape grow
at different rates than in FVS; it's too hard to try to hit 2" diameter
classes, so we use much wider 6" diameter classes to try to catch more
trees.

```{targets cficop_hvst_tpa, tar_simple = TRUE}
tmp_harvested <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    YearCut = VisitCycle - YearsSinceLastCut
  ) |>
  filter(
    !is.na(YearCut) & !is.na(dbh_prior)
  ) |>
  # When harvest and inventory occur in the same year,
  # harvest might occur before or after the inventory.
  # To simplify things, we'll make all harvest happen
  # before the inventory.
  # Select the post-harvest inventory as the VisitCycle
  mutate(
    VisitCycle = ceiling(YearCut / 10) * 10,
    DBH_MIN = floor(conv_unit(dbh_prior, "cm", "in") / 6) * 6
  ) |>
  group_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN) |>
  summarize(YearCut = max(YearCut), .groups = "drop")

cficop_hvst_tpa <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  filter(!is.na(dbh_prior)) |>
  mutate(
    DBH_MIN = floor(conv_unit(dbh_prior, "cm", "in") / 6) * 6
  ) |>
  semi_join(
    tmp_harvested,
    by = join_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN)
  ) |>
  group_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN) |>
  summarize(
    TPA = n() * 5, # Each plot is 1/5 acre
    .groups = "drop"
  ) |>
  # Add back in the things that were harvested down to 0 TPA
  # This will also add YearCut to the things harvested to >0 TPA
  full_join(
    tmp_harvested,
    by = join_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN)
  ) |>
  # The harvest blocks added back will have TPA == NA
  mutate(
    DBH_MAX = DBH_MIN + 6,
    TPA = if_else(is.na(TPA), 0, TPA),
    BA = 0
  ) |>
  # For prescription-based harvest, our schema is:
  # STAND_CN, TREE_CN, PREV_TRE_CN, YEAR, PRESCRIPTION
  # For DBH-based harvest, our schema is:
  # STAND_CN, YEAR, DBH_MIN, DBH_MAX, SPCD, TPA, BA
  mutate(
    YEAR = YearCut
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  left_join(
    cfigro_plot |> filter(INV_YEAR == 1970) |> select(STAND_CN, STAND_ID),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  select(
    STAND_CN,
    YEAR = YearCut,
    DBH_MIN,
    DBH_MAX,
    SPCD = FVS_SPCD,
    TPA,
    BA
  )
```

### Post-Harvest Regeneration

Post-harvest there's a bunch of regeneration; find it and insert it.

## Run FVS through the UI

FVS really doesn't like to do harvest. Let's load up the UI and see if
we can convince it.

### FvS_StandInit and FVS_PlotInit

From DBSUserGuide.pdf, Table 3.0.1, we learn the following about table
structure:

-   Stand_CN - optional, but used by queries, so we'll include it
-   Stand_ID - required
-   StandPlot_ID - optional; plot identification code; required by the
    user interface when populating stand lists from the plot initiation
    table (PlotInit)
-   Variant - required; set to "ne"
-   Inv_Year - required; set to 1970
-   Latitude - optional
-   Longitude - optional
-   Location - recommended; nearest national forest; takes precedence
    over Region, forest, District, and Compartment
-   Apsect - recommended; (0 = no meaningful aspect; 360 = north)
-   Slope - recommended
-   ElevFt - recommended; Elevation in feet
-   Basal_Area_Factor - required
-   Inv_Plot_Size - required
-   Brk_DBH - required
-   Num_Plots - recommended; if 0, the number of plots in the stand is
    determined by counting the numbers [sic] of unique plot
    identification codes on the tree record data

All other fields are optional.

### FVS_TreeInit

From DBSUserGuide.pdf, Table 3.0.2, we learn the following about table
structure:

-   Stand_CN - optional, but used by queries, so we'll include it
-   Stand_ID - required
-   StandPlot_ID - required to run plots as stands
-   Plot_ID - recommended; plot identification, should be unique for an
    individual stand but can be repeated for plots in different stands
-   Tree_ID - optional
-   Tree_Count - recommended - the number of trees represented by the
    tree record; if left blank FVS assumes the tree record represents a
    single sample tree
-   Species - required
-   DBH or Diameter - required
-   Ht - recommended; total height in feet
-   CrRatio - recommended; values 1-9 are read as crown ratio codes;
    10-99 as percent

All other fields are optional.

### Generate Tables for UI

```{r, eval = FALSE}
blank_db <- DBI::dbConnect(
  RSQLite::SQLite(),
  file.path("data/fvs/BlankDatabase.db"),
  flags = RSQLite::SQLITE_RO
)

fvs_data_db <- DBI::dbConnect(
  RSQLite::SQLite(),
  file.path("data/fvs/FVS_Data.db"),
  flags = RSQLite::SQLITE_RWC
)

# There are three tables: FVS_PlotInit, FVS_StandInit, and FVS_TreeInit

# FVS_PlotInit
# We aren't supporting by-plot
DBI::dbWriteTable(
  fvs_data_db,
  "FVS_PlotInit",
  tbl(blank_db, "FVS_PlotInit") |> collect(),
  overwrite = TRUE
)

# FVS_StandInit
# Grab the structure of the table from the blank database,
# and make cfigro_plot conform to it
fvs_standinit <- bind_rows(
  tbl(blank_db, "FVS_StandInit") |>
    collect() |>
    mutate(
      Stand_CN = character()
    ),
  cfigro_plot |>
    filter(INV_YEAR == 1970) |>
    mutate(
      Stand_ID = as.character(STAND_ID),
      Variant = "ne",
      Aspect = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
      INV_PLOT_SIZE = 5, # We don't use microplot data
      Sam_Wt = 10
    ) |>
    select(
      Stand_CN = STAND_CN,
      Stand_ID,
      StandPlot_ID = STAND_ID,
      Variant,
      Inv_Year = INV_YEAR,
      Latitude = LATITUDE,
      Longitude = LONGITUDE,
      Region = REGION,
      Forest = FOREST,
      Aspect,
      Slope = SLOPE,
      ElevFt = ELEVFT,
      Basal_Area_Factor = BASAL_AREA_FACTOR,
      Inv_Plot_Size = INV_PLOT_SIZE,
      Brk_DBH = BRK_DBH,
      Num_Plots = NUM_PLOTS,
      Sam_Wt
  )
)

DBI::dbWriteTable(
  fvs_data_db,
  "FVS_StandInit",
  fvs_standinit,
  overwrite = TRUE
)

# FVS_TreeInit
# Grab the structure of the table from the blank database,
# and make cfigro_trees conform to it
fvs_treeinit <- bind_rows(
  tbl(blank_db, "FVS_TreeInit") |>
    collect(),
  cfigro_trees |>
    semi_join(fvs_standinit |> distinct(Stand_CN), by = join_by(STAND_CN == Stand_CN)) |>
    mutate(
      Stand_ID = as.character(STAND_ID),
      Variant = "ne",
      Aspect = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
      Tree_Count = 1
    ) |>
    select(
      Stand_CN = STAND_CN,
      Stand_ID,
      Plot_ID = STAND_ID,
      Tree_CN = TREE_CN,
#      Tree_ID = TREE_ID,
      Tree_Count,
      Species = SPECIES,
      Diameter = DIAMETER,
      Ht = HT,
      Slope = SLOPE,
      Aspect = ASPECT,
      TopoCode = TOPOCODE
  )
)

DBI::dbWriteTable(
  fvs_data_db,
  "FVS_TreeInit",
  fvs_treeinit,
  overwrite = TRUE
)

DBI::dbDisconnect(blank_db)
DBI::dbDisconnect(fvs_data_db)
```

## Run FVS

Make our best effort at replicating what actually happend in Quabbin.

```{targets cficop_hvst}
tar_target(
  cficop_hvst,
  {
    timestep <- 10 # years; determined by FVSne variant
    fvsbin_dir <- "/fvs/fvsbin" # TODO: put these in a config file
    fvs_variant <- "fvsne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "HVST"
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = cfigro_plot |>
        # Filter out the 6 plots that do not get a sibling
        semi_join(
          cficop_plot_sibling,
          by = join_by(STAND_ID == MasterPlotID)
        ) |>
        mutate(
          ASPECT = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
          INV_PLOT_SIZE = 5, # We don't use microplot data
          SAM_WT = 10
        ),
      trees = cfigro_trees,
      regen = cficop_hvst_estab, # cficop_dflt_estab,  
      harvest = cficop_hvst_tpa,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

## Load Results

Carbon

```{targets cficop_hvst_carbon, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_hvst_summary, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_Summary2_East") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

Cut trees

```{targets cficop_hvst_cutlist, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_CutList_East") |>
  filter(random_seed == min(random_seed))
```

### Did it do what we asked?

FVS will cancel a harvest if it can't meet the target.

How many THINDBH directives did we give?

```{bash, eval = FALSE}
# I've done this in R before, but don't feel like going and figuring
# it out again
grep "ThinDBH" data/fvs/FVS_CFICarbonOffsetProgram_HVST/*R43517.key | wc -l
```

> 3873

```{bash, eval = FALSE}
grep "DELETED OR CANCELED" data/fvs/FVS_CFICarbonOffsetProgram_HVST/*R43517.out | grep THINDBH | wc -l
```
> 2242 with 1980 compensatio, 2252 without - so canceled 8 more harvests

So, a little over half the harvest was canceled.
This seems to be:
1. Trees where actual diameter growth was different enough from modeled
   diameter growth that it ended up in a different diameter class than
   harvest specified (lessened by using 6" diameter classes), and
2. At the low end, different ingrowth rates between actual and modeled.

Is it a problem?

Let's compare harvested and remaining TPA and BA.

> Deleted: Analysis of MasterPlotID 10153

### Projected vs Observed TPA

```{r, eval = FALSE}
observed_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    Scenario = "Observed",
    MasterPlotID = as.character(MasterPlotID)
  ) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(Tpa = n() * 5, .groups = "drop") |>
  select(
    StandID = MasterPlotID,
    Year = VisitCycle,
    Observed = Tpa
  )

observed_trees |>
  full_join(
    cficop_hvst_summary |>
      filter(Year %% 10 == 0) |>
      select(StandID, Year, Projected = Tpa),
    by = join_by(StandID, Year)
  ) |>
  mutate(Year = sprintf("%d", Year)) |>
  pivot_longer(cols = c(Observed, Projected)) |>
  ggplot(aes(Year, value, fill = name)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  ggtitle("Projected vs. Observed Tree Density") +
  ylab(bquote(trees %.% acre^-1))

observed_trees |>
  full_join(
    cficop_hvst_summary |>
      filter(Year %% 10 == 0) |>
      select(StandID, Year, Projected = Tpa),
    by = join_by(StandID, Year)
  ) |>
  mutate(
    Year = sprintf("%d", Year),
    Surplus = Projected - Observed
  ) |>
  ggplot(aes(Year, Surplus)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  ggtitle("Surplus trees by inventory") +
  ylab(bquote(trees %.% acre^-1))
```

So overall TPA in the modeled plots goes up a lot compared to what's observed.
A lot ov this is likely because the model can see all trees down to seedlings,
whereas CFI inventory only captures trees >5.6"

### Measured vs. Projected Harvest

How does harvest compare to survey?

```{r hvst vs srvy, eval = FALSE}
measured_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    Year = VisitCycle - YearsSinceLastCut,
    ba = pi * (conv_unit(dbh_prior, "cm", "m") / 2)^2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(!is.na(ba) & ba > 0 & StatusB == "C" | Status6 == "B") |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  group_by(GENUS) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(ba, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

projected_harvest <- cficop_hvst_cutlist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS)
  ) |>
  pivot_longer(cols = c(Measured_BA, Projected_BA)) |>
  filter(!is.na(value)) |>
  group_by(GENUS) |>
  filter(sum(value) > 5) |>
  ungroup() |>
  ggplot(aes(GENUS, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  #facet_wrap("name") +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Basal Area Harvested") +
  xlab("Genus") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Basal Area harvested" ~(m^2)))

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS)
  ) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  filter(!is.na(value)) |>
  group_by(GENUS) |>
  filter(sum(value) > 100) |>
  ungroup() |>
  ggplot(aes(GENUS, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  #facet_wrap("name") +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Harvested") +
  xlab("Genus") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Trees harvested" ~ (trees)))
```

```{r, eval = FALSE}
measured_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(StatusB == "C" | Status6 == "B") |>
  group_by(VisitCycle) |>
  summarize(
    Observed = n() * 5,
    .groups = "keep"
  ) |>
  ungroup() |>
  select(Year = VisitCycle, Observed)

projected_harvest <- cficop_hvst_cutlist |>
  mutate(Year = ceiling(Year / 10) * 10) |>
  group_by(Year) |>
  summarize(
    Projected = sum(TPA),
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(Year)
  ) |>
  pivot_longer(cols = c(Observed, Projected)) |>
  mutate(Year = sprintf("%d", Year)) |>
  ggplot(aes(Year, value, fill = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  ggtitle("Projected Vs Observed Trees Harvested") +
  ylab("Number of Trees Harvested")
```

### Measured vs. Projected Post-Harvest Trees

cficop_hvst_tpa is the tpa that we requested that FVS leave behind.

Ugh; stand 119700442 was harvested twice between 1970 and 1980. Let's
ignore it for now; probably the right thing to do is to consolidate into
a single harvest - the earlier one to give regeneration more time, or
the later one to give regeneration less time.

```{r hvst vs srvy remainders, eval = FALSE}
measured_remainders <- cficop_hvst_tpa |>
  filter(!is.na(SPCD)) |>
  left_join(
    species_crosswalk |> select(FVS_SPCD, GENUS),
    by = join_by(SPCD == FVS_SPCD)
  ) |>
  group_by(GENUS, DBH_MIN) |>
  summarize(
    Measured_Trees = sum(TPA), # TPA is already scaled to full acres
    .groups = "drop"
  )

post_harvest_inventories <- cficop_hvst_tpa |>
  mutate(YEAR = 10 * ceiling(YEAR / 10)) |>
  distinct(STAND_CN, YEAR) |>
  left_join(cfigro_plot |> distinct(STAND_CN, STAND_ID), by = join_by(STAND_CN)) |>
  select(
    StandID = STAND_ID,
    Year = YEAR
  ) |>
  mutate(StandID = as.character(StandID))

post_harvest_treelist <- fvs_read_output(cficop_hvst, "FVS_TreeList_East") |>
  filter(random_seed == min(random_seed)) |>
  semi_join(post_harvest_inventories, by = join_by(StandID, Year))
  
projected_remainders <- post_harvest_treelist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    DBH_MIN = floor(DBH / 6) * 6
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, DBH_MIN) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees.
    Projected_Trees = sum(TPA),
    .groups = "keep"
  ) |>
  ungroup()

measured_remainders |>
  full_join(
    projected_remainders,
    by = join_by(GENUS, DBH_MIN)
  ) |>
  filter(DBH_MIN > 0) |>
  mutate(DBH_MIN = ordered(DBH_MIN)) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  filter(!is.na(GENUS) & !is.na(value)) |>
  # Remove uninteresting genera
  group_by(GENUS) |>
  filter(sum(value) > 2000) |>
  ungroup() |>
  ggplot(aes(DBH_MIN, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap("GENUS") +
  scale_x_discrete(drop = FALSE) +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Post-Harvest Trees") +
  xlab("6\" Diameter Class") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Post-Harvest Trees" ~ (trees)))
```

# Carbon on the Landscape

-   Treat the 277 plots as a sample (by area?)
-   For the sample, we can compute the Standard Error of the Mean (SEM)
    as:
    -   SEM = s / sqrt(n) where s = sample standard deviation and n =
        number of points in the sample
    -   Note that if sample size \> 5% of the population size, also
        apply Finite Population Correction (FPC), which says
    -   SEM = s / sqrt(n) \* (N - n) / (N - 1) where N = population size
-   Then a 95% CI for the population mean is x_bar +/- (1.96 \* SEM)
    where x_bar is the sample mean

What does this mean for total carbon on the landscape?

I think we just multiply the FVS projected vs. corrected AGLCF times
years times acres (converted to hectares).

from above, estimated 56816 forested acres; we have 277 \* 0.2 = 55.4
acres, so 0.1%

## Summary Dataframe

I use this a lot; capture it as a thing:
```{r cficop_cmp_carbon, eval = FALSE}
cficop_cmp_carbon <- cficop_none_carbon |>
  rename(`Site Calibrated` = Aboveground_Total_Live) |>
  left_join(
    cficop_dflt_carbon |>
      rename(Default = Aboveground_Total_Live),
    by = join_by(StandID, Year)
  ) |>
  left_join(
    cficop_srvy_carbon |>
      rename(`All Plots` = Aboveground_Total_Live),
    by = join_by(StandID, Year)
  ) |>
  left_join(
    cficop_none_bai |>
      select(StandID, Starting_Carbon, Projected_Carbon, Adjusted_AGLCF) |>
      group_by(StandID) |>
      summarize(
        Starting_Carbon = min(Starting_Carbon),
        Adjusted_AGLCF = mean(Adjusted_AGLCF),
        .groups = "drop"
      ),
    by = join_by(StandID)
  ) |>
  mutate(
    `Best Estimate` = Starting_Carbon + Adjusted_AGLCF * (Year - 1970)
  ) |>
  select(-c("Starting_Carbon", "Adjusted_AGLCF")) |>
  pivot_longer(-c("StandID", "Year")) |>
  union_all(
    cfigro_srvy_carbon |>
      select(StandID, Year, `Undisturbed (31)` = Aboveground_Total_Live) |>
      pivot_longer(-c("StandID", "Year"))
  ) |>
  mutate(
    method = name |> recode_values(
      "Default" ~ "Projected",
      "Site Calibrated" ~ "Projected",
      "Best Estimate" ~ "Projected",
      "Undisturbed (31)" ~ "Measured",
      "All Plots" ~ "Measured"
    ),
    name = ordered(
      name,
      levels = c("Default", "Site Calibrated", "Undisturbed (31)", "Best Estimate", "All Plots")
    )
  )

cficop_cmp_carbon_summary <- cficop_cmp_carbon |>
  group_by(Year, name, method) |>
  summarize(
    n = n(),
    mean = mean(value),
    ymin = mean(value) - 1.96 * sd(value) / sqrt(n),
    ymax = mean(value) + 1.96 * sd(value) / sqrt(n),
    .groups = "drop"
  )
```

## Population Mean

So, let's look at distribution of biomass in our 277 plot sample:

```{r population_mean, eval = FALSE}
hectares <- conv_unit(56816, "acre", "hectare")

population_mean <- cficop_none_carbon |>
  rename(
    Projected_Carbon = Aboveground_Total_Live
  ) |>
  left_join(
    cficop_none_bai |>
      select(StandID, Starting_Carbon, Projected_Carbon, Adjusted_AGLCF) |>
      group_by(StandID) |>
      summarize(
        Starting_Carbon = min(Starting_Carbon),
        Adjusted_AGLCF = mean(Adjusted_AGLCF),
        .groups = "keep"
      ) |>
      ungroup(),
    by = join_by(StandID)
  ) |>
  mutate(
    Adjusted_Carbon = Starting_Carbon + Adjusted_AGLCF * (Year - 1970)
  ) |>
  group_by(Year) |>
  summarize(
    projected_carbon_mean = mean(Projected_Carbon),
    projected_carbon_sd = sd(Projected_Carbon),
    adjusted_carbon_mean = mean(Adjusted_Carbon),
    adjusted_carbon_sd = sd(Adjusted_Carbon),
    n = n(),
    .groups = "keep"
  ) |>
  ungroup() |>
  mutate(
    # convert sd to ci via 1.96 / sqrt(n)
    projected_carbon = projected_carbon_mean,
    projected_carbon_min = projected_carbon - projected_carbon_sd * 1.96 / sqrt(n),
    projected_carbon_max = projected_carbon + projected_carbon_sd * 1.96 / sqrt(n),
    adjusted_carbon = adjusted_carbon_mean,
    adjusted_carbon_min = adjusted_carbon - adjusted_carbon_sd * 1.96 / sqrt(n),
    adjusted_carbon_max = adjusted_carbon + adjusted_carbon_sd * 1.96 / sqrt(n)
  ) |>
  select(
    Year,
    projected_carbon, projected_carbon_min, projected_carbon_max,
    adjusted_carbon, adjusted_carbon_min, adjusted_carbon_max
  ) |>
  pivot_longer(-Year) |>
  mutate(
    value = value * hectares,
    group = if_else(grepl("projected", name), "projected", "adjusted"),
    kind  = if_else(grepl("_min", name), "min", if_else(grepl("_max", name), "max", "mean"))
  ) |>
  pivot_wider(id_cols = c("group", "Year"), names_from = "kind")
```

Plot it:

```{r plot population mean, eval = FALSE}
population_mean |>
  mutate(group = ordered(group, levels = c("projected", "adjusted"))) |>
  ggplot(aes(Year, mean)) +
  with_blur(
    geom_ribbon(aes(ymin = min, ymax = max, fill = group), alpha = 0.25),
    sigma = 3
    
  ) +
  geom_line(aes(color = group)) +
  coord_cartesian(ylim = c(0, NA)) +
  guides(
    color = guide_legend(order = 1),
    fill = guide_legend(order = 2)
  ) +
  theme_bw() +
  ggtitle("Projected vs. Measured Total Carbon, 0.1% sample") +
  ylab(bquote("Total Carbon" ~(`Mg C`))) +
  scale_color_discrete(name = "Estimated\nPopulation\nMean") +
  scale_fill_discrete(name = "95% CI")
```

```{r population mean values, eval = FALSE}
population_mean |>
  filter(Year == 1970 | Year == 2020) |>
  mutate(ci = mean - min) |>
  select(group, Year, mean, ci)
```

In 1970, there were 1170155 +- 57272 Mg C / ha In 2020, projected was
2839915 +- 80259 In 2020, adjusted was 2713476 +- 68425 2839915 -
2713476 = 126439, 126439 / 2839915 = 0.04452210718, so off by 4.5%

2713476 - 1170155 = 1543321 Mg C growth 126439 / 1543321 =
0.08192657263, so over-projects growth by 8%

## Calibrated vs Uncalibrated vs Observed

```{r, eval = FALSE}
for (max_scenario in 1:5) {
  plt <- cficop_cmp_carbon_summary |>
    filter(name %in% levels(name)[1:max_scenario]) |>
    ggplot(aes(Year, mean, color = name)) +
    geom_line(aes(linetype = method)) +
    geom_point() +
    theme_bw(base_size = 14) +
    theme(aspect.ratio = 1) +
    scale_color_manual(
      name = "Scenario",
      values = c(
        "Default" = series_colors[1],
        "Site Calibrated" = series_colors[2],
        "Undisturbed (31)" = series_colors[3],
        "Best Estimate" = series_colors[4],
        "All Plots" = series_colors[5]
      ),
      guide = guide_legend(order = 1)
    ) +
    scale_linetype_manual(
      name = "Method",
      values = c("Projected" = "dashed", "Measured" = "solid"),
      guide = guide_legend(order = 2)
    ) +
    scale_y_continuous(limits = c(48, 150)) +
    labs(
      title = "Projected vs. Measured Carbon"
    ) +
    xlab("Year") +
    ylab(bquote("Aboveground Live Carbon" ~(`Mg C` %.% ha^-1)))
  print(plt)
}
```

Q: Why does "Reference" go above Calibrated in this picture, but
originally it did not?

A: Because originally, we plotted the 31 plot Reference projection
against a 31 plot projection; now we have a 277 plot projection, and lo
and behold it is different!

In numbers:

```{r, eval = FALSE}
cficop_cmp_carbon_summary |>
  filter(Year == 2020 | Year == 1970)
```

How comparable are these?

```{r, eval = FALSE}
x <- cficop_none_carbon |>
  rename(Calibrated = Aboveground_Total_Live) |>
  left_join(
    cficop_dflt_carbon |>
      rename(Uncalibrated = Aboveground_Total_Live),
    by = join_by(StandID, Year)
  ) |>
  left_join(
    cficop_srvy_carbon |>
      rename(Measured = Aboveground_Total_Live),
    by = join_by(StandID, Year)
  ) |>
  left_join(
    cficop_none_bai |>
      select(StandID, Starting_Carbon, Projected_Carbon, Adjusted_AGLCF) |>
      group_by(StandID) |>
      summarize(
        Starting_Carbon = min(Starting_Carbon),
        Adjusted_AGLCF = mean(Adjusted_AGLCF),
        .groups = "keep"
      ) |>
      ungroup(),
    by = join_by(StandID)
  ) |>
  mutate(
    Adjusted = Starting_Carbon + Adjusted_AGLCF * (Year - 1970)
  ) |>
  select(-c("Starting_Carbon", "Adjusted_AGLCF")) |>
  pivot_longer(-c("StandID", "Year")) |>
  union_all(
    cfigro_srvy_carbon |>
      select(StandID, Year, Reference = Aboveground_Total_Live) |>
      pivot_longer(-c("StandID", "Year"))
  ) |>
  filter(Year == 2020) |>
  pivot_wider(id_cols = c(StandID, Year))

t.test(x$Adjusted, x$Reference, paired=TRUE)
t.test(x$Uncalibrated, x$Adjusted, paired = TRUE)
t.test(x$Calibrated, x$Adjusted, paired = TRUE)
```

## Diameter Distribution

Look at distribution of trees by diameter class across time.

Treelists are big (millions of trees), so we don't hold on to the raw
list.

```{r dimater class distribution, eval = FALSE}
add_diameter_class <- function(x) {
  x |>
    mutate(diameter_class = case_when(
      DBH < 3 ~ 0,
      DBH < 5 ~ 3,
      DBH < 9 ~ 5,
      DBH < 15 ~ 9,
      DBH < 21 ~ 15,
      DBH < 40 ~ 21,
      .default = 40
    )
    )
}

fill_dbhlist <- function(x) {
  all_diameter_classes <- c(0, 3, 5, 9, 15, 21, 40)
  all_years <- unique(x$Year)
  all_stands <- unique(x$StandID)

  everything <- tibble(diameter_class = all_diameter_classes) |>
    cross_join(tibble(StandID = all_stands)) |>
    cross_join(tibble(Year = all_years))

  x |>
    full_join(everything, by = join_by(Year, diameter_class, StandID)) |>
    replace_na(list(tpa = 0))
}

cficop_none_dbhdist <- fvs_read_output(cficop_none, "FVS_TreeList_East") |>
  filter(Year %% 10 == 0) |>
  add_diameter_class() |>
  group_by(Year, diameter_class, StandID, random_seed) |>
  summarize(tpa = sum(TPA), .groups = "keep") |>
  group_by(Year, diameter_class, StandID) |>
  summarize(tpa = mean(tpa), .groups = "keep") |>
  ungroup() |>
  fill_dbhlist()

cficop_srvy_dbhdist <- fvs_read_output(cficop_srvy, "FVS_TreeList_East") |>
  filter(Year %% 10 == 0) |>
  add_diameter_class() |>
  group_by(Year, diameter_class, StandID, random_seed) |>
  summarize(tpa = sum(TPA), .groups = "keep") |>
  group_by(Year, diameter_class, StandID) |>
  summarize(tpa = mean(tpa), .groups = "keep") |>
  ungroup() |>
  fill_dbhlist()

cficop_dflt_dbhdist <- fvs_read_output(cficop_dflt, "FVS_TreeList_East") |>
  filter(Year %% 10 == 0) |>
  add_diameter_class() |>
  group_by(Year, diameter_class, StandID, random_seed) |>
  summarize(tpa = sum(TPA), .groups = "keep") |>
  group_by(Year, diameter_class, StandID) |>
  summarize(tpa = mean(tpa), .groups = "keep") |>
  ungroup() |>
  fill_dbhlist()

cfigro_srvy_dbhdist <- fvs_read_output(cfigro_none, "FVS_TreeList_East") |>
  filter(Year %% 10 == 0) |>
  add_diameter_class() |>
  group_by(Year, diameter_class, StandID, random_seed) |>
  summarize(tpa = sum(TPA), .groups = "keep") |>
  group_by(Year, diameter_class, StandID) |>
  summarize(tpa = mean(tpa), .groups = "keep") |>
  ungroup() |>
  fill_dbhlist()
```

```{r plot dbhdist, eval = FALSE}
all_dbhdist <- cficop_none_dbhdist |>
  rename(Adjusted = tpa) |>
  left_join(
    cficop_srvy_dbhdist |>
      rename(Measured = tpa),
    by = join_by(Year, diameter_class, StandID)
  ) |>
  left_join(
    cficop_dflt_dbhdist |>
      rename(Uncalibrated = tpa),
    by = join_by(Year, diameter_class, StandID)
  ) |>
  pivot_longer(-c(Year, diameter_class, StandID)) |>    
  union_all(
    cfigro_srvy_dbhdist |>
      rename(Reference = tpa) |>
      pivot_longer(-c(Year, diameter_class, StandID))
  ) |>
  mutate(name = ordered(
    name,
    levels = c("Uncalibrated", "Calibrated", "Adjusted", "Reference", "Measured")
  ))

all_dbhdist |>
  filter(diameter_class >= 5 & Year == 1970) |>
  mutate(diameter_class = factor(diameter_class)) |>
  ggplot(aes(diameter_class, value, fill = name)) +
  geom_boxplot(outlier.size = 0.1) +
#  scale_fill_manual(values = c(
#    "Uncalibrated" = colors[1],
#    "Calibrated" = colors[2],
#    "Adjusted" = colors[3],
#    "Reference" = colors[4],
#    "Measured" = colors[5]
#  )) +
  facet_wrap("Year") +
  coord_cartesian(ylim = c(0, 200)) +
  scale_fill_discrete() +
  theme_bw() +
  ggtitle("CFI Diameter Class Frequency by Plot") +
  xlab("Diameter Class (inches)") +
  ylab(bquote(trees %.% acre^-1))
```

# Bootstrapping

## 95% CI for 2020 Carbon

Quabbin has 55760 forested acres. By FIA standards, we need one plot
per 6,000 acres, so that's 9.3 plots, round up to 10. From our 277 plots,
we'll take repeated samples of 10 plots (with replacement) to get a
distribution of the mean.

```{r, eval = FALSE}
tmp_carbon <- cficop_cmp_carbon |>
  pivot_wider(id_cols = c("StandID", "Year"))

# We can get a correlated sample for the series that use all plots
plot_carbon_all_plots <- lapply(1:10000, \(x) {
  tmp_carbon |>
    group_by(Year) |>
    sample_n(size = 10, replace = TRUE) |>
    summarize(
      Default = mean(Default),
      `Site Calibrated` = mean(`Site Calibrated`),
      `Best Estimate` = mean(`Best Estimate`),
      `All Plots` = mean(`All Plots`)
    ) |>
    mutate(Sample = x)
}) |>
  do.call(rbind, args = _)

# Add in an independent sample for Grow Only
tmp_carbon_grow_only <- tmp_carbon |>
  filter(!is.na(`Undisturbed (31)`))
plot_carbon_grow_only <- lapply(1:10000, \(x) {
  tmp_carbon_grow_only |>
    group_by(Year) |>
    sample_n(size = 10, replace = TRUE) |>
    summarize(
      `Undisturbed (31)` = mean(`Undisturbed (31)`)
    ) |>
    mutate(Sample = x)
}) |>
  do.call(rbind, args = _)

plot_carbon_bootstrap <- plot_carbon_all_plots |>
  left_join(plot_carbon_grow_only, by = join_by(Year, Sample)) |>
  pivot_longer(-c("Year", "Sample")) |>
  mutate(
    name = ordered(
      name,
      levels = c("Default", "Site Calibrated", "Undisturbed (31)", "Best Estimate", "All Plots")
    )
  )

# Compute Percentile boostrap CIs
plot_carbon_ci <- plot_carbon_bootstrap |>
  group_by(Year, name) |>
  summarize(
    lower_ci = quantile(value, probs = 0.025),
    middle_ci = quantile(value, probs = 0.5),
    upper_ci = quantile(value, probs = 0.975),
    .groups = "drop"
  ) |>
  mutate(
    dummy = "95% CI"
  )

plot_carbon_bootstrap |>
  filter(name %in% c("Default", "Site Calibrated", "Best Estimate")) |>
#  filter(name %in% c("All Plots", "Grow Only")) |>
  ggplot(aes(value, color = name)) +
  geom_density() +
  geom_vline(
    data = plot_carbon_ci |>
      filter(name %in% c("Default", "Site Calibrated", "Best Estimate")),
#      filter(name %in% c("All Plots", "Grow Only")),
    mapping = aes(xintercept = lower_ci, color = name, linetype = dummy)
  ) +
  geom_vline(
    data = plot_carbon_ci |>
      filter(name %in% c("Default", "Site Calibrated", "Best Estimate")),
#      filter(name %in% c("All Plots", "Grow Only")),
    mapping = aes(xintercept = upper_ci, color = name, linetype = dummy),
    linetype = "dashed"
  ) +
  scale_color_manual(
    name = "Scenario",
    values = c(
      "Default" = series_colors[1],
      "Site Calibrated" = series_colors[2],
      "Undisturbed (31)" = series_colors[3],
      "Best Estimate" = series_colors[4],
      "All Plots" = series_colors[5]
    ),
    guide = guide_legend(order = 1)
  ) +
  scale_linetype_manual(
    name = NULL,
    values = c("95% CI" = "dashed"),
    guide = guide_legend(order = 2)
  ) +
  facet_wrap(~Year) +
  theme_bw() + ggtitle("Bootstrap CI (sample size=10, n=10000)") +
  xlab(bquote("Aboveground Live Carbon" ~ (`Mg C` %.% hectare^-1)))
```


# Imaginary COP

We want two lines:
- Stored Carbon
- Baseline

For stored carbon, we also want bars
- Base of the bar is at the lag(1) stored carbon level
- Red bar for any growth in teh baseline
- Black bar for any growth in stored carbon - baseline
- top of the stacked bar should touch stored carbon here

## Simple

First, let's build sequences for stored carbon and baseline.

We can use the actual carbon summary for the modeled / observed,
but we need a baseline.

The 2.5%ile of the Monte Carlo for All Plots lands at about 74, whereas
the mean lands at about 98. Since actual management represents IFM already,
let's create a fake baseline by migrating from mean to the 2.5%ile over
50 years, which is subtracting 5 for each decade after 1970.

```{r}
imaginary_cop_baseline <- cficop_cmp_carbon_summary |>
  filter(name == "All Plots") |>
  mutate(
    name = "Baseline",
    mean = mean - 5 * (Year - 1970) / 10,
    ymin = ymin - 5 * (Year - 1970) / 10,
    ymax = ymax - 5 * (Year - 1970) / 10
  )
imaginary_cop <- cficop_cmp_carbon_summary |>
  union_all(imaginary_cop_baseline) |>
  mutate(
    name = ordered(
      name,
      levels = c(
        "Default", "Site Calibrated", "Undisturbed (31)",
        "Best Estimate", "All Plots", "Baseline"
      )
    )
  )
```

## Base Plot

```{r}
base_cop_plot <- imaginary_cop |>
  filter(
    name %in% c("Baseline", "Default")
  ) |>
  ggplot(aes(Year, mean, color = name)) +
  geom_line(linetype = "dashed") +
  scale_color_manual(
    name = "Scenario",
    values = c(
      "Baseline" = "darkblue",
      "Default" = series_colors[1],
      "Site Calibrated" = series_colors[2],
      "Undisturbed (31)" = series_colors[3],
      "Best Estimate" = series_colors[4],
      "All Plots" = series_colors[5]
    ),
    guide = guide_legend(order = 1)
  ) +
  scale_fill_manual(
    name = "Credit/Debit",
    values = c(
      "Credit" = "black",
      "Debit" = "red",
      "Reversal" = "cyan"
    )
  ) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1) +
  coord_cartesian(ylim = c(48, 150)) +
  ggtitle("Imaginary Quabbin IFM COP") +
  ylab(bquote("AGL" ~ (`Mg C` %.% hectare^-1)))

print(base_cop_plot)
```

## Time 0: Estimated Value

```{r}
overall_value <- imaginary_cop |>
  filter(Year == 2020) |>
  pivot_wider(id_cols = "Year", values_from = mean) |>
  mutate(
    overall_value = Default - Baseline,
    credit_debit = "Credit"
  )

base_cop_plot +
  geom_point(
    data = imaginary_cop |>
      filter(name == "All Plots" & Year == 1970)
  ) +
  geom_rect(
    data = overall_value,
    aes(
      xmin = Year - 0.5,
      xmax = Year,
      ymax = Default,
      ymin = Default - overall_value,
      fill = credit_debit
    ),
    inherit.aes = FALSE
  ) +
  labs(
    caption = paste0("Estimated Credits = ", round(overall_value$overall_value), "/ha")
  )
```

## Years 1-9 Desk Reports

```{r}
# We need annual credit / debit.
credit_debit <- imaginary_cop |>
  filter(Year <= 1980) |>
  filter(name %in% c("Default", "Baseline")) |>
  # Create a table with a column per Year
  pivot_wider(
    id_cols = "name", names_from = "Year", values_from = "mean"
  ) |>
  # Compute the annual accumulation for each
  mutate(
    starting_value = `1970`,
    annual_credits = (`1980` - `1970`) / 10
  ) |>
  distinct(name, starting_value, annual_credits) |>
  full_join(
    expand_grid(
      Year = 1971:1980,
      name = c("Default", "Baseline")
    ),
    by = join_by(name)
  ) |>
  mutate(
    value = starting_value + (Year - 1970) * annual_credits,
    credit_debit = if_else(name == "Baseline", "Debit", "Credit")
  )

total_credits <- credit_debit |>
  filter(credit_debit == "Credit") |>
  pull(annual_credits) |>
  sum()

total_debits <- credit_debit |>
  filter(credit_debit == "Debit") |>
  pull(annual_credits) |>
  sum()

credit_caption <- paste0(
  round(total_credits),
  " Credits/ha - ",
  round(total_debits),
  " Debits/ha = ",
  round(total_credits - total_debits),
  " Credits/ha"
)

base_cop_plot +
  geom_rect(
    data = credit_debit |>
      filter(name == "Default" & Year < 1980),
    aes(
      xmin = Year - 0.2,
      xmax = Year,
      ymax = value,
      ymin = value - annual_credits,
      fill = credit_debit
    ),
    inherit.aes = FALSE
  ) +
  geom_rect(
    data = credit_debit |>
      filter(name == "Baseline" & Year < 1980),
    aes(
      xmin = Year,
      xmax = Year + 0.2,
      ymax = value,
      ymin = value - annual_credits,
      fill = credit_debit
    ),
    inherit.aes = FALSE
  ) +
  geom_point(
    data = imaginary_cop |>
      filter(name == "All Plots" & Year == 1970)
  ) +
  theme(
    aspect.ratio = 1,
    axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)
  ) +
  coord_cartesian(xlim = c(1970, 1979), ylim = c(48, 75)) +
  scale_x_continuous(breaks = 1970:1980) +
  labs(
    caption = credit_caption
  )

```

## Year 10: Re-Survey

```{r}
adjusted_credits <- imaginary_cop |>
  filter(name == "All Plots" & Year == 1980) |>
  pull(mean)

credit_debit <- credit_debit |>
  arrange(name, Year) |>
  mutate(
    value = if_else(
      name == "Default" & Year == 1980,
      adjusted_credits,
      value
    ),
    annual_credits = if_else(
      name == "Default" & Year == 1980,
      value - lag(value),
      annual_credits
    )
  )


total_credits <- credit_debit |>
  filter(credit_debit == "Credit") |>
  pull(annual_credits) |>
  sum()

total_debits <- credit_debit |>
  filter(credit_debit == "Debit") |>
  pull(annual_credits) |>
  sum()

credit_caption <- paste0(
  round(total_credits),
  " Credits/ha - ",
  round(total_debits),
  " Debits/ha = ",
  round(total_credits - total_debits),
  " Credits/ha"
)


base_cop_plot +
  geom_rect(
    data = credit_debit |>
      filter(name == "Default"),
    aes(
      xmin = Year - 0.2,
      xmax = Year,
      ymax = value,
      ymin = value - annual_credits,
      fill = credit_debit
    ),
    inherit.aes = FALSE
  ) +
  geom_rect(
    data = credit_debit |>
      filter(name == "Baseline"),
    aes(
      xmin = Year,
      xmax = Year + 0.2,
      ymax = value,
      ymin = value - annual_credits,
      fill = credit_debit
    ),
    inherit.aes = FALSE
  ) +
  geom_point(
    data = imaginary_cop |>
      filter(name == "All Plots" & Year <= 1980)
  ) +
  theme(
    aspect.ratio = 1,
    axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)
  ) +
  coord_cartesian(xlim = c(1970, 1980), ylim = c(48, 75)) +
  scale_x_continuous(breaks = 1970:1980) +
  labs(
    caption = credit_caption
  )

```

## Subsequent Re-Surveys

```{r}
credit_debit <- imaginary_cop |>
  filter(name %in% c("Default", "All Plots", "Baseline")) |>
  # Create a table with a column per Year
  pivot_wider(
    id_cols = "Year", names_from = "name", values_from = "mean"
  ) |>
  arrange(Year) |>
  mutate(
    Baseline_Credits = Baseline - lag(Baseline),
    Baseline_Direction = if_else(Baseline_Credits > 0, "Debit", "Credit"),
    All_Plots_Credits = `All Plots` - lag(`All Plots`),
    All_Plots_Direction = if_else(All_Plots_Credits < 0, "Debit", "Credit"),
    Default_Credits = Default - lag(Default) - All_Plots_Credits,
    Default_Direction = "Reversal", 
  )

starting_credits <- credit_debit |>
  filter(Year == 1970) |>
  pull(`All Plots`)
ending_credits <- credit_debit |>
  filter(Year == 2020) |>
  pull(`All Plots`)
total_credits <- ending_credits - starting_credits

starting_debits <- credit_debit |>
  filter(Year == 1970) |>
  pull(Baseline)
ending_debits <- credit_debit |>
  filter(Year == 2020) |>
  pull(Baseline)
total_debits <- ending_debits - starting_debits

credit_caption <- paste0(
  round(total_credits),
  " Credits/ha - ",
  round(total_debits),
  " Debits/ha = ",
  round(total_credits - total_debits),
  " Credits/ha"
)

base_cop_plot +
  geom_rect(
    data = credit_debit |>
      filter(Year > 1970),
    aes(
      xmin = Year - 1,
      xmax = Year,
      ymax = `All Plots`,
      ymin = `All Plots` - All_Plots_Credits,
      fill = All_Plots_Direction
    ),
    inherit.aes = FALSE
  ) +
  geom_rect(
    data = credit_debit |>
      filter(Year > 1970),
    aes(
      xmin = Year,
      xmax = Year + 1,
      ymax = Baseline,
      ymin = Baseline - Baseline_Credits,
      fill = Baseline_Direction
    ),
    inherit.aes = FALSE
  ) +
  geom_rect(
    data = credit_debit |>
      filter(Year > 1990),
    aes(
      xmin = Year,
      xmax = Year + 1,
      ymax = Default,
      ymin = Default - Default_Credits,
      fill = Default_Direction
    ),
    inherit.aes = FALSE
  ) +
  geom_line(
    data = imaginary_cop |>
      filter(name == "All Plots")
  ) +
  geom_point(
    data = imaginary_cop |>
      filter(name == "All Plots")
  ) +
  labs(
    caption = credit_caption
  )
```
