---
title: "12_CFICOP"
output: html_document
---

# Setup

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(targets)
  library(tarchetypes)
  library(tidyverse)
  library(readxl)
  library(terra)
  library(sf)
  library(rnaturalearth)
  library(measurements)
  library(mlr3verse) # Machine Learning in R framework, larger universe
  library(ggspatial) # annotation_scale, annotation_north_arrow
})
tar_source()
```

## Load targets results from other documents

```{r load tar objects, eval = FALSE}
tar_load(fiadb)
tar_load(species_crosswalk)
tar_load(fvs_num_partitions)
tar_load(fvs_partition)
tar_load(fvs_randseed)
tar_load(fvsne_states)
tar_load(cfigro_bai_model)
tar_load(cfigro_estab)
tar_load(cfigro_none_growth)
```

## Preload everything from 11_CFIGrowOnly.Rmd

```{r load 11 cfigrowonly, eval = FALSE}
get_this_rmd_file() |>
  str_replace(basename(get_this_rmd_file()), "11_CFIGrowOnly.Rmd") |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```

## Preload targets results from this document

When tar_make() has already built the objects defined in this document,
run this block to pre-load them. You can then skip running ```{targets} blocks.

```{r load tar objects defined here, eval = FALSE}
get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```

# MODEL ALL* THE PLOTS!!!

We want to:
- Model all the plots grow-only
- Apply the model of residuals to get an idea of how much FVS might have over-projected
- Look at harvest, and how much carbon is actually on the landscape
- Multiply by area to get an idea of what COP would be, etc.

* We don't have harvest information except on Audrey's plots, so we'll actually
just use her 283 plots. So it's not _all_ of all the plots, but it's all the
plots for which we have good data.

## Plots

cfigro_plot has all 283 plots.

## Trees

cfigro_tree has trees for all 283 plots.

## Regeneration

We ignored harvest in CFIGRO; can we continue to ignore it?

```{r plot establishment rate, eval = FALSE}
cfigro_estab |>
  mutate(STAND_CN = as.character(STAND_CN)) |>
  group_by(STAND_CN) |>
  summarize(DENSITY = sum(DENSITY), .groups = "keep") |>
  left_join(cfigro_plot |> select(STAND_CN, STAND_ID), by = join_by(STAND_CN)) |>
  mutate(MasterPlotID = as.numeric(STAND_ID)) |>
  left_join(cfigro_plot_harvested, by = join_by(MasterPlotID)) |>
  left_join(cfigro_plot_disturbed, by = join_by(MasterPlotID)) |>
  mutate(Status = case_when(
    CFIDisturbed & CFIHarvested ~ "Disturbed & Harvested",
    CFIDisturbed ~ "Disturbed",
    CFIHarvested ~ "Harvested",
    .default = "Grow-Only"
  )) |>
  group_by(Status) |>
  mutate(Status = paste0(Status, ", n=", n())) |>
  ggplot(aes(Status, DENSITY)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("Establishment rate by plot status") +
  ylab(bquote("Establishment rate" ~ (stems %.% acre^-1 %.% year^-1)))
#  scale_y_continuous(transform = "log10")
  
```

non-Grow-Only plots show double and more the establishment rate of grow-only
plots.

Let's match by BA of the major species for establishment rate instead.

### Plot Matching

Start simple - minimize the delta in the three major species. We can leave "Other" alone.

```{targets cficop_plot_sibling, tar_simple = TRUE}
tmp_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(VisitCycle == 1970) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  # Quercus rubra and Quercus velutina are often conflated in surveys; instead
  # of treating them as two separate important species, combine them into one and
  # treat them as a single species.
  # Replace Quercus velutina (837) with Quercus rubra (833)
  mutate(
    SpeciesCode = if_else(SpeciesCode == 837, 833, SpeciesCode)
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, SCIENTIFIC_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    TPA = 5,
    BA = VisitTreeDIAM
  )

tmp_by_plot <- tmp_trees |>
  group_by(MasterPlotID) |>
  summarize(
    TPA_Plot = sum(TPA, na.rm = TRUE),
    BA_Plot = sum(BA, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ungroup() |>
  select(MasterPlotID, TPA_Plot, BA_Plot)

tmp_top_species <- c(
  "Acer rubrum", "Quercus rubra", "Pinus strobus"
)

tmp_plot_fingerprint <- tmp_trees |>
  group_by(MasterPlotID, SCIENTIFIC_NAME) |>
  summarize(
    TPA_Species = sum(TPA, na.rm = TRUE),
    BA_Species = sum(BA, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  left_join(tmp_by_plot, by = join_by(MasterPlotID)) |>
  mutate(
    TPA_Frac = TPA_Species / TPA_Plot,
    BA_Frac = BA_Species / BA_Plot
  ) |>
  select(-ends_with("_Species")) |>
#  filter(BA_Frac > 0.1) |>
  filter(SCIENTIFIC_NAME %in% tmp_top_species) |>
  group_by(MasterPlotID) |>
  mutate(
    Species = do.call(
      paste,
      append(sort(SCIENTIFIC_NAME), list(sep = ", "))
    )
  ) |>
  ungroup() |>
  mutate(SCIENTIFIC_NAME = gsub(" ", "_", SCIENTIFIC_NAME)) |>
  pivot_wider(names_from = "SCIENTIFIC_NAME", values_from = c("TPA_Frac", "BA_Frac"))

tmp_plots_grow_only <- cfigro_plot_harvested |>
  filter(!CFIHarvested) |>
  inner_join(
    cfigro_plot_disturbed |> filter(!CFIDisturbed),
    by = join_by(MasterPlotID)
  ) |>
  select(MasterPlotID)

pct_diff <- function(a, b) {
  abs(a - b) / ((a + b) / 2)
}

most_similar_plot <- function(target_plots) {
  target_plots <- target_plots |>
    # Penalize things that differ on whether they have
    # a species on the landscape
    replace_na(list(
      BA_Frac_Pinus_strobus = -1,
      BA_Frac_Acer_rubrum = -1,
      BA_Frac_Quercus_rubra = -1
    ))

  lapply(1:nrow(target_plots), \(n) {
    # Minimize RMS difference in BA
    ba_plot <- target_plots[[n, "BA_Plot"]]
    ba_pist <- target_plots[[n, "BA_Frac_Pinus_strobus"]]
    ba_acru <- target_plots[[n, "BA_Frac_Acer_rubrum"]]
    ba_quru <- target_plots[[n, "BA_Frac_Quercus_rubra"]]
    tmp_plot_fingerprint |>
      inner_join(tmp_plots_grow_only, by = join_by(MasterPlotID)) |>
      replace_na(list(
        BA_Frac_Pinus_strobus = -1,
        BA_Frac_Acer_rubrum = -1,
        BA_Frac_Quercus_rubra = -1
      )) |>
      mutate(
        BA_RMSE = sqrt(
          pct_diff(BA_Plot, ba_plot)^2 +
            (BA_Frac_Pinus_strobus - ba_pist)^2 +
            (BA_Frac_Acer_rubrum - ba_acru)^2 +
            (BA_Frac_Quercus_rubra - ba_quru)^2
        )
      ) |>
      arrange(BA_RMSE) |>
      head(1) |>
      pull(MasterPlotID)    
  }) |>
    unlist()
}

tmp_plot_fingerprint |>
  mutate(sibling_plot = most_similar_plot(tmp_plot_fingerprint)) |>
  select(MasterPlotID, sibling_plot)
```

How's that looks:

```{r plot sibling plots, eval = FALSE}
tmp_top_species <- c(
  "Acer rubrum", "Quercus rubra", "Pinus strobus"
)

tmp_species <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(VisitCycle == 1970) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |> # Only Live Trees
  mutate(
    SpeciesCode = if_else(SpeciesCode == 837, 833, SpeciesCode)
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, SCIENTIFIC_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(SCIENTIFIC_NAME %in% tmp_top_species) |>
  distinct(MasterPlotID, SCIENTIFIC_NAME) |>
  group_by(MasterPlotID) |>
  summarize(
    Species = do.call(
      paste,
      append(sort(SCIENTIFIC_NAME), list(sep = ", "))
    ),
    .groups = "drop"
  ) |>
  select(MasterPlotID, Species)

matched_plots <- cficop_plot_sibling |>
  mutate(
    status = if_else(MasterPlotID == sibling_plot, "reference", "matched")
  ) |>
  left_join(
    cfigro_plot,
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  rename(
    lat = LATITUDE,
    long = LONGITUDE
  ) |>
  mutate(
    group = as.character(sibling_plot)
  ) |>
  select(
    MasterPlotID, status, lat, long, group
  ) |>
  left_join(
    tmp_species, by = join_by(MasterPlotID)
  ) |>
  group_by(group) |>
  mutate(n = n(), group = paste0(group, ", n=", n())) |>
  ungroup() |>
  filter(n > 6)

matched_plot_group_order <- matched_plots |>
  distinct(group, n) |>
  arrange(desc(n))

matched_plots$group <- ordered(matched_plots$group, levels = matched_plot_group_order$group)

cfi_bbox <- st_as_sfc(
  st_bbox(vect(matched_plots, geom = c("long", "lat")))
)
sf_use_s2(FALSE)
st_crs(cfi_bbox) <- "WGS84"
lakes_north_america <- ne_download(
  scale = 10, type = "lakes_north_america", category = "physical"
)
lakes <- lakes_north_america[
  st_intersects(lakes_north_america, cfi_bbox, sparse = FALSE),
]
rivers_north_america <- ne_download(
  scale = 10, type = "rivers_north_america", category = "physical"
)
rivers <- rivers_north_america[
  st_intersects(rivers_north_america, cfi_bbox, sparse = FALSE),
]
ggplot() +
  geom_sf(data = lakes, fill = "blue", inherit.aes = FALSE) +
#  geom_sf(data = rivers, fill = "blue", inherit.aes = FALSE) +
  geom_point(
    data = matched_plots |> filter(status == "reference"),
    aes(long, lat, color = Species, shape = status), size = 2.5
  ) +
  geom_point(
    data = matched_plots |> filter(status == "matched"),
    aes(long, lat, color = Species, shape = status), size = 1.75
  ) +
  facet_wrap("group") +
#  scale_color_discrete(guide = "none") +
  scale_shape_manual(values = list(reference = 17, matched = 16)) +
#  annotation_scale(location = "br") +
#  annotation_north_arrow(
#    location = "tl",
#    width = unit(1, "cm"),
#    pad_x = unit(0.75, "cm"),
#    pad_y = unit(0.5, "cm")
#  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ggtitle("Matching harvested and disturbed CFI plots to grow-only plots")
```


### Copy CFIGRO Estab

We'll duplicate cfigro_estab from the reference plots to their sibling plots.

```{targets cficop_estab, tar_simple = TRUE}
# Each plot has a sibling plot;
# find the estab associated with the sibling plot.
cficop_plot_sibling |>
  left_join(
    cfigro_plot |> select(STAND_CN, MasterPlotID = STAND_ID),
    by = join_by(sibling_plot == MasterPlotID)
  ) |>
  mutate(STAND_CN = as.numeric(STAND_CN)) |>
  left_join(
    cfigro_estab,
    by = join_by(STAND_CN),
    # Each row on the left is a stand; many stands have the same STAND_CN.
    # Each row on the right is a stand-species; many rows on the right will
    # match a single stand. Therefore this is a many-to-many join.
    relationship = "many-to-many"
  ) |>
  # STAND_CN is for the sibling plot; swap that out
  # with the STAND_CN for the matched plot.
  select(-STAND_CN) |>
  left_join(
    cfigro_plot |> select(STAND_CN, MasterPlotID = STAND_ID),
    by = join_by(MasterPlotID)
  )
```

### Growth

cfigro_none_growth is tree-specific; for each tree within a plot, provide the diameter
and height 10 years from the simulation start time.

### Mortality

cfigro_none_death is tree-specific; for each tree within a plot, provide the status
at simulation start time.

## Watershed Area

I think I need to get a polygon for the watershed / districts, then use
land cover to find the forested area ... in 1970. Which I can't. Unless
they have literature that tells me.

But we can start with, what if it's 100% forested.

```{targets cficop_sam_wt, tar_simple = TRUE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  group_by(MasterPlotID) |>
  arrange(VisitCycle, MasterTreeID) |>
  filter(row_number() == 1) |>
  ungroup() |>
  left_join(
    tblSuppPlotForestrySubBasins,
    by = join_by(DWSPForestrySubWatershed == MasterForestryBasinID)
  ) |>
  group_by(EQDISTRICT, EQSUBDISTRICT) |>
  mutate(
    TotalAcres = max(DWSPOwnedAcres, na.rm = TRUE),
    NumPlots = n(),
    AcresPerPlot = TotalAcres / NumPlots,
    .groups = "keep"
  ) |>
  ungroup() |>
  select(
    EQDISTRICT,
    EQSUBDISTRICT,
    MasterPlotID,
    TotalAcres,
    NumPlots,
    AcresPerPlot
  )
```

```{r plot scaling cfi plots to watershed, eval = FALSE}
cficop_sam_wt |>
  group_by(EQSUBDISTRICT) |>
  filter(row_number() == 1) |>
  ggplot(aes(EQSUBDISTRICT, NumPlots, size = AcresPerPlot)) +
  geom_point() +
  coord_cartesian(ylim = c(0, 80)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -15, hjust = 0, vjust = 1)) +
  ggtitle("Scaling CFI plots to the Watershed") +
  xlab("Subwatershed") +
  ylab("Number of Plots")
```

```{r map cfi plot scaling to watershed, eval = FALSE}
tmp_state_map <- fvsne_states |>
  filter(STATE_NAME == "Massachusetts") |>
  rename(region = STATE_NAME) |>
  _$region |>
  map_data("state", region = _)

tmp_plot_loc <- cfigro_plot |>
  rename(lat = LATITUDE, long = LONGITUDE, group = STAND_CN) |>
  left_join(
    cficop_sam_wt |> select(MasterPlotID, EQSUBDISTRICT, AcresPerPlot),
    by = join_by(STAND_ID == MasterPlotID)
  )

ggplot(tmp_state_map, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  geom_point(
    data = tmp_plot_loc,
    aes(size = AcresPerPlot, color = EQSUBDISTRICT)
  ) +
  coord_cartesian(xlim = c(-72.45, -72.15), ylim = c(42.25, 42.55)) +
  theme_bw()
```

# CFICOP_NONE

## Run FVS

```{targets cficop_none}
tar_target(
  cficop_none,
  {
    timestep <- 10 # years; determined by FVSne variant
    fvsbin_dir <- "/fvs/fvsbin" # TODO: put these in a config file
    fvs_variant <- "fvsne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "NONE"
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = cfigro_plot, # |>
#        left_join(
#          cficop_sam_wt |> select(STAND_ID = MasterPlotID, SAM_WT = AcresPerPlot),
#          by = join_by(STAND_ID)
#        ),
      trees = cfigro_trees,
      regen = cficop_estab,
#      calibration = cfigro_none_growth,
#      calib_mort = cfigro_none_death,
#      calib_years = 10,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

## Load Results

Carbon

```{targets cficop_none_carbon, tar_simple = TRUE}
fvs_read_output(cficop_none, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_none_summary, tar_simple = TRUE}
fvs_read_output(cficop_none, "FVS_Summary2_East") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

## BA Over Time

```{r plot ba over time, eval = FALSE}
cficop_none_summary |>
  filter(Year %% 10 == 0) |>
#  group_by(Year) |>
#  summarize(BA = sum(BA) / n()) |>
#  ungroup() |>
  mutate(Series = "Projected") |>
  union_all(
    cfigro_srvy_summary |>
      filter(Year %% 10 == 0) |>
#      group_by(Year) |>
#      summarize(BA = sum(BA) / n()) |>
#      ungroup() |>
      mutate(Series = "Grow-Only")
  ) |>
  mutate(Year = as.character(Year)) |>
  ggplot(aes(Year, BA, color = Series)) +
  geom_boxplot()
```

## Carbon Over Time

```{r plot carbon over time, eval = FALSE}
cficop_none_carbon |>
#  group_by(Year) |>
#  summarize(Aboveground_Total_Live = sum(Aboveground_Total_Live) / n()) |>
#  ungroup() |>
  mutate(Series = "Projection") |>
  union_all(
    cfigro_srvy_carbon |>
#      group_by(Year) |>
#      summarize(Aboveground_Total_Live = sum(Aboveground_Total_Live) / n()) |>
#      ungroup() |>
      mutate(Series = "Reference")
  ) |>
  mutate(Year = as.character(Year)) |>
  ggplot(aes(Year, Aboveground_Total_Live, fill = Series)) +
  geom_boxplot(outlier.size = 0.5) +
  theme_bw() +
  ggtitle("CFI Grow-Only Projection") +
  ylab(bquote("Aboveground Live Carbon " ~ (`Mg C` %.% hectare^-1)))
```

## Zero-Growth Outliers

There's a handful of plotat that FVS won't grow. Let's see what's up with them.

```{r zero growth outliers, eval = FALSE}
cfigro_trees |>
  semi_join(
    cficop_none_carbon |>
      filter(Year == 2020) |>
      filter(Aboveground_Total_Live < 25) |>
      mutate(STAND_ID = as.numeric(StandID)),
    by = join_by(STAND_ID)
  ) |>
  View()
```

Looks like some trees are not getting HISTORY filled in.
