---
title: "13_CFICOP_HVST"
output: html_document
---

# Intro

Attempt to reproduce actual harvest in Quabbin, 1970 - 2000.

# Setup

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(targets)
  library(tarchetypes)
  library(tidyverse)
  library(readxl)
  library(terra)
  library(tidyterra)
  library(sf)
  library(rnaturalearth)
  library(measurements)
  library(mlr3verse) # Machine Learning in R framework, larger universe
  library(ggspatial) # annotation_scale, annotation_north_arrow
  library(ggfx) # for fancy ggplot stuff like with_blur
  library(patchwork) # for n-up ggplot
})
tar_source()

## Load targets results from other documents ----

tar_load(fiadb)
tar_load(species_crosswalk)
tar_load(fvs_num_partitions)
tar_load(fvs_partition)
tar_load(fvs_randseed)
tar_load(fvsne_states)
tar_load(cfigro_bai_model)
tar_load(cfigro_estab)
tar_load(cfigro_none_growth)
tar_load(cficop_estab)
tar_load(cficop_dflt_estab)

## Preload everything from 11_CFIGrowOnly.Rmd ----

get_this_rmd_file() |>
  str_replace(basename(get_this_rmd_file()), "11_CFIGrowOnly.Rmd") |>
  tar_objects_defined_in_rmd() |>
  tar_load()

## Preload targets results from this document ----

# When tar_make() has already built the objects defined in this document,
# run this block to pre-load them. You can then skip running ```{targets}
# blocks.

get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()

eves_colors <- c(
  "Other" = "#B4B432",
  "Fraxinus" = "#CD5C5C",
  "Betula" = "#8B4513",
  "Tsuga" = "#5F9EA0",
  "Acer" = "#DEB887",
  "Pinus" = "#556B2F",
  "Quercus" = "#D2691E"
)
```

# Observed Harvest

We're interested in a few things:

1.  What trees get cut? (when / where / species / size / tpa)
2.  What trees are left? (when / where / species / size / tpa)

NB: this does not start from cfigro_trees because those are, by definition,
live trees, and we're interested in cut / killed trees.

```{targets cficop_hvst_observed_harvest, tar_simple = TRUE}
cficop_hvst_observed_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  # Deal with species that FVS doesn't handle
  mutate(SpeciesCode = replace_values(
    SpeciesCode,
    320 ~ 317, # norway maple -> sugar maple
    402 ~ 403, # bitternut hickory -> pignut hickory
    740 ~ 743  # unknown aspen -> bigtooth aspen
  )) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(!is.na(FVS_SPCD)) |>
  # At this point we have fully-decorated tree records
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # If either of StatusB or Status6 is "C", they both are, but
  # this shows the intent
  filter(StatusB == "C" | Status6 == "C") |>
  # Accept dbh_prior or dbhcm for harvest dbh; ignore harvested
  # trees whose diameter can't be determined.
  mutate(
    dbh_harvest = coalesce(dbh_prior, dbhcm),
    DBH_MIN = floor(conv_unit(dbh_harvest, "cm", "in") / 5) * 5
  ) |>
  filter(!is.na(dbh_harvest)) |>
  mutate(
    basal_area = (conv_unit(dbh_harvest, "cm", "m") / 2)^2 * pi,
    HarvestYear = VisitCycle - YearsSinceLastCut,
    PreHarvestVisitCycle = floor(HarvestYear / 10) * 10
  )
```

## By Year

```{r, eval = FALSE}
cficop_hvst_observed_harvest |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, basal_area, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("CFI Observed Harvest (BA Per Plot)") +
  ylab(bquote("Harvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

```{r, eval = FALSE}
cficop_hvst_observed_harvest |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, tpa, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  scale_y_continuous(transform = "log10") +
  ggtitle("CFI Observed Harvest (TPA Per Plot)") +
  ylab(bquote("Harvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

## By Genus

```{r, eval = FALSE}
cficop_hvst_observed_harvest |>
  mutate(
    Diameter_Class = paste0(DBH_MIN, " - ", DBH_MIN + 5),
    VisitCycle = paste0(VisitCycle - 10, " - ", VisitCycle),
    GENUS = fct_other(GENUS, keep = c("Acer", "Betula", "Pinus", "Quercus", "Tsuga"))
  ) |>
  group_by(VisitCycle, GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(basal_area, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  complete(
    VisitCycle, Diameter_Class, GENUS,
    fill = list(Measured_Trees = 1, Measured_BA = 0)
  ) |>
  pivot_longer(cols = Measured_BA) |>
  ggplot(aes(Diameter_Class, value, fill = GENUS, group = GENUS)) +
  geom_col(position = "dodge") +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  facet_wrap(~VisitCycle) +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Basal Area Harvested") +
  xlab("Diameter (inches)") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Basal Area harvested" ~(m^2)))
```

```{r, eval = FALSE}
cficop_hvst_observed_harvest |>
  mutate(
    Diameter_Class = paste0(DBH_MIN, " - ", DBH_MIN + 5),
    VisitCycle = paste0(VisitCycle - 10, " - ", VisitCycle),
    GENUS = fct_other(GENUS, keep = c("Acer", "Betula", "Pinus", "Quercus", "Tsuga"))
  ) |>
  group_by(VisitCycle, GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(basal_area, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  complete(
    VisitCycle, Diameter_Class, GENUS,
    fill = list(Measured_Trees = 1, Measured_BA = 0)
  ) |>
  pivot_longer(cols = Measured_Trees) |>
  ggplot(aes(Diameter_Class, value, fill = GENUS, group = GENUS)) +
  geom_col(position = "dodge") +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  facet_wrap(~VisitCycle) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Harvested") +
  xlab("Diameter (inches)") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Trees harvested" ~(trees)))
```

## Multiple Harvest

Are any stands harvested in multiple cycles?

```{r, eval = FALSE}
cficop_hvst_observed_harvest |>
  distinct(MasterPlotID, VisitCycle) |>
  group_by(MasterPlotID) |>
  summarise(num_harvests = n()) |>
  ungroup() |>
  filter(num_harvests > 1) |>
  nrow()
```

So, 54 plots are harvested more than once.

Any stands harvested multiple times in a single cycle?

```{r, eval = FALSE}
cficop_hvst_observed_harvest |>
  distinct(MasterPlotID, VisitCycle, HarvestYear) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(num_harvests = n(), .groups = "keep") |>
  ungroup() |>
  filter(num_harvests > 1) |>
  nrow()
```

No; when a plot is harvested during a visitcycle, all trees are
harvested in the same year.

# Observed Unharvest

> It's not enough for a plot to be harvested and some trees left over;
> we need to know what kind of trees were harvested, and how many of
> the kinds harvested were left over.

When a plot is harvested, what is left unharvested?

To figure this out, we'll:

- Filter to plot/cycles with harvest
- sum the ba of what's left

Sawtimber / Poletimber
- Hardwood break diameter = 11 (or)
- Softwood break diameter = 9

```{targets cficop_hvst_observed_remainders, tar_simple = TRUE}
cficop_hvst_observed_remainders <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  # Deal with species that FVS doesn't handle
  mutate(SpeciesCode = replace_values(
    SpeciesCode,
    320 ~ 317, # norway maple -> sugar maple
    402 ~ 403, # bitternut hickory -> pignut hickory
    740 ~ 743  # unknown aspen -> bigtooth aspen
  )) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(!is.na(FVS_SPCD)) |>
  # Live trees, not recruits
  filter(StatusB == "L" | Status6 == "L") |>
  mutate(
    dbh_harvest = coalesce(dbh_prior, dbhcm),
    DBH_MIN = floor(conv_unit(dbh_harvest, "cm", "in") / 5) * 5
  ) |>
  filter(!is.na(dbh_harvest)) |>
  inner_join(
    # Which sizes of which species are harvested on which plots, in
    # which visit cycles? Keep only those ones.
    cficop_hvst_observed_harvest |>
      distinct(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN, HarvestYear),
    by = join_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi
  )
```

## By Year

```{r, eval = FALSE}
cficop_hvst_observed_remainders |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, basal_area, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  ggtitle("CFI Observed Remainders (BA Per Plot)") +
  ylab(bquote("Harvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```


```{r, eval = FALSE}
cficop_hvst_observed_remainders |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, tpa, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  scale_y_continuous(transform = "log10") +
  ggtitle("CFI Observed Remainders (TPA Per Plot)") +
  ylab(bquote("Harvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

## By Genus

```{r, eval = FALSE}
cficop_hvst_observed_remainders |>
  mutate(
    Diameter_Class = paste0(DBH_MIN, " - ", DBH_MIN + 5),
    VisitCycle = paste0(VisitCycle - 10, " - ", VisitCycle),
    GENUS = fct_other(GENUS, keep = c("Acer", "Betula", "Pinus", "Quercus", "Tsuga"))
  ) |>
  group_by(VisitCycle, GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(basal_area, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  complete(
    VisitCycle, Diameter_Class, GENUS,
    fill = list(Measured_Trees = 1, Measured_BA = 0)
  ) |>
  pivot_longer(cols = Measured_Trees) |>
  ggplot(aes(Diameter_Class, value, fill = GENUS, group = GENUS)) +
  geom_col(position = "dodge") +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  facet_wrap(~VisitCycle) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Unharvested") +
  xlab("Diameter (inches)") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Trees harvested" ~(trees)))
```

# Regeneration

## 1980 Establishment

The 1970 inventory does not have saplings or seedlings, which will
cause under-recruitement in 1980.

Rather than try to turn 1980 ingrowth into 1970 saplings, which would
require figuring out a survival rate, we can represent the 1970 sapling
population by actual 1980 ingrowth. By 1980, the 1970 seedlings in the
background establishment should have created a sapling pool that FVS can
track, so we don't need to do anything beyond 1980. So let's add 1980
establishment as a one-off. We use actual trees, with height when
available.

Note that we do not subtract the background regeneration from this;
background regeneration is put in as seedlings, these trees are put
in as saplings (with height) and so will establish much sooner.

```{targets cficop_hvst_estab_1980, tar_simple = TRUE}
# Find 1970 trees; we'll remove these from 1980
tmp_trees_1970 <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(
    VisitCycle == 1970
  ) |>
  select(
    MasterPlotID, MasterTreeID
  )

# Find 1980 trees, less those that were around in 1970
cficop_hvst_estab_1980 <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  # Live trees and recruits
  filter(StatusB == "L" | StatusB == "R") |>
  # Only trees that were around in 1980
  filter(
    VisitCycle == 1980
  ) |>
  # Not trees that were around in 1970
  anti_join(
    tmp_trees_1970,
    by = join_by(MasterPlotID, MasterTreeID)
  ) |>
  # Deal with species that FVS doesn't handle
  mutate(SpeciesCode = replace_values(
    SpeciesCode,
    320 ~ 317, # norway maple -> sugar maple
    402 ~ 403, # bitternut hickory -> pignut hickory
    740 ~ 743  # unknown aspen -> bigtooth aspen
  )) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(!is.na(FVS_SPCD)) |>
  # Tree height - some trees have it, some don't.
  # Impute mean height of those that have it for missing values
  group_by(FVS_SPCD) |>
  mutate(
    # A height of 0 is a missing value; fix that.
    VisitTreeTotalHeight = if_else(
      VisitTreeTotalHeight == 0, NA, VisitTreeTotalHeight
    ),
    VisitTreeTotalHeight = floor(if_else(
      is.na(VisitTreeTotalHeight),
      mean(VisitTreeTotalHeight, na.rm = TRUE),
      VisitTreeTotalHeight
    )),
    DENSITY = 5  # trees per acre for one tree on a 1/5 acre plot
  ) |>
  ungroup() |>
  # Consolidate redundant records
  group_by(MasterPlotID, VisitCycle, FVS_SPCD, VisitTreeTotalHeight) |>
  summarize(DENSITY = sum(DENSITY), .groups = "drop") |>
  ungroup() |>
  # We need STAND_CN for the 1970 stand
  left_join(
    cfigro_plot |>
      filter(INV_YEAR == 1970) |>
      select(STAND_ID, STAND_CN),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  select(
    STAND_ID = MasterPlotID,
    STAND_CN,
    YEAR = VisitCycle,
    SPECIES = FVS_SPCD,
    DENSITY,
    HEIGHT = VisitTreeTotalHeight
  )
```

## Post-Harvest Establishment

It can take a while for post-harvest trees to reach establishment
diameter. For each plot harvested, look at establishment in the
two inventories post-harvest, and turn those into post-harvest
establishment.

It doesn't matter what species was harvested; harvest'll open up space
for everything.

As we did for 1980 establishment, we do not subtract the background
regeneration from this; background regeneration is put in as seedlings,
these trees are put in as saplings (with height) and so will establish
much sooner.

```{targets cficop_hvst_estab_post_harvest, tar_simple = TRUE}
 tmp_post_harvest_plots <- cficop_hvst_observed_harvest |>
  distinct(MasterPlotID, VisitCycle) |>
  # Add in the post-post-harvest visit cycle
  mutate(VisitCycleAfter = VisitCycle + 10) |>
  pivot_longer(cols = c(VisitCycle, VisitCycleAfter), values_to = "VisitCycle") |>
  # This will have introduced duplicates if a plot was harvested in two
  # consecutive VisitCycles
  distinct(MasterPlotID, VisitCycle) |>
  # Don't do 1980; those are captured above
  filter(VisitCycle != 1980)

cficop_hvst_estab_post_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  # Only recruits
  filter(StatusB == "R") |>
  # Only trees that establish in post-harvest plots
  semi_join(
    tmp_post_harvest_plots,
    by = join_by(MasterPlotID, VisitCycle)
  ) |>
  # Deal with species that FVS doesn't handle
  mutate(SpeciesCode = replace_values(
    SpeciesCode,
    320 ~ 317, # norway maple -> sugar maple
    402 ~ 403, # bitternut hickory -> pignut hickory
    740 ~ 743  # unknown aspen -> bigtooth aspen
  )) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(!is.na(FVS_SPCD)) |>
  # Tree height - some trees have it, some don't.
  # Impute mean height of those that have it for missing values
  group_by(FVS_SPCD) |>
  mutate(
    # A height of 0 is a missing value; fix that.
    VisitTreeTotalHeight = if_else(
      VisitTreeTotalHeight == 0, NA, VisitTreeTotalHeight
    ),
    VisitTreeTotalHeight = floor(if_else(
      is.na(VisitTreeTotalHeight),
      mean(VisitTreeTotalHeight, na.rm = TRUE),
      VisitTreeTotalHeight
    )),
    DENSITY = 5  # trees per acre for one tree on a 1/5 acre plot
  ) |>
  # Replase NaN HEIGHTs with 0 to avoid errors
  replace_na(list(VisitTreeTotalHeight = 0)) |> 
  ungroup() |>
  # Consolidate redundant records
  group_by(MasterPlotID, VisitCycle, FVS_SPCD, VisitTreeTotalHeight) |>
  summarize(DENSITY = sum(DENSITY), .groups = "drop") |>
  ungroup() |>
  # We need STAND_CN for the appropriate stand
  # We need STAND_CN for the 1970 stand
  left_join(
    cfigro_plot |>
      filter(INV_YEAR == 1970) |>
      select(STAND_ID, STAND_CN),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  select(
    STAND_ID = MasterPlotID,
    STAND_CN,
    YEAR = VisitCycle,
    SPECIES = FVS_SPCD,
    DENSITY,
    HEIGHT = VisitTreeTotalHeight
  )
```

## Combined Establishment

Combine:

- establishment from Fred Hunt's thesis
- 1980 establishment
- post-harvest establishemtn

to create one big table of establishment for the harvest scenarios.

```{targets cficop_hvst_estab, tar_simple = TRUE}
cficop_hvst_estab <- cficop_dflt_estab |>
  mutate(
    STAND_ID = NA,
    HEIGHT = 0
  ) |>
  union_all(
    cficop_hvst_estab_1980
  ) |>
  union_all(
    cficop_hvst_estab_post_harvest
  )
```

# Harvest

FVS provides the ThinDBH keyword to schedule thinning from throughout a
dbh range for any or all species to a residual basal area (BA) or trees
per acre (TPA) target.

We will provide multiple ThinDBH keywords per stand per timestep,
describing the target residual trees per acre for that species for that
dbh range for that stand for that havest; the rest will be harvested.
Ideally, we'd specify the number of trees to remove, but between
ingrowth, mortality, and multiple rounds of harvest, we don't actually
know how many trees FVS will have in the stand at any given time, so we
can't compute the residual number of trees that would cause the right
number to be removed. We assume that harvest is to a target resudual
TPA, and that the amount of harvest is the free variable.

## Harvest Prescription

Let's build a list of the number of live trees left after each harvest.

We need:

- In what plot, in what year, what species was harvested
- for the plot/year/species combo, how many trees were left behind
- lump the trees left behind into dbh classes for ThinDBH

Trees on the landscape grow at different rates than in FVS; it's not
possible to align actual trees in 2" diameter classes with projected
trees in 2" diameter classes over several decades, so we use much wider
5" diameter classes to align more actual harvest with simulated harvest.

```{targets cficop_hvst_tpa, tar_simple = TRUE}
# Find the trees left after each harvest
cficop_hvst_tpa <- cficop_hvst_observed_remainders |>
  group_by(MasterPlotID, VisitCycle, FVS_SPCD, DBH_MIN, HarvestYear) |>
  summarize(
    TPA = n() * 5,
    .groups = "drop"
  ) |>
  # Add back in the things that were harvested down to 0 TPA
  # This will also add HarvestYear to the things harvested to >0 TPA
  full_join(
    cficop_hvst_observed_harvest |>
      distinct(MasterPlotID, VisitCycle, FVS_SPCD, DBH_MIN, HarvestYear),
    by = join_by(MasterPlotID, VisitCycle, FVS_SPCD, DBH_MIN)
  ) |>
  # Never harvest things with DBH < 5"
  filter(
    DBH_MIN > 0
  ) |>
  mutate(
    HarvestYear = coalesce(HarvestYear.x, HarvestYear.y),
    DBH_MAX = DBH_MIN + 5,
    # Harvests added by the full_join will nave TPA==NA;
    # there were not trees of this species left in this size range,
    # so set the residual TPA to 0.
    TPA = coalesce(TPA, 0),
    BA = 0 # We're not using BA; TODO make this NA
  ) |>
  # For prescription-based harvest, our schema is:
  # STAND_CN, TREE_CN, PREV_TRE_CN, YEAR, PRESCRIPTION
  # For DBH-based harvest, our schema is:
  # STAND_CN, YEAR, DBH_MIN, DBH_MAX, SPCD, TPA, BA
  left_join(
    cfigro_plot |>
      filter(INV_YEAR == 1970) |>
      select(STAND_CN, STAND_ID),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  # Remove records for cohorts that weren't harvested
  filter(
    !is.na(HarvestYear)
  ) |>
  select(
    STAND_CN,
    YEAR = HarvestYear,
    DBH_MIN,
    DBH_MAX,
    SPCD = FVS_SPCD,
    TPA,
    BA
  ) |>
  # Arrange this for humans to be able to digest
  arrange(YEAR, SPCD, DBH_MIN)
```

# Run FVS

Make a best effort at replicating what actually happend in Quabbin.

## A Note on SAM_WT

There are a few things that play together here:

- Aspect, slope, lat/lon, forest type, etc are per-stand
- A stand can have multiple plots
- The STAND_CN is really about the stand
  - PLOT ID is observed by FVS to figure out how many plots are in the stand
- Plot design is specified for all plots in the stand
  - Fixed radius, number of subplots, etc.
- The acreage of the STAND is specified
- SAM_WT is (the inverse of) how each plot in the stand should be weighted
  - If there are 10 plots in a 1000 acre stand, then SAM_WT is 10
    which will cause each plot to be "expansion factor"ed to 100 acres
- Since FVS can figure all this out, we can omit SAM_WT

```{targets cficop_hvst}
tar_target(
  cficop_hvst,
  {
    timestep <- 10 # years; determined by FVSne variant
    fvsbin_dir <- "~/bin" # TODO: put these in a config file
    fvs_variant <- "FVSne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "HVST"
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = cfigro_plot |>
        # Filter out the 6 plots that do not get a sibling
        semi_join(
          cficop_plot_sibling,
          by = join_by(STAND_ID == MasterPlotID)
        ) |>
        mutate(
          ASPECT = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
          # We don't use microplot data, but there are 4x6' radius
          # microplots
          INV_PLOT_SIZE = 385/4
        ),
      trees = cfigro_trees |>
        filter(DIAMETER > 0),
      regen = cficop_hvst_estab,
      harvest = cficop_hvst_tpa,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

# Fault Analysis

What errors and warnings do we get?
```{r cficop_hvst_warnings, eval = FALSE}
title <- "CFICarbonOffsetProgram"
mgmt_id <- "HVST"

cficop_hvst_outfiles <- list.files(
  path = file.path(
    "data/fvs",
    paste0("FVS_", title, "_", mgmt_id)
  ),
  pattern = paste0("FVS_", title, "_", mgmt_id, "_P.*_R", fvs_randseed[1], ".out"),
  full.names = TRUE
)

cficop_hvst_warnings <- lapply(cficop_hvst_outfiles, \(filename) {
  partition = filename |>
    sub("^.*_P(\\d+)_R.*$", "\\1", x = _)
  read_lines(filename) |>
    as.data.frame(nm = "text") |>
    # Filter to just initial STAND ID and WARNING lines
    filter(
      grepl("^ +STAND ID=", text) |
        grepl("WARNING:", text) |
        grepl("ERROR:", text)
    ) |>
    # Parse out the STAND ID and fill it into the following lines
    mutate(
      STAND_ID = case_when(
        grepl("^ +STAND ID=", text) ~ sub("^ +STAND ID= *([^ ]+) .*$", "\\1", text)
      ),
      partition = !!partition
    ) |>
    fill(STAND_ID) |>
    # Remove everything but warnings (which now have STAND_ID)
    filter(
      grepl("WARNING:", text) |
        grepl("ERROR:", text)
    )
}) |>
  bind_rows() |>
  # See below about ignoring this warning
  filter(text != "********   WARNING:  DATE OF DISTURBANCE IS IGNORED.") |>
  # See below about ignoring this warning
  filter(text != " *** FFE MODEL WARNING: NO INITIAL BASAL AREA")

cficop_hvst_warnings |>
  nrow()
```

## DATE OF DISTURBANCE IS IGNORED

We are using IF / THEN to trigger establishment based on cycle,
not disturbance. So this is correct, the date of disturbance is
ignored, and that's what we intended.

## FFE MODEL WARNING: NO INITIAL BASAL AREA

This error does not mean there is no initial basal area; it means
that the cover type could not be determined from crown area.

From fire/ne/fmcba.f:
```{fortran, eval = FALSE}
C----------
C   IF THERE ARE NO TREES (COVTYP=0) IN CYCLE 1,
C   SET COVTYP TO A DEFAULT BALSAM FIR COVER. AFTER THE
C   FIRST CYCLE, USE THE PREVIOUS COVER TYPE IF NO COVER IS
C   PRESENT.
C----------
      IF (COVTYP .EQ. 0) THEN
        IF (IYR .EQ. IY(1)) THEN
             WRITE (JOSTND,"(/1X,'*** FFE MODEL WARNING: NO INITIAL ',
     &       'BASAL AREA',
     &       /1X,'*** COVER TYPE SET TO BALSAM FIR',/1X)")
            CALL RCDSET (2,.TRUE.)
            COVTYP = 1
        ELSE
          COVTYP = OLDCOVTYP
        ENDIF
      ENDIF

```
This code was introduced in 2022; no idea why the error started to
show up with the 2026 release, maybe has to do with how COVTYP is
determined.

Anyway, we can ignore these errors, especially since cover type is
not used by the NE variant.

## Harvest Directives

FVS will cancel a harvest if it can't meet the target.

How many THINDBH directives did we give?

```{r, eval = FALSE}
cficop_hvst_directives <- lapply(cficop_hvst_outfiles, \(filename) {
  partition = filename |>
    sub("^.*_P(\\d+)_R.*$", "\\1", x = _)
  read_lines(filename) |>
    as.data.frame(nm = "text") |>
    # Filter to just initial STAND ID and WARNING lines
    filter(
      grepl("^ +STAND ID=", text) |
        grepl("BASE     THINDBH .* (DONE|DELETED)", text)
    ) |>
    # Parse out the STAND ID and fill it into the following lines
    mutate(
      STAND_ID = case_when(
        grepl("^ +STAND ID=", text) ~ sub("^ +STAND ID= *([^ ]+) .*$", "\\1", text)
      ),
      partition = !!partition
    ) |>
    fill(STAND_ID) |>
    # Remove everything but ThinDBH directives
    filter(
      grepl("BASE     THINDBH", text)
    ) |>
    mutate(
      STATUS = case_when(
        grepl("DONE", text) ~ "DONE",
        grepl("DELETED", text) ~ "DELETED"
      )
    )
}) |>
  bind_rows()

cficop_hvst_directives |>
  nrow()
```

> 1161

## Canceled Harvest

```{r, eval = FALSE}
cficop_hvst_directives |>
  filter(STATUS == "DELETED") |>
  nrow()
```

> 341

that's about 30% of harvest is deleted or canceled.


# Results

## Load Results

Carbon

```{targets cficop_hvst_carbon, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_hvst_summary, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_Summary2") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

Cut trees

```{targets cficop_hvst_cutlist, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_CutList") |>
  filter(random_seed == min(random_seed))
```

## Projected vs Observed TPA

```{r, eval = FALSE}
observed_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    Scenario = "Observed",
    MasterPlotID = as.character(MasterPlotID)
  ) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(Tpa = n() * 5, .groups = "drop") |>
  select(
    StandID = MasterPlotID,
    VisitCycle,
    Observed = Tpa
  )

observed_trees |>
  full_join(
    cficop_hvst_summary |>
      filter(Year %% 10 == 0) |>
      select(StandID, VisitCycle = Year, Projected = Tpa),
    by = join_by(StandID, VisitCycle)
  ) |>
  mutate(VisitCycle = as.character(VisitCycle)) |>
  complete(StandID, VisitCycle, fill = list(Observed = 0, Projected = 0)) |>
  pivot_longer(cols = c(Observed, Projected)) |>
  ggplot(aes(VisitCycle, value, fill = name)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  ggtitle("Projected vs. Observed Tree Density") +
  ylab(bquote(trees %.% acre^-1))

observed_trees |>
  full_join(
    cficop_hvst_summary |>
      filter(Year %% 10 == 0) |>
      select(StandID, VisitCycle = Year, Projected = Tpa),
    by = join_by(StandID, VisitCycle)
  ) |>
  complete(StandID, VisitCycle, fill = list(Observed = 0, Projected = 0)) |>
  mutate(
    VisitCycle = as.character(VisitCycle),
    Surplus = Projected - Observed
  ) |>
  ggplot(aes(VisitCycle, Surplus)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  ggtitle("Surplus trees by inventory") +
  ylab(bquote(trees %.% acre^-1))
```

So overall TPA in the modeled plots goes up a lot compared to what's observed.
A lot of this is likely because the model can see all trees down to seedlings,
whereas CFI inventory only captures trees >5.6"

## Meas vs. Proj harvested

### By Genus

How does harvest compare to survey?

```{r hvst vs srvy, eval = FALSE}
measured_harvest <- cficop_hvst_observed_harvest |>
  group_by(GENUS) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(basal_area, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

projected_harvest <- cficop_hvst_cutlist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS)
  ) |>
  filter(Measured_Trees > 100 | Projected_Trees > 100) |>
  ggplot(aes(fct_reorder(GENUS, desc(Measured_Trees)))) +
  geom_col(
    aes(y = Measured_BA, fill = "Measured", group = "Measured"),
    width = 0.4,
    position = position_nudge(x = -0.2)
  ) +
  geom_col(
    aes(y = Projected_BA, fill = "Projected", group = "Projected"),
    width = 0.4,
    position = position_nudge(x = 0.2)
  ) +
  theme_bw() +
  #facet_wrap("name") +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Basal Area Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Genus") +
  ylab(bquote("Basal Area harvested" ~(m^2))) +
  labs(fill = NULL)

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS)
  ) |>
  filter(Measured_Trees > 100 | Projected_Trees > 100) |>
  ggplot(aes(fct_reorder(GENUS, desc(Measured_Trees)))) +
  geom_col(
    aes(y = Measured_Trees, fill = "Measured", group = "Measured"),
    width = 0.4,
    position = position_nudge(x = -0.2)
  ) +
  geom_col(
    aes(y = Projected_Trees, fill = "Projected", group = "Projected"),
    width = 0.4,
    position = position_nudge(x = 0.2)
  ) +
  theme_bw() +
  #facet_wrap("name") +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Genus") +
  ylab(bquote("Trees harvested" ~ (trees))) +
  labs(fill = NULL)
```


### By Diameter Class


```{r, eval = FALSE}
measured_harvest <- cficop_hvst_observed_harvest |>
  mutate(
    Diameter_Class = paste0(DBH_MIN, " - ", DBH_MIN + 5)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(basal_area, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

projected_harvest <- cficop_hvst_cutlist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    Diameter_Class = floor(DBH / 5) * 5,
    Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5),
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  pivot_longer(cols = c(Measured_BA, Projected_BA)) |>
  filter(!is.na(value)) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Basal Area Harvested") +
  xlab("Diameter (inches)") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Basal Area harvested" ~(m^2)))

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  complete(
    Diameter_Class, GENUS,
    fill = list(Measured_Trees = 1, Projected_Trees = 1)
  ) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  filter(!is.na(value)) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Diameter (inches)") +
  ylab("Trees Harvested (trees)")
```

Looks OK, but we're generally harvesting less than measured.

Let's see what size trees FVS thought were there at harvest.

### Pre-Harvest Trees

Treelist is big, ~4G

```{r, eval = FALSE}
cficop_hvst_treelist <- fvs_read_output(cficop_hvst, "FVS_TreeList") |>
  filter(random_seed == min(random_seed))
```

```{r, eval = FALSE}
pre_harvest_year <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    HarvestYear = VisitCycle - YearsSinceLastCut,
    PreHarvestYear = floor(HarvestYear / 10) * 10
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(!is.na(dbh_prior) & dbh_prior > 0 & StatusB == "C" | Status6 == "B") |>
  filter(!is.na(PreHarvestYear)) |>
  distinct(MasterPlotID, PreHarvestYear)

measured_pre_harvest_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  semi_join(
    pre_harvest_year,
    by = join_by(MasterPlotID, VisitCycle == PreHarvestYear)
  ) |>
  mutate(
    Diameter_Class = floor(conv_unit(dbh_prior, "cm", "in") / 5) * 5,
    ba = pi * (conv_unit(dbh_prior, "cm", "m") / 2)^2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(
    !is.na(ba) & ba > 0 & (
      StatusB == "L" |
        StatusB == "R" |
        Status6 == "L" |
        Status6 == "R"
    )
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(ba, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

projected_pre_harvest_trees <- cficop_hvst_treelist |>
   mutate(StandID = as.numeric(StandID)) |>
  semi_join(
    pre_harvest_year,
    by = join_by(StandID == MasterPlotID, Year == PreHarvestYear)
  ) |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    Diameter_Class = floor(DBH / 5) * 5,
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0 & DBH >= 5) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_pre_harvest_trees |>
  full_join(
    projected_pre_harvest_trees,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  # There's one 45" maple
  filter(Diameter_Class < 30) |>
  mutate(
        Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5)
  ) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    Measured_BA = sum(Measured_BA, na.rm = TRUE),
    Projected_BA = sum(Projected_BA, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ungroup() |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30", "30 - 35")
  )) |>
  pivot_longer(cols = c(Measured_BA, Projected_BA)) |>
  replace_na(list(value = 0)) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Pre-Harvest Basal Area") +
  xlab("Diameter (inches)") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Basal Area" ~(m^2)))


measured_pre_harvest_trees |>
  full_join(
    projected_pre_harvest_trees,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  # There's one 45" maple
  filter(Diameter_Class < 30) |>
  mutate(
        Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5)
  ) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30", "30 - 35")
  )) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  replace_na(list(value = 1)) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Pre-Harvest Trees") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Diameter (inches)") +
  ylab("Trees (trees)")
```

### By Year

```{r, eval = FALSE}
measured_harvest <- cficop_hvst_observed_harvest |>
  group_by(GENUS, VisitCycle) |>
  summarize(
    Observed = n() * 5, # TPA
    .groups = "keep"
  ) |>
  ungroup() |>
  select(GENUS, VisitCycle, Observed)

projected_harvest <- cficop_hvst_cutlist |>
  mutate(
    # CAVEAT:
    # FVS Cuts trees in year 0 of a cycle;
    # CFI VisitCycle is year 10 of a cycle
    # Therefore FVS year cut is CFI VisitCycle - 10
    VisitCycle = Year + 10
  ) |>
  left_join(
    species_crosswalk |> select(FVS_SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesFVS == FVS_SPCD)
  ) |>
  group_by(GENUS, VisitCycle) |>
  summarize(
    Projected = sum(TPA, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup() |>
  select(GENUS, VisitCycle, Projected)

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS, VisitCycle)
  ) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  pivot_longer(cols = c(Observed, Projected)) |>
  replace_na(list(value = 0)) |>
  mutate(Year = sprintf("%d", VisitCycle)) |>
  mutate(value = if_else(value == 0, 1, value)) |>
  ggplot(aes(Year, value, fill = name)) +
  geom_col(position = "dodge") +
#  scale_y_continuous(transform = "log10") +
  facet_wrap(~GENUS) +
  theme_bw() +
  ggtitle("Projected Vs Observed Trees Harvested") +
  ylab("Number of Trees Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1))

```

### Measured vs. Projected Post-Harvest Trees

cficop_hvst_tpa is the tpa that we requested that FVS leave behind.

TODO: Ugh; stand 119700442 was harvested twice between 1970 and 1980. Let's
ignore it for now; probably the right thing to do is to consolidate into
a single harvest - the earlier one to give regeneration more time, or
the later one to give regeneration less time.

```{r hvst vs srvy remainders, eval = FALSE}
measured_remainders <- cficop_hvst_tpa |>
  filter(!is.na(SPCD)) |>
  left_join(
    species_crosswalk |> select(FVS_SPCD, GENUS),
    by = join_by(SPCD == FVS_SPCD)
  ) |>
  group_by(GENUS, DBH_MIN) |>
  summarize(
    Measured_Trees = sum(TPA), # TPA is already scaled to full acres
    .groups = "drop"
  )

post_harvest_inventories <- cficop_hvst_tpa |>
  mutate(YEAR = 10 * ceiling(YEAR / 10)) |>
  distinct(STAND_CN, YEAR) |>
  left_join(cfigro_plot |> distinct(STAND_CN, STAND_ID), by = join_by(STAND_CN)) |>
  select(
    StandID = STAND_ID,
    Year = YEAR
  ) |>
  mutate(StandID = as.character(StandID))

post_harvest_treelist <- fvs_read_output(cficop_hvst, "FVS_TreeList") |>
  filter(random_seed == min(random_seed)) |>
  semi_join(post_harvest_inventories, by = join_by(StandID, Year))
  
projected_remainders <- post_harvest_treelist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    DBH_MIN = floor(DBH / 5) * 5
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, DBH_MIN) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees.
    Projected_Trees = sum(TPA),
    .groups = "keep"
  ) |>
  ungroup()

measured_remainders |>
  full_join(
    projected_remainders,
    by = join_by(GENUS, DBH_MIN)
  ) |>
  filter(DBH_MIN > 0) |>
  mutate(DBH_MIN = ordered(DBH_MIN)) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  filter(!is.na(GENUS) & !is.na(value)) |>
  # Remove uninteresting genera
  group_by(GENUS) |>
  filter(sum(value) > 100) |>
  ungroup() |>
  ggplot(aes(DBH_MIN, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap("GENUS") +
  scale_x_discrete(drop = FALSE) +
  #scale_y_continuous(transform = "log10") +
  #coord_cartesian(ylim = c(0, 100)) +
  ggtitle("Quabbin CFI Post-Harvest Trees") +
  xlab("5\" Diameter Class") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Post-Harvest Trees" ~ (trees)))
```

