# CFICOP_HVST

Let's look at the distribution of harvest, and what we'll need to model.

## Observed Harvest

We're interested in a few things:

1.  What trees get cut? (when / where / species / size / tpa)
2.  What trees are left? (when / where / species / size / tpa)

```{r observed harvest ba, eval = FALSE}
tmp_observed_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(StatusB == "C" | Status6 == "B") |>
  filter(!is.na(basal_area)) |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

tmp_observed_harvest |>
  ggplot(aes(VisitCycle, basal_area, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("Harvested BA Per Plot") +
  ylab(bquote("Harvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

tmp_observed_harvest |>
  ggplot(aes(VisitCycle, tpa, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
#  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Harvested TPA Per Plot") +
  ylab(bquote("Harvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

```

Are any stands harvested in multiple cycles?

```{r observed multiple harvest, eval = FALSE}
tmp_observed_harvest |>
  distinct(MasterPlotID, VisitCycle) |>
  group_by(MasterPlotID) |>
  summarise(num_harvests = n()) |>
  ungroup() |>
  filter(num_harvests > 1) |>
  nrow()
```

So, 54 plots are harvested more than once.

Any stands harvested multiple times in a single cycle?

```{r observed multiple year cut, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    YearCut = VisitCycle - YearsSinceLastCut
  ) |>
  distinct(MasterPlotID, VisitCycle, YearCut) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(num_harvests = n(), .groups = "keep") |>
  ungroup() |>
  filter(num_harvests > 1) |>
  nrow()
```

Stand 10202 was harvested in 1980 and 1982.

No; when a plot is harvested during a visitcycle, all trees are
harvested in the same year.

## Harvest Parameters

### Who got left behind?

Unharvested:

- Filter to plot/cycles with harvest
- sum the ba of what's left

```{r observed unharvested ba, eval = FALSE}

# Which plots are harvested, in which visit cycles?

tmp_harvest_history <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_abp(cfiabp_trees) |>
  # cfi_harvested() |> # Can't use this, doesn't retain VisitCycle
  group_by(MasterPlotID, VisitCycle) |>
  summarize(
    CFIHarvested = any(CutSinceLastVisit, na.rm = TRUE),
    .groups = "drop"
  )

# What trees are left behind after harvest?

tmp_observed_remainders <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | Status6 == "L") |>
  filter(!is.na(basal_area)) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  mutate(
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other"),
    VisitCycle = ordered(VisitCycle)
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    basal_area = sum(basal_area, na.rm = TRUE),
    tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

tmp_observed_remainders |>
  ggplot(aes(VisitCycle, basal_area, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 2)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("Unharvested BA Per Plot") +
  ylab(bquote("Unarvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

tmp_observed_remainders |>
  ggplot(aes(VisitCycle, tpa, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(5, 100)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
#  scale_y_continuous(transform = "log10") +
  ggtitle("Unharvested TPA Per Plot") +
  ylab(bquote("Unarvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

tmp_observed_remainders |>
  ggplot(aes(VisitCycle, tpa)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(1, 50)) +
  theme_bw() +
#  scale_fill_manual(values = eves_colors) +
#  scale_y_continuous(transform = "log10") +
  ggtitle("Unharvested TPA Per Plot") +
  ylab(bquote("Unarvested TPA" ~ (trees %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

```

### Harvest Efficiency

What fraction of trees were harvested?

> Take the 1980 pulse (delta in 6" fron 1970 - 1980?) and
> turn those into extra trees in 1970? 4" trees? 2/3 the height? 2/3 the
> DBH? (2/3 because 4 is 2/3 of 6) also, i feel like i was looking at a
> spot in harvest where we're looking at fvs trees without summing tpa
> Also also, instead of ThinDBH, try ThinBTA - thin from below to a TPA
> target. - DBH range - everything - TPA target - what's left on the
> plot - Efficiency - calculate ratio of initial to residual tpa in each
> diameter class, take the max? mean? median? ThinBTA doesn't allow
> species selection; ThinDBH does, can maybe use ThinDBH with Cuteff to
> achieve the same goal? Note that ThinDBH ignores CutEFF if target TPA
> or BA is specified. ThinQFA also allows species to be specified; can
> thin to TPA, specified in supplemental record.

```{r observed harvested ba, eval = FALSE}
# What got harvested?

tmp_observed_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(StatusB == "C" | Status6 == "B") |>
  filter(!is.na(basal_area)) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    hvst_basal_area = sum(basal_area, na.rm = TRUE),
    hvst_tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

# Which plots are harvested, in which visit cycles?

tmp_harvest_history <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_abp(cfiabp_trees) |>
  # cfi_harvested() |> # Can't use this, doesn't retain VisitCycle
  group_by(MasterPlotID, VisitCycle) |>
  summarize(
    CFIHarvested = any(CutSinceLastVisit, na.rm = TRUE),
    .groups = "drop"
  )

# What trees are left behind after harvest?

tmp_observed_remainders <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | Status6 == "L") |>
  filter(!is.na(basal_area)) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    rmdr_basal_area = sum(basal_area, na.rm = TRUE),
    rmdr_tpa = n() * 5,
    .groups = "keep"
  ) |>
  ungroup()

tmp_harvest_fraction <- tmp_observed_harvest |>
  left_join(
    tmp_observed_remainders,
    by = join_by(MasterPlotID, VisitCycle, GENUS)
  ) |>
  replace_na(list(
    rmdr_basal_area = 0,
    rmdr_tpa = 0,
    hvst_basal_area = 0,
    hvst_tpa = 0
  )) |>
  mutate(
    tpa_fraction = hvst_tpa / (hvst_tpa + rmdr_tpa),
    ba_fraction = hvst_basal_area / (hvst_basal_area + rmdr_basal_area)
  ) |>
  mutate(
    group = sprintf("%s %s", VisitCycle, GENUS)
  )

tmp_harvest_fraction |>
  ggplot(aes(VisitCycle, tpa_fraction, fill = GENUS, group = group)) +
  geom_boxplot() +
  scale_fill_manual(values = eves_colors)


```

### 1980 Pulse

Let's see how many extra trees recruit in 1980

```{r excess recruits in 1980, eval = FALSE}
excess_recruits <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees and recruits
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    diameter_class = as.integer(VisitTreeDIAM / 2) * 2
  ) |>
  filter(
    diameter_class == 6
  ) |>
  group_by(MasterPlotID, FVS_SPCD, VisitCycle) |>
  summarize(recruits = n(), .groups = "keep") |>
  left_join(
    cficop_estab |> select(MasterPlotID, SPECIES, DENSITY),
    by = join_by(MasterPlotID, FVS_SPCD == SPECIES)
  ) |>
  mutate(excess_recruits = recruits - DENSITY) |>
  ungroup() |>
  filter(!is.na(excess_recruits) & excess_recruits >= 0) |>
  mutate(Year = ordered(VisitCycle))

excess_recruits |>
  ggplot(aes(Year, excess_recruits, group = Year)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 5))
```

Rather than derive a principled way to turn 1980 ingrowth into 1970
saplings, we can represent the 1970 sapling population by actual 1980
ingrowth. By 1980, the 1970 seedling should have created a sapling pool
that FVS can track, so we don't need to do anything beyond 1980. So
let's add 1980 establishment as a 1-off. We use actual trees and height
when available.

```{targets cficop_hvst_estab, tar_simple = TRUE}
# Find 1970 trees
tmp_trees_1970 <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(
    VisitCycle == 1970
  ) |>
  select(
    MasterPlotID, MasterTreeID
  )

tmp_estab_1980 <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  # Remove any trees that were around in 1970
  anti_join(
    tmp_trees_1970,
    by = join_by(MasterPlotID, MasterTreeID)
  ) |>
  mutate(SpeciesCode = replace_values(
    SpeciesCode,
    320 ~ 317, # norway maple -> sugar maple
    402 ~ 403, # bitternut hickory -> pignut hickory
    740 ~ 743  # unknown aspen -> bigtooth aspen
  )) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(!is.na(FVS_SPCD)) |>
  # Live trees and recruits
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    diameter_class = as.integer(VisitTreeDIAM / 2) * 2
  ) |>
  filter(
    VisitCycle == 1980
  ) |>
  group_by(FVS_SPCD) |>
  # Impute mean height for missing values
  mutate(
    VisitTreeTotalHeight = if_else(
      is.na(VisitTreeTotalHeight) | VisitTreeTotalHeight == 0,
      mean(VisitTreeTotalHeight, na.rm = TRUE),
      VisitTreeTotalHeight
    ),
    DENSITY = 5  # trees per acre
  ) |>
  left_join(
    cfigro_plot |> select(STAND_ID, STAND_CN),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  select(
    STAND_ID = MasterPlotID,
    STAND_CN,
    YEAR = VisitCycle,
    SPECIES = FVS_SPCD,
    DENSITY,
    HEIGHT = VisitTreeTotalHeight
  )

cficop_hvst_estab <- cficop_dflt_estab |>
  mutate(
    STAND_ID = NA,
    HEIGHT = 0
  ) |>
  union_all(
    tmp_estab_1980
  )
```

### Diameter Class

What's the distribution of trees left behind, by diameter class?

```{r harvest by diameter class, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = ordered(as.integer(conv_unit(dbh_prior, "cm", "in") / 5) * 5)
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | StatusB == "R" | Status6 == "L" | Status6 == "R") |>
  filter(!is.na(diameter_class)) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(basal_area = sum(basal_area, na.rm = TRUE), .groups = "keep") |>
  ungroup() |>
  ggplot(aes(VisitCycle, basal_area, fill = diameter_class)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() +
  scale_fill_discrete() +
  ggtitle("Unharvested BA Per Plot by 5\" diameter class") +
  ylab(bquote("Unharvested BA" ~ (m^2))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

```{r remaining tpa by diameter class, eval = FALSE}
remaining_tpa_by_diameter_class <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = as.integer(conv_unit(dbh_prior, "cm", "in") / 2) * 2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | StatusB == "R" | Status6 == "L" | Status6 == "R") |>
  filter(!is.na(diameter_class)) |>
  filter(diameter_class <= 40) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(
    tpa = n() * 5, .groups = "keep"
  ) |>
  ungroup()

remaining_tpa_by_diameter_class |>
  mutate(
    VisitCycle = ordered(VisitCycle),
    diameter_class = ordered(diameter_class)
  ) |>
  ggplot(aes(VisitCycle, tpa, fill = diameter_class)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(5, 100)) +
  theme_bw() +
  scale_y_continuous(transform = "log2") +
  scale_fill_discrete() +
  ggtitle("Unharvested TPA Per Plot by 2\" diameter class") +
  ylab(bquote("Unharvested TPA" ~ (stems %.% acre^-1))) +
  xlab("Visit Cycle (Number of Plots Harvested)")

ideal_tpa_by_diameter_class <- tibble(
  diameter_class = seq(3, 35, 2),
  tpa = sapply(0:16, \(i) as.integer(42 / 1.4^i))
)


remaining_tpa_by_diameter_class |>
  ggplot(aes(diameter_class, tpa)) +
  geom_boxplot(
    aes(group = diameter_class),
    width = 0.7,
    outlier.size = 0.1
  ) +
  geom_step(
    data = ideal_tpa_by_diameter_class,
    color = "red"
  ) +
  scale_x_continuous(breaks = seq(2,40,2)) +
  coord_cartesian(xlim = c(6, 20), ylim = c(0, 100)) +
  annotate("text", x = 18, y = 0, label = "DBq, q=1.4", color = "red") +
  theme_bw() +
  ggtitle("Unharvested TPA Per Plot by 2\" diameter class") +
  ylab(bquote("Unharvested TPA" ~ (stems %.% acre^-1))) +
  xlab("Diameter Class")

```

```{r remaining tpa by diameter class 2, eval = FALSE}
remaining_tpa_by_diameter_class <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = as.integer(conv_unit(dbh_prior, "cm", "in") / 2) * 2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(StatusB == "L" | StatusB == "R" | Status6 == "L" | Status6 == "R") |>
  filter(!is.na(diameter_class)) |>
  filter(diameter_class <= 40) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(
    tpa = n() * 5, .groups = "keep"
  ) |>
  ungroup()

starting_tpa <- remaining_tpa_by_diameter_class |>
  filter(diameter_class == 6) |>
  ungroup() |>
  select(MasterPlotID, VisitCycle, starting_tpa = tpa) |>
  mutate(
    starting_tpa = 25 * floor(starting_tpa / 25),
    tpa_class = if_else(
      starting_tpa < 75, as.character(starting_tpa),
      "75+"
    )
  )

ideal_tpa_by_diameter_class <- lapply(
  c(0, 25, 50, 75), \(x) {
    tibble(
      diameter_class = seq(5, 35, 2),
      tpa = sapply(0:15, \(i) as.integer((x + 12.5) / 1.4^i)),
      starting_tpa = x
    ) |>
      mutate(
        tpa_class = if_else(
          starting_tpa < 75, as.character(starting_tpa),
          "75+"
        )
      )
  }
) |>
  do.call(bind_rows, args = _)

remaining_tpa_by_diameter_class |>
  left_join(
    starting_tpa, by = join_by(MasterPlotID, VisitCycle)
  ) |>
  replace_na(list(tpa_class = "0")) |>
  mutate(group = paste(tpa_class, diameter_class)) |>
  ggplot(aes(diameter_class, tpa)) +
  geom_boxplot(
    aes(fill = tpa_class, group = group),
    outlier.size = 0.1
  ) +
  geom_step(
    data = ideal_tpa_by_diameter_class,
    aes(color = tpa_class),
    show.legend = FALSE
  ) +
  scale_x_continuous(breaks = seq(2,40,2)) +
  coord_cartesian(xlim = c(6, 20), ylim = c(0, 100)) +
  theme_bw() +
  labs(fill = "6\" TPA") +
  ggtitle("Unharvested TPA Per Plot by 2\" diameter class") +
  ylab(bquote("Unharvested TPA" ~ (stems %.% acre^-1))) +
  xlab("Diameter Class")
```

### Before / After Ratio

```{r harvest ratio by diameter class, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi,
    diameter_class = ordered(as.integer(conv_unit(dbh_prior, "cm", "in") / 5) * 5)
  ) |>
  # Live trees, not recruits
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(
    StatusB == "L" | StatusB == "R" | StatusB == "C" |
      Status6 == "L" | Status6 == "R" | Status6 == "C"
  ) |>
  filter(!is.na(diameter_class)) |>
  group_by(VisitCycle) |>
  mutate(VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")) |>
  ungroup() |>
  group_by(MasterPlotID, VisitCycle, diameter_class) |>
  summarize(
    pre_harvest_ba = sum(basal_area, na.rm = TRUE),
    harvested_ba = sum(if_else(StatusB == "C" | Status6 == "C", 0, basal_area), na.rm = TRUE),
    harvest_fraction = harvested_ba / pre_harvest_ba,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, harvest_fraction, fill = diameter_class)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() +
  scale_fill_discrete() +
  ggtitle("Harvested BA Ratio Per Plot by 5\" diameter class") +
  ylab(bquote("Harvested BA" ~ (fraction))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

```{r harvest ratio by genus, eval = FALSE}
qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  left_join(tmp_harvest_history, by = join_by(MasterPlotID, VisitCycle)) |>
  filter(CFIHarvested) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  mutate(
    # Judgment: basal_area intended to be left in harvest is better
    # represented by ba PRIOR to harvest, than POST harvest. We don't
    # have BA at the time of harvest, which would be ideal.
    basal_area = (conv_unit(dbh_prior, "cm", "m") / 2)^2 * pi
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  # Live trees, not recruits
  filter(
    StatusB == "L" | StatusB == "R" | StatusB == "C" |
      Status6 == "L" | Status6 == "R" | Status6 == "C"
  ) |>
  filter(!is.na(basal_area)) |>
  group_by(VisitCycle) |>
  mutate(
    VisitCycle = paste0(VisitCycle, " (n=", n_distinct(MasterPlotID), ")")
  ) |>
  ungroup() |>
  mutate(
    VisitCycle = ordered(VisitCycle),
    GENUS = fct_lump_n(GENUS, n = 5, w = basal_area, other_level = "Other")
  ) |>
  group_by(MasterPlotID, VisitCycle, GENUS) |>
  summarize(
    pre_harvest_ba = sum(basal_area, na.rm = TRUE),
    harvested_ba = sum(if_else(StatusB == "C" | Status6 == "C", 0, basal_area), na.rm = TRUE),
    harvest_fraction = harvested_ba / pre_harvest_ba,
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(VisitCycle, harvest_fraction, fill = GENUS)) +
  geom_boxplot(outlier.size = 0.1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_bw() +
  scale_fill_manual(values = eves_colors) +
  ggtitle("Harvested BA Ratio Per Plot by Genus") +
  ylab(bquote("Harvested BA" ~ (fraction))) +
  xlab("Visit Cycle (Number of Plots Harvested)")
```

### Theoretical Q

```{r theoretical description of what is on the landscape, eval = FALSE}
x <- tibble(
  diameter_class = seq(6, 40, 2),
  tpa = sapply(0:17, \(i) (40 / 1.4^i))
)

x |>
  ggplot(aes(diameter_class, tpa)) +
  geom_line() +
  geom_point()
```

## Harvest using ThinDBH

After talking with folks, we thought we were going to harvest by:

- basal area removed, by
- diameter class (6"), by
- species, by
- plot, by
- time

FVS provides the THINDBH keyword to schedule thinning from throughout a
dbh range for any or all species to a residual basal area (BA) or trees per
acre (TPA) target.

We will provide multiple THINDBH keywords per stand per timestep,
describing the target residual trees per acre of that species in that
dbh range for that stand; the rest will be harvested. Ideally, we'd
specify the number of trees to remove, but between ingrowth, mortality,
and multiple rounds of harvest, we don't actually know how many trees
FVS will have in the stand at any given time, so we can't compute the
residual number of trees that would cause the right number to be
removed.

We could also do this by telling FVS the fraction of trees to remove; we
know how many were harvested and when, but we don't know how many trees
were on the landscape at the time of harvest, so we can't compute the
correct fraction. By looking at cficop_dflt_treelist and
cficop_none_treelist, we can get a pretty good idea of what FVS thinks
is there.

### Harvest List

Let's build a list of the number of live trees after each harvest:

We need:

- In what plot, in what year, what species was harvested
- for the plot/year/species combo, what trees were left behind?
- lump those left behind into dbh classes for ThinDBH
- Trees on the landscape grow at different rates than in FVS;

it's too hard to try to hit 2" diameter classes, so we use much wider
6" diameter classes to try to catch more trees.

> Why 6, making the 5" BRK_DBH awkward? Can we use 5" diameter classes?

```{targets cficop_hvst_tpa, tar_simple = TRUE}
tmp_harvested <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    YearCut = VisitCycle - YearsSinceLastCut
  ) |>
  filter(
    !is.na(YearCut) & !is.na(dbh_prior)
  ) |>
  # When harvest and inventory occur in the same year,
  # harvest might occur before or after the inventory.
  # To simplify things, we'll make all harvest happen
  # before the inventory.
  # Select the post-harvest inventory as the VisitCycle
  mutate(
    VisitCycle = ceiling(YearCut / 10) * 10,
    DBH_MIN = floor(conv_unit(dbh_prior, "cm", "in") / 5) * 5
  ) |>
  group_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN) |>
  summarize(YearCut = max(YearCut), .groups = "drop")

cficop_hvst_tpa <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  filter(!is.na(dbh_prior)) |>
  filter(MasterPlotID == 10226) |>
  mutate(
    DBH_MIN = floor(conv_unit(dbh_prior, "cm", "in") / 5) * 5
  ) |>
  semi_join(
    tmp_harvested,
    by = join_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN)
  ) |>
  group_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN) |>
  summarize(
    TPA = n() * 5, # Each plot is 1/5 acre
    .groups = "drop"
  ) |>
  # Add back in the things that were harvested down to 0 TPA
  # This will also add YearCut to the things harvested to >0 TPA
  full_join(
    tmp_harvested,
    by = join_by(MasterPlotID, VisitCycle, SpeciesCode, DBH_MIN)
  ) |>
  # The harvest blocks added back will have TPA == NA
  mutate(
    DBH_MAX = DBH_MIN + 5,
    TPA = if_else(is.na(TPA), 0, TPA),
    BA = 0
  ) |>
  # Except seedlings / saplings - remove those from the harvest list
  filter(
    DBH_MIN > 0
  ) |>
  # For prescription-based harvest, our schema is:
  # STAND_CN, TREE_CN, PREV_TRE_CN, YEAR, PRESCRIPTION
  # For DBH-based harvest, our schema is:
  # STAND_CN, YEAR, DBH_MIN, DBH_MAX, SPCD, TPA, BA
  mutate(
    YEAR = YearCut
  ) |>
  mutate(SpeciesCode = replace_values(
    SpeciesCode,
    320 ~ 317, # norway maple -> sugar maple
    402 ~ 403, # bitternut hickory -> pignut hickory
    740 ~ 743  # unknown aspen -> bigtooth aspen
  )) |>
  left_join(
    species_crosswalk |> select(SPCD, FVS_SPCD),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  filter(!is.na(FVS_SPCD)) |>
  left_join(
    cfigro_plot |> filter(INV_YEAR == 1970) |> select(STAND_CN, STAND_ID),
    by = join_by(MasterPlotID == STAND_ID)
  ) |>
  select(
    STAND_CN,
    YEAR = YearCut,
    DBH_MIN,
    DBH_MAX,
    SPCD = FVS_SPCD,
    TPA,
    BA
  )
```

### Post-Harvest Regeneration

TODO: Post-harvest there's a bunch of regeneration; find it and insert it.

### FvS_StandInit and FVS_PlotInit

From DBSUserGuide.pdf, Table 3.0.1, we learn the following about table
structure:

-   Stand_CN - optional, but used by queries, so we'll include it
-   Stand_ID - required
-   StandPlot_ID - optional; plot identification code; required by the
    user interface when populating stand lists from the plot initiation
    table (PlotInit)
-   Variant - required; set to "ne"
-   Inv_Year - required; set to 1970
-   Latitude - optional
-   Longitude - optional
-   Location - recommended; nearest national forest; takes precedence
    over Region, forest, District, and Compartment
-   Apsect - recommended; (0 = no meaningful aspect; 360 = north)
-   Slope - recommended
-   ElevFt - recommended; Elevation in feet
-   Basal_Area_Factor - required
-   Inv_Plot_Size - required
-   Brk_DBH - required
-   Num_Plots - recommended; if 0, the number of plots in the stand is
    determined by counting the numbers [sic] of unique plot
    identification codes on the tree record data

All other fields are optional.

### FVS_TreeInit

From DBSUserGuide.pdf, Table 3.0.2, we learn the following about table
structure:

-   Stand_CN - optional, but used by queries, so we'll include it
-   Stand_ID - required
-   StandPlot_ID - required to run plots as stands
-   Plot_ID - recommended; plot identification, should be unique for an
    individual stand but can be repeated for plots in different stands
-   Tree_ID - optional
-   Tree_Count - recommended - the number of trees represented by the
    tree record; if left blank FVS assumes the tree record represents a
    single sample tree
-   Species - required
-   DBH or Diameter - required
-   Ht - recommended; total height in feet
-   CrRatio - recommended; values 1-9 are read as crown ratio codes;
    10-99 as percent

All other fields are optional.

### Generate Tables for UI

```{r, eval = FALSE}
blank_db <- DBI::dbConnect(
  RSQLite::SQLite(),
  file.path("data/fvs/BlankDatabase.db"),
  flags = RSQLite::SQLITE_RO
)

fvs_data_db <- DBI::dbConnect(
  RSQLite::SQLite(),
  file.path("data/fvs/FVS_Data.db"),
  flags = RSQLite::SQLITE_RWC
)

# There are three tables: FVS_PlotInit, FVS_StandInit, and FVS_TreeInit

# FVS_PlotInit
# We aren't supporting by-plot
DBI::dbWriteTable(
  fvs_data_db,
  "FVS_PlotInit",
  tbl(blank_db, "FVS_PlotInit") |> collect(),
  overwrite = TRUE
)

# FVS_StandInit
# Grab the structure of the table from the blank database,
# and make cfigro_plot conform to it
fvs_standinit <- bind_rows(
  tbl(blank_db, "FVS_StandInit") |>
    collect() |>
    mutate(
      Stand_CN = character()
    ),
  cfigro_plot |>
    filter(INV_YEAR == 1970) |>
    mutate(
      Stand_ID = as.character(STAND_ID),
      Variant = "ne",
      Aspect = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
      INV_PLOT_SIZE = 385/4, # We don't use microplot data, but there
      # are 4x6' radius microplots.
      Sam_Wt = 10
    ) |>
    select(
      Stand_CN = STAND_CN,
      Stand_ID,
      StandPlot_ID = STAND_ID,
      Variant,
      Inv_Year = INV_YEAR,
      Latitude = LATITUDE,
      Longitude = LONGITUDE,
      Region = REGION,
      Forest = FOREST,
      Aspect,
      Slope = SLOPE,
      ElevFt = ELEVFT,
      Basal_Area_Factor = BASAL_AREA_FACTOR,
      Inv_Plot_Size = INV_PLOT_SIZE,
      Brk_DBH = BRK_DBH,
      Num_Plots = NUM_PLOTS,
      Sam_Wt
  )
)

DBI::dbWriteTable(
  fvs_data_db,
  "FVS_StandInit",
  fvs_standinit,
  overwrite = TRUE
)

# FVS_TreeInit
# Grab the structure of the table from the blank database,
# and make cfigro_trees conform to it
fvs_treeinit <- bind_rows(
  tbl(blank_db, "FVS_TreeInit") |>
    collect(),
  cfigro_trees |>
    semi_join(fvs_standinit |> distinct(Stand_CN), by = join_by(STAND_CN == Stand_CN)) |>
    mutate(
      Stand_ID = as.character(STAND_ID),
      Variant = "ne",
      Aspect = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
      Tree_Count = 1
    ) |>
    select(
      Stand_CN = STAND_CN,
      Stand_ID,
      Plot_ID = STAND_ID,
      Tree_CN = TREE_CN,
#      Tree_ID = TREE_ID,
      Tree_Count,
      Species = SPECIES,
      Diameter = DIAMETER,
      Ht = HT,
      Slope = SLOPE,
      Aspect = ASPECT,
      TopoCode = TOPOCODE
  )
)

DBI::dbWriteTable(
  fvs_data_db,
  "FVS_TreeInit",
  fvs_treeinit,
  overwrite = TRUE
)

DBI::dbDisconnect(blank_db)
DBI::dbDisconnect(fvs_data_db)
```

## Run FVS

Make our best effort at replicating what actually happend in Quabbin.

```{targets cficop_hvst}
tar_target(
  cficop_hvst,
  {
    timestep <- 10 # years; determined by FVSne variant
    fvsbin_dir <- "~/bin" # TODO: put these in a config file
    fvs_variant <- "FVSne"      # TODO: put these in a config file
    data_dir <- "data/fvs"
    title <- "CFICarbonOffsetProgram"
    mgmt_id <- "HVST"
    project_dir <- file.path(data_dir, paste0("FVS_", title, "_", mgmt_id))
    if (!dir.exists(project_dir)) {
      dir.create(project_dir)
    }
    
    fvs_run(
      fvsbin_dir = fvsbin_dir,
      fvs_variant = fvs_variant,
      project_dir = project_dir,
      fiadb = NULL,
      title = title,
      mgmt_id = mgmt_id,
      stands = cfigro_plot |>
        # Filter out the 6 plots that do not get a sibling
        semi_join(
          cficop_plot_sibling,
          by = join_by(STAND_ID == MasterPlotID)
        ) |>
        mutate(
          ASPECT = if_else(SLOPE > 0 & ASPECT == 0, 360, ASPECT),
          INV_PLOT_SIZE = 385/4, # We don't use microplot data, but
          # there are 4x6' radius microplots
          SAM_WT = 10
        ),
      trees = cfigro_trees |>
        filter(DIAMETER > 0),
      regen = cficop_hvst_estab, # cficop_dflt_estab,  
      harvest = cficop_hvst_tpa,
      num_partitions = fvs_num_partitions,
      partition = fvs_partition,
      random_seed = fvs_randseed
    )
  },
  # iteration = "vector" branches execution for each partition value (see below)
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here.
  # cross() ensures that every combination of values for its arguments is processed
  # map() distributes each value of its argument to a separate sub-target (branch)
  # so cross(randseed, map(partition)) will run each partition in a separate branch,
  # and each branch will run with each value of randseed
  pattern = cross(fvs_randseed, map(fvs_partition))
)
```

## Load Results

Carbon

```{targets cficop_hvst_carbon, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_Carbon") |>
  group_by(StandID, Year) |> # Combine results from different random seeds
  summarize(
    Aboveground_Total_Live = mean(Aboveground_Total_Live),
    .groups = "keep"
  ) |>
  ungroup()
```

Summary

```{targets cficop_hvst_summary, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_Summary2") |>
  group_by(StandID, Year) |>
  summarize(
    BA = mean(BA),
    Tpa = mean(Tpa),
    .groups = 'keep'
  ) |>
  ungroup()
```

Cut trees

```{targets cficop_hvst_cutlist, tar_simple = TRUE}
fvs_read_output(cficop_hvst, "FVS_CutList") |>
  filter(random_seed == min(random_seed))
```

### Parse Problems

What errors and warnings do we get?
```{r cficop_hvst_warnings, eval = FALSE}
title <- "CFICarbonOffsetProgram"
mgmt_id <- "HVST"

cficop_hvst_outfiles <- list.files(
  path = file.path(
    "data/fvs",
    paste0("FVS_", title, "_", mgmt_id)
  ),
  pattern = paste0("FVS_", title, "_", mgmt_id, "_P.*_R", fvs_randseed[1], ".out"),
  full.names = TRUE
)

cficop_hvst_warnings <- lapply(cficop_hvst_outfiles, \(filename) {
  partition = filename |>
    sub("^.*_P(\\d+)_R.*$", "\\1", x = _)
  read_lines(filename) |>
    as.data.frame(nm = "text") |>
    # Filter to just initial STAND ID and WARNING lines
    filter(
      grepl("^ +STAND ID=", text) |
        grepl("WARNING:", text) |
        grepl("ERROR:", text)
    ) |>
    # Parse out the STAND ID and fill it into the following lines
    mutate(
      STAND_ID = case_when(
        grepl("^ +STAND ID=", text) ~ sub("^ +STAND ID= *([^ ]+) .*$", "\\1", text)
      ),
      partition = !!partition
    ) |>
    fill(STAND_ID) |>
    # Remove everything but warnings (which now have STAND_ID)
    filter(
      grepl("WARNING:", text) |
        grepl("ERROR:", text)
    )
}) |>
  bind_rows() |>
  # See below about ignoring this warning
  filter(text != "********   WARNING:  DATE OF DISTURBANCE IS IGNORED.") |>
  # See below about ignoring this warning
  filter(text != " *** FFE MODEL WARNING: NO INITIAL BASAL AREA")
```

Find the distinct warnings
```{r distinct fvs warnings, eval = FALSE}
fvs_warnings_distinct <- fvs_warnings |>
  # Warnings have some leading text, the tag WARNING:,
  # then descriptive text; if there's a number in the descriptive text,
  # it's probably pulled from the data, so we chop off the descriptive
  # text starting at the first number to find the truly distinct warnings
  mutate(
    PROBLEM = case_when(
      grepl("WARNING:", text) ~ sub("^[*. ]*(.*WARNING: [^0-9]+)([0-9].*)?$", "\\1", text),
      grepl("ERROR:", text) ~ sub("^[*. ]*(.*ERROR: [^0-9]+)([0-9].*)?$", "\\1", text)
    ),
    PROBLEM = sub("\\s+$", "", PROBLEM)
  ) |>
  distinct(PROBLEM) |>
  arrange(PROBLEM)
knitr::kable(fvs_warnings_distinct)
```

### DATE OF DISTURBANCE IS IGNORED

We are using IF / THEN to trigger establishment based on cycle,
not disturbance. So this is correct, the date of disturbance is
ignored, and that's what we intended.

### FFE MODEL WARNING: NO INITIAL BASAL AREA

This error does not mean there is no initial basal area; it means
that the cover type could not be determined from crown area.

From fire/ne/fmcba.f:
```{fortran, eval = FALSE}
C----------
C   IF THERE ARE NO TREES (COVTYP=0) IN CYCLE 1,
C   SET COVTYP TO A DEFAULT BALSAM FIR COVER. AFTER THE
C   FIRST CYCLE, USE THE PREVIOUS COVER TYPE IF NO COVER IS
C   PRESENT.
C----------
      IF (COVTYP .EQ. 0) THEN
        IF (IYR .EQ. IY(1)) THEN
             WRITE (JOSTND,"(/1X,'*** FFE MODEL WARNING: NO INITIAL ',
     &       'BASAL AREA',
     &       /1X,'*** COVER TYPE SET TO BALSAM FIR',/1X)")
            CALL RCDSET (2,.TRUE.)
            COVTYP = 1
        ELSE
          COVTYP = OLDCOVTYP
        ENDIF
      ENDIF

```
This code was introduced in 2022; no idea why the error started to
show up with the 2026 release, maybe has to do with how COVTYP is
determined.

Anyway, we can ignore these errors, especially since cover type is
not used by the NE variant.

### Did it do what we asked?

FVS will cancel a harvest if it can't meet the target.

How many THINDBH directives did we give?

```{bash, eval = FALSE}
# I've done this in R before, but don't feel like going and figuring
# it out again
grep "ThinDBH" data/fvs/FVS_CFICarbonOffsetProgram_HVST/*R43517.key | wc -l
```

> 3873

```{bash, eval = FALSE}
grep "DELETED OR CANCELED" data/fvs/FVS_CFICarbonOffsetProgram_HVST/*R43517.out | grep THINDBH | wc -l
```
> 2242 with 1980 compensation, 2252 without - so canceled 8 more harvests

So, a little over half the harvest was canceled.
This seems to be:
1. Trees where actual diameter growth was different enough from modeled
   diameter growth that it ended up in a different diameter class than
   harvest specified (lessened by using 6" diameter classes), and
2. At the low end, different ingrowth rates between actual and modeled.

Is it a problem?

Let's compare harvested and remaining TPA and BA.

> Deleted: Analysis of MasterPlotID 10153

### Projected vs Observed TPA

```{r, eval = FALSE}
observed_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(cfi_status_live(VisitTreeStatusCode)) |>
  mutate(
    Scenario = "Observed",
    MasterPlotID = as.character(MasterPlotID)
  ) |>
  group_by(MasterPlotID, VisitCycle) |>
  summarize(Tpa = n() * 5, .groups = "drop") |>
  select(
    StandID = MasterPlotID,
    Year = VisitCycle,
    Observed = Tpa
  )

observed_trees |>
  full_join(
    cficop_hvst_summary |>
      filter(Year %% 10 == 0) |>
      select(StandID, Year, Projected = Tpa),
    by = join_by(StandID, Year)
  ) |>
  mutate(Year = sprintf("%d", Year)) |>
  pivot_longer(cols = c(Observed, Projected)) |>
  ggplot(aes(Year, value, fill = name)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  ggtitle("Projected vs. Observed Tree Density") +
  ylab(bquote(trees %.% acre^-1))

observed_trees |>
  full_join(
    cficop_hvst_summary |>
      filter(Year %% 10 == 0) |>
      select(StandID, Year, Projected = Tpa),
    by = join_by(StandID, Year)
  ) |>
  mutate(
    Year = sprintf("%d", Year),
    Surplus = Projected - Observed
  ) |>
  ggplot(aes(Year, Surplus)) +
  geom_boxplot(outlier.size = 0.1) +
  theme_bw() +
  ggtitle("Surplus trees by inventory") +
  ylab(bquote(trees %.% acre^-1))
```

So overall TPA in the modeled plots goes up a lot compared to what's observed.
A lot of this is likely because the model can see all trees down to seedlings,
whereas CFI inventory only captures trees >5.6"

### Meas vs. Proj harvested

#### By Genus

How does harvest compare to survey?

```{r hvst vs srvy, eval = FALSE}
measured_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    Year = VisitCycle - YearsSinceLastCut,
    ba = pi * (conv_unit(dbh_prior, "cm", "m") / 2)^2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(!is.na(ba) & ba > 0 & StatusB == "C" | Status6 == "B") |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  group_by(GENUS) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(ba, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

projected_harvest <- cficop_hvst_cutlist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS)
  ) |>
  filter(Measured_Trees > 100 | Projected_Trees > 100) |>
  ggplot(aes(fct_reorder(GENUS, desc(Measured_Trees)))) +
  geom_col(
    aes(y = Measured_BA, fill = "Measured", group = "Measured"),
    width = 0.4,
    position = position_nudge(x = -0.2)
  ) +
  geom_col(
    aes(y = Projected_BA, fill = "Projected", group = "Projected"),
    width = 0.4,
    position = position_nudge(x = 0.2)
  ) +
  theme_bw() +
  #facet_wrap("name") +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Basal Area Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Genus") +
  ylab(bquote("Basal Area harvested" ~(m^2))) +
  labs(fill = NULL)

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS)
  ) |>
  filter(Measured_Trees > 100 | Projected_Trees > 100) |>
  ggplot(aes(fct_reorder(GENUS, desc(Measured_Trees)))) +
  geom_col(
    aes(y = Measured_Trees, fill = "Measured", group = "Measured"),
    width = 0.4,
    position = position_nudge(x = -0.2)
  ) +
  geom_col(
    aes(y = Projected_Trees, fill = "Projected", group = "Projected"),
    width = 0.4,
    position = position_nudge(x = 0.2)
  ) +
  theme_bw() +
  #facet_wrap("name") +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Genus") +
  ylab(bquote("Trees harvested" ~ (trees))) +
  labs(fill = NULL)
```


#### By Diameter Class


```{r, eval = FALSE}
measured_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    Year = VisitCycle - YearsSinceLastCut,
    Diameter_Class = floor(conv_unit(dbh_prior, "cm", "in") / 5) * 5,
    Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5),
    ba = pi * (conv_unit(dbh_prior, "cm", "m") / 2)^2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(!is.na(ba) & ba > 0 & StatusB == "C" | Status6 == "B") |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(ba, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

projected_harvest <- cficop_hvst_cutlist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    Diameter_Class = floor(DBH / 5) * 5,
    Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5),
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  pivot_longer(cols = c(Measured_BA, Projected_BA)) |>
  filter(!is.na(value)) |>
  group_by(GENUS) |>
  filter(sum(value) > 5) |>
  ungroup() |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Basal Area Harvested") +
  xlab("Diameter (inches)") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Basal Area harvested" ~(m^2)))

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30")
  )) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  filter(!is.na(value)) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Trees Harvested") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Diameter (inches)") +
  ylab("Trees Harvested (trees)")

```

Looks OK, but we're generally harvesting less than measured.

Let's see what size trees FVS thought were there at harvest.

#### Pre-Harvest Trees

```{r, eval = FALSE}
pre_harvest_year <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  mutate(
    HarvestYear = VisitCycle - YearsSinceLastCut,
    PreHarvestYear = floor(HarvestYear / 10) * 10
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(!is.na(dbh_prior) & dbh_prior > 0 & StatusB == "C" | Status6 == "B") |>
  filter(!is.na(PreHarvestYear)) |>
  distinct(MasterPlotID, PreHarvestYear)

measured_pre_harvest_trees <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  semi_join(
    pre_harvest_year,
    by = join_by(MasterPlotID, VisitCycle == PreHarvestYear)
  ) |>
  mutate(
    Diameter_Class = floor(conv_unit(dbh_prior, "cm", "in") / 5) * 5,
    ba = pi * (conv_unit(dbh_prior, "cm", "m") / 2)^2
  ) |>
  # Replace na's in status to keep them from propagating
  replace_na(list(StatusB = "X", Status6 = "X")) |>
  filter(
    !is.na(ba) & ba > 0 & (
      StatusB == "L" |
        StatusB == "R" |
        Status6 == "L" |
        Status6 == "R"
    )
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SpeciesCode == SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    Measured_Trees = n(),
    Measured_BA = sum(ba, na.rm = TRUE),
    .groups = "keep"
  ) |>
  ungroup()

cficop_hvst_treelist <- fvs_read_output(cficop_hvst, "FVS_TreeList")

projected_pre_harvest_trees <- cficop_hvst_cutlist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    Diameter_Class = floor(DBH / 5) * 5,
    ba = pi * (conv_unit(DBH, "in", "m") / 2)^2
  ) |>
  filter(!is.na(ba) & ba > 0 & DBH >= 5) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees. Divide by 5
    # because FVS provides a per-acre value and CFI plots are 1/5 acre.
    Projected_Trees = sum(TPA) / 5,
    # Since FVS can have fractional trees per acre, we need
    # to scale BA by the number of trees
    Projected_BA = sum(ba * TPA) / 5,
    .groups = "keep"
  ) |>
  ungroup()

measured_pre_harvest_trees |>
  full_join(
    projected_pre_harvest_trees,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  mutate(
        Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5)
  ) |>
  group_by(GENUS, Diameter_Class) |>
  summarize(
    Measured_BA = sum(Measured_BA),
    Projected_BA = sum(Projected_BA),
    .groups = "drop"
  ) |>
  ungroup() |>
  # There's one 45" maple
  filter(Diameter_Class < 30) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30", "30 - 35")
  )) |>
  pivot_longer(cols = c(Measured_BA, Projected_BA)) |>
  replace_na(list(value = 0)) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  #scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Pre-Harvest Basal Area") +
  xlab("Diameter Class") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Basal Area" ~(m^2)))


measured_pre_harvest_trees |>
  full_join(
    projected_pre_harvest_trees,
    by = join_by(GENUS, Diameter_Class)
  ) |>
  # There's one 45" maple
  filter(Diameter_Class < 30) |>
  mutate(
        Diameter_Class = paste0(Diameter_Class, " - ", Diameter_Class + 5)
  ) |>
  filter(GENUS %in% c("Acer", "Betula", "Fraxinus", "Pinus", "Quercus", "Tsuga")) |>
  mutate(Diameter_Class = ordered(
    Diameter_Class,
    c("5 - 10", "10 - 15", "15 - 20", "20 - 25", "25 - 30", "30 - 35")
  )) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  replace_na(list(value = 1)) |>
  ggplot(aes(Diameter_Class, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap(~GENUS) +
  scale_y_continuous(transform = "log10") +
  ggtitle("Quabbin CFI Pre-Harvest Trees") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  xlab("Diameter (inches)") +
  ylab("Trees (trees)")
```

#### By Year

```{r, eval = FALSE}
measured_harvest <- qryDWSPCFIPlotVisitTreeDetail |>
  cfi_with_visit_info(tblDWSPCFIPlotVisitsComplete) |>
  cfi_with_tree_info(tblDWSPCFITreesComplete) |>
  cfi_with_plot_info(tblDWSPCFIPlotsComplete) |>
  cfi_abp(cfiabp_trees) |>
  filter(StatusB == "C" | Status6 == "B") |>
  group_by(VisitCycle) |>
  summarize(
    Observed = n() * 5,
    .groups = "keep"
  ) |>
  ungroup() |>
  select(Year = VisitCycle, Observed)

projected_harvest <- cficop_hvst_cutlist |>
  mutate(Year = ceiling(Year / 10) * 10) |>
  group_by(Year) |>
  summarize(
    Projected = sum(TPA),
    .groups = "keep"
  ) |>
  ungroup()

measured_harvest |>
  full_join(
    projected_harvest,
    by = join_by(Year)
  ) |>
  pivot_longer(cols = c(Observed, Projected)) |>
  mutate(Year = sprintf("%d", Year)) |>
  ggplot(aes(Year, value, fill = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  ggtitle("Projected Vs Observed Trees Harvested") +
  ylab("Number of Trees Harvested")
```

### Measured vs. Projected Post-Harvest Trees

cficop_hvst_tpa is the tpa that we requested that FVS leave behind.

TODO: Ugh; stand 119700442 was harvested twice between 1970 and 1980. Let's
ignore it for now; probably the right thing to do is to consolidate into
a single harvest - the earlier one to give regeneration more time, or
the later one to give regeneration less time.

```{r hvst vs srvy remainders, eval = FALSE}
measured_remainders <- cficop_hvst_tpa |>
  filter(!is.na(SPCD)) |>
  left_join(
    species_crosswalk |> select(FVS_SPCD, GENUS),
    by = join_by(SPCD == FVS_SPCD)
  ) |>
  group_by(GENUS, DBH_MIN) |>
  summarize(
    Measured_Trees = sum(TPA), # TPA is already scaled to full acres
    .groups = "drop"
  )

post_harvest_inventories <- cficop_hvst_tpa |>
  mutate(YEAR = 10 * ceiling(YEAR / 10)) |>
  distinct(STAND_CN, YEAR) |>
  left_join(cfigro_plot |> distinct(STAND_CN, STAND_ID), by = join_by(STAND_CN)) |>
  select(
    StandID = STAND_ID,
    Year = YEAR
  ) |>
  mutate(StandID = as.character(StandID))

post_harvest_treelist <- fvs_read_output(cficop_hvst, "FVS_TreeList") |>
  filter(random_seed == min(random_seed)) |>
  semi_join(post_harvest_inventories, by = join_by(StandID, Year))
  
projected_remainders <- post_harvest_treelist |>
  mutate(
    SPCD = as.integer(SpeciesFIA),
    DBH_MIN = floor(DBH / 5) * 5
  ) |>
  left_join(
    species_crosswalk |> select(SPCD, GENUS, SPECIES, COMMON_NAME),
    by = join_by(SPCD)
  ) |>
  group_by(GENUS, DBH_MIN) |>
  summarize(
    # Note that FVS can have fractional trees per acre, so it's
    # important to sum TPA rather than count trees.
    Projected_Trees = sum(TPA),
    .groups = "keep"
  ) |>
  ungroup()

measured_remainders |>
  full_join(
    projected_remainders,
    by = join_by(GENUS, DBH_MIN)
  ) |>
  filter(DBH_MIN > 0) |>
  mutate(DBH_MIN = ordered(DBH_MIN)) |>
  pivot_longer(cols = c(Measured_Trees, Projected_Trees)) |>
  filter(!is.na(GENUS) & !is.na(value)) |>
  # Remove uninteresting genera
  group_by(GENUS) |>
  filter(sum(value) > 100) |>
  ungroup() |>
  ggplot(aes(DBH_MIN, value, fill = name, group = name)) +
  geom_col(position = "dodge") +
  theme_bw() +
  facet_wrap("GENUS") +
  scale_x_discrete(drop = FALSE) +
  #scale_y_continuous(transform = "log10") +
  coord_cartesian(ylim = c(0, 100)) +
  ggtitle("Quabbin CFI Post-Harvest Trees") +
  xlab("5\" Diameter Class") +
  theme(axis.text.x = element_text(angle = -30, hjust = 0, vjust = 1)) +
  ylab(bquote("Post-Harvest Trees" ~ (trees)))
```

