---
title: "FVS Carbon Modeling Analysis"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

# Setup

```{r setup, include=FALSE}
library(targets)
library(tarchetypes)
library(tidyverse)
library(reshape2) # melt()
library(measurements) # conv_unit and conv_multiunit
library(maps)
library(ggspatial) # for maps
library(ggpmisc) # for stat_poly_eq
library(rFIA)
library(pbapply)
options(scipen = 9) # Use integer notation for numbers under 9 digits
tar_source()
```


## Load targets results from other documents

```{r load tar objects defined elsewhere, eval = FALSE}
tar_load(fiadb)
tar_load(species_crosswalk)
```

## Preload targets results from this document

When tar_make() has already built the objects defined in this document,
run this block to pre-load them. You can then skip running ```{targets} blocks.

```{r load tar objects defined here, eval = FALSE}
get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```

# Grow-Only Plots in the United States

49 continental states
```{targets fiagro_states, tar_simple = TRUE}
fia_plots_filtered(fiadb, plots = NULL, \(.data, con) {
  fia_state <- tbl(con, "REF_RESEARCH_STATION") |>
    distinct(STATECD, STATE_NAME, STATE_ABBR)
  .data |>
    filter(STATECD <= 56) |> # Wyoming is 56 because there are gaps
    distinct(STATECD) |>
    left_join(fia_state, by = join_by(STATECD)) |>
    filter(STATE_NAME != "Hawaii")
})
```

Modern, forested plots that grew undisturbed for at least 10 years.
Note: this takes about 5 minutes on a 2024 MacBook.
```{targets fiagro_plot, tar_simple = TRUE}
fia_plots_filtered(
  fiadb, filter = \(.data, con) {
    .data |>
      filter(INVYR >= 1999) |> # Do this first, it impacts later filters
      filter(STATECD %in% fiagro_states$STATECD) |>
      filter_plots_modern(con) |>
      filter_plots_trees(con) |>
      filter_plots_long_measurement(con) |>
      filter_plots_forested(con) |>
      filter_plots_undisturbed(con) |>
      filter_plots_untreated(con)
  }) |>
  # Filter down to just the first and last inventory
  group_by(STATECD, COUNTYCD, PLOT) |>
  # find first and last by MEASYEAR; some INVYR are 9999, which is unhelpful
  filter(MEASYEAR == min(MEASYEAR) | MEASYEAR == max(MEASYEAR)) |>
  ungroup()
```

Gather additional statistics about the plots
```{targets fiagro_plot_stats, tar_simple = TRUE}
fia_plots_filtered(fiadb, fiagro_plot, \(.data, con) {
    plots <- .data |> distinct(STATECD, COUNTYCD, PLOT, INVYR)
    plots_join_by <- join_by(STATECD, COUNTYCD, PLOT, INVYR)

    forest_type <- tbl(con, "REF_FOREST_TYPE") |>
      select(VALUE, TYPGRPCD, MEANING) |>
      rename(
        FORTYPCD = VALUE,
        FOREST_TYPE = MEANING,
      )
    
    forest_type_group <- tbl(con, "REF_FOREST_TYPE_GROUP") |>
      select(VALUE, MEANING) |>
      rename(
        TYPGRPCD = VALUE,
        FOREST_TYPE_GROUP = MEANING
      )
    
    condprop_unadj <- tbl(con, "COND") |>
      inner_join(plots, by = plots_join_by) |>
      distinct(STATECD, COUNTYCD, PLOT, CONDID, INVYR, CONDPROP_UNADJ)
    
    tree_stats <- tbl(con, "TREE") |>
      inner_join(plots, by = plots_join_by) |>
      left_join(
        condprop_unadj,
        by = join_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR)
      ) |>
      select(STATECD, COUNTYCD, PLOT, INVYR, DIA, CARBON_AG, TPA_UNADJ, CONDPROP_UNADJ) |>
      mutate(TPA = TPA_UNADJ * CONDPROP_UNADJ) |>
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      summarize(
        CARBON_AG = sum(CARBON_AG, na.rm = TRUE),
        CPA = sum(CARBON_AG * TPA, na.rm = TRUE),
        # BA_TREES is a number of trees, not basal area.
        # Restrict to trees that match those used in QMD computation, below.
        BA_TREES = sum(if_else(DIA >= 1, TPA, 0), na.rm = TRUE),
        .groups = "keep"
      )

    cond_stats <- tbl(con, "COND") |>
      inner_join(plots, by = plots_join_by) |>
      select(STATECD, COUNTYCD, PLOT, INVYR, STDAGE, BALIVE, CONDPROP_UNADJ, FORTYPCD) |>
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      # QMD = sqrt(sum(DIA^2) / n)
      # Which is equivalent to
      # QMD = sqrt(sum(BALIVE * TPA_UNADJ) / (n * k)), where n is number of trees,
      # and k is π/576 ≅ 0.005454 for B in square feet and QMD in inches;
      # and k is π/40000 ≅ 0.0000785 for B in square meters and QMD in centimeters.
      # (see https://doi.org/10.1016/j.fecs.2023.100114 )
      # When computing QMD from BA we need to use the same trees for BA and n;
      # FIADB data dictionary 2.5.51 BALIVE says "Basal area in square feet per
      # acre of all live trees ω1.0 inch d.b.h/d.r.c sampled in the condition."
      summarize(
        BALIVE = sum(BALIVE * CONDPROP_UNADJ, na.rm = TRUE),
        FORTYPCD = max(FORTYPCD, na.rm = TRUE),
        STDAGE = max(STDAGE, na.rm = TRUE),
        .groups = "keep"
      ) |>
      ungroup()

    .data |>
      select(CN, STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR) |>
      left_join(cond_stats, by = plots_join_by) |>
      left_join(tree_stats, by = plots_join_by) |>
      left_join(forest_type, by = join_by(FORTYPCD)) |>
      left_join(forest_type_group, by = join_by(TYPGRPCD)) |>
      mutate(
        STDAGE = if_else(STDAGE < 0, NA, STDAGE)
      )
  }) |>
  filter_decode_forest_type_group() |>
  mutate(
    STAND_ID = sprintf("%04d%03d%05d", STATECD, COUNTYCD, PLOT),
    BALIVE_METRIC = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    QMD = sqrt(BALIVE / (BA_TREES * (pi / 576))),
    QMD_METRIC = sqrt(BALIVE_METRIC / (BA_TREES * (pi / 40000))),
    CARBON_METRIC = conv_multiunit(CPA, "lbs / acre", "Mg / hectare")
  ) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(
    BALIVE_START = if_else(MEASYEAR == min(MEASYEAR, na.rm = TRUE), BALIVE_METRIC, NA),
    BALIVE_DELTA = BALIVE_METRIC - max(BALIVE_START, na.rm = TRUE),
    YEARS = MEASYEAR - min(MEASYEAR, na.rm = TRUE)
  ) |> 
  ungroup()
```

```{targets fiagro_bai, tar_simple = TRUE}
prev_plot_tmp <- fiagro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == min(INVYR)) |>
  ungroup() |>
  left_join(
    fiagro_plot_stats |> select(STATECD, COUNTYCD, PLOT, INVYR, BALIVE),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  select(STATECD, COUNTYCD, PLOT, LAT, LON, ELEV, REMPER, MEASYEAR, BALIVE) |>
  rename(
    PREV_MEASYEAR = MEASYEAR,
    PREV_BALIVE = BALIVE
  )

fiagro_plot_stats |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == max(INVYR)) |>
  ungroup() |>
  left_join(prev_plot_tmp, by = join_by(STATECD, COUNTYCD, PLOT)) |>
  mutate(
    BALIVE = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    PREV_BALIVE = conv_multiunit(PREV_BALIVE, "ft2 / acre", "m2 / hectare"),
    BA_DELTA = BALIVE - PREV_BALIVE,
    BAI = BA_DELTA / (MEASYEAR - PREV_MEASYEAR)
  )  
```

```{r measured bai by latitude and forest type group, eval = FALSE}
fiagro_bai |>
  filter(BAI > quantile(fiagro_bai$BAI, 0.01) &
    BAI < quantile(fiagro_bai$BAI, 0.99)) |>
  ggplot(aes(LAT, BAI)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = y ~ x, se = FALSE) +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x) +
  theme_bw() +
  ggtitle("FIA Measured BAI by Latitude and Forest Type Group") +
  ylab(bquote("BAI" ~ (m^2 %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```

```{r measured bai by latitude, eval = FALSE}
fiagro_bai |>
  filter(
    BAI > quantile(fiagro_bai$BAI, 0.01),
    BAI < quantile(fiagro_bai$BAI, 0.99)
  ) |>
  ggplot(aes(LAT, BAI)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = y ~ x, se = FALSE, linewidth = 0.5) +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x) +
  geom_hline(yintercept = 0, linewidth = 0.25) +
  facet_wrap("FOREST_TYPE_GROUP") +
  theme_bw() +
  coord_cartesian(ylim = c(-1, 2)) +
  ggtitle("FIA Measured BAI in grow-only plots by Latitude, 1999 - 2024") +
  ylab(bquote("BAI" ~ (m^2 %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```

```{r qmd hist, eval = FALSE}
fiagro_bai |>
  filter(
    BAI > quantile(fiagro_bai$BAI, 0.01),
    BAI < quantile(fiagro_bai$BAI, 0.99)
  ) |>
  ggplot(aes(QMD)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  ggtitle("FIA Measured QMD in grow-only plots") +
  ylab("Count") +
  xlab("QMD (cm)") +
  labs(caption = "1% outliers removed")
```

```{r measured bai by latitude and qmd, eval = FALSE}
formula <- y ~ x
fiagro_bai |>
  filter(
    BAI > quantile(fiagro_bai$BAI, 0.01),
    BAI < quantile(fiagro_bai$BAI, 0.99)
  ) |>
  mutate(
    QMD_Range = cut(QMD, quantile(fiagro_bai$QMD, 0:9 / 9), include.lowest = TRUE)
  ) |>
  ggplot(aes(LAT, BAI)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = formula, se = FALSE, linewidth = 0.5) +
  stat_poly_eq(use_label("eq", "R2"), formula = formula) +
  geom_hline(yintercept = 0, linewidth = 0.25) +
  facet_wrap("QMD_Range", labeller = "label_both") +
  theme_bw() +
  coord_cartesian(ylim = c(-1, 2)) +
  ggtitle("FIA Measured BAI in grow-only plots by Latitude, 1999 - 2024") +
  ylab(bquote("BAI" ~ (m^2 %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```


```{r balive_start histogram, eval=FALSE}
fiagro_bai |>
  filter(
    BAI > quantile(fiagro_bai$BAI, 0.01),
    BAI < quantile(fiagro_bai$BAI, 0.99)
  ) |>
  ggplot(aes(PREV_BALIVE)) +
  geom_histogram(bins = 100) +
  theme_bw() +
  ggtitle("FIA Measured PREV_BALIVE in grow-only plots") +
  ylab("Count") +
  xlab("PREV_BALIVE (m2 / hectare)") +
  labs(caption = "1% outliers removed")
```


```{r measured bai by latitude and starting ba range, eval = FALSE}
formula <- y ~ x
fiagro_bai |>
  filter(
    BAI > quantile(fiagro_bai$BAI, 0.01),
    BAI < quantile(fiagro_bai$BAI, 0.99)
  ) |>
  mutate(
    BA_Range = cut(PREV_BALIVE, quantile(fiagro_bai$PREV_BALIVE, 0:10 / 10), include.lowest = TRUE)
  ) |>
  ggplot(aes(LAT, BAI)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = formula, se = FALSE, linewidth = 0.5) +
  stat_poly_eq(use_label("eq", "R2"), formula = formula) +
  geom_hline(yintercept = 0, linewidth = 0.25) +
  facet_wrap("BA_Range", ncol = 2, labeller = "label_both") +
  theme_bw() +
  coord_cartesian(ylim = c(-1, 2)) +
  ggtitle("FIA Measured BAI in grow-only plots by Latitude, 1999 - 2024") +
  ylab(bquote("BAI" ~ (m^2 %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```

# FIA Biomass

use rFIA to compute biomass
see https://doserlab.com/files/rfia/reference/biomass

Intersection of state and year(s) is an evalid: https://doserlab.com/files/rfia/reference/findevalid
Then use clipFIA to clip to a state and evalid: https://doserlab.com/files/rfia/reference/clipfia
Then use biomass to get biomass for that year: https://doserlab.com/files/rfia/reference/biomass
note: clipFIA returns a FIA.Database, which can be passed to other things

```{targets fiagro_biomass}
tar_target(
  fiagro_biomass,
  {
    do.call(
      rbind,
      pbapply(fiagro_states, 1, \(state) {
        fia_biomass_for_state(fiadb, state[["STATECD"]])
      })
    )
  },
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here:
  pattern = map(fiagro_states)
)
```

Align that with plots so we can do spatial and temporal things:
```{r align, eval = FALSE}
plots_with_carbon <- fiagro_plot |>
  left_join(fiagro_biomass, by = join_by(CN == PLT_CN, MEASYEAR == YEAR)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(all(!is.na(CARB_HECTARE))) |>
  ungroup()
prev_plot_carbon <- plots_with_carbon |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == min(INVYR)) |>
  ungroup() |>
  select(STATECD, COUNTYCD, PLOT, MEASYEAR, CARB_HECTARE, BIO_HECTARE) |>
  rename(
    PREV_MEASYEAR = MEASYEAR,
    PREV_BIO_HECTARE = BIO_HECTARE,
    PREV_CARB_HECTARE = CARB_HECTARE
  )
fia_carbon_flux <- plots_with_carbon |>
  select(STATECD, COUNTYCD, PLOT, LAT, LON, ELEV, REMPER, MEASYEAR, CARB_HECTARE, BIO_HECTARE) |>
  left_join(prev_plot_carbon, by = join_by(STATECD, COUNTYCD, PLOT)) |>
  mutate(
    BIO_FLUX = (BIO_HECTARE - PREV_BIO_HECTARE) / (MEASYEAR - PREV_MEASYEAR),
    CARB_FLUX = (CARB_HECTARE - PREV_CARB_HECTARE) / (MEASYEAR - PREV_MEASYEAR)
  ) |>
  filter(!is.na(CARB_FLUX)) |>
  left_join(
    fiagro_plot_stats |>
      group_by(STATECD, COUNTYCD, PLOT) |>
      arrange(MEASYEAR) |>
      filter(row_number() == 1) |>
      ungroup() |>
      select(STATECD, COUNTYCD, PLOT, FOREST_TYPE_GROUP, STDAGE),
    by = join_by(STATECD, COUNTYCD, PLOT)
  )
```

plot it
```{r plot carbon flux by latitude, eval = FALSE}
fia_carbon_flux |>
  filter(CARB_FLUX > quantile(fia_carbon_flux$CARB_FLUX, 0.01) &
    CARB_FLUX < quantile(fia_carbon_flux$CARB_FLUX, 0.99)) |>
  ggplot(aes(LAT, CARB_FLUX)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = y ~ x, se = FALSE) +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x) +
  facet_wrap("FOREST_TYPE_GROUP") +
  theme_bw() +
  ggtitle("rFIA computed annual carbon flux by Latitude") +
  ylab(bquote("Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```

```{r plot carbon flux by stand age, eval = FALSE}
fia_carbon_flux |>
  filter(!is.na(STDAGE)) |>
  mutate(STDAGE_RANGE = cut(STDAGE, quantile(fia_carbon_flux$STDAGE, 0:9 / 9, na.rm = TRUE), include.lowest = TRUE)) |>
  filter(CARB_FLUX > quantile(fia_carbon_flux$CARB_FLUX, 0.01) &
    CARB_FLUX < quantile(fia_carbon_flux$CARB_FLUX, 0.99)) |>
  ggplot(aes(LAT, CARB_FLUX)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = y ~ x, se = FALSE) +
  stat_poly_eq(use_label("eq", "R2"), formula = y ~ x) +
  facet_wrap("STDAGE_RANGE", labeller = "label_both") +
  theme_bw() +
  ggtitle("rFIA computed annual carbon flux by Stand Age") +
  ylab(bquote("Carbon Flux" ~ (`Mg C` %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```


# FIA TPA and BAA

It is interesting to look at how number of stems and basal area change over time.
This shows, for example, how the stand density index for a plot changes over time.

Here's the thought:
- compute BAA and TPA for each stand for each inventory.
- create a heat map - BAA and TPA, with heat the number of years spent there

we'll want to break BAA and TPA into chunks - maybe start with 50 each?

```{r tpa_for_state, eval = FALSE}
tpa_for_state <- function(fiadb, statecd) {
  con <- DBI::dbConnect(RSQLite::SQLite(), fiadb, flags = RSQLite::SQLITE_RO)
  on.exit(DBI::dbDisconnect(con), add = TRUE, after = FALSE)

  state <- tbl(con, "REF_RESEARCH_STATION") |>
    filter(STATECD == statecd) |>
    distinct(STATECD, STATE_NAME, STATE_ABBR) |>
    collect()

  state_abbr <- state[["STATE_ABBR"]]
  state_data <- rFIA::readFIA(
    con = con,
    schema = "main",
    states = state_abbr, # state abbreviation, e.g. "MA", or c("MA", "CT")
    inMemory = TRUE # findEVALID doesn't work with inMemory = FALSE
  )

  state_name <- state[["STATE_NAME"]]
  evalids <- rFIA::findEVALID(
    state_data,
    mostRecent = FALSE,
    state = state_name, # full names of states of interest
    year = 1999:2024, # Evidence suggests this wants an INVYR, not MEASYEAR
    type = NULL #  ('ALL', 'CURR', 'VOL', 'GROW', 'MORT', 'REMV', 'CHANGE', 'DWM', 'REGEN')
  )

  # Clip to inventories of interest
  # This is necessary for biomass() to return anything other than just
  # the most recent year.
  state_data_clipped <- rFIA::clipFIA(state_data, mostRecent = FALSE, evalid = evalids)

  # note that byPlot=TRUE causes the result's YEAR to be MEASYEAR
  rFIA::tpa(state_data_clipped, byPlot = TRUE) |>
    mutate(
      # TPA - estimate of mean trees per acre
      # we are converting from 1/acre to 1/hectare, which is
      # the same as converting hectares to acres.
      TPH = conv_unit(TPA, "hectare", "acre"),
      # BAA - estimate of mean basal area (sq. ft.) per acre
      BAH = conv_multiunit(BAA, "ft2 / acre", "m2 / hectare")
    )
}
```

```{targets fiagro_tpa}
tar_target(
  fiagro_tpa,
  {
    do.call(
      rbind,
      pbapply(fiagro_states, 1, \(state) {
        fia_tpa_for_state(fiadb, state[["STATECD"]])
      })
    )
  },
  iteration = "vector",
  # cross() and map() are unparsed targets:: functions here:
  pattern = map(fiagro_states)
)
```

```{r plot tph vs bah, eval = FALSE}
fiagro_tpa |>
  full_join(
    expand_grid(
      pltID = unique(fiagro_tpa$pltID),
      YEAR = unique(fiagro_tpa$YEAR)
    ),
    by = c("pltID", "YEAR")
  ) |>
  group_by(pltID) |>
  arrange(YEAR) |>
  fill(TPA, BAA, TPH, BAH, .direction = "down") |>
  ungroup() |>
  filter(!is.na(BAH) & BAH > 0, !is.na(TPH) & TPH > 0) |>
  filter(YEAR > 2001 & YEAR < 2022) |>
  ggplot(aes(BAH, TPH)) +
  geom_bin2d(bins = 25) +
  scale_fill_viridis_c() +
  scale_x_continuous(transform = "log10") +
  scale_y_continuous(transform = "log10") +
  facet_wrap("YEAR", labeller = label_both) +
  ggtitle("Heatmap of basal area and tree count") +
  ylab(bquote("Trees per hectare")) +
  xlab(bquote("Basal area" ~ (m^2 %.% hectare^-1))) +
  theme_bw()
```

Do it by stand age, instead.
```{r plot tph and bah by stand age, eval = FALSE}
tmp_tpa <- fiagro_tpa |>
  separate_wider_delim(pltID, "_", names = c("UNITCD", "STATECD", "COUNTYCD", "PLOT")) |>
  mutate(
    STATECD = as.integer(STATECD),
    UNITCD = as.integer(UNITCD),
    COUNTYCD = as.integer(COUNTYCD),
    PLOT = as.integer(PLOT)
  ) |>
  rename(MEASYEAR = YEAR) |>
  # there are somehow duplicates? remove them!
  group_by(STATECD, COUNTYCD, PLOT, MEASYEAR) |>
  filter(row_number() == 1) |>
  ungroup()

plots_with_tpa <- fiagro_plot |>
  left_join(
    tmp_tpa |> select(STATECD, COUNTYCD, PLOT, MEASYEAR, TPH, BAH),
    by = join_by(STATECD, COUNTYCD, PLOT, MEASYEAR)
  ) |>
  select(STATECD, COUNTYCD, PLOT, MEASYEAR, TPH, BAH) |>
  left_join(
    fiagro_plot_stats |>
      group_by(STATECD, COUNTYCD, PLOT) |>
      arrange(MEASYEAR) |>
      filter(row_number() == 1) |>
      ungroup() |>
      select(STATECD, COUNTYCD, PLOT, FOREST_TYPE_GROUP, STDAGE),
    by = join_by(STATECD, COUNTYCD, PLOT)
  )
```

```{r plot tpa and bah by stdage, eval=FALSE}
plots_with_tpa |>
  filter(!is.na(BAH) & !is.na(TPH)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(desc(MEASYEAR)) |>
  filter(row_number() == 1) |>
  ungroup() |>
  ggplot(aes(TPH, BAH)) +
  geom_bin2d(bins = 50) +
  theme_bw() +
  scale_fill_viridis_c(name = "Number of Plots") + # DOn't know how to get STDAGE in here
  scale_x_continuous(transform = "log10") +
  scale_y_continuous(transform = "log10") +
  ggtitle("Heatmap of basal area and tree count") +
  xlab(bquote("Stem Density" ~ (trees %.% ha^-1))) +
  ylab(bquote("Basal Area" ~ (m^2 %.% ha^-1)))
```

```{r plot tpa and bah by median stdage, eval = FALSE}
tmp_tpa <- plots_with_tpa |>
  filter(!is.na(BAH) & !is.na(TPH) & !is.na(STDAGE)) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  arrange(desc(MEASYEAR)) |>
  filter(row_number() == 1) |>
  ungroup()

tmp_tpa |>
  mutate(
    BAH_RANGE = cut(BAH, 0:20 * max(BAH) / 20, include.lowest = TRUE),
    TPH_RANGE = cut(TPH, 0:20 * max(TPH) / 20, include.lowest = TRUE, dig.lab = 5)
  ) |>
  group_by(BAH_RANGE, TPH_RANGE) |>
  summarize(MEDIAN_STDAGE = median(STDAGE), .groups = "keep") |>
  ungroup() |>
  ggplot(aes(TPH_RANGE, BAH_RANGE, color = MEDIAN_STDAGE)) +
  geom_point(size = 8, shape = "square") +
  scale_color_viridis_c(name = "Median Stand Age\n(Years)", transform = "log10") +
  theme_bw() +
  # rotate x axis labels by 30 degrees
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  ggtitle("Median Stand Age by Basal Area and Stem Density") +
  xlab(bquote("Stem Density" ~ (trees %.% ha^-1))) +
  ylab(bquote("Basal Area" ~ (m^2 %.% ha^-1)))
```

```{r plot tpa vs bah using heatmap, eval = FALSE}
tmp_tpa |>
  # scale axes logarighmicaply
  mutate(
    TPH_BIN = 10^(trunc(10 * (log10(TPH) + 0.5)) / 10),
    BAH_BIN = 10^(trunc(10 * (log10(BAH) + 0.5)) / 10)
  ) |>
  group_by(TPH_BIN, BAH_BIN) |>
  summarize(
    N = n(),
    STDAGE = quantile(STDAGE, 0.75),
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(TPH_BIN, BAH_BIN)) +
  geom_raster(aes(fill = N)) +
  scale_x_continuous(name = bquote("Stem Density" ~ (trees %.% ha^-1)), transform = "log10") +
  scale_y_continuous(name = bquote("Basal Area" ~ (m^2 %.% ha^-1)), transform = "log10") +
  scale_fill_viridis_c(name = "Number of Stands") +
  theme_bw()
```

```{r bah and tph by stand-years, eval=FALSE}
fiagro_tpa |>
  full_join(
    expand_grid(
      pltID = unique(fiagro_tpa$pltID),
      YEAR = unique(fiagro_tpa$YEAR)
    ),
    by = c("pltID", "YEAR")
  ) |>
  group_by(pltID) |>
  arrange(YEAR) |>
  fill(TPA, BAA, TPH, BAH, .direction = "down") |>
  ungroup() |>
  filter(!is.na(BAH) & BAH > 0, !is.na(TPH) & TPH > 0) |>
  filter(YEAR > 2001 & YEAR < 2022) |>
  # scale axes logarighmicaply
  mutate(
    TPH_BIN = 10^(trunc(10 * (log10(TPH) + 0.5)) / 10),
    BAH_BIN = 10^(trunc(10 * (log10(BAH) + 0.5)) / 10)
  ) |>
  group_by(TPH_BIN, BAH_BIN) |>
  summarize(
    N = n(),
    .groups = "keep"
  ) |>
  ungroup() |>
  ggplot(aes(TPH_BIN, BAH_BIN)) +
  geom_tile(aes(fill = N)) +
  scale_x_continuous(name = bquote("Stem Density" ~ (trees %.% ha^-1)), transform = "log10") +
  scale_y_continuous(name = bquote("Basal Area" ~ (m^2 %.% ha^-1)), transform = "log10") +
  scale_fill_viridis_c(name = "Stand-Years") +
  theme_bw() +
  ggtitle("Stand-Years by basal area and stem density") +
  labs(caption = "FIA Grow-Only plots 2002-2021")
```

```{r bah and tph by stand-years by year, eval=FALSE}
x <- fiagro_tpa |>
  full_join(
    expand_grid(
      pltID = unique(fiagro_tpa$pltID),
      YEAR = unique(fiagro_tpa$YEAR)
    ),
    by = c("pltID", "YEAR")
  ) |>
  group_by(pltID) |>
  arrange(YEAR) |>
  fill(TPA, BAA, TPH, BAH, .direction = "down") |>
  ungroup() |>
  filter(!is.na(BAH) & BAH > 0, !is.na(TPH) & TPH > 0) |>
  filter(YEAR > 2001 & YEAR < 2022) |>
  # scale axes logarithmically
  mutate(
    TPH_BIN = trunc(10^(trunc(10 * (log10(TPH) + 0.5)) / 10)),
    BAH_BIN = trunc(10^(trunc(10 * (log10(BAH) + 0.5)) / 10))
  ) |>
  group_by(TPH_BIN, BAH_BIN, YEAR) |>
  summarize(
    GROUP = paste0(TPH_BIN, "_", BAH_BIN),
    N = n(),
    .groups = "keep"
  ) |>
  ungroup()
```
```{r}
x |>  
  ggplot(aes(TPH_BIN, BAH_BIN, group = GROUP)) +
  geom_point(aes(color = N), shape = "square", size = 3) +
  scale_x_continuous(name = bquote("Stem Density" ~ (trees %.% ha^-1)), transform = "log10") +
  scale_y_continuous(name = bquote("Basal Area" ~ (m^2 %.% ha^-1)), transform = "log10") +
  scale_fill_viridis_c(name = "Stand-Years") +
  theme_bw() +
  ggtitle("Stand-Years by basal area and stem density in {closest_state}") +
  labs(caption = "FIA Grow-Only plots 2002-2021") +
  transition_states(YEAR, state_length = 3)
```

Can we color this by:
- From this cell, did stands in aggregate increase or decrease BAH?
- From this cell, did stands in aggregate increase or decrease TPH?

Use the color wheel for direction:
    N
 NW   NE
W       E
 SW   SE
    S

        Red
  Magenta  Orange
Purple        Yellow
  Blue     Green
       Cyan
       
ArcGIS has 8 direction flow diagrams (D8): https://pro.arcgis.com/en/pro-app/latest/tool-reference/spatial-analyst/how-flow-direction-works.htm

Their colors are:
East - Green
Southeast - Light Green
South - Light Blue
Southwest - Blue
West - Dark Blue
Northwest - Lavender
North - Red
Northeast - Orange

```{r direction_colors, eval = FALSE}
direction_colors <- tibble(
  TPH_DIR = c(
    -1, 0, 1,
    -1, 0, 1,
    -1, 0, 1
  ),
  BAH_DIR = c(
    -1, -1, -1,
    0, 0, 0,
    1, 1, 1
  ),
  COLOR = c(
    "magenta", "red", "orange",
    "purple", "lightgray", "yellow",
    "blue", "cyan", "green")
  )
direction_colors |>
  ggplot(aes(TPH_DIR, BAH_DIR, fill = COLOR)) +
  scale_fill_manual(breaks = direction_colors$COLOR, values = direction_colors$COLOR) +
  geom_raster() +
  theme_bw()
```

with white in the center

Really what I want, though, is a plot of arrows pointing in the right direction

```{r tpa with direction, eval = FALSE}
tph_bah_dir <- fiagro_tpa |>
  group_by(pltID) |>
  arrange(YEAR) |>
  mutate(
    TPH_DELTA = TPH - lag(TPH),
    BAH_DELTA = BAH - lag(BAH),
  ) |>
  ungroup() |>
  filter(!is.na(TPH_DELTA) & !is.na(BAH_DELTA)) |>
  mutate(
    TPH_BIN = 10^(trunc(10 * (log10(TPH) + 0.5)) / 10),
    BAH_BIN = 10^(trunc(10 * (log10(BAH) + 0.5)) / 10)
  ) |>
  group_by(TPH_BIN, BAH_BIN) |>
  summarize(
    TPH_DELTA = mean(TPH_DELTA),
    BAH_DELTA = mean(BAH_DELTA),
    .groups = "keep"
  ) |>
  ungroup() |>
  mutate(
    TPH_DIR = case_when(
      TPH_DELTA > 20 ~ -1,
      TPH_DELTA < -100 ~  1,
      .default = 0
    ),
    BAH_DIR = case_when(
      BAH_DELTA > 2 ~ -1,
      BAH_DELTA < -2 ~  1,
      .default = 0
    )
  ) |>
  left_join(direction_colors, by = join_by(TPH_DIR, BAH_DIR))
```

```{r}
tph_bah_dir |>
  ggplot(aes(TPH_BIN, BAH_BIN, color = COLOR)) +
  geom_point(shape = "square", size = 3.75) +
  #geom_raster() +
  scale_color_manual(breaks = direction_colors$COLOR, values = direction_colors$COLOR) +
  scale_x_continuous(name = bquote("Stem Density" ~ (trees %.% ha^-1)), transform = "log10") +
  scale_y_continuous(name = bquote("Basal Area" ~ (m^2 %.% ha^-1)), transform = "log10") +
  theme_bw()
```
