---
title: "FVS Carbon Modeling Analysis"
author: "Nikolaus Bates-Haus"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

# Setup

```{r setup, include=FALSE}
library(targets)
library(tarchetypes)
library(tidyverse)
library(reshape2) # melt()
library(measurements) # conv_unit and conv_multiunit
library(maps)
library(ggspatial) # for maps
library(ggpmisc) # for stat_poly_eq
options(scipen = 9) # Use integer notation for numbers under 9 digits
tar_source()
```


## Load targets results from other documents

```{r load tar objects defined elsewhere, eval = FALSE}
tar_load(fiadb)
tar_load(species_crosswalk)
```

## Preload targets results from this document

When tar_make() has already built the objects defined in this document,
run this block to pre-load them. You can then skip running ```{targets} blocks.

```{r load tar objects defined here, eval = FALSE}
get_this_rmd_file() |>
  tar_objects_defined_in_rmd() |>
  tar_load()
```

# Grow-Only Plots in the United States

49 continental states
```{targets fiagro_states, tar_simple = TRUE}
fia_plots_filtered(fiadb, plots = NULL, \(.data, con) {
  fia_state <- tbl(con, "REF_RESEARCH_STATION") |>
    distinct(STATECD, STATE_NAME, STATE_ABBR)
  .data |>
    filter(STATECD <= 56) |> # Wyoming is 56 because there are gaps
    distinct(STATECD) |>
    left_join(fia_state, by = join_by(STATECD)) |>
    filter(STATE_NAME != "Hawaii")
})
```

Modern, forested plots that grew undisturbed for at least 10 years.
Note: this takes about 5 minutes on a 2024 MacBook.
```{targets fiagro_plot, tar_simple = TRUE}
fia_plots_filtered(
  fiadb, filter = \(.data, con) {
    .data |>
      filter(INVYR >= 1999) |> # Do this first, it impacts later filters
      filter(STATECD %in% fiagro_states$STATECD) |>
      filter_plots_modern(con) |>
      filter_plots_trees(con) |>
      filter_plots_long_measurement(con) |>
      filter_plots_forested(con) |>
      filter_plots_undisturbed(con) |>
      filter_plots_untreated(con)
  }) |>
  # Filter down to just the first and last inventory
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == min(INVYR) | INVYR == max(INVYR)) |>
  ungroup()
```

Gather additional statistics about the plots
```{targets fiagro_plot_stats, tar_simple = TRUE}
fia_plots_filtered(fiadb, fiagro_plot, \(.data, con) {
    plots <- .data |> distinct(STATECD, COUNTYCD, PLOT, INVYR)
    plots_join_by <- join_by(STATECD, COUNTYCD, PLOT, INVYR)

    forest_type <- tbl(con, "REF_FOREST_TYPE") |>
      select(VALUE, MEANING) |>
      rename(FORTYPCD = VALUE) |>
      rename(FORTYPE = MEANING)
    
    condprop_unadj <- tbl(con, "COND") |>
      inner_join(plots, by = plots_join_by) |>
      distinct(STATECD, COUNTYCD, PLOT, CONDID, INVYR, CONDPROP_UNADJ)
    
    tree_stats <- tbl(con, "TREE") |>
      inner_join(plots, by = plots_join_by) |>
      left_join(
        condprop_unadj,
        by = join_by(STATECD, COUNTYCD, PLOT, CONDID, INVYR)
      ) |>
      select(STATECD, COUNTYCD, PLOT, INVYR, DIA, CARBON_AG, TPA_UNADJ, CONDPROP_UNADJ) |>
      mutate(TPA = TPA_UNADJ * CONDPROP_UNADJ) |>
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      summarize(
        CARBON_AG = sum(CARBON_AG, na.rm = TRUE),
        CPA = sum(CARBON_AG * TPA, na.rm = TRUE),
        # BA_TREES is a number of trees, not basal area.
        # Restrict to trees that match those used in QMD computation, below.
        BA_TREES = sum(if_else(DIA >= 1, TPA, 0), na.rm = TRUE),
        .groups = "keep"
      )

    cond_stats <- tbl(con, "COND") |>
      inner_join(plots, by = plots_join_by) |>
      select(STATECD, COUNTYCD, PLOT, INVYR, STDAGE, BALIVE, CONDPROP_UNADJ, FORTYPCD) |>
      group_by(STATECD, COUNTYCD, PLOT, INVYR) |>
      # QMD = sqrt(sum(DIA^2) / n)
      # Which is equivalent to
      # QMD = sqrt(sum(BALIVE * TPA_UNADJ) / (n * k)), where n is number of trees,
      # and k is π/576 ≅ 0.005454 for B in square feet and QMD in inches;
      # and k is π/40000 ≅ 0.0000785 for B in square meters and QMD in centimeters.
      # (see https://doi.org/10.1016/j.fecs.2023.100114 )
      # When computing QMD from BA we need to use the same trees for BA and n;
      # FIADB data dictionary 2.5.51 BALIVE says "Basal area in square feet per
      # acre of all live trees ω1.0 inch d.b.h/d.r.c sampled in the condition."
      summarize(
        BALIVE = sum(BALIVE * CONDPROP_UNADJ, na.rm = TRUE),
        FORTYPCD = max(FORTYPCD, na.rm = TRUE),
        STDAGE = max(STDAGE, na.rm = TRUE),
        .groups = "keep"
      ) |>
      ungroup()

    .data |>
      select(CN, STATECD, COUNTYCD, PLOT, INVYR, MEASYEAR) |>
      left_join(cond_stats, by = plots_join_by) |>
      left_join(tree_stats, by = plots_join_by) |>
      left_join(forest_type, by = join_by(FORTYPCD)) |>
      rename(FOREST_TYPE = FORTYPE) |>
      mutate(
        STDAGE = if_else(STDAGE < 0, NA, STDAGE),
        FRTYGRCD = floor(FORTYPCD / 10) * 10
      ) |>
      left_join(forest_type, by = join_by(FRTYGRCD == FORTYPCD)) |>
      rename(FOREST_TYPE_GROUP = FORTYPE)
  }) |>
  filter_decode_forest_type_group() |>
  mutate(
    STAND_ID = sprintf("%04d%03d%05d", STATECD, COUNTYCD, PLOT),
    BALIVE_METRIC = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    QMD = sqrt(BALIVE / (BA_TREES * (pi / 576))),
    QMD_METRIC = sqrt(BALIVE_METRIC / (BA_TREES * (pi / 40000))),
    CARBON_METRIC = conv_multiunit(CPA, "lbs / acre", "Mg / hectare")
  ) |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  mutate(
    BALIVE_START = if_else(MEASYEAR == min(MEASYEAR, na.rm = TRUE), BALIVE_METRIC, NA),
    BALIVE_DELTA = BALIVE_METRIC - max(BALIVE_START, na.rm = TRUE),
    YEARS = MEASYEAR - min(MEASYEAR, na.rm = TRUE)
  ) |> 
  ungroup()
```

```{targets fiagro_bai, tar_simple = TRUE}
prev_plot_tmp <- fiagro_plot |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == min(INVYR)) |>
  ungroup() |>
  left_join(
    fiagro_plot_stats |> select(STATECD, COUNTYCD, PLOT, INVYR, BALIVE),
    by = join_by(STATECD, COUNTYCD, PLOT, INVYR)
  ) |>
  select(STATECD, COUNTYCD, PLOT, LAT, LON, ELEV, REMPER, MEASYEAR, BALIVE) |>
  rename(
    PREV_MEASYEAR = MEASYEAR,
    PREV_BALIVE = BALIVE
  )

fiagro_plot_stats |>
  group_by(STATECD, COUNTYCD, PLOT) |>
  filter(INVYR == max(INVYR)) |>
  ungroup() |>
  left_join(prev_plot_tmp, by = join_by(STATECD, COUNTYCD, PLOT)) |>
  mutate(
    BALIVE = conv_multiunit(BALIVE, "ft2 / acre", "m2 / hectare"),
    PREV_BALIVE = conv_multiunit(PREV_BALIVE, "ft2 / acre", "m2 / hectare"),
    BA_DELTA = BALIVE - PREV_BALIVE,
    BAI = BA_DELTA / (MEASYEAR - PREV_MEASYEAR)
  )  
```

```{r measured bai by latitude and forest type group, eval = FALSE}
formula <- y ~ x

fiagro_bai |>
  filter(BAI > quantile(fiagro_bai$BAI, 0.01) &
         BAI < quantile(fiagro_bai$BAI, 0.99)) |>
  ggplot(aes(LAT, BAI)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = formula, se = FALSE) +
  stat_poly_eq(use_label("eq", "R2"), formula = formula) +
  facet_wrap("FOREST_TYPE_GROUP") +
  theme_bw() +
  ggtitle("FIA Measured BAI by Latitude and Forest Type Group") +
  ylab(bquote("BAI" ~ (m^2 %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")

```

```{r measured bai by latitude, eval = FALSE}
formula <- y ~ x

fiagro_bai |>
  filter(
    BAI > quantile(fiagro_bai$BAI, 0.01),
    BAI < quantile(fiagro_bai$BAI, 0.99)
  ) |>
  ggplot(aes(LAT, BAI)) +
  geom_point(size = 0.1) +
  stat_poly_line(formula = formula, se = FALSE, linewidth = 0.5) +
  stat_poly_eq(use_label("eq", "R2"), formula = formula) +
  geom_hline(yintercept = 0, linewidth = 0.25) +
  theme_bw() +
  coord_cartesian(ylim = c(-1, 2)) +
  ggtitle("FIA Measured BAI in grow-only plots by Latitude, 1999 - 2024") +
  ylab(bquote("BAI" ~ (m^2 %.% ha^-1 %.% yr^-1))) +
  xlab("Latitude (degrees)") +
  labs(caption = "1% outliers removed")
```
